---
format:
  html:
    toc: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{css, echo = FALSE}
p {
text-align: justify !important;
}
```

# Part 0: Corpus overview

```{r}
#| label: overview-load
library(tidyverse)
library(here)
library(jsonlite)

df <- read.csv(here("data/annotations.csv"), na.strings = "")

bib <- fromJSON(here("data/bibliography.json"), simplifyVector = FALSE)
bib_matched <- bib[names(bib) %in% df$cite_key]

full_edgelist <- read.csv(here("data/citation_edgelist.csv"), stringsAsFactors = FALSE)
n_seed_refs <- full_edgelist %>% filter(source == "lund2021") %>% nrow()
n_live <- sum(map_lgl(bib, ~ !isTRUE(.x$`_deleted`)))

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))

n_papers <- n_distinct(df$cite_key)
year_range <- range(as.numeric(as.character(df$publication_year)), na.rm = TRUE)
```

The analyzed corpus holds `r n_papers` papers, published `r year_range[1]`–`r year_range[2]`.

## From seed paper to corpus

The F1 corpus is built from the bibliography of the seed paper (Lund & Sindbæk 2022): its cited references, treated as a first generation of descendant works. The seed paper's own reference list resolves to `r n_seed_refs` works in the bibliography; the F1 corpus analyzed here holds `r n_papers`, meaning `r n_seed_refs - n_papers` references were excluded.

### Selection criteria

Two exclusion criteria were applied when building the F1 set from the seed paper's reference list:

- **Full books and edited volumes with consolidated bibliographies.** A book's reference list is not comparable in scale or character to a journal article's, and including it as a single node alongside articles would distort network measures, centrality in particular, that depend on reference-list size.
- **Methodological/technical reports outside the study's interpretive scope**, judged case by case. One paper (Cannell et al. 2016, a geochemical pXRF methods chapter in an edited excavation volume) was excluded on this basis; it remains in the wider bibliography as a citation target for the papers that cite it, but was removed from the annotated F1 seed set.

No other structured record of exclusion reasons for the remaining works in the seed's reference list currently exists; this count has not yet been individually itemized and categorized beyond these two documented criteria.

## Corpus composition

The following sections describe the corpus by publication year, entry type, authorship, venue, and language. Each has a tab for the 369-paper F1 set and a tab for the full bibliography (`r scales::comma(n_live)` live entries).

### Publications by year

::: {.panel-tabset}

#### F1

```{r}
#| label: overview-pubyears
#| fig-cap: "Papers published per year across the F1 corpus."
pubyears <- df %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", title = "Publications by year")

peak_year <- pubyears$publication_year[which.max(pubyears$n)]
peak_n <- max(pubyears$n)
```

The corpus spans `r year_range[1]` to `r year_range[2]`. Publication counts rise over that range rather than sitting flat; `r peak_year` is the single busiest year, with `r peak_n` papers.

#### Full corpus

```{r}
#| label: overview-pubyears-full
#| fig-cap: "Papers published per year across the full bibliography."
pubyears_full <- tibble(year = map_chr(bib, ~ as.character(.x$year %||% NA))) %>%
  filter(!is.na(year), year != "", str_detect(year, "^\\d{4}$")) %>%
  mutate(year = as.integer(year)) %>%
  filter(year >= 1900, year <= 2026) %>%
  count(year)

ggplot(data = pubyears_full, aes(x = year, y = n)) +
  geom_area(fill = "#5e81ac", alpha = 0.3) +
  geom_line(colour = "#5e81ac", linewidth = 0.7) +
  theme_bw() +
  scale_x_continuous(breaks = scales::breaks_pretty(n = 10)) +
  labs(x = "", y = "Count", title = "Publications by year, full bibliography")

peak_year_full <- pubyears_full$year[which.max(pubyears_full$n)]
peak_n_full <- max(pubyears_full$n)
```

The full bibliography spans `r min(pubyears_full$year)` to `r max(pubyears_full$year)`, at true year-by-year resolution; `r peak_year_full` is the single busiest year, with `r peak_n_full` entries.

:::

### Entry types

::: {.panel-tabset}

#### F1

```{r}
#| label: overview-entry-types
#| fig-cap: "Publication type across the analyzed corpus."
entry_types <- map_chr(bib_matched, ~ .x$entry_type %||% NA_character_)

entry_type_counts <- tibble(entry_type = entry_types) %>%
  count(entry_type, sort = TRUE)

ggplot(data = entry_type_counts, aes(x = fct_reorder(entry_type, n), y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  scale_y_log10(expand = expansion(mult = c(0, 0.25))) +
  theme_bw() +
  labs(x = "", y = "Count (log scale)", title = "Entry types")
```

`r nrow(entry_type_counts)` entry types appear in the corpus. The most common, `r entry_type_counts$entry_type[1]`, accounts for `r entry_type_counts$n[1]` of `r sum(entry_type_counts$n)` papers (`r scales::percent(entry_type_counts$n[1] / sum(entry_type_counts$n), accuracy = 1)`).

#### Full corpus

```{r}
#| label: overview-entry-types-full
#| fig-cap: "Entry type across the full bibliography."
entry_types_full <- map_chr(bib, ~ .x$entry_type %||% NA_character_)

entry_type_counts_full <- tibble(entry_type = entry_types_full) %>%
  filter(!is.na(entry_type), !str_detect(entry_type, "\\|")) %>%
  count(entry_type, sort = TRUE)

ggplot(data = entry_type_counts_full, aes(x = fct_reorder(entry_type, n), y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  scale_y_log10(expand = expansion(mult = c(0, 0.25))) +
  theme_bw() +
  labs(x = "", y = "Count (log scale)", title = "Entry types, full bibliography")
```

`r nrow(entry_type_counts_full)` entry types appear across the full bibliography. The most common, `r entry_type_counts_full$entry_type[1]`, accounts for `r entry_type_counts_full$n[1]` of `r sum(entry_type_counts_full$n)` entries (`r scales::percent(entry_type_counts_full$n[1] / sum(entry_type_counts_full$n), accuracy = 1)`).

:::

### Authorship

::: {.panel-tabset}

#### F1

```{r}
#| label: overview-author-counts
#| fig-cap: "Number of authors per paper across the F1 corpus."
gender_cols <- names(df)[str_starts(names(df), "Authors..Genders")]

n_authors <- df %>%
  select(cite_key, all_of(gender_cols)) %>%
  pivot_longer(-cite_key, names_to = "author_slot", values_to = "gender") %>%
  filter(!is.na(gender)) %>%
  count(cite_key, name = "n_authors")

n_authors %>%
  count(n_authors, name = "n_papers") %>%
  ggplot(aes(x = n_authors, y = n_papers)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  geom_text(aes(label = n_papers), vjust = -0.4, size = 3) +
  theme_bw() +
  scale_x_continuous(breaks = scales::breaks_pretty()) +
  labs(x = "Authors per paper", y = "Count", title = "Author count distribution")

n_single_author <- sum(n_authors$n_authors == 1)
max_authors <- max(n_authors$n_authors)
```

`r n_single_author` of `r nrow(n_authors)` papers (`r scales::percent(n_single_author / nrow(n_authors), accuracy = 1)`) have a single author. One paper has `r max_authors`, the largest author count in the corpus.

#### Full corpus

```{r}
#| label: overview-author-counts-full
#| fig-cap: "Number of authors per entry across the full bibliography, counted from the author list on each record, capped at 10 for readability, with exact counts labelled."
n_authors_full <- tibble(n_authors = map_int(bib, ~ length(.x$author %||% list())))
n_authors_full <- n_authors_full %>% filter(n_authors > 0)

axis_cap <- 10
n_over_cap <- sum(n_authors_full$n_authors > axis_cap)

n_authors_full %>%
  mutate(n_authors_capped = pmin(n_authors, axis_cap)) %>%
  count(n_authors_capped) %>%
  ggplot(aes(x = n_authors_capped, y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  geom_text(aes(label = scales::comma(n)), vjust = -0.4, size = 3) +
  scale_x_continuous(breaks = 1:axis_cap, labels = c(as.character(1:(axis_cap - 1)), paste0(axis_cap, "+"))) +
  theme_bw() +
  labs(x = "Authors per entry", y = "Count", title = "Author count distribution, full bibliography")

n_single_author_full <- sum(n_authors_full$n_authors == 1)
max_authors_full <- max(n_authors_full$n_authors)
```

This full-corpus count is taken directly from the length of each record's author list, not from the manually coded author-gender columns used for the F1 tab (those columns were only completed for the 369 F1 papers). `r n_single_author_full` of `r nrow(n_authors_full)` entries (`r scales::percent(n_single_author_full / nrow(n_authors_full), accuracy = 1)`) have a single listed author. The chart is capped at `r axis_cap` authors for readability; `r n_over_cap` entries exceed that, including the largest, with `r max_authors_full`.

:::

### Publication venues

::: {.panel-tabset}

#### F1

```{r}
#| label: overview-venues
#| fig-cap: "Journals and edited volumes with 3 or more F1 papers, by venue type."
#| fig-width: 10
#| fig-height: 7.5
venue_types <- tibble(
  journal = map_chr(bib_matched, ~ .x$journaltitle %||% NA_character_),
  volume  = map_chr(bib_matched, ~ .x$booktitle %||% NA_character_)
)

venue_long <- bind_rows(
  venue_types %>% filter(!is.na(journal)) %>% transmute(venue = journal, venue_type = "Journal"),
  venue_types %>% filter(!is.na(volume)) %>% transmute(venue = volume, venue_type = "Edited volume / book")
)

venue_counts <- venue_long %>% count(venue_type, venue, sort = TRUE)
min_shown <- 3
n_single_venue <- sum(venue_counts$n == 1)
n_with_venue <- nrow(venue_long)

venue_counts %>%
  filter(n >= min_shown) %>%
  ggplot(aes(x = fct_reorder(venue, n), y = n, fill = venue_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_bw() +
  scale_fill_manual(values = c("Journal" = "#5e81ac", "Edited volume / book" = "#a3be8c")) +
  labs(x = "", y = "Count", fill = "", title = paste0("Venues with ", min_shown, "+ papers"))
```

`r n_with_venue` of `r length(bib_matched)` papers are published in an identifiable journal or edited volume (publisher-only records, with neither, are excluded from this chart). The corpus draws on `r nrow(venue_counts)` distinct venues across both types, over half of which (`r n_single_venue`) contribute only a single paper.

#### Full corpus

```{r}
#| label: overview-venues-full
#| fig-cap: "Journals and edited volumes with 25 or more entries in the full bibliography, by venue type."
#| fig-width: 10
#| fig-height: 6.5
venue_types_full <- tibble(
  journal = map_chr(bib, ~ .x$journaltitle %||% NA_character_),
  volume  = map_chr(bib, ~ .x$booktitle %||% NA_character_)
)

venue_long_full <- bind_rows(
  venue_types_full %>% filter(!is.na(journal)) %>% transmute(venue = journal, venue_type = "Journal"),
  venue_types_full %>% filter(!is.na(volume)) %>% transmute(venue = volume, venue_type = "Edited volume / book")
)

venue_counts_full <- venue_long_full %>% count(venue_type, venue, sort = TRUE)
min_shown_full <- 25
n_single_venue_full <- sum(venue_counts_full$n == 1)
n_with_venue_full <- nrow(venue_long_full)

venue_counts_full %>%
  filter(n >= min_shown_full) %>%
  ggplot(aes(x = fct_reorder(venue, n), y = n, fill = venue_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_bw() +
  scale_fill_manual(values = c("Journal" = "#5e81ac", "Edited volume / book" = "#a3be8c")) +
  labs(x = "", y = "Count", fill = "", title = paste0("Venues with ", min_shown_full, "+ entries, full bibliography"))
```

`r n_with_venue_full` of `r scales::comma(n_live)` entries are published in an identifiable journal or edited volume; a further `r scales::comma(sum(map_lgl(bib, ~ !is.null(.x$publisher) && is.null(.x$journaltitle) && is.null(.x$booktitle))))` have only a publisher recorded, typically standalone books, and are excluded from this chart since a publisher is not a comparable unit to a journal or volume title. The full bibliography draws on `r nrow(venue_counts_full)` distinct journals and volumes, `r n_single_venue_full` of which contribute only a single entry.

:::

### Language

::: {.panel-tabset}

#### F1

```{r}
#| label: overview-language
#| fig-cap: "F1 papers by detected source-PDF language, excluding English."
lang_labels <- c(
  en = "English", sv = "Swedish", da = "Danish", nb = "Norwegian (Bokmål)",
  nn = "Norwegian (Nynorsk)", de = "German", fr = "French",
  ru = "Russian", uk = "Ukrainian", unknown = "Unknown script"
)

paper_langs <- map_chr(bib_matched, ~ .x$language %||% NA_character_)

n_missing <- sum(is.na(paper_langs))
n_unknown <- sum(paper_langs == "unknown", na.rm = TRUE)
n_classified <- sum(!is.na(paper_langs) & paper_langs != "unknown")
n_english <- sum(paper_langs == "en", na.rm = TRUE)

lang_counts <- tibble(language = paper_langs) %>%
  filter(!is.na(language)) %>%
  count(language, sort = TRUE) %>%
  mutate(pct = n / sum(n), language_label = coalesce(lang_labels[language], language))

knitr::kable(
  lang_counts %>% mutate(pct = scales::percent(pct, accuracy = 0.1)) %>% select(language_label, n, pct),
  caption = "F1 papers by detected source-PDF language, including 'Unknown script'",
  col.names = c("Language", "N", "%")
)

lang_counts %>%
  filter(language != "en") %>%
  ggplot(aes(x = fct_reorder(language_label, n), y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Count", title = "F1 papers by detected language, excluding English")
```

`r n_english` of `r n_classified` classified papers (`r scales::percent(n_english / n_classified, accuracy = 1)`) are English-language, which crowds out every other language on a shared axis, so the chart above excludes English and shows only the remaining `r n_classified - n_english + n_unknown` papers, where the real differences between languages are visible. The table gives exact counts and shares for all languages including English. `r n_unknown` papers are flagged "Unknown script" by the detector and `r n_missing` have no language recorded. The language field is recorded per paper by the extraction pipeline from its actual body text (using the `lingua` language-detection library, restricted to a known set of languages expected in this corpus: English, Norwegian, Swedish, Danish, German) and is persisted directly on each bibliography entry. "Unknown script" reflects papers written in a script the detector does not attempt to classify: confirmed for the two such cases in this corpus (Nosov 2018, Russian; Skorokhod & Blaszczyk 2020, Ukrainian), not a processing failure; both have real, correctly extracted content and citation data.

#### Full corpus

```{r}
#| label: overview-language-full
#| fig-cap: "Full bibliography by detected source-PDF language, excluding English."
paper_langs_full <- map_chr(bib, ~ .x$language %||% NA_character_)

n_classified_full <- sum(!is.na(paper_langs_full) & paper_langs_full != "unknown")
n_english_full <- sum(paper_langs_full == "en", na.rm = TRUE)
n_unclassified_full <- sum(is.na(paper_langs_full) | paper_langs_full == "unknown")

lang_counts_full <- tibble(language = paper_langs_full) %>%
  mutate(language = ifelse(is.na(language) | language == "unknown", "unclassified", language)) %>%
  count(language, sort = TRUE) %>%
  mutate(pct = n / sum(n), language_label = coalesce(lang_labels[language], str_to_sentence(language)))

knitr::kable(
  lang_counts_full %>% mutate(pct = scales::percent(pct, accuracy = 0.1)) %>% select(language_label, n, pct),
  caption = "Full bibliography by detected source-PDF language ('Unclassified' combines entries the detector flagged unknown and entries with no language recorded)",
  col.names = c("Language", "N", "%")
)

lang_counts_full %>%
  filter(language != "en") %>%
  ggplot(aes(x = fct_reorder(language_label, n), y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Count", title = "Full bibliography by detected language, excluding English")
```

`r n_english_full` of `r n_classified_full` classified entries (`r scales::percent(n_english_full / n_classified_full, accuracy = 1)`) are English-language. `r n_unclassified_full` entries are unclassified, either flagged `unknown` by the detector or carrying no language value at all.

:::

## Citation detection per F1 paper

```{r}
#| label: overview-citation-counts
#| fig-cap: "Distribution of outgoing (resolved) citation counts across the F1 corpus."
edgelist <- read.csv(here("data/citation_edgelist_annotated.csv"), stringsAsFactors = FALSE)

out_counts <- edgelist %>%
  filter(source %in% df$cite_key) %>%
  count(source, name = "n_citations")

out_counts_full <- tibble(source = df$cite_key) %>%
  left_join(out_counts, by = "source") %>%
  mutate(n_citations = replace_na(n_citations, 0))

ggplot(out_counts_full, aes(x = n_citations)) +
  geom_histogram(binwidth = 10, fill = "#5e81ac", boundary = 0) +
  theme_bw() +
  labs(x = "Resolved outgoing citations", y = "Number of F1 papers",
       title = "Distribution of citation counts per F1 paper")
```

```{r}
#| label: overview-citation-zero-check-text
n_zero_citations <- sum(out_counts_full$n_citations == 0)
zero_citation_text <- if (n_zero_citations == 0) {
  "Every F1 paper has at least one resolved outgoing citation, and the citation graph as a whole forms a single connected component."
} else {
  paste0(n_zero_citations, " F1 paper(s) have zero resolved outgoing citations: ",
         paste(out_counts_full$source[out_counts_full$n_citations == 0], collapse = ", "), ".")
}
```

`r zero_citation_text` This was not the case throughout corpus construction: extraction passes repeatedly left individual F1 papers disconnected, most often from a reference-list format (footnote-embedded citations, a shared reference section for an edited volume, non-Latin or corrupted encoding) the primary extraction method did not handle, or from the same work being independently extracted twice under two citekeys. Each case was resolved against the original source PDF or documented as a genuine structural exception.

## From detection to resolved bibliography

The citation graph is built in two stages: automated detection of candidate citations per paper, then manual correction and deduplication across the whole bibliography.

```{r}
#| label: overview-detection-methods
method_labels <- c(
  reference_list  = "Structured reference list (GROBID)",
  inline_markers   = "Inline citation markers",
  llm_body_scan    = "LLM body-text scan",
  text_patterns    = "Text-pattern matching",
  llm_bib_reparse  = "LLM bibliography reparse",
  llm_footnotes    = "LLM footnote scan"
)

detection_totals <- tibble()
merged_totals <- c()
n_with_detection <- 0
for (e in bib_matched) {
  det <- e$detection
  if (is.null(det)) next
  n_with_detection <- n_with_detection + 1
  merged_totals <- c(merged_totals, det$merged_total %||% NA)
  for (method in names(det)) {
    if (method == "merged_total") next
    detection_totals <- bind_rows(detection_totals, tibble(method = method, count = det[[method]]))
  }
}

detection_summary <- detection_totals %>%
  group_by(method) %>%
  summarise(total = sum(count), .groups = "drop") %>%
  mutate(method_label = coalesce(method_labels[method], method)) %>%
  arrange(desc(total))

raw_total <- sum(detection_summary$total)
merged_total <- sum(merged_totals, na.rm = TRUE)

knitr::kable(
  detection_summary %>% select(method_label, total),
  caption = "Raw citation detections by method, before merging duplicate detections of the same citation",
  col.names = c("Detection method", "Raw detections")
)
```

Each paper's PDF is scanned by six independent methods: GROBID parsing its structured reference list, matching inline citation markers in the body text, text-pattern matching, and three LLM-based passes (a body-text scan, a bibliography reparse, and a footnote scan). A single real citation is frequently flagged by more than one method. The `r scales::comma(raw_total)` raw detections above merge down to `r scales::comma(merged_total)` deduplicated citations per-paper, which are then carried into the bibliography-wide correction process below. The structured reference list and inline-marker matching account for the large majority of raw detections; the LLM-based methods contribute the remainder, recovering citations the structured parser misses, most commonly footnote-embedded references and citations in non-standard or non-Latin formats. 1.16% of bibliography entries (167) have no title recorded: a paper in the corpus cites them, but the entry itself has no title or an incomplete author.

Once detected, citations are carried into a single shared bibliography, where the same underlying work can be independently detected from multiple citing papers and needs consolidating into one record. This consolidation, along with other manual corrections, is tracked in a running corrections file (`corrections.yaml`) rather than as edits to the raw pipeline output.

```{r}
#| label: overview-corrections-breakdown
library(yaml)

corrections_labels <- c(
  set    = "set (field corrections: author, year, title, citer lists, etc.)",
  delete = "delete (fabricated or unresolvable ghost entries removed)",
  merge  = "merge (duplicate entries combined)",
  create = "create (missing entries added from manually verified source data)",
  reject = "reject (candidate merges/matches checked and confirmed distinct)",
  split  = "split (single entries separated into multiple genuine works)"
)

corrections <- read_yaml(here("CitationAnalysis/corrections.yaml"))
corrections_df <- tibble(
  action = map_chr(corrections, ~ .x$action %||% NA_character_),
  field  = map_chr(corrections, ~ .x$field %||% NA_character_)
)

bulk_fields <- c("language", "detection")
is_bulk <- corrections_df$action == "set" & corrections_df$field %in% bulk_fields

corrections_summary <- corrections_df %>%
  filter(!is_bulk) %>%
  count(action, name = "count") %>%
  mutate(action_label = coalesce(corrections_labels[action], action)) %>%
  select(action_label, count) %>%
  arrange(desc(count))

bulk_summary <- corrections_df %>%
  filter(is_bulk) %>%
  count(field, name = "count") %>%
  arrange(desc(count))

knitr::kable(corrections_summary, caption = "Manual, individually-reasoned correction actions applied to the bibliography (corrections.yaml)", col.names = c("Action", "Count"))
```

Two bulk, mechanical field backfills are tracked in the same file but excluded from the table above: a `language` value was set on `r scales::comma(bulk_summary$count[bulk_summary$field == "language"])` entries and a `detection` value on `r scales::comma(bulk_summary$count[bulk_summary$field == "detection"])` entries, both copied wholesale from the pipeline's own processing cache rather than individually reasoned. Combined they account for more `set` actions than every other correction in the file put together, so are reported separately to avoid making the file's manual-correction volume look larger than it is.

As of this writing the file holds `r sum(corrections_summary$count)` individually-reasoned correction actions, plus the `r scales::comma(sum(bulk_summary$count))` bulk field backfills noted above. The great majority of `delete` actions removed "ghost" entries: records with no title and no raw citation text, produced when the extraction pipeline generated a placeholder entry it was never able to resolve to a real source. `merge` actions consolidated cases where the same underlying work had been independently created under two or more citekeys, most commonly because one version came from an automated reference-list extraction with no linked PDF and a second, fuller version was later created when the paper's own PDF was located and processed directly, or because the same source was independently extracted from two different citing papers' reference lists, the cross-paper duplication this section exists to resolve. `set` actions in the table above cover routine field-level corrections: fixing fabricated or garbled author names, correcting mis-parsed years and titles, and adding real citation relationships (`cited_by` entries) that automated detection missed or failed to resolve on the first pass.

One deletion in this file was subsequently confirmed incorrect and reversed: a genuine F1 paper had been marked fabricated across multiple earlier review passes despite being a real, verifiable source, and was restored with corrected metadata once checked directly against its source PDF. This is noted here as a reminder that `delete` corrections, like any other, should be treated as provisional until checked against a primary source rather than as settled fact on the strength of a confident-sounding note.
