---
title: "BibVik — Part 0a: Corpus overview"
format:
  html:
    toc: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{css, echo = FALSE}
p {
text-align: justify !important;
}
```

# Part 0a — Corpus overview

Placeholder outline, not a finished section. Corpus size, date range,
and the publications-per-year chart, moved out of `01_annotation.qmd`.

Motivating question and methodological stance: see `00_plan.qmd`,
pending its own revision.

```{r}
#| label: overview-load
library(tidyverse)
library(here)

df <- read.csv(here("data/annotations.csv"), na.strings = "")

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))
```

## Corpus size and date range

```{r}
#| label: overview-basic-facts
n_papers <- n_distinct(df$cite_key)
year_range <- range(as.numeric(as.character(df$publication_year)), na.rm = TRUE)
cat("Papers:", n_papers, "\n")
cat("Date range:", year_range[1], "-", year_range[2], "\n")
```

## From seed paper to corpus

The F1 corpus is built from the bibliography of the seed paper (Lund &
Sindbæk 2022) — its cited references, treated as a first generation of
descendant works. The seed paper's own reference list runs to over 500
works; the F1 corpus analyzed here holds 372.

Full books and edited volumes with consolidated bibliographies were
excluded. A book's reference list is not comparable in scale or
character to a journal article's, and including it as a single node
alongside articles would distort network measures — centrality in
particular — that depend on reference-list size.

**Not yet done** — a precise, confirmed account of the full
inclusion/exclusion criteria applied to reach 372 from the 500+
starting point.

## Publications per year

```{r}
#| label: overview-pubyears
#| fig-cap: "Papers published per year across the F1 corpus."
pubyears <- df %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat = "identity", fill = "#5e81ac") +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", title = "Publications by year")
```

## Not yet done

- Placeholder outline only — holds what moved out of `01_annotation.qmd`
  (corpus size, date range, publications-per-year chart) plus a first
  pass at the seed-paper-to-corpus filtering explanation.
- No restatement of the "camps" question or no-baseline stance here;
  both live in `00_plan.qmd` only. Revisit once that file is rewritten.
- No corpus-level gender, topic, method, or source summary — those
  currently start inside `01_annotation.qmd`'s own sections.
- No region/geographic preview — currently only a word cloud at the end
  of `01_annotation.qmd`.
- No pointers forward to `01_annotation.qmd`, `02_network_structure.qmd`,
  or `03_joined_descriptions.qmd`.
