---
title: "BibVik — Part 1: Annotation descriptives (F1 corpus)"
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

Strategy and reasoning for this analysis: see `00_plan.qmd`.

This section computes the annotation cross-tab entirely independently of
the graph work above — no reference to citation structure, communities, or
centrality. Per the no-baseline stance, this is not a check against Part 1;
it stands on its own, and Part 3 (below) is where the two are placed
side by side.

```{r}
#| label: annotation-libraries
library(here)
library(tidyverse)
library(patchwork)
library(nord)
library(tidytext)
library(khroma)
library(ggh4x)
```

```{r}
#| label: annotation-palettes
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868"
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]
gender_colors <- c("Woman" = nord_palettes$red_mountain[1],
                   "Man" = nord_palettes$algoma_forest[4])
```

```{r}
#| label: annotation-load
df <- read.csv(here("data/BibVik_Data_Entry_cite_keys.csv"), na.strings = "")

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))

# Exported so 03_joined_descriptions.qmd can load it when rendered
# standalone (outside the combined index.qmd include chain), without
# recomputing Part 1 — consistent with 03's own stated design ("does
# not recompute either" upstream part).
write_csv(df, here("analysis", "bibvik_annotation_df.csv"))
```

## Publications per year

```{r}
#| label: pubyears
pubyears <- df %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat="identity", fill = nord_palettes$aurora[1]) +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE ) +
  labs(x = "", y = "Count", title = "Publications by year")
```

## Gender by first author and all authors

Publications by gender for each year. Single-author work vs. collaborations
have not yet been examined separately.

```{r}
#| label: gender-by-year
#| fig-width: 10
#| fig-height: 8
gender_pubyears <- df %>%
  group_by(Authors..Genders..Author.1.) %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

men_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Man",]) / nrow(gender_pubyears) * 100)
women_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Woman",]) / nrow(gender_pubyears) * 100)

plt1 <-
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1.)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = gender_colors) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = paste0("First author gender: Women = ", women_prcnt, "% Men = ", men_prcnt, "%")) +
  theme(legend.position = "bottom")

all_authors_gender_year <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(publication_year, gender) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na()

all_authors_gender <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(gender) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na()

men_aa <- round(all_authors_gender[all_authors_gender$gender == "Man", "n"] / sum(all_authors_gender$n) * 100)
women_aa <- round(all_authors_gender[all_authors_gender$gender == "Woman", "n"] / sum(all_authors_gender$n) * 100)
uncertain_aa <- round(all_authors_gender[all_authors_gender$gender == "Uncertain", "n"] / sum(all_authors_gender$n) * 100)

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender",
       title = paste0("All authors gender: Uncertain = ", uncertain_aa, "% Women = ", women_aa, "% Men = ", men_aa, "%")) +
  theme(legend.position = "bottom")

plt1 / plt2
```

## Multi-author gender composition and author order

The analyses above use first-author gender only (for the topic/method/source
cross-tabs later in this document) or a flattened, order-blind pool of all
author genders (the "all authors" plot). Neither captures gender
**composition per paper** (all-male vs. all-female vs. mixed) or **order**
within a paper, both of which are needed for the co-authorship questions
raised in planning (do co-authorships cross or reinforce whatever divisions
appear elsewhere; does the most senior/lead author's gender matter
independently of the full author-list composition).

**Author order is genuinely ambiguous in this corpus and must be handled
carefully, not with a single fixed rule.** Field convention varies:
first-author-as-lead is standard for excavation/fieldwork-based
publications, but last-author-as-lead (PI convention) is standard for
lab-based work. Which convention applies depends on the paper's own
method — one of the coded variables, not something inferable from the
author list alone. The lookup below (`fieldwork_lead_methods`) is a
placeholder mapping specific coded method categories to the convention
that applies; **it must be filled in with the actual method category
names from the codebook before this analysis is meaningful.** Left
unmapped, methods default to "ambiguous" and are excluded from any
lead-author-specific analysis (though still included in the
order-blind composition analysis below).

This section works generically across however many
`Authors..Genders..Author.N.` columns exist in the data — no fixed
column count is assumed.

```{r}
#| label: author-gender-cols
author_gender_cols <- df %>%
  select(starts_with("Authors..Genders..Author.")) %>%
  names()

cat("Detected", length(author_gender_cols), "author-position gender columns:\n")
print(author_gender_cols)
```

```{r}
#| label: gender-sequence
# Per-paper ordered gender sequence (e.g. "Man, Woman, Man"), preserving
# author order and author count, rather than pooling into an unordered set.
gender_sequences <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(
    cols = all_of(author_gender_cols),
    names_to = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  # Extract numeric author position from the column name for correct
  # ordering regardless of how many positions exist.
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(
    n_authors        = n(),
    gender_sequence  = paste(gender, collapse = ", "),
    first_author_gender = first(gender),
    last_author_gender  = last(gender),
    .groups = "drop"
  )

gender_sequences |> count(n_authors) |> knitr::kable(caption = "Author-count distribution")
```

### Candidate gender-composition categorizations

Several ways of collapsing a per-paper gender sequence into a single
cross-tabbable category are computed side by side below, since it is not
yet decided which is most useful. This is a case of comparing several
independent categorization schemes against the same underlying data, not
picking the "correct" one — inspect the resulting distributions before
choosing.

```{r}
#| label: gender-composition-candidates
composition_candidates <- gender_sequences %>%
  rowwise() %>%
  mutate(
    genders_list = list(str_split(gender_sequence, ", ")[[1]]),
    n_men       = sum(genders_list == "Man"),
    n_women     = sum(genders_list == "Woman"),
    n_uncertain = sum(genders_list == "Uncertain"),
  ) %>%
  ungroup() %>%
  mutate(
    # Candidate 1: strict all-same vs. mixed
    composition_strict = case_when(
      n_uncertain > 0            ~ "includes uncertain",
      n_men > 0 & n_women == 0   ~ "all-male",
      n_women > 0 & n_men == 0   ~ "all-female",
      n_men > 0 & n_women > 0    ~ "mixed",
      TRUE                       ~ NA_character_
    ),
    # Candidate 2: majority-based (ties = balanced)
    composition_majority = case_when(
      n_uncertain > 0     ~ "includes uncertain",
      n_men > n_women     ~ "majority-male",
      n_women > n_men     ~ "majority-female",
      n_men == n_women    ~ "balanced",
      TRUE                ~ NA_character_
    ),
    # Candidate 3: single-author papers split out separately, since
    # "all-male"/"all-female" for n=1 is a different phenomenon
    # (individual authorship) than for n>1 (a genuinely uniform team)
    composition_singleauthor_split = case_when(
      n_authors == 1 & n_men == 1   ~ "solo-male",
      n_authors == 1 & n_women == 1 ~ "solo-female",
      n_authors == 1                ~ "solo-uncertain",
      n_uncertain > 0                ~ "includes uncertain",
      n_men > 0 & n_women == 0       ~ "all-male (multi-author)",
      n_women > 0 & n_men == 0       ~ "all-female (multi-author)",
      TRUE                           ~ "mixed"
    )
  ) %>%
  select(-genders_list)

composition_candidates |> count(composition_strict) |> knitr::kable(caption = "Candidate 1: strict (all-same vs. mixed)")
composition_candidates |> count(composition_majority) |> knitr::kable(caption = "Candidate 2: majority-based")
composition_candidates |> count(composition_singleauthor_split) |> knitr::kable(caption = "Candidate 3: solo-authorship split out")
```

### Lead-author gender, conditioned on field convention

```{r}
#| label: lead-author-convention
# PLACEHOLDER — fill in with actual method category names from the
# codebook. Maps each coded primary method to whether first-author or
# last-author is the conventional "lead" position for that kind of work.
# Methods not listed default to "ambiguous" and are excluded from the
# lead-author-specific comparison below (they remain in every other
# analysis in this document).
fieldwork_lead_methods <- c(
  # "Excavation"        = "first",
  # "Survey"            = "first",
  # "Isotope analysis"  = "last",
  # "aDNA analysis"     = "last"
)

primary_method_lookup <- df %>%
  select(cite_key, contains("Methods")) %>%
  pivot_longer(!cite_key, names_to = "method", values_to = "value") %>%
  filter(value == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  select(cite_key, primary_method = method)

lead_author_gender <- gender_sequences %>%
  left_join(primary_method_lookup, by = "cite_key") %>%
  mutate(
    # Indexing an EMPTY named vector (fieldwork_lead_methods currently
    # has zero entries — see the placeholder above) returns an
    # unnamed NA vector, which broke replace_na() downstream and halted
    # the whole render. unname() + explicit NA_character_ fallback
    # makes this run safely while the mapping is still unfilled; every
    # row will show "ambiguous" until real method names are added
    # above, same as the placeholder's own stated behavior.
    lead_convention = if (length(fieldwork_lead_methods) > 0) {
      unname(fieldwork_lead_methods[primary_method])
    } else {
      rep(NA_character_, nrow(.))
    },
    lead_convention = replace_na(lead_convention, "ambiguous"),
    lead_author_gender = case_when(
      lead_convention == "first" ~ first_author_gender,
      lead_convention == "last"  ~ last_author_gender,
      TRUE                        ~ NA_character_
    )
  )

cat("Papers by lead convention (based on placeholder mapping above — fill in before trusting this):\n")
lead_author_gender |> count(lead_convention) |> knitr::kable()
```

### Gender composition by primary topic, method, and source of evidence

Unlike the first-author-only cross-tabs later in this document, this uses
the composition categories above — bringing multi-author, order-aware
gender data into the same kind of comparison already done for first-author
gender. `composition_strict` is used here as a starting point; swap in
`composition_majority` or `composition_singleauthor_split` to compare.

```{r}
#| label: composition-by-topic
#| fig-width: 8
composition_topic <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(count == "Primary") %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic), !is.na(composition_strict))

composition_topic %>%
  count(topic, composition_strict) %>%
  group_by(topic) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(topic = fct_reorder(topic, prop * (composition_strict == "all-female"), .fun = sum),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")")) %>%
  ggplot(aes(x = topic_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Author gender composition",
       title = "Primary topic by full-author-list gender composition")
```

```{r}
#| label: composition-by-method
#| fig-width: 8
composition_method <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(count == "Primary") %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method), !is.na(composition_strict))

composition_method %>%
  count(method, composition_strict) %>%
  group_by(method) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")")) %>%
  ggplot(aes(x = method_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Author gender composition",
       title = "Primary method by full-author-list gender composition")
```

```{r}
#| label: composition-by-source
#| fig-width: 8
composition_source <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(count == "Primary") %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source), !is.na(composition_strict))

composition_source %>%
  count(source, composition_strict) %>%
  group_by(source) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")")) %>%
  ggplot(aes(x = source_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Author gender composition",
       title = "Primary source of evidence by full-author-list gender composition")
```

**Not yet done, worth noting:** the topic/method/source cross-tabs above
use *primary* values only, same limitation as the first-author-only
cross-tabs later in this document — secondary topic/method/source values
are not yet brought into any gender comparison. Extending
`composition_topic`/`composition_method`/`composition_source` to include
secondary values would mean joining on both Primary and Secondary counts
rather than filtering to Primary only, and deciding how to weight or
display a paper that contributes to multiple topics at once.

## Primary topic, in aggregate and over time

```{r}
#| label: topics
#| fig-width: 10
#| fig-height: 6
topic_count <- df %>%
  select(publication_year, contains("Topics")) %>%
  pivot_longer(!publication_year,
               names_to = "topic",
               values_to = "count")

totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n)

primary_count_year <- topic_count %>%
  group_by(topic, publication_year) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(publication_year) %>%
  mutate(prop_primary = n / sum(n))

primary_count <- topic_count %>%
  group_by(topic) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_count_year, aes(x = publication_year, y = n, fill = topic)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count_year, aes(x = publication_year, y = prop_primary, fill = topic)) +
  geom_bar(stat="identity")+
  geom_text(data = totals_by_year,
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count |> mutate(topic = str_wrap(topic, width = 25)),
       aes(x = reorder(topic, n), y = n)) +
  geom_bar(stat="identity", fill = topic_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary topic", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("topics_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

## Primary method, in aggregate and over time

```{r}
#| label: methods
#| fig-width: 10
#| fig-height: 6
method_count <- df %>%
  select(publication_year, contains("Methods")) %>%
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")

primary_method_year <- method_count %>%
  group_by(method, publication_year) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(publication_year) %>%
  mutate(prop_primary = n / sum(n))

primary_count_method <- method_count %>%
  group_by(method) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_method_year, aes(x = publication_year, y = n, fill = method)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Count", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_method_year, aes(x = publication_year, y = prop_primary, fill = method)) +
  geom_bar(stat="identity") +
  geom_text(data = totals_by_year,
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count_method |> mutate(method = str_wrap(method, width = 25)),
       aes(x = reorder(method, n), y = n)) +
  geom_bar(stat="identity", fill = method_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary method", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("methods_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

## Primary sources of evidence, in aggregate and over time

```{r}
#| label: sources
#| fig-width: 10
#| fig-height: 6
totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n) %>%
  mutate(publication_year = factor(publication_year, levels = sort(unique(as.numeric(as.character(publication_year))))))

source_count <- df %>%
  select(publication_year, contains("Sources.of.Evidence")) %>%
  mutate(publication_year = factor(publication_year, levels = levels(totals_by_year$publication_year))) %>%
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")

primary_source_year <- source_count %>%
  group_by(source, publication_year) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(publication_year) %>%
  mutate(prop_primary = n / sum(n)) %>%
  ungroup() %>%
  mutate(publication_year = droplevels(publication_year))

primary_source_count <- source_count %>%
  group_by(source) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_source_year, aes(x = publication_year, y = n, fill = source)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 2))
ggsave("sources_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_source_year, aes(x = publication_year, y = prop_primary, fill = source)) +
  geom_bar(stat="identity")+
  geom_text(data = totals_by_year %>% filter(publication_year %in% unique(primary_source_year$publication_year)),
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("sources_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_source_count |> mutate(source = str_wrap(source, width = 25)),
       aes(x = reorder(source, n), y = n)) +
  geom_bar(stat="identity", fill = source_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary source of evidence", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("sources_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

## Secondary topics by primary topic

```{r}
#| label: topics-secondary
#| fig-width: 30
#| fig-height: 20
topics <- df %>%
  select(contains("Topics")) %>%
  rename_with(~str_remove(., "Topics.."))

coocurrences <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary topic by primary topic")
```

## Secondary methods by primary method

```{r}
#| label: methods-secondary
#| fig-width: 20
#| fig-height: 20
methods <- df %>%
  select(contains("Methods")) %>%
  rename_with(~str_remove(., "Methods.."))

coocurrences <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary method by primary method")
```

## Secondary sources of evidence by primary source of evidence

```{r}
#| label: sources-secondary
#| fig-width: 20
#| fig-height: 20
sources <- df %>%
  select(contains("Sources.of.Evidence")) %>%
  rename_with(~str_remove(., "Sources.of.Evidence.."))

coocurrences <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary source of evidence by primary source of evidence")
```

## Primary topic by first author gender

This is the first of the three plots most directly relevant to the seed
paper's tripartite (gender/topic/method) supposition. Normalizing each
topic's proportions against the corpus-wide gender base rate has been
discussed as a follow-up refinement but not yet implemented — the raw
proportions below do not yet control for the overall gender imbalance in
the corpus.

```{r}
#| label: topic-by-gender
#| fig-width: 8
topic_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Topics")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_topic_gender <- topic_count_gender %>%
  group_by(topic, gender_first_author) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(topic) %>%
  mutate(prop_primary = n / sum(n))

primary_topic_gender |>
  group_by(topic) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(topic = fct_reorder(topic, ratio),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary topic by first author gender")
```

## Primary method by first author gender

```{r}
#| label: method-by-gender
#| fig-width: 8
method_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Methods")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_method_gender <- method_count_gender %>%
  group_by(method, gender_first_author) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(method) %>%
  mutate(prop_primary = n / sum(n))

primary_method_gender |>
  group_by(method) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(method = fct_reorder(method, ratio),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary method by first author gender")
```

## Primary source of evidence by first author gender

```{r}
#| label: source-by-gender
#| fig-width: 8
source_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Sources")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_source_gender <- source_count_gender %>%
  group_by(source, gender_first_author) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(source) %>%
  mutate(prop_primary = n / sum(n))

primary_source_gender |>
  group_by(source) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(source = fct_reorder(source, ratio),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary source of evidence by first author gender")
```

## Region

```{r}
#| label: region
library(ggwordcloud)

region_counts <- df %>%
  pull(Region) %>%
  na.omit() %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Region", "n")) %>%
  mutate(n = as.numeric(n),
         n_scaled = n + 3) %>%
  arrange(desc(n))

ggplot(region_counts, aes(label = Region, size = n_scaled, color = n)) +
  geom_text_wordcloud(area_corr = TRUE, rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  scale_color_gradient(low = topic_col, high = "darkred") +
  theme_minimal()
```

---
