---
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{css, echo = FALSE}
p {
text-align: justify !important;
}
```

# Part 1 — Annotation descriptives

```{r}
#| label: annotation-libraries
library(here)
library(tidyverse)
library(patchwork)
library(nord)
library(tidytext)
library(khroma)
library(ggh4x)
```

```{r}
#| label: annotation-palettes
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868"
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]
trend_col <- "#3d3d3d"  # neutral, fixed across sections -- a rolling trend
                         # line/band is a different kind of thing from the
                         # section's own data, so it stays visually distinct
                         # from topic_col/method_col/source_col everywhere
gender_colors <- c("Woman" = nord_palettes$red_mountain[1],
                   "Man" = nord_palettes$algoma_forest[4])
```

```{r}
#| label: annotation-load
df <- read.csv(here("data/annotations.csv"), na.strings = "")

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))

# Exported so 03_joined_descriptions.qmd can load it when rendered
# standalone (outside the combined index.qmd include chain), without
# recomputing Part 1 — consistent with 03's own stated design ("does
# not recompute either" upstream part).
write_csv(df, here("analysis", "bibvik_annotation_df.csv"))
```

```{r}
#| label: annotation-shared-helpers
# Shared across every "by year" and Wilson-faded chart in this
# document (Gender, Topic, Method, Source of Evidence) -- defined once
# here so all of them draw on the same functions and the same
# per-year paper counts.
wilson_lower_bound <- function(x, n, conf = 0.95) {
  z <- qnorm(1 - (1 - conf) / 2)
  phat <- x / n
  denom <- 1 + z^2 / n
  centre <- phat + z^2 / (2 * n)
  adj <- z * sqrt(phat * (1 - phat) / n + z^2 / (4 * n^2))
  (centre - adj) / denom
}
wilson_upper_bound <- function(x, n, conf = 0.95) {
  z <- qnorm(1 - (1 - conf) / 2)
  phat <- x / n
  denom <- 1 + z^2 / n
  centre <- phat + z^2 / (2 * n)
  adj <- z * sqrt(phat * (1 - phat) / n + z^2 / (4 * n^2))
  (centre + adj) / denom
}

# Base-R centered rolling sum (window of 3, shrinking at the edges to
# whatever's available rather than requiring a full window) -- avoids
# adding a package dependency (slider) for one small operation.
roll_sum3 <- function(x) {
  n <- length(x)
  vapply(seq_len(n), function(i) sum(x[max(1, i - 1):min(n, i + 1)]), numeric(1))
}

n_per_year <- df %>% count(publication_year, name = "n_papers_year")
n_total_papers <- n_distinct(df$cite_key)
```

## Gender

Author gender, examined two ways: first-author gender and all-author
gender by year, and gender composition across the full author list per
paper.

### By year

```{r}
#| label: gender-by-year
#| fig-cap: >
#|   Top: first author only, one bar per gender per year — faint bars
#|   mean few papers published that year. Bottom: every author
#|   position pooled (order not preserved), so a multi-author paper
#|   contributes more than one tally — faint bars there mean few
#|   author-slots that year, not few papers. Same Wilson-interval fade
#|   used throughout this document, computed against the relevant
#|   denominator for each panel.
#| fig-width: 10
#| fig-height: 8
gender_pubyears <- df %>%
  count(publication_year, Authors..Genders..Author.1., name = "n") %>%
  left_join(n_per_year, by = "publication_year") %>%
  mutate(
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  )

plt1 <-
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1., alpha = -ci_width)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = gender_colors) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = "First author gender by year") +
  theme(legend.position = "bottom")

all_authors_gender_year <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na(gender) %>%
  count(publication_year, gender)

# Denominator here is total author-slots that year, not paper count --
# a multi-author paper contributes more than one gender tally, so
# pooled counts routinely exceed n_papers_year and phat = n/n_papers_year
# would exceed 1, which breaks the Wilson formula (undefined variance).
n_author_slots_per_year <- all_authors_gender_year %>%
  group_by(publication_year) %>%
  summarise(n_slots_year = sum(n), .groups = "drop")

all_authors_gender_year <- all_authors_gender_year %>%
  left_join(n_author_slots_per_year, by = "publication_year") %>%
  mutate(
    ci_lower = wilson_lower_bound(n, n_slots_year),
    ci_upper = wilson_upper_bound(n, n_slots_year),
    ci_width = ci_upper - ci_lower
  )

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender, alpha = -ci_width)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender", title = "All authors' gender by year") +
  theme(legend.position = "bottom")

plt1 / plt2
```

### Composition across the full author list

First-author gender and the order-blind pool above don't capture gender
**composition per paper** (all-male vs. all-female vs. mixed) or **order**
within a paper.

```{r}
#| label: gender-sequence
#| fig-width: 6
#| fig-height: 3
author_gender_cols <- df %>%
  select(starts_with("Authors..Genders..Author.")) %>%
  names()

# Per-paper ordered gender sequence (e.g. "Man, Woman, Man"), preserving
# author order and author count, rather than pooling into an unordered set.
gender_sequences <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(
    cols = all_of(author_gender_cols),
    names_to = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  # Extract numeric author position from the column name for correct
  # ordering regardless of how many positions exist.
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(
    n_authors        = n(),
    gender_sequence  = paste(gender, collapse = ", "),
    first_author_gender = first(gender),
    last_author_gender  = last(gender),
    .groups = "drop"
  )

gender_sequences |>
  count(n_authors) |>
  ggplot(aes(x = factor(n_authors), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Authors per paper", y = "Papers") +
  theme_bw()
```

Several ways of collapsing a per-paper gender sequence into a single
cross-tabbable category are compared below — a strict all-same-vs-mixed
split, a majority-based split, and a version that separates solo
authorship out from multi-author teams (since "all-male" means something
different for n=1 than for a genuinely uniform team).

```{r}
#| label: gender-composition-candidates
#| fig-width: 10
#| fig-height: 3.5
composition_candidates <- gender_sequences %>%
  rowwise() %>%
  mutate(
    genders_list = list(str_split(gender_sequence, ", ")[[1]]),
    n_men       = sum(genders_list == "Man"),
    n_women     = sum(genders_list == "Woman"),
    n_uncertain = sum(genders_list == "Uncertain"),
  ) %>%
  ungroup() %>%
  mutate(
    # Strict: all-same vs. mixed
    composition_strict = case_when(
      n_uncertain > 0            ~ "includes uncertain",
      n_men > 0 & n_women == 0   ~ "all-male",
      n_women > 0 & n_men == 0   ~ "all-female",
      n_men > 0 & n_women > 0    ~ "mixed",
      TRUE                       ~ NA_character_
    ),
    # Majority-based (ties = balanced)
    composition_majority = case_when(
      n_uncertain > 0     ~ "includes uncertain",
      n_men > n_women     ~ "majority-male",
      n_women > n_men     ~ "majority-female",
      n_men == n_women    ~ "balanced",
      TRUE                ~ NA_character_
    ),
    # Single-author papers split out separately, since "all-male"/
    # "all-female" for n=1 is a different phenomenon than for n>1
    composition_singleauthor_split = case_when(
      n_authors == 1 & n_men == 1   ~ "solo-male",
      n_authors == 1 & n_women == 1 ~ "solo-female",
      n_authors == 1                ~ "solo-uncertain",
      n_uncertain > 0                ~ "includes uncertain",
      n_men > 0 & n_women == 0       ~ "all-male (multi-author)",
      n_women > 0 & n_men == 0       ~ "all-female (multi-author)",
      TRUE                           ~ "mixed"
    )
  ) %>%
  select(-genders_list)

bind_rows(
  composition_candidates |> count(value = composition_strict) |> mutate(scheme = "Strict"),
  composition_candidates |> count(value = composition_majority) |> mutate(scheme = "Majority"),
  composition_candidates |> count(value = composition_singleauthor_split) |> mutate(scheme = "Solo split out")
) |>
  mutate(scheme = factor(scheme, levels = c("Strict", "Majority", "Solo split out"))) |>
  ggplot(aes(x = fct_reorder(value, n), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), hjust = -0.3, size = 3) +
  facet_wrap(vars(scheme), scales = "free", ncol = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  coord_flip() +
  labs(x = "", y = "Papers") +
  theme_bw()
```

## Topic

A paper can carry a Primary-coded topic (the coder's judgment of its main
focus) and any number of Secondary-coded topics (also present, but not
the main focus) — see the codebook for the exact scope of each topic
category. Charts below count a paper toward a topic if it carries either
value, so a paper coded with one Primary and two Secondary topics
contributes to three bars — counts and proportions are not mutually
exclusive across topics and will not sum to the corpus total. This
applies throughout this document, not just to Topic.

```{r}
#| label: topics-setup
topic_count <- df %>%
  select(publication_year, contains("Topics")) %>%
  pivot_longer(!publication_year,
               names_to = "topic",
               values_to = "count")

totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n)
```

### Totals

```{r}
#| label: topics-totals
#| fig-cap: >
#|   A paper counts toward a topic once, regardless of how many values
#|   it carries (Primary or Secondary — see above).
#| fig-width: 8
#| fig-height: 7
merged_count <- topic_count %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic) %>%
  mutate(pct = n / sum(n))

ggplot(merged_count, aes(x = reorder(str_wrap(topic, width = 25), n), y = n)) +
  geom_col(fill = topic_col) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Topic", y = "Count") +
  theme_bw() +
  coord_flip()
ggsave("topics_aggregate.png", plot = last_plot(), width = 8, height = 7, dpi = 300)
```

### Over time

```{r}
#| label: topics-over-time-setup
# Built directly from df (not topic_count above, which was pivoted
# without cite_key and can't be used to count distinct papers per topic).
topic_long_with_key <- df %>%
  select(cite_key, publication_year, contains("Topics")) %>%
  pivot_longer(cols = contains("Topics"), names_to = "topic", values_to = "count") %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " "))

topic_corpus_rate <- topic_long_with_key %>%
  distinct(cite_key, topic) %>%
  count(topic, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

topic_year_wilson <- topic_long_with_key %>%
  distinct(cite_key, publication_year, topic) %>%
  count(publication_year, topic, name = "n") %>%
  complete(publication_year, topic, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(topic_corpus_rate, by = "topic") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(topic, publication_year)

# Rolling: pool counts over a 3-year window, then compute the Wilson CI
# on the POOLED proportion directly (not by averaging per-year CIs).
# ylim(0,4) on the enrichment charts below clips extreme ratios that
# come from topics with a very low corpus_rate (e.g. Heritage)
# inflating the enrichment-scaled interval -- clipping is done
# per-panel via coord_cartesian, not via a shared alpha scale, so one
# topic's wide interval cannot distort the visual reading of every
# other topic's panel.
topic_year_roll <- topic_year_wilson %>%
  arrange(topic, publication_year) %>%
  group_by(topic) %>%
  mutate(
    roll_n = roll_sum3(n),
    roll_n_papers_year = roll_sum3(n_papers_year),
    roll_enrichment = (roll_n / roll_n_papers_year) / corpus_rate,
    roll_ci_lower = wilson_lower_bound(roll_n, roll_n_papers_year) / corpus_rate,
    roll_ci_upper = wilson_upper_bound(roll_n, roll_n_papers_year) / corpus_rate
  ) %>%
  ungroup()
```

::: {.column-margin}
**Count.** Raw papers per year, nothing normalized.

**Confidence.** Bar and band opacity come from a 95% Wilson interval —
faint means too few papers behind that estimate to trust it,
solid means there's enough. Fades continuously, no hard cutoff.
:::

```{r}
#| label: topics-over-time-count
#| fig-cap: "Papers per year, per topic. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 10
ggplot(topic_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = "#6a9870") +
  facet_wrap(~ topic, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("topics_count_by_year.png", plot = last_plot(), width = 11, height = 10, dpi = 300)
```

::: {.panel-tabset}

#### By year

```{r}
#| label: topics-over-time-enrichment
#| fig-cap: "Enrichment per year, per topic. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 10
ggplot(topic_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = topic_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ topic, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this topic's own corpus-wide rate", title = "Enrichment — by year") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("topics_enrichment_by_year.png", plot = last_plot(), width = 11, height = 10, dpi = 300)
```

#### Smoothed trend

```{r}
#| label: topics-over-time-enrichment-smoothed
#| fig-cap: "Same enrichment bars, with a 3-year rolling average and its own confidence band."
#| fig-width: 11
#| fig-height: 10
ggplot(topic_year_roll, aes(x = publication_year)) +
  geom_col(aes(y = enrichment, alpha = -ci_width), fill = topic_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  geom_ribbon(aes(ymin = pmax(roll_ci_lower, 0), ymax = pmin(roll_ci_upper, 4), group = topic),
              fill = trend_col, alpha = 0.18) +
  geom_line(aes(y = roll_enrichment, group = topic), colour = trend_col, linewidth = 0.8) +
  facet_wrap(~ topic, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this topic's own corpus-wide rate", title = "Enrichment — smoothed trend") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("topics_enrichment_by_year_smoothed.png", plot = last_plot(), width = 11, height = 10, dpi = 300)
```

:::

::: {.column-margin}
**Enrichment.** That year's coverage of the topic, divided by the
topic's own average coverage across the whole corpus (1.0x, dashed
line = normal). Each topic is checked against its own average, not
against other topics.

**Smoothed trend.** Replaces individual years with a 3-year rolling
average and its own confidence band, to make direction easier to read
through the year-to-year noise.

**Y-axis.** Capped at 4x on both tabs so the bars stay readable — a
few bars run higher than that and get cut off at the top.
:::

### Co-occurrence

```{r}
#| label: topics-cooccurrence-setup
topics <- df %>%
  select(contains("Topics")) %>%
  rename_with(~str_remove(., "Topics.."))
```

::: {.column-margin}
**Reading a facet.** Each facet is one topic; its bars are every
other topic, ranked by how often it co-occurs with the facet topic.
Each facet's bars sum to 1 on their own, so proportions aren't
comparable across facets without also checking each facet's n.
:::

```{r}
#| label: topics-cooccurrence-directional
#| fig-cap: "Directional: given a paper's Primary-coded topic (the facet), how often is each other topic marked Secondary on the same paper."
#| fig-width: 30
#| fig-height: 20
coocurrences <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

```{r}
#| label: topics-cooccurrence-undirected
#| fig-cap: "Undirected: which topics appear together at all, regardless of rank — present if either value is carried, pairing symmetric."
#| fig-width: 30
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Topic.A", "Topic.B")) %>%
  filter(Topic.A != Topic.B) %>%
  mutate(Topic.A = str_wrap(str_replace_all(Topic.A, fixed("."), " "), width = 20)) %>%
  mutate(Topic.B = str_replace_all(Topic.B, fixed("."), " ")) %>%
  group_by(Topic.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Topic.A, Prop, Topic.B), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Topic.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

### By first-author gender

```{r}
#| label: topic-by-gender
#| fig-cap: >
#|   Proportion of papers by first-author gender, within each topic.
#|   Topics are labelled with their n; horizontal lines mark 25/50/75%
#|   for reference. Raw proportions, not corrected for the corpus-wide
#|   gender split.
#| fig-width: 10
#| fig-height: 8
topic_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Topics")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

merged_topic_gender <- topic_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, gender_first_author) %>%
  group_by(topic) %>%
  mutate(prop_merged = n / sum(n))

merged_topic_gender |>
  group_by(topic) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(topic = fct_reorder(topic, ratio),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("topic_by_gender_merged.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### By full-author-list gender composition

Every paper's authors, sorted into one of five groups: solo female
author, solo male author, an all-female team, an all-male team, or a
mixed-gender team. Solo authorship is kept separate from team
composition — a single female author isn't the same thing as a team
that happens to be all women, and lumping them together would hide
that difference.

```{r}
#| label: composition-by-topic
#| fig-cap: >
#|   Proportion of papers by author-list gender composition (solo
#|   female/male, all-female/all-male team, mixed team), within each
#|   topic. Horizontal lines mark 25/50/75% for reference.
#| fig-width: 10
#| fig-height: 8
composition_topic <- composition_candidates %>%
  select(cite_key, composition_singleauthor_split) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic), !is.na(composition_singleauthor_split))

comp_topic_summary <- composition_topic %>%
  count(topic, composition_singleauthor_split) %>%
  group_by(topic) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

comp_topic_female_share <- comp_topic_summary %>%
  filter(composition_singleauthor_split %in% c("solo-female", "all-female (multi-author)")) %>%
  group_by(topic) %>%
  summarise(female_share = sum(prop), .groups = "drop")

comp_topic_summary %>%
  left_join(comp_topic_female_share, by = "topic") %>%
  mutate(female_share = replace_na(female_share, 0),
         topic = fct_reorder(topic, female_share),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, female_share),
         composition_singleauthor_split = factor(composition_singleauthor_split,
           levels = c("solo-female", "all-female (multi-author)", "mixed",
                      "all-male (multi-author)", "solo-male"))) %>%
  ggplot(aes(x = topic_label, y = prop, fill = composition_singleauthor_split)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "black", linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "solo-female" = "#bf616a", "all-female (multi-author)" = "#d89898",
    "mixed" = "#a0a8b8",
    "all-male (multi-author)" = "#8fbc8f", "solo-male" = "#4f6a31")) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("composition_by_topic.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### Mixed-gender authorship only

::: {.column-margin}
**Green dot.** Share of that topic's mixed-gender papers with a woman
first author, sized by number of mixed papers.

**Tan dots.** Every distinct team size that occurs among that topic's
mixed papers, sized by how many papers share that size — no jitter,
no averaging, so a big tan dot at one team size means many papers,
not several dots scattered near it.
:::

```{r}
#| label: mixed-only-topic
#| fig-cap: >
#|   Each topic has two overlaid series on the same row, using different
#|   x-axes. Green dots (bottom axis): share of that topic's mixed-gender
#|   papers with a woman first author, dot size is the number of mixed
#|   papers behind that topic. Tan dots (top axis): every distinct
#|   author-team size found among that topic's mixed papers, dot size is
#|   how many papers have that size. Dashed vertical line marks 50% on
#|   the green (bottom) axis only.
#| fig-width: 11.5
#| fig-height: 9.5
mixed_topic_raw <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(all_of(author_gender_cols), names_to = "position", values_to = "gender") %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(n_authors = n(), n_women = sum(gender == "Woman"), n_men = sum(gender == "Man"),
            first = first(gender), .groups = "drop") %>%
  filter(n_women > 0, n_men > 0) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic)) %>%
  mutate(is_woman_first = as.numeric(first == "Woman"))

mixed_topic_summary <- mixed_topic_raw %>%
  group_by(topic) %>%
  summarise(prop = mean(is_woman_first), total = n(), .groups = "drop") %>%
  arrange(prop) %>%
  mutate(topic_label = paste0(str_wrap(topic, 22), " (n=", total, ")"),
         topic_label = factor(topic_label, levels = topic_label))

mixed_topic_teamsize_counts <- mixed_topic_raw %>%
  count(topic, n_authors, name = "n_at_size") %>%
  left_join(mixed_topic_summary %>% select(topic, topic_label), by = "topic")

mixed_topic_size_range <- range(mixed_topic_raw$n_authors)

ggplot() +
  geom_vline(xintercept = 0.5, colour = "grey70", linetype = "dashed") +
  geom_point(data = mixed_topic_teamsize_counts,
             aes(x = (n_authors - mixed_topic_size_range[1]) / diff(mixed_topic_size_range),
                 y = topic_label, size = n_at_size),
             colour = "#c8a87a", alpha = 0.8) +
  geom_point(data = mixed_topic_summary, aes(x = prop, y = topic_label, size = total), colour = topic_col) +
  scale_x_continuous(labels = scales::percent, limits = c(-0.02, 1.02),
    sec.axis = sec_axis(~ . * diff(mixed_topic_size_range) + mixed_topic_size_range[1],
                         name = "Team size (tan)",
                         breaks = c(mixed_topic_size_range[1], 5, 10, 15, mixed_topic_size_range[2]))) +
  scale_size_continuous(range = c(1.5, 8), name = "papers at\nthat size") +
  theme_bw() +
  labs(x = "Share of mixed papers with a woman first author (green)", y = "")
ggsave("mixed_firstauthor_by_topic.png", plot = last_plot(), width = 11.5, height = 9.5, dpi = 300)
```

## Method

```{r}
#| label: methods-setup
method_count <- df %>%
  select(publication_year, contains("Methods")) %>%
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")
```

### Totals

```{r}
#| label: methods-totals
#| fig-cap: "A paper counts toward a method once, regardless of Primary/Secondary status."
#| fig-width: 8
#| fig-height: 7
merged_count_method <- method_count %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method) %>%
  mutate(pct = n / sum(n))

ggplot(merged_count_method, aes(x = reorder(str_wrap(method, width = 25), n), y = n)) +
  geom_col(fill = method_col) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Method", y = "Count") +
  theme_bw() +
  coord_flip()
ggsave("methods_aggregate.png", plot = last_plot(), width = 8, height = 7, dpi = 300)
```

### Over time

```{r}
#| label: methods-over-time-setup
method_long_with_key <- df %>%
  select(cite_key, publication_year, contains("Methods")) %>%
  pivot_longer(cols = contains("Methods"), names_to = "method", values_to = "count") %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " "))

method_corpus_rate <- method_long_with_key %>%
  distinct(cite_key, method) %>%
  count(method, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

method_year_wilson <- method_long_with_key %>%
  distinct(cite_key, publication_year, method) %>%
  count(publication_year, method, name = "n") %>%
  complete(publication_year, method, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(method_corpus_rate, by = "method") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(method, publication_year)

method_year_roll <- method_year_wilson %>%
  arrange(method, publication_year) %>%
  group_by(method) %>%
  mutate(
    roll_n = roll_sum3(n),
    roll_n_papers_year = roll_sum3(n_papers_year),
    roll_enrichment = (roll_n / roll_n_papers_year) / corpus_rate,
    roll_ci_lower = wilson_lower_bound(roll_n, roll_n_papers_year) / corpus_rate,
    roll_ci_upper = wilson_upper_bound(roll_n, roll_n_papers_year) / corpus_rate
  ) %>%
  ungroup()
```

::: {.column-margin}
**Count.** Raw papers per year, nothing normalized.

**Confidence.** Bar and band opacity come from a 95% Wilson interval —
faint means too few papers behind that estimate to trust it,
solid means there's enough. Fades continuously, no hard cutoff.
:::

```{r}
#| label: methods-over-time-count
#| fig-cap: "Papers per year, per method. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 8
ggplot(method_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = method_col) +
  facet_wrap(~ method, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("methods_count_by_year.png", plot = last_plot(), width = 11, height = 8, dpi = 300)
```

::: {.panel-tabset}

#### By year

```{r}
#| label: methods-over-time-enrichment
#| fig-cap: "Enrichment per year, per method. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 8
ggplot(method_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = method_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ method, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this method's own corpus-wide rate", title = "Enrichment — by year") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("methods_enrichment_by_year.png", plot = last_plot(), width = 11, height = 8, dpi = 300)
```

#### Smoothed trend

```{r}
#| label: methods-over-time-enrichment-smoothed
#| fig-cap: "Same enrichment bars, with a 3-year rolling average and its own confidence band."
#| fig-width: 11
#| fig-height: 8
ggplot(method_year_roll, aes(x = publication_year)) +
  geom_col(aes(y = enrichment, alpha = -ci_width), fill = method_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  geom_ribbon(aes(ymin = pmax(roll_ci_lower, 0), ymax = pmin(roll_ci_upper, 4), group = method),
              fill = trend_col, alpha = 0.18) +
  geom_line(aes(y = roll_enrichment, group = method), colour = trend_col, linewidth = 0.8) +
  facet_wrap(~ method, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this method's own corpus-wide rate", title = "Enrichment — smoothed trend") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("methods_enrichment_by_year_smoothed.png", plot = last_plot(), width = 11, height = 8, dpi = 300)
```

:::

::: {.column-margin}
**Enrichment.** That year's coverage of the method, divided by the
method's own average coverage across the whole corpus (1.0x, dashed
line = normal). Each method is checked against its own average, not
against other methods.

**Smoothed trend.** Replaces individual years with a 3-year rolling
average and its own confidence band, to make direction easier to read
through the year-to-year noise.

**Y-axis.** Capped at 4x on both tabs so the bars stay readable — a
few bars run higher than that and get cut off at the top.
:::

### Co-occurrence

```{r}
#| label: methods-cooccurrence-setup
methods <- df %>%
  select(contains("Methods")) %>%
  rename_with(~str_remove(., "Methods.."))
```

::: {.column-margin}
**Reading a facet.** Each facet is one method; its bars are every
other method, ranked by how often it co-occurs with the facet method.
Each facet's bars sum to 1 on their own, so proportions aren't
comparable across facets without also checking each facet's n.
:::

```{r}
#| label: methods-cooccurrence-directional
#| fig-cap: "Directional: given a paper's Primary-coded method (the facet), how often is each other method marked Secondary on the same paper."
#| fig-width: 20
#| fig-height: 20
coocurrences <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

```{r}
#| label: methods-cooccurrence-undirected
#| fig-cap: "Undirected: which methods appear together at all, regardless of rank — present if either value is carried, pairing symmetric."
#| fig-width: 20
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Method.A", "Method.B")) %>%
  filter(Method.A != Method.B) %>%
  mutate(Method.A = str_wrap(str_replace_all(Method.A, fixed("."), " "), width = 20)) %>%
  mutate(Method.B = str_replace_all(Method.B, fixed("."), " ")) %>%
  group_by(Method.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Method.A, Prop, Method.B), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Method.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

### By first-author gender

```{r}
#| label: method-by-gender
#| fig-cap: >
#|   Proportion of papers by first-author gender, within each method.
#|   Methods are labelled with their n; horizontal lines mark 25/50/75%
#|   for reference. Raw proportions, not corrected for the corpus-wide
#|   gender split.
#| fig-width: 10
#| fig-height: 8
method_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Methods")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

merged_method_gender <- method_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, gender_first_author) %>%
  group_by(method) %>%
  mutate(prop_merged = n / sum(n))

merged_method_gender |>
  group_by(method) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(method = fct_reorder(method, ratio),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("method_by_gender_merged.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### By full-author-list gender composition

Every paper's authors, sorted into one of five groups: solo female
author, solo male author, an all-female team, an all-male team, or a
mixed-gender team. Solo authorship is kept separate from team
composition, so a single author's gender isn't treated as if it were
a team's composition.

```{r}
#| label: composition-by-method
#| fig-cap: >
#|   Proportion of papers by author-list gender composition (solo
#|   female/male, all-female/all-male team, mixed team), within each
#|   method. Horizontal lines mark 25/50/75% for reference.
#| fig-width: 10
#| fig-height: 8
composition_method <- composition_candidates %>%
  select(cite_key, composition_singleauthor_split) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method), !is.na(composition_singleauthor_split))

comp_method_summary <- composition_method %>%
  count(method, composition_singleauthor_split) %>%
  group_by(method) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

comp_method_female_share <- comp_method_summary %>%
  filter(composition_singleauthor_split %in% c("solo-female", "all-female (multi-author)")) %>%
  group_by(method) %>%
  summarise(female_share = sum(prop), .groups = "drop")

comp_method_summary %>%
  left_join(comp_method_female_share, by = "method") %>%
  mutate(female_share = replace_na(female_share, 0),
         method = fct_reorder(method, female_share),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, female_share),
         composition_singleauthor_split = factor(composition_singleauthor_split,
           levels = c("solo-female", "all-female (multi-author)", "mixed",
                      "all-male (multi-author)", "solo-male"))) %>%
  ggplot(aes(x = method_label, y = prop, fill = composition_singleauthor_split)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "black", linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "solo-female" = "#bf616a", "all-female (multi-author)" = "#d89898",
    "mixed" = "#a0a8b8",
    "all-male (multi-author)" = "#8fbc8f", "solo-male" = "#4f6a31")) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("composition_by_method.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### Mixed-gender authorship only

::: {.column-margin}
**Navy dot.** Share of that method's mixed-gender papers with a woman
first author, sized by number of mixed papers.

**Tan dots.** Every distinct team size that occurs among that method's
mixed papers, sized by how many papers share it — no jitter, no
averaging.
:::

```{r}
#| label: mixed-only-method
#| fig-cap: >
#|   Each method has two overlaid series on the same row, using different
#|   x-axes. Navy dots (bottom axis): share of that method's mixed-gender
#|   papers with a woman first author, dot size is the number of mixed
#|   papers behind that method. Tan dots (top axis): every distinct
#|   author-team size found among that method's mixed papers, dot size is
#|   how many papers have that size. Dashed vertical line marks 50% on
#|   the navy (bottom) axis only.
#| fig-width: 11
#| fig-height: 7
mixed_method_raw <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(all_of(author_gender_cols), names_to = "position", values_to = "gender") %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(n_authors = n(), n_women = sum(gender == "Woman"), n_men = sum(gender == "Man"),
            first = first(gender), .groups = "drop") %>%
  filter(n_women > 0, n_men > 0) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method)) %>%
  mutate(is_woman_first = as.numeric(first == "Woman"))

mixed_method_summary <- mixed_method_raw %>%
  group_by(method) %>%
  summarise(prop = mean(is_woman_first), total = n(), .groups = "drop") %>%
  arrange(prop) %>%
  mutate(method_label = paste0(str_wrap(method, 22), " (n=", total, ")"),
         method_label = factor(method_label, levels = method_label))

mixed_method_teamsize_counts <- mixed_method_raw %>%
  count(method, n_authors, name = "n_at_size") %>%
  left_join(mixed_method_summary %>% select(method, method_label), by = "method")

mixed_method_size_range <- range(mixed_method_raw$n_authors)

ggplot() +
  geom_vline(xintercept = 0.5, colour = "grey70", linetype = "dashed") +
  geom_point(data = mixed_method_teamsize_counts,
             aes(x = (n_authors - mixed_method_size_range[1]) / diff(mixed_method_size_range),
                 y = method_label, size = n_at_size),
             colour = "#c8a87a", alpha = 0.8) +
  geom_point(data = mixed_method_summary, aes(x = prop, y = method_label, size = total), colour = method_col) +
  scale_x_continuous(labels = scales::percent, limits = c(-0.02, 1.02),
    sec.axis = sec_axis(~ . * diff(mixed_method_size_range) + mixed_method_size_range[1],
                         name = "Team size (tan)",
                         breaks = c(mixed_method_size_range[1], 5, 10, 15, mixed_method_size_range[2]))) +
  scale_size_continuous(range = c(1.5, 8), name = "papers at\nthat size") +
  theme_bw() +
  labs(x = "Share of mixed papers with a woman first author (navy)", y = "")
ggsave("mixed_firstauthor_by_method.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

## Source of evidence

```{r}
#| label: sources-setup
source_count <- df %>%
  select(publication_year, contains("Sources.of.Evidence")) %>%
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")
```

### Totals

```{r}
#| label: sources-totals
#| fig-cap: "A paper counts toward a source once, regardless of Primary/Secondary status."
#| fig-width: 8
#| fig-height: 7
merged_source_count <- source_count %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source) %>%
  mutate(pct = n / sum(n))

ggplot(merged_source_count, aes(x = reorder(str_wrap(source, width = 25), n), y = n)) +
  geom_col(fill = source_col) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Source of evidence", y = "Count") +
  theme_bw() +
  coord_flip()
ggsave("sources_aggregate.png", plot = last_plot(), width = 8, height = 7, dpi = 300)
```

### Over time

```{r}
#| label: sources-over-time-setup
source_long_with_key <- df %>%
  select(cite_key, publication_year, contains("Sources.of.Evidence")) %>%
  pivot_longer(cols = contains("Sources.of.Evidence"), names_to = "source", values_to = "count") %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " "))

source_corpus_rate <- source_long_with_key %>%
  distinct(cite_key, source) %>%
  count(source, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

source_year_wilson <- source_long_with_key %>%
  distinct(cite_key, publication_year, source) %>%
  count(publication_year, source, name = "n") %>%
  complete(publication_year, source, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(source_corpus_rate, by = "source") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(source, publication_year)

source_year_roll <- source_year_wilson %>%
  arrange(source, publication_year) %>%
  group_by(source) %>%
  mutate(
    roll_n = roll_sum3(n),
    roll_n_papers_year = roll_sum3(n_papers_year),
    roll_enrichment = (roll_n / roll_n_papers_year) / corpus_rate,
    roll_ci_lower = wilson_lower_bound(roll_n, roll_n_papers_year) / corpus_rate,
    roll_ci_upper = wilson_upper_bound(roll_n, roll_n_papers_year) / corpus_rate
  ) %>%
  ungroup()
```

::: {.column-margin}
**Count.** Raw papers per year, nothing normalized.

**Confidence.** Bar and band opacity come from a 95% Wilson interval —
faint means too few papers behind that estimate to trust it,
solid means there's enough. Fades continuously, no hard cutoff.
:::

```{r}
#| label: sources-over-time-count
#| fig-cap: "Papers per year, per source. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 7
ggplot(source_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = source_col) +
  facet_wrap(~ source, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("sources_count_by_year.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

::: {.panel-tabset}

#### By year

```{r}
#| label: sources-over-time-enrichment
#| fig-cap: "Enrichment per year, per source. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 7
ggplot(source_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = source_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ source, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this source's own corpus-wide rate", title = "Enrichment — by year") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("sources_enrichment_by_year.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

#### Smoothed trend

```{r}
#| label: sources-over-time-enrichment-smoothed
#| fig-cap: "Same enrichment bars, with a 3-year rolling average and its own confidence band."
#| fig-width: 11
#| fig-height: 7
ggplot(source_year_roll, aes(x = publication_year)) +
  geom_col(aes(y = enrichment, alpha = -ci_width), fill = source_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  geom_ribbon(aes(ymin = pmax(roll_ci_lower, 0), ymax = pmin(roll_ci_upper, 4), group = source),
              fill = trend_col, alpha = 0.18) +
  geom_line(aes(y = roll_enrichment, group = source), colour = trend_col, linewidth = 0.8) +
  facet_wrap(~ source, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this source's own corpus-wide rate", title = "Enrichment — smoothed trend") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("sources_enrichment_by_year_smoothed.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

:::

::: {.column-margin}
**Enrichment.** That year's coverage of the source, divided by the
source's own average coverage across the whole corpus (1.0x, dashed
line = normal). Each source is checked against its own average, not
against other sources.

**Smoothed trend.** Replaces individual years with a 3-year rolling
average and its own confidence band, to make direction easier to read
through the year-to-year noise.

**Y-axis.** Capped at 4x on both tabs so the bars stay readable — a
few bars run higher than that and get cut off at the top.
:::

### Co-occurrence

```{r}
#| label: sources-cooccurrence-setup
sources <- df %>%
  select(contains("Sources.of.Evidence")) %>%
  rename_with(~str_remove(., "Sources.of.Evidence.."))
```

::: {.column-margin}
**Reading a facet.** Each facet is one source; its bars are every
other source, ranked by how often it co-occurs with the facet source.
Each facet's bars sum to 1 on their own, so proportions aren't
comparable across facets without also checking each facet's n.
:::

```{r}
#| label: sources-cooccurrence-directional
#| fig-cap: "Directional: given a paper's Primary-coded source (the facet), how often is each other source marked Secondary on the same paper."
#| fig-width: 20
#| fig-height: 20
coocurrences <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

```{r}
#| label: sources-cooccurrence-undirected
#| fig-cap: "Undirected: which sources appear together at all, regardless of rank — present if either value is carried, pairing symmetric."
#| fig-width: 20
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Source.A", "Source.B")) %>%
  filter(Source.A != Source.B) %>%
  mutate(Source.A = str_wrap(str_replace_all(Source.A, fixed("."), " "), width = 20)) %>%
  mutate(Source.B = str_replace_all(Source.B, fixed("."), " ")) %>%
  group_by(Source.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Source.A, Prop, Source.B), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Source.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

### By first-author gender

```{r}
#| label: source-by-gender
#| fig-cap: >
#|   Proportion of papers by first-author gender, within each source.
#|   Sources are labelled with their n; horizontal lines mark 25/50/75%
#|   for reference. Raw proportions, not corrected for the corpus-wide
#|   gender split.
#| fig-width: 10
#| fig-height: 8
source_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Sources")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

merged_source_gender <- source_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, gender_first_author) %>%
  group_by(source) %>%
  mutate(prop_merged = n / sum(n))

merged_source_gender |>
  group_by(source) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(source = fct_reorder(source, ratio),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("source_by_gender_merged.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### By full-author-list gender composition

Every paper's authors, sorted into one of five groups: solo female
author, solo male author, an all-female team, an all-male team, or a
mixed-gender team. Solo authorship is kept separate from team
composition, so a single author's gender isn't treated as if it were
a team's composition.

```{r}
#| label: composition-by-source
#| fig-cap: >
#|   Proportion of papers by author-list gender composition (solo
#|   female/male, all-female/all-male team, mixed team), within each
#|   source of evidence. Horizontal lines mark 25/50/75% for reference.
#| fig-width: 10
#| fig-height: 8
composition_source <- composition_candidates %>%
  select(cite_key, composition_singleauthor_split) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source), !is.na(composition_singleauthor_split))

comp_source_summary <- composition_source %>%
  count(source, composition_singleauthor_split) %>%
  group_by(source) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

comp_source_female_share <- comp_source_summary %>%
  filter(composition_singleauthor_split %in% c("solo-female", "all-female (multi-author)")) %>%
  group_by(source) %>%
  summarise(female_share = sum(prop), .groups = "drop")

comp_source_summary %>%
  left_join(comp_source_female_share, by = "source") %>%
  mutate(female_share = replace_na(female_share, 0),
         source = fct_reorder(source, female_share),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, female_share),
         composition_singleauthor_split = factor(composition_singleauthor_split,
           levels = c("solo-female", "all-female (multi-author)", "mixed",
                      "all-male (multi-author)", "solo-male"))) %>%
  ggplot(aes(x = source_label, y = prop, fill = composition_singleauthor_split)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "black", linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "solo-female" = "#bf616a", "all-female (multi-author)" = "#d89898",
    "mixed" = "#a0a8b8",
    "all-male (multi-author)" = "#8fbc8f", "solo-male" = "#4f6a31")) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("composition_by_source.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### Mixed-gender authorship only

::: {.column-margin}
**Red dot.** Share of that source's mixed-gender papers with a woman
first author, sized by number of mixed papers.

**Tan dots.** Every distinct team size that occurs among that source's
mixed papers, sized by how many papers share it — no jitter, no
averaging.
:::

```{r}
#| label: mixed-only-source
#| fig-cap: >
#|   Each source of evidence has two overlaid series on the same row,
#|   using different x-axes. Red dots (bottom axis): share of that
#|   source's mixed-gender papers with a woman first author, dot size is
#|   the number of mixed papers behind that source. Tan dots (top axis):
#|   every distinct author-team size found among that source's mixed
#|   papers, dot size is how many papers have that size. Dashed vertical
#|   line marks 50% on the red (bottom) axis only.
#| fig-width: 11
#| fig-height: 6.5
mixed_source_raw <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(all_of(author_gender_cols), names_to = "position", values_to = "gender") %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(n_authors = n(), n_women = sum(gender == "Woman"), n_men = sum(gender == "Man"),
            first = first(gender), .groups = "drop") %>%
  filter(n_women > 0, n_men > 0) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source)) %>%
  mutate(is_woman_first = as.numeric(first == "Woman"))

mixed_source_summary <- mixed_source_raw %>%
  group_by(source) %>%
  summarise(prop = mean(is_woman_first), total = n(), .groups = "drop") %>%
  arrange(prop) %>%
  mutate(source_label = paste0(str_wrap(source, 22), " (n=", total, ")"),
         source_label = factor(source_label, levels = source_label))

mixed_source_teamsize_counts <- mixed_source_raw %>%
  count(source, n_authors, name = "n_at_size") %>%
  left_join(mixed_source_summary %>% select(source, source_label), by = "source")

mixed_source_size_range <- range(mixed_source_raw$n_authors)

ggplot() +
  geom_vline(xintercept = 0.5, colour = "grey70", linetype = "dashed") +
  geom_point(data = mixed_source_teamsize_counts,
             aes(x = (n_authors - mixed_source_size_range[1]) / diff(mixed_source_size_range),
                 y = source_label, size = n_at_size),
             colour = "#c8a87a", alpha = 0.8) +
  geom_point(data = mixed_source_summary, aes(x = prop, y = source_label, size = total), colour = source_col) +
  scale_x_continuous(labels = scales::percent, limits = c(-0.02, 1.02),
    sec.axis = sec_axis(~ . * diff(mixed_source_size_range) + mixed_source_size_range[1],
                         name = "Team size (tan)",
                         breaks = c(mixed_source_size_range[1], 5, 10, 15, mixed_source_size_range[2]))) +
  scale_size_continuous(range = c(1.5, 8), name = "papers at\nthat size") +
  theme_bw() +
  labs(x = "Share of mixed papers with a woman first author (red)", y = "")
ggsave("mixed_firstauthor_by_source.png", plot = last_plot(), width = 11, height = 6.5, dpi = 300)
```

## Region

```{r}
#| label: region
library(ggwordcloud)

region_counts <- df %>%
  pull(Region) %>%
  na.omit() %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Region", "n")) %>%
  mutate(n = as.numeric(n),
         n_scaled = n + 3) %>%
  arrange(desc(n))

ggplot(region_counts, aes(label = Region, size = n_scaled, color = n)) +
  geom_text_wordcloud(area_corr = TRUE, rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  scale_color_gradient(low = topic_col, high = "darkred") +
  theme_minimal()
```

### Matching to Wikidata

Region is free text drawn directly from each paper, not a controlled
vocabulary: spelling variants and mixed specificity (continent, country,
named site) count as distinct terms in the word cloud above. Wikidata is
used to consolidate these to real places with coordinates. Pleiades,
Pelagios/Peripleo, and iDAI.gazetteer were considered and set aside: all
three are institutionally anchored in Mediterranean/classical-world
antiquity, with no confirmed non-trivial coverage of Viking-Age
Scandinavia. Wikidata was used instead for its dense, well-curated
coverage of Scandinavian archaeological sites and structured multilingual
labels/aliases, which handle spelling variants (Danmark/Denmark) close to
for free.

The chunks below call Wikidata's SPARQL endpoint and so require outbound
network access; they default to `eval: false` and do not run on a normal
render of this document. Run them explicitly (flip `eval: true`) when the
lookup table needs building or refreshing, then leave them off again —
the rest of this section reads the resulting static CSV, same as Topic,
Method, and Source read their own derived data.

```{r}
#| label: region-tokens
# One row per (cite_key, region_token) pair, trimmed, blanks dropped.
region_tokens <- df %>%
  select(cite_key, Region) %>%
  filter(!is.na(Region), Region != "") %>%
  separate_rows(Region, sep = ",") %>%
  mutate(Region = str_trim(Region)) %>%
  filter(Region != "") %>%
  # A trailing period on an entry ("Mälaren.") is a citation-formatting
  # artifact, not part of the place name — strip it so it doesn't
  # produce a separate, unmatched token alongside the real one.
  mutate(Region = str_remove(Region, "\\.$"))

# Known spelling/language variants, folded to one canonical form before
# matching. Add to this list as more are found during review — this is
# not exhaustive, just the duplicates visible in the corpus-wide token
# frequency table.
variant_map <- c(
  "Danmark" = "Denmark"
)

region_tokens <- region_tokens %>%
  mutate(Region_normalized = coalesce(variant_map[Region], Region))

unique_tokens <- region_tokens %>%
  distinct(Region_normalized) %>%
  arrange(Region_normalized) %>%
  pull(Region_normalized)
```

**Coverage probe.** Before matching all tokens, check hit rate on the
most frequent ones — this is the actual empirical check for whether
Wikidata is workable here, not an assumption.

```{r}
#| label: region-probe
#| eval: false
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  install.packages("WikidataQueryServiceR", repos = "https://cloud.r-project.org")
}
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  # CRAN install failed (blocked mirror, network issue, etc.) — fall back
  # to installing directly from the package's GitHub source, which
  # bypasses CRAN entirely.
  if (!requireNamespace("remotes", quietly = TRUE)) {
    install.packages("remotes", repos = "https://cloud.r-project.org")
  }
  remotes::install_github("bearloga/WikidataQueryServiceR")
}
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  stop(
    "WikidataQueryServiceR failed to install from both CRAN and GitHub. ",
    "This points to an environment issue rather than a package issue — most ",
    "likely no outbound network access at all from this machine. Run ",
    "install.packages(\"WikidataQueryServiceR\", repos = \"https://cloud.r-project.org\") ",
    "directly in the console (outside this render) to see the real underlying error."
  )
}
library(WikidataQueryServiceR)
library(httr)

top_tokens <- region_tokens %>%
  count(Region_normalized, sort = TRUE) %>%
  slice_max(n, n = 40, with_ties = FALSE) %>%
  pull(Region_normalized)

# SPARQL: label or alias match (English, Swedish, Danish, Norwegian,
# Icelandic — the languages a Scandinavian place name is most likely to
# appear under in Wikidata), filtered to place-like classes: archaeological
# site (Q839954), human settlement (Q486972), instance/subclass of
# location (Q17334923) generally, island (Q23442), geographic region
# (Q82794), and tumulus/burial mound (Q34023). The tumulus class was added
# after an initial probe run missed Gokstad and Oseberg — both real,
# well-known Viking Age ship burial sites — because Wikidata classifies
# them as tumuli (Gokstad's actual item is "Gokstad Mound"), not as
# archaeological sites or settlements under the original filter.
# Deliberately broad on class, narrow on label match — better to
# over-fetch candidates and filter in R than to miss a real match by
# being too restrictive in SPARQL.
probe_query_template <- '
SELECT ?item ?itemLabel ?coord ?instanceOfLabel WHERE {
  VALUES ?searchterm { "%s"@en "%s"@sv "%s"@da "%s"@nb "%s"@is }
  ?item rdfs:label|skos:altLabel ?searchterm .
  ?item wdt:P31/wdt:P279* ?instanceOf .
  VALUES ?instanceOf { wd:Q839954 wd:Q486972 wd:Q17334923 wd:Q23442 wd:Q82794 wd:Q34023 }
  OPTIONAL { ?item wdt:P625 ?coord . }
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}
LIMIT 5
'

# Fallback for tokens with no exact label/alias match: a CONTAINS-style
# substring match against English labels only. Slower (a full label scan,
# not an index lookup) and lower-precision, so it only runs on tokens the
# exact-match query already missed, and results from this path are
# flagged separately rather than treated as equally confident. This is
# what actually catches a case like "Gokstad": the real Wikidata item is
# titled "Gokstad Mound", so an exact match on the bare token "Gokstad"
# never hits, even with the right class included.
#
# A SPARQL CONTAINS scan was tried first and confirmed genuinely too slow
# at Wikidata's scale, timing out even after restructuring the query to
# apply the class filter before the string filter — SPARQL's CONTAINS
# does an unindexed substring scan, and is documented as not built for
# this kind of fuzzy lookup. Wikidata's own wbsearchentities REST API is
# the purpose-built tool for partial/fuzzy label search (it is what
# Wikidata's own search box uses), genuinely fast because it is backed
# by a real search index. It cannot filter by entity class directly, so
# the class check happens as a second step on each candidate it returns.
search_entities <- function(token, limit = 5) {
  res <- tryCatch(
    httr::GET(
      "https://www.wikidata.org/w/api.php",
      query = list(
        action = "wbsearchentities", search = token, language = "en",
        type = "item", limit = limit, format = "json"
      ),
      httr::timeout(15)
    ),
    error = function(e) NULL
  )
  if (is.null(res) || httr::http_error(res)) return(tibble())
  parsed <- tryCatch(httr::content(res, as = "parsed", type = "application/json"),
                      error = function(e) NULL)
  if (is.null(parsed) || length(parsed$search) == 0) return(tibble())
  tibble(
    qid = map_chr(parsed$search, ~ .x$id %||% NA_character_),
    label = map_chr(parsed$search, ~ .x$label %||% NA_character_)
  )
}

# Second step: for each candidate QID from search, check whether it is
# actually one of the place-like classes used throughout this section —
# wbsearchentities has no class filter of its own, so this keeps the
# fallback path just as class-restricted as the exact-match SPARQL query.
check_place_class <- function(qids) {
  if (length(qids) == 0) return(tibble())
  values_clause <- paste(paste0("wd:", qids), collapse = " ")
  q <- sprintf('
SELECT ?item ?itemLabel ?coord ?instanceOfLabel WHERE {
  VALUES ?item { %s }
  ?item wdt:P31/wdt:P279* ?instanceOf .
  VALUES ?instanceOf { wd:Q839954 wd:Q486972 wd:Q17334923 wd:Q23442 wd:Q82794 wd:Q34023 }
  OPTIONAL { ?item wdt:P625 ?coord . }
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}', values_clause)
  tryCatch(query_wikidata(q), error = function(e) tibble())
}

probe_one <- function(token) {
  cat("Querying:", token, "... ")
  q <- sprintf(probe_query_template, token, token, token, token, token)
  res <- tryCatch(
    {
      if (!requireNamespace("R.utils", quietly = TRUE)) {
        install.packages("R.utils", repos = "https://cloud.r-project.org")
      }
      R.utils::withTimeout(query_wikidata(q), timeout = 20, onTimeout = "silent")
    },
    error = function(e) NULL
  )
  if (!is.null(res) && nrow(res) > 0) {
    cat(nrow(res), "candidate(s) [exact]\n")
    return(tibble(token = token, hit = TRUE, n_candidates = nrow(res), match_type = "exact"))
  }

  # Exact match failed — try wbsearchentities fallback before giving up.
  candidates <- search_entities(token)
  if (nrow(candidates) == 0) {
    cat("no match\n")
    return(tibble(token = token, hit = FALSE, n_candidates = 0, match_type = "none"))
  }
  res_c <- check_place_class(candidates$qid)
  if (nrow(res_c) == 0) {
    cat("no match (search hit, wrong class)\n")
    return(tibble(token = token, hit = FALSE, n_candidates = 0, match_type = "none"))
  }
  cat(nrow(res_c), "candidate(s) [search]\n")
  tibble(token = token, hit = TRUE, n_candidates = nrow(res_c), match_type = "search")
}

probe_results <- map_dfr(top_tokens, probe_one)
cat("Coverage probe: ", sum(probe_results$hit), " / ", nrow(probe_results),
    " of the top ", length(top_tokens), " tokens returned at least one candidate\n", sep = "")
cat("  exact:", sum(probe_results$match_type == "exact"),
    " | search fallback:", sum(probe_results$match_type == "search"),
    " | no match:", sum(probe_results$match_type == "none"), "\n")
print(probe_results, n = 50)
```

**Check `probe_results` before continuing.** If the hit rate on real,
frequent Viking-Age placenames (Birka, Hedeby, Kaupang, Ribe, Uppåkra,
Tissø, etc.) is low, the class filter above may be too narrow — Wikidata
items for archaeological sites are not always tagged with exactly these
`instance of` values — and the SPARQL query needs revising before running
the full match. A high count under `contains fallback` is expected and
fine — it means the class filter is right but the corpus token is a
shortened form of the item's real label (e.g. "Gokstad" for "Gokstad
Mound"), not a coverage problem.

**Full match**, once the probe above confirms acceptable coverage. This
single chunk matches every token, separates out macro-region terms that
can't resolve to a point, and writes the resulting lookup table plus a
manual-review export, all together (previously three separate chunks;
merged into one so a single `eval` flag controls the whole matching-to-
saved-file pipeline, since matching without saving the result was wasted
work run for nothing):

```{r}
#| label: region-match
#| eval: false
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  install.packages("WikidataQueryServiceR", repos = "https://cloud.r-project.org")
}
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  if (!requireNamespace("remotes", quietly = TRUE)) {
    install.packages("remotes", repos = "https://cloud.r-project.org")
  }
  remotes::install_github("bearloga/WikidataQueryServiceR")
}
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  stop(
    "WikidataQueryServiceR failed to install from both CRAN and GitHub. ",
    "This points to an environment issue rather than a package issue — most ",
    "likely no outbound network access at all from this machine. Run ",
    "install.packages(\"WikidataQueryServiceR\", repos = \"https://cloud.r-project.org\") ",
    "directly in the console (outside this render) to see the real underlying error."
  )
}
library(WikidataQueryServiceR)
library(httr)

match_query_template <- '
SELECT ?item ?itemLabel ?coord ?instanceOfLabel WHERE {
  VALUES ?searchterm { "%s"@en "%s"@sv "%s"@da "%s"@nb "%s"@is }
  ?item rdfs:label|skos:altLabel ?searchterm .
  ?item wdt:P31/wdt:P279* ?instanceOf .
  VALUES ?instanceOf { wd:Q839954 wd:Q486972 wd:Q17334923 wd:Q23442 wd:Q82794 wd:Q34023 }
  OPTIONAL { ?item wdt:P625 ?coord . }
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}
LIMIT 5
'

# Fallback for tokens with no exact label/alias match — see the probe
# chunk above for the full rationale (catches cases like "Gokstad" vs.
# the real item title "Gokstad Mound"). Flagged as match_method
# "search_fallback" so it's visible which matches came from the
# lower-precision path during manual review.
#
# Redefined here (identical to the probe chunk's versions) rather than
# relying on them existing from a prior run of region-probe in the same
# session — this chunk may be run on its own.
search_entities <- function(token, limit = 5) {
  res <- tryCatch(
    httr::GET(
      "https://www.wikidata.org/w/api.php",
      query = list(
        action = "wbsearchentities", search = token, language = "en",
        type = "item", limit = limit, format = "json"
      ),
      httr::timeout(15)
    ),
    error = function(e) NULL
  )
  if (is.null(res) || httr::http_error(res)) return(tibble())
  parsed <- tryCatch(httr::content(res, as = "parsed", type = "application/json"),
                      error = function(e) NULL)
  if (is.null(parsed) || length(parsed$search) == 0) return(tibble())
  tibble(
    qid = map_chr(parsed$search, ~ .x$id %||% NA_character_),
    label = map_chr(parsed$search, ~ .x$label %||% NA_character_)
  )
}

check_place_class <- function(qids) {
  if (length(qids) == 0) return(tibble())
  values_clause <- paste(paste0("wd:", qids), collapse = " ")
  q <- sprintf('
SELECT ?item ?itemLabel ?coord ?instanceOfLabel WHERE {
  VALUES ?item { %s }
  ?item wdt:P31/wdt:P279* ?instanceOf .
  VALUES ?instanceOf { wd:Q839954 wd:Q486972 wd:Q17334923 wd:Q23442 wd:Q82794 wd:Q34023 }
  OPTIONAL { ?item wdt:P625 ?coord . }
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}', values_clause)
  tryCatch(query_wikidata(q), error = function(e) tibble())
}
#
# Uses the same wbsearchentities + class-check pattern as the probe
# chunk above, rather than a SPARQL CONTAINS scan, which was confirmed
# to time out at Wikidata's scale even after restructuring the query —
# see the probe chunk's comment for the full rationale. search_entities()
# and check_place_class() are defined there; this chunk reuses them.

match_one <- function(token) {
  q <- sprintf(match_query_template, token, token, token, token, token)
  res <- tryCatch(query_wikidata(q), error = function(e) NULL)

  if (is.null(res) || nrow(res) == 0) {
    # Exact match failed — try wbsearchentities fallback before giving up.
    candidates <- search_entities(token)
    res_c <- if (nrow(candidates) > 0) check_place_class(candidates$qid) else tibble()
    if (nrow(res_c) == 0) {
      return(tibble(
        region_token = token, wikidata_qid = NA_character_,
        normalized_label = NA_character_, lat = NA_real_, lon = NA_real_,
        place_type = NA_character_, match_method = "no_match"
      ))
    }
    res <- res_c
    fallback_used <- TRUE
  } else {
    fallback_used <- FALSE
  }

  # Multiple candidates: prefer the one whose returned label exactly
  # equals the token over an arbitrary first result, since an
  # exact-label match among several candidates is a real signal, not
  # a coin flip. Falls back to the first row only when no candidate's
  # label matches the token exactly (that case still gets flagged
  # ambiguous below, since the pick is genuinely uncertain).
  exact_label_match <- which(str_to_lower(res$itemLabel) == str_to_lower(token))
  best <- if (length(exact_label_match) > 0) res[exact_label_match[1], ] else res[1, ]
  resolved_by_label <- length(exact_label_match) > 0

  coord <- best$coord
  lat <- NA_real_; lon <- NA_real_
  if (!is.na(coord) && str_detect(coord, "^Point")) {
    nums <- str_match(coord, "Point\\(([-0-9.]+) ([-0-9.]+)\\)")
    lon <- as.numeric(nums[2]); lat <- as.numeric(nums[3])
  }
  is_ambiguous <- nrow(res) > 1 && !resolved_by_label
  method <- if (fallback_used) {
    if (is_ambiguous) "search_fallback_ambiguous" else "search_fallback"
  } else {
    if (is_ambiguous) "exact_ambiguous" else "exact"
  }
  tibble(
    region_token = token,
    wikidata_qid = str_extract(best$item, "Q[0-9]+$"),
    normalized_label = best$itemLabel,
    lat = lat, lon = lon,
    place_type = best$instanceOfLabel,
    match_method = method
  )
}

region_matches <- map_dfr(unique_tokens, match_one)

cat("Matched:", sum(region_matches$match_method != "no_match"), "/", nrow(region_matches), "\n")
cat("Ambiguous (multiple candidates, first taken):",
    sum(region_matches$match_method == "exact_ambiguous"), "\n")
cat("No match:", sum(region_matches$match_method == "no_match"), "\n")

unmatched <- region_matches %>% filter(match_method == "no_match") %>% pull(region_token)
print(unmatched)

# Write the raw lookup table. Manual corrections (known-bad matches) and
# macro-region flagging are NOT applied here — they live in region-joined
# below instead, which runs on every normal render (no network needed)
# and reads straight from this file. That way, fixing a bad match or
# adding a term to the macro-region list never requires re-running this
# chunk (or its ~10 minutes of Wikidata queries) again — only editing
# region-joined and re-rendering, which is instant.
write_csv(region_matches, here("data", "region_lookup.csv"))
cat("Written to data/region_lookup.csv:", nrow(region_matches), "rows\n")

region_matches %>%
  filter(match_method %in% c("exact_ambiguous", "search_fallback",
                              "search_fallback_ambiguous", "no_match")) %>%
  write_csv(here("data", "region_lookup_needs_review.csv"))
```

**Nesting (optional).** Every matched place is currently flat: "Birka,"
"Vestfold," "Sweden," and "Scandinavia" are independent tokens with no
recorded relationship between them, even though Birka is genuinely in
Uppland, Sweden, Scandinavia. This chunk resolves three nesting levels
for each matched place, each as its own column:

- **`admin_region`** — the first `P131` ("located in the administrative
  territorial entity") ancestor above the site itself, before reaching
  the country: a county, province, or similar (e.g. Birka → Uppland).
  Not every place has a clean single admin-region ancestor in Wikidata;
  those are left `NA` rather than guessed.
- **`country`** — walked further up `P131*` to the first ancestor that
  is itself a country (Wikidata class Q6256).
- **`super_region`** — not derived from Wikidata at all. Unlike country,
  there is no single reliable Wikidata class for "super-region" in the
  sense this corpus uses the term (Scandinavia, British Isles, Baltic
  states); these are real, well-understood historical/cultural
  groupings, not administrative ones, so this is a small, explicit,
  hand-authored country → super-region mapping instead of an
  automatically-derived one. Extend `super_region_map` below as more
  countries appear in the resolved data.

Like the matching chunks above, this calls Wikidata and needs network
access; it stays off by default.

Country resolution here runs on the raw matched coordinates, before the
`known_bad_matches` correction is applied (that correction lives in
region-joined below, and only nulls the bad matches' `lat`/`lon`/etc.
there, not here). So resolving nesting right after this chunk can
produce a real `country` value for a token later corrected to
`no_match` — e.g. "Repton" resolving to "United States" here, even
though `region-joined` clears it entirely downstream. That correction
was extended to also null `admin_region`/`country`/`super_region` when
present, so this is handled automatically as long as `region-joined`
runs after nesting, which it always does on a normal render.

```{r}
#| label: region-nesting
#| eval: false
# This chunk runs independently of region-match above and does not
# inherit its library load — confirmed the actual cause of every past
# empty result: query_wikidata() was never defined here, so every batch
# failed immediately with "could not find function" and silently
# returned zero rows, which then looked like a query or network problem
# rather than a missing library call.
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  install.packages("WikidataQueryServiceR", repos = "https://cloud.r-project.org")
}
if (!requireNamespace("WikidataQueryServiceR", quietly = TRUE)) {
  if (!requireNamespace("remotes", quietly = TRUE)) {
    install.packages("remotes", repos = "https://cloud.r-project.org")
  }
  remotes::install_github("bearloga/WikidataQueryServiceR")
}
library(WikidataQueryServiceR)

region_lookup_for_nesting <- read_csv(here("data", "region_lookup.csv"), show_col_types = FALSE) %>%
  # Drop any nesting columns from a prior run before this one starts —
  # otherwise left_join() below collides with them and produces
  # admin_region.x/admin_region.y-style duplicates instead of clean
  # single columns, compounding further on every re-run.
  select(-any_of(c("admin_region", "country", "super_region")))

# Apply the same manual overrides used in region-joined before resolving
# nesting — without this, nesting would resolve admin_region/country/
# super_region from the ORIGINAL wrong QIDs on disk (this file is the
# raw match output; the override only ever lived in region-joined's
# in-memory region_lookup, never written back here), producing correct
# lat/lon downstream but stale, wrong country values for exactly the
# nine tokens this override exists to fix. Kept in sync manually with
# the table in region-joined below — if one changes, update the other.
known_bad_match_overrides <- tibble::tribble(
  ~region_token,   ~wikidata_qid,  ~normalized_label,          ~lat,      ~lon,       ~place_type,
  "East Norway",   "Q216065",      "Eastern Norway (Østlandet)", 60.25,   10.67,      "region",
  "Repton",        "Q1018787",     "Repton",                    52.83983, -1.55061,  "village",
  "Baltic",        "Q545",         "Baltic Sea",                58,       20,         "sea",
  "North Norway",  "Q214041",      "Northern Norway",           69.667,   19.0,       "region",
  "Scotland",      "Q22",          "Scotland",                  56,       -4,         "country",
  "Masham",        "Q2330678",     "Masham",                    54.22,   -1.656,      "town",
  "West Sweden",   "Q4992082",     "West Sweden (Västsverige)", 57.558,   12.602,     "region",
  "East Denmark",  "Q131281582",   "Eastern Denmark (general, administrative region est. 2027)", 55.44, 11.56, "region"
)

region_lookup_for_nesting <- region_lookup_for_nesting %>%
  left_join(known_bad_match_overrides %>% select(region_token, wikidata_qid),
            by = "region_token", suffix = c("", "_override")) %>%
  mutate(wikidata_qid = coalesce(wikidata_qid_override, wikidata_qid)) %>%
  select(-any_of("wikidata_qid_override"))

qids_to_resolve <- region_lookup_for_nesting %>%
  filter(!is.na(wikidata_qid)) %>%
  distinct(wikidata_qid) %>%
  pull(wikidata_qid)

cat("Resolving nesting for", length(qids_to_resolve), "matched places...\n")

# One query per batch resolves both levels at once: the immediate P131
# parent (candidate admin_region, whatever it turns out to be) and the
# walked-up-to-country ancestor (country). A batch item's admin_region
# is only kept if it is NOT itself the country (i.e. the site has a real
# intermediate level); otherwise left NA rather than duplicating the
# country value into both columns.
resolve_nesting_batch <- function(qids) {
  values_clause <- paste(paste0("wd:", qids), collapse = " ")
  # Both admin_region and country wrapped in OPTIONAL — some matched
  # items (islands, ocean regions, and similar) have no P131 ancestor
  # that is itself a country, and without OPTIONAL here those items
  # would silently drop out of the result entirely rather than
  # returning with country = NA, which could shrink the whole batch's
  # result to zero rows and break the code downstream.
  q <- sprintf('
SELECT ?item ?adminLabel ?countryLabel WHERE {
  VALUES ?item { %s }
  OPTIONAL { ?item wdt:P131 ?admin . }
  OPTIONAL {
    ?item wdt:P131* ?countryCandidate .
    ?countryCandidate wdt:P31 wd:Q6256 .
    BIND(?countryCandidate AS ?country)
  }
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}', values_clause)
  res <- tryCatch(query_wikidata(q), error = function(e) tibble())
  # Defensive check: an empty or malformed result (no "item" column)
  # must not reach distinct()/transmute() downstream — return a
  # correctly-shaped zero-row tibble instead so map_dfr can still bind
  # it with real batches.
  if (!"item" %in% names(res)) {
    return(tibble(item = character(), adminLabel = character(), countryLabel = character()))
  }
  res
}

batch_size <- 20
qid_batches <- split(qids_to_resolve, ceiling(seq_along(qids_to_resolve) / batch_size))

nesting_results <- map_dfr(seq_along(qid_batches), function(i) {
  batch <- qid_batches[[i]]
  cat("  batch", i, "of", length(qid_batches), "...\n")
  resolve_nesting_batch(batch)
})

if (nrow(nesting_results) == 0) {
  cat("No nesting results returned at all — check network access and the query above directly.\n")
  nesting_lookup <- tibble(wikidata_qid = character(), admin_region = character(), country = character())
} else {
  # Wikidata sometimes labels a country item with its formal name
  # ("Kingdom of Denmark," "Kingdom of the Netherlands") rather than the
  # common one ("Denmark," "Netherlands") — normalized here, once, at
  # the source, so every downstream use (this column, super_region
  # mapping, the country totals chart) sees one consistent value instead
  # of two country labels for the same real country.
  country_name_normalize <- c(
    "Kingdom of Denmark" = "Denmark",
    "Kingdom of the Netherlands" = "Netherlands",
    "Kingdom of Sweden" = "Sweden",
    "Kingdom of Norway" = "Norway",
    "Republic of Ireland" = "Ireland",
    "Kingdom of Belgium" = "Belgium"
  )

  nesting_lookup <- nesting_results %>%
    distinct(item, .keep_all = TRUE) %>%
    transmute(
      wikidata_qid = str_extract(item, "Q[0-9]+$"),
      admin_region = if_else(is.na(adminLabel) | is.na(countryLabel) | adminLabel == countryLabel,
                              NA_character_, adminLabel),
      country = coalesce(country_name_normalize[countryLabel], countryLabel)
    )
}

cat("Resolved country for", sum(!is.na(nesting_lookup$country)), "/", length(qids_to_resolve), "matched places\n")
cat("Resolved a distinct admin_region for", sum(!is.na(nesting_lookup$admin_region)), "of those\n")

# Super-region: hand-authored, not derived from Wikidata — see the
# rationale above. Extend as more countries show up in country above.
# Keys are common country names only — formal-name variants (Kingdom of
# Denmark, Kingdom of the Netherlands, etc.) are already normalized to
# the common form in country_name_normalize above, before this map is
# ever consulted.
super_region_map <- c(
  "Sweden" = "Scandinavia", "Norway" = "Scandinavia", "Denmark" = "Scandinavia",
  "Iceland" = "Scandinavia", "Finland" = "Scandinavia", "Greenland" = "Scandinavia",
  "United Kingdom" = "British Isles", "Ireland" = "British Isles",
  "Russia" = "Eastern Europe", "Ukraine" = "Eastern Europe",
  "Poland" = "Eastern Europe", "Lithuania" = "Baltic states",
  "Latvia" = "Baltic states", "Estonia" = "Baltic states",
  "Germany" = "Continental Europe", "Netherlands" = "Continental Europe",
  "Belgium" = "Continental Europe",
  "France" = "Continental Europe", "Canada" = "North Atlantic"
)

nesting_lookup <- nesting_lookup %>%
  mutate(super_region = super_region_map[country])

n_no_super_region <- sum(!is.na(nesting_lookup$country) & is.na(nesting_lookup$super_region))
if (n_no_super_region > 0) {
  cat(n_no_super_region, "resolved countries have no super_region mapping yet — add them to",
      "super_region_map above:\n")
  nesting_lookup %>% filter(!is.na(country) & is.na(super_region)) %>%
    distinct(country) %>% print()
}

region_lookup_nested <- region_lookup_for_nesting %>%
  left_join(nesting_lookup, by = "wikidata_qid")

write_csv(region_lookup_nested, here("data", "region_lookup.csv"))
cat("Written data/region_lookup.csv with admin_region, country, super_region columns added:",
    nrow(region_lookup_nested), "rows\n")
```

```{r}
#| label: region-joined
region_lookup_path <- here("data", "region_lookup.csv")

if (file.exists(region_lookup_path)) {
  region_lookup <- read_csv(region_lookup_path, show_col_types = FALSE)

  # Manual corrections, found during review: these tokens resolved (via
  # exact match or search fallback) to a real Wikidata place that shares
  # the corpus token's name but is a different, wrong place — a same-name
  # town elsewhere, not the actual site the paper means. Each override
  # below is a directly verified, real Wikidata QID and coordinate for
  # what the token actually means in this corpus, checked individually
  # (not guessed) before being hardcoded — see the region-match/
  # region-joined commit history for the sourcing on each. Applied here
  # rather than in region-match above, so fixing or adding a correction
  # never requires re-running the ~10-minute network match — only
  # editing this table and re-rendering, which is instant.
  known_bad_match_overrides <- tibble::tribble(
    ~region_token,   ~wikidata_qid,  ~normalized_label,          ~lat,      ~lon,       ~place_type,
    "East Norway",   "Q216065",      "Eastern Norway (Østlandet)", 60.25,   10.67,      "region",
    "Repton",        "Q1018787",     "Repton",                    52.83983, -1.55061,  "village",
    "Baltic",        "Q545",         "Baltic Sea",                58,       20,         "sea",
    "North Norway",  "Q214041",      "Northern Norway",           69.667,   19.0,       "region",
    "Scotland",      "Q22",          "Scotland",                  56,       -4,         "country",
    "Masham",        "Q2330678",     "Masham",                    54.22,   -1.656,      "town",
    "West Sweden",   "Q4992082",     "West Sweden (Västsverige)", 57.558,   12.602,     "region",
    "East Denmark",  "Q131281582",   "Eastern Denmark (general, administrative region est. 2027)", 55.44, 11.56, "region"
  )

  region_lookup <- region_lookup %>%
    left_join(known_bad_match_overrides, by = "region_token", suffix = c("", "_override")) %>%
    mutate(
      wikidata_qid = coalesce(wikidata_qid_override, wikidata_qid),
      normalized_label = coalesce(normalized_label_override, normalized_label),
      lat = coalesce(lat_override, lat),
      lon = coalesce(lon_override, lon),
      place_type = coalesce(place_type_override, place_type),
      match_method = if_else(!is.na(wikidata_qid_override), "manual_override", match_method)
    ) %>%
    select(-ends_with("_override"))

  # The override above only replaces wikidata_qid/lat/lon/place_type —
  # it does NOT touch admin_region/country/super_region, which (if
  # region-nesting has already been run) were resolved from the
  # *original wrong* coordinates and are now stale for these nine
  # tokens. Left alone, this is exactly the same class of bug already
  # fixed once for East Norway: a chart reading country directly would
  # still show the old, wrong value even though lat/lon are now correct.
  # Cleared here so they read as genuinely unresolved rather than
  # silently wrong; re-run region-nesting to populate them correctly
  # from the corrected coordinates if admin_region/country/super_region
  # values are wanted for these nine.
  region_lookup <- region_lookup %>%
    mutate(across(any_of(c("admin_region", "country", "super_region")),
                  ~ if_else(region_token %in% known_bad_match_overrides$region_token, NA, .)))

  # "East Baltic" is a genuine exception: its Wikidata entry (Q5327781)
  # is a disambiguation page, not a single resolvable place (it covers
  # the eastern Baltic region generally, East Baltic languages, and a
  # historical race classification, with no one coordinate that honestly
  # represents all three) — moved to macro_region_terms below instead of
  # given a fabricated single point.
  #
  # "Eurasia" is a second, different exception: a real Wikidata match
  # exists (50°N, 80°E — lands in Kazakhstan on a map), but the concept
  # itself spans the whole landmass, so that single coordinate is
  # coincidental placement, not a meaningful site — moved to
  # macro_region_terms below for the same reason as East Baltic.

  # Macro-region terms: "Scandinavia," "North Atlantic," "Viking world,"
  # "South Scandinavia," "East Baltic," "Eurasia" (see notes above), and
  # similar will not resolve to a single sensible coordinate regardless
  # of gazetteer — they describe a broad area or a genuinely ambiguous
  # concept, not a point. Flagged here, also for the same instant-edit
  # reason above, rather than force-matched to a single Wikidata item
  # with a misleading coordinate. Extend this list after reviewing
  # region_lookup_needs_review.csv.
  macro_region_terms <- c(
    "Scandinavia", "South Scandinavia", "North Atlantic", "North Europe",
    "Northwest Europe", "East Europe", "Viking world", "British Isles",
    "Scandinavian diaspora", "the North", "Arctic Canada", "Arctic Norway",
    "Central East Europe", "Central East Sweden", "East Arctic",
    "Middle Sweden", "North Norway and Sweden", "North Scandinavia",
    "Northwest Slavic area", "Norway and Sweden", "Sami areas",
    "Southeast Norway", "Southwest Norway", "Southwest Sweden",
    "West Scandinavia", "West-Balts", "East Baltic", "Eurasia",
    # Second pass, found by checking every matched row for the same
    # "genuine Wikidata match, but country/continent-scale concept
    # plotted at a coincidental centroid" signature that caught Eurasia
    # — country and continent-level terms, previously left plottable as
    # single misleading points even though they are not site-level.
    "Britain", "British isles", "Denmark", "England", "Europe",
    "Finland", "Germany", "Norge", "Norway", "Northern Europe",
    "Northern Hemisphere", "Oceania", "Pacific Ocean", "Poland",
    "Russia", "South Norway", "South Sweden", "Southwest Finland",
    "Sverige", "Sweden", "West Europe"
  )

  region_lookup <- region_lookup %>%
    mutate(is_macro_region = region_token %in% macro_region_terms)

  # Every macro-region term must also be cleared of any coordinate it
  # may carry from the raw match — the is_macro_region flag alone only
  # excludes them from the point map, but a leftover coordinate would
  # still count toward country/super_region rollups, which is exactly
  # the same category of error the manual overrides above exist to fix.
  # Real, confirmed example: "Eurasia" genuinely matched a real Wikidata
  # item at 50°N, 80°E (landing in Kazakhstan on a map) — a real match,
  # but for a continent-scale concept, so that coordinate is
  # coincidental placement, not a meaningful site. The same logic
  # applies to every other entry on the list above (country- and
  # continent-level terms), not just the two found first.
  region_lookup <- region_lookup %>%
    mutate(across(any_of(c("wikidata_qid", "normalized_label", "lat", "lon", "place_type",
                            "admin_region", "country", "super_region")),
                  ~ if_else(is_macro_region, NA, .)),
           match_method = if_else(is_macro_region, "no_match", match_method))

  region_df <- region_tokens %>%
    left_join(region_lookup, by = c("Region_normalized" = "region_token"))
} else {
  cat("data/region_lookup.csv not found yet — run region-match (eval: true) first.\n")
}
```

Once `region_lookup.csv` exists, the resulting `region_df` joins region to
gender, topic, method, source, and year the same way those variables
already join to each other elsewhere in this document — by `cite_key`,
using the same Wilson-interval and by-year machinery.

### Where the corpus is set

Most of the real signal in Region is simply where papers are concentrated
geographically — a stacked gender-composition bar chart like the ones
used for Topic, Method, and Source needs several well-populated
categories to say anything real, and Region has one (Birka, 20 papers)
with a steep drop to a long tail of 1-3-paper tokens, so that chart type
doesn't suit this variable the way it suits the others. Two more direct
views instead: where the named, point-matched places actually are, and
how concentrated the corpus really is by raw count.

**What the map includes.** A region token appears as a point only if it
has a real matched coordinate, is not a macro-region term (Scandinavia,
Viking world, East Baltic, and the rest of that list — excluded even
when Wikidata does return a coordinate for the concept, since a single
point can't honestly represent an area that broad), and falls within a
fixed plotting bounding box (lat 25-80, lon -60 to 100) that excludes a
couple of real but geographically irrelevant matches (Oceania, Pacific
Ocean) which would otherwise force the whole map out to a global scale.
Dot size is the number of papers carrying that exact region token — 20
papers all tagged "Birka" produce one dot sized 20, not 20 separate
dots. The map's crop is computed from the real spread of whatever points
pass all three filters, with a fixed margin, so it always fits what's
actually shown rather than a hardcoded guess.

```{r}
#| label: region-map
#| fig-cap: >
#|   Named places mentioned in the corpus, sized by number of papers.
#|   Macro-region terms (Scandinavia, Viking world, and similar) are
#|   excluded even when Wikidata returns a coordinate for them, since a
#|   single point cannot honestly represent an area that broad — see the
#|   totals chart below for those. Cropped to the real spread of the
#|   remaining matched coordinates with a margin; a handful of correct
#|   but far-flung matches (e.g. Oceania, Pacific Ocean — real tokens in
#|   the corpus, real Wikidata matches) fall outside this crop and are
#|   not shown on the map.
#| fig-width: 11
#| fig-height: 8
if (!requireNamespace("rnaturalearth", quietly = TRUE)) {
  install.packages("rnaturalearth", repos = "https://cloud.r-project.org")
}
if (!requireNamespace("rnaturalearthdata", quietly = TRUE)) {
  install.packages("rnaturalearthdata", repos = "https://cloud.r-project.org")
}
library(rnaturalearth)

# is_macro_region must be excluded explicitly here — some macro-region
# terms (Scandinavia among them) DO get a real Wikidata coordinate (the
# country/concept item has one), but a single point still cannot
# honestly represent an area that broad, which is the whole reason
# these terms were flagged as macro-region in the first place. Filtering
# only on !is.na(lat) is not enough; that flag has to be checked too.
region_points <- region_df %>%
  filter(!is.na(lat), !is.na(lon), !is_macro_region) %>%
  count(region = Region_normalized, lat, lon, name = "n_papers")

# A handful of matches are real but far outside the North Atlantic/
# European scope this corpus is actually about (Oceania, Pacific Ocean —
# genuine tokens, genuine Wikidata matches, just not relevant to a map of
# this corpus's geography). Excluded from the crop calculation so they
# don't force the whole map out to a near-global scale; still excluded
# from the plotted points themselves for the same reason.
region_points_mapped <- region_points %>%
  filter(lat >= 25, lat <= 80, lon >= -60, lon <= 100)

map_margin <- 8
map_xlim <- c(min(region_points_mapped$lon) - map_margin, max(region_points_mapped$lon) + map_margin)
map_ylim <- c(min(region_points_mapped$lat) - map_margin, max(region_points_mapped$lat) + map_margin)

world <- ne_countries(scale = "medium", returnclass = "sf")

ggplot() +
  geom_sf(data = world, fill = "grey95", colour = "grey80", linewidth = 0.2) +
  geom_point(data = region_points_mapped, aes(x = lon, y = lat, size = n_papers),
             colour = "#bf616a", alpha = 0.7) +
  coord_sf(xlim = map_xlim, ylim = map_ylim, expand = FALSE) +
  scale_size_continuous(range = c(1.5, 10), name = "papers") +
  theme_bw() +
  labs(x = NULL, y = NULL)
```

```{r}
#| label: region-totals
#| fig-cap: >
#|   Every region token with at least 3 tagged papers, point-matched
#|   places and macro-region terms together, ranked by raw count.
#| fig-width: 9
#| fig-height: 8
region_totals <- region_tokens %>%
  count(region = Region_normalized, sort = TRUE) %>%
  filter(n >= 3)

ggplot(region_totals, aes(x = fct_reorder(region, n), y = n)) +
  geom_col(fill = "#bf616a") +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  theme_bw() +
  labs(x = "", y = "Papers")
```

**By country (optional, requires country nesting).** The chart above
keeps every region token flat, "Birka" and "Sweden" and "Scandinavia"
are three separate bars even though Birka is in Sweden — deliberately,
since the paper's own choice of a site-level or country-level term is
itself part of what the data says, not noise to collapse away. This
alternate view rolls site-level tags up to the country they're actually
in, using the `country` column added by the region-nesting chunk above
(only present once that chunk has been run). A paper tagged "Birka"
counts toward Sweden here even if it was never separately tagged
"Sweden." Macro-region terms (Scandinavia, Viking world) still have no
single country and are excluded from this view.

```{r}
#| label: region-totals-by-country
#| fig-cap: "Papers per country, site-level tags rolled up to their containing country."
#| fig-width: 9
#| fig-height: 6
# Check for actual data, not just column presence — a prior failed or
# empty region-nesting run can leave a "country" column that exists but
# is entirely NA, which would otherwise reach fct_reorder() on an empty
# vector and crash rather than print a useful message.
has_country_data <- "country" %in% names(region_df) && any(!is.na(region_df$country))

if (has_country_data) {
  region_country_totals <- region_df %>%
    filter(!is.na(country)) %>%
    distinct(cite_key, country) %>%
    count(country, sort = TRUE)

  ggplot(region_country_totals, aes(x = fct_reorder(country, n), y = n)) +
    geom_col(fill = "#5e81ac") +
    geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
    coord_flip() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
    theme_bw() +
    labs(x = "", y = "Papers")
} else if ("country" %in% names(region_df)) {
  cat("country column exists but has no non-NA values — region-nesting ran",
      "but returned nothing. Check network access and the query directly.\n")
} else {
  cat("No country column yet — run region-nesting (eval: true) first.\n")
}
```

### Topic, method, and source by region

Same problem as gender composition above: most region tokens have too
few papers to support a per-region proportion chart. Enrichment solves
this differently — restricted to regions with at least 5 tagged papers,
each cell shows that region's coverage of a topic/method/source divided
by that category's own rate across the whole corpus (1.0x = corpus
average), the same measure already used for cluster structure in Part 3.
A region's low paper count still means less certainty behind its cells,
but enrichment at least says something real about *direction* even on a
modest sample, where a stacked-proportion chart on the same sample would
mostly just show noise. All three heatmaps below share the same region
row order, so a region's profile can be compared directly across tabs.

::: {.panel-tabset}

#### Topic

```{r}
#| label: region-topic-enrichment
#| fig-cap: >
#|   Topic enrichment by region (regions with 5+ tagged papers only).
#|   White = topic at corpus average rate for that region; blue = above
#|   average; red = below. Cell shows raw count behind the percentage.
#|   Capped at 3x so a handful of extreme small-n ratios don't wash out
#|   the rest of the scale.
#| fig-width: 11
#| fig-height: 9
region_n_papers <- region_tokens %>%
  distinct(cite_key, region = Region_normalized) %>%
  count(region, name = "n_region_papers") %>%
  filter(n_region_papers >= 5)

# Fixed row order, computed once and shared across all three heatmaps
# below, so the same region always sits at the same row whether looking
# at topic, method, or source — the whole point of putting these in one
# tabset is comparing a region's profile across the three, and that only
# works if the rows don't silently reorder between tabs (fct_reorder on
# each chart's own filtered subset can do exactly that if a region has
# zero rows in one category but not another). Labels carry the paper
# count behind each row's percentages directly — n_region_papers IS
# the full count here (unlike the GN cluster heatmaps in Part 3, a
# region has no separate "true size" beyond its own tagged-paper count,
# so there is no annotated-vs-true-size gap to show).
region_order_df <- region_n_papers %>% arrange(n_region_papers) %>%
  mutate(region_label = paste0(region, " (n=", n_region_papers, ")"))
region_order <- region_order_df$region_label
region_label_lookup <- setNames(region_order_df$region_label, region_order_df$region)

region_topic <- region_tokens %>%
  select(cite_key, region = Region_normalized) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic))

region_topic_enrichment <- region_topic %>%
  filter(region %in% region_n_papers$region) %>%
  distinct(cite_key, region, topic) %>%
  count(region, topic, name = "n") %>%
  left_join(region_n_papers, by = "region") %>%
  left_join(topic_corpus_rate, by = "topic") %>%
  mutate(prop = n / n_region_papers, enrichment = prop / corpus_rate,
         region = factor(region_label_lookup[region], levels = region_order)) %>%
  # Drop topics with no coverage in any of the shown regions, so the
  # chart doesn't carry empty columns for categories nobody here was
  # coded with — same as the cluster-vs-topic heatmap in Part 3.
  group_by(topic) %>%
  filter(sum(n) > 0) %>%
  ungroup()

region_topic_heatmap <- ggplot(region_topic_enrichment, aes(x = topic, y = region, fill = enrichment)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = n), size = 2.8) +
  # Diverging scale capped at 3x, same as the cluster-vs-topic heatmap in
  # Part 3 — a handful of extreme small-n ratios would otherwise wash
  # out the rest of the scale.
  scale_fill_gradient2(low = "#bf616a", mid = "white", high = "#5e81ac", midpoint = 1,
                        limits = c(0, 3), oob = scales::squish,
                        name = "Enrichment", labels = scales::label_number(suffix = "x")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "")
region_topic_heatmap
```

#### Method

```{r}
#| label: region-method-enrichment
#| fig-cap: >
#|   Method enrichment by region (regions with 5+ tagged papers only).
#|   Same reading as the topic heatmap above.
#| fig-width: 11
#| fig-height: 9
region_method <- region_tokens %>%
  select(cite_key, region = Region_normalized) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method))

region_method_enrichment <- region_method %>%
  filter(region %in% region_n_papers$region) %>%
  distinct(cite_key, region, method) %>%
  count(region, method, name = "n") %>%
  left_join(region_n_papers, by = "region") %>%
  left_join(method_corpus_rate, by = "method") %>%
  mutate(prop = n / n_region_papers, enrichment = prop / corpus_rate,
         region = factor(region_label_lookup[region], levels = region_order)) %>%
  group_by(method) %>%
  filter(sum(n) > 0) %>%
  ungroup()

region_method_heatmap <- ggplot(region_method_enrichment, aes(x = method, y = region, fill = enrichment)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = n), size = 2.8) +
  scale_fill_gradient2(low = "#bf616a", mid = "white", high = "#5e81ac", midpoint = 1,
                        limits = c(0, 3), oob = scales::squish,
                        name = "Enrichment", labels = scales::label_number(suffix = "x")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "")
region_method_heatmap
```

#### Source of evidence

```{r}
#| label: region-source-enrichment
#| fig-cap: >
#|   Source-of-evidence enrichment by region (regions with 5+ tagged
#|   papers only). Same reading as the topic heatmap above.
#| fig-width: 11
#| fig-height: 9
region_source <- region_tokens %>%
  select(cite_key, region = Region_normalized) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source))

region_source_enrichment <- region_source %>%
  filter(region %in% region_n_papers$region) %>%
  distinct(cite_key, region, source) %>%
  count(region, source, name = "n") %>%
  left_join(region_n_papers, by = "region") %>%
  left_join(source_corpus_rate, by = "source") %>%
  mutate(prop = n / n_region_papers, enrichment = prop / corpus_rate,
         region = factor(region_label_lookup[region], levels = region_order)) %>%
  group_by(source) %>%
  filter(sum(n) > 0) %>%
  ungroup()

region_source_heatmap <- ggplot(region_source_enrichment, aes(x = source, y = region, fill = enrichment)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = n), size = 2.8) +
  scale_fill_gradient2(low = "#bf616a", mid = "white", high = "#5e81ac", midpoint = 1,
                        limits = c(0, 3), oob = scales::squish,
                        name = "Enrichment", labels = scales::label_number(suffix = "x")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "")
region_source_heatmap
```

:::

### Topic, method, and source by administrative level

The tabset above groups by the raw region token as written in each
paper, "Birka" and "Sweden" stay separate rows even though one is inside
the other. This section repeats the same topic/method/source enrichment
analysis, but grouped at each of the three nesting levels resolved by
the region-nesting chunk above instead: admin region, country, and
super-region. Requires that chunk to have been run at least once — each
tab below checks for its column and explains what to do if it's missing.

The enrichment baseline here is different from the tabset above, on
purpose: each cell compares a group's rate against the rate among only
the papers that resolved to *some* value at that nesting level, not
against the whole 369-paper F1 corpus. The flat-token tabset above uses
the whole-corpus rate, since every F1 paper is a fair comparison point
for a raw region tag. Nesting-level resolution is a narrower subset
(only papers whose region successfully resolved up to a country, say),
so comparing against the whole corpus would dilute the baseline with
papers that never entered this analysis at all and understate real
enrichment within the resolved set. Enrichment numbers are therefore not
directly comparable between the two tabsets, even when they group the
same underlying papers.

```{r}
#| label: region-nesting-heatmap-fn
# One reusable function instead of writing the topic/method/source
# enrichment logic three more times per nesting level — takes the
# grouping column name as a parameter and returns the three heatmaps
# (or a message) for that level.
make_nesting_heatmaps <- function(group_col, min_papers = 5) {
  if (!group_col %in% names(region_df)) {
    cat("No", group_col, "column yet — run region-nesting (eval: true) first.\n")
    return(invisible(NULL))
  }

  nested <- region_df %>%
    filter(!is.na(.data[[group_col]])) %>%
    rename(group = all_of(group_col))

  n_nested_papers <- n_distinct(nested$cite_key)

  group_n_papers <- nested %>%
    distinct(cite_key, group) %>%
    count(group, name = "n_group_papers") %>%
    filter(n_group_papers >= min_papers)

  if (nrow(group_n_papers) == 0) {
    cat("No", group_col, "value has at least", min_papers, "tagged papers — nothing to show.\n")
    return(invisible(NULL))
  }

  group_order_df <- group_n_papers %>% arrange(n_group_papers) %>%
    mutate(group_label = paste0(group, " (n=", n_group_papers, ")"))
  group_order <- group_order_df$group_label
  group_label_lookup <- setNames(group_order_df$group_label, group_order_df$group)

  # Baseline for enrichment here is the rate among papers that actually
  # have a value for this nesting level (n_nested_papers), not the whole
  # 369-paper F1 corpus. The whole-corpus rate mixes in papers with no
  # region tag at all, or a region tag that never resolved to this
  # nesting level, which would understate enrichment for anything
  # genuinely concentrated within the resolved subset. long_df is
  # already restricted to nested's cite_keys by the caller, so the rate
  # only needs recomputing against n_nested_papers, not whole-corpus
  # corpus_rate_df.
  make_one <- function(long_df, category_col, fill_label) {
    category_rate <- long_df %>%
      distinct(cite_key, .data[[category_col]]) %>%
      count(.data[[category_col]], name = "n_subset_corpus") %>%
      mutate(subset_corpus_rate = n_subset_corpus / n_nested_papers)

    enrichment <- long_df %>%
      filter(group %in% group_n_papers$group) %>%
      distinct(cite_key, group, .data[[category_col]]) %>%
      count(group, .data[[category_col]], name = "n") %>%
      left_join(group_n_papers, by = "group") %>%
      left_join(category_rate, by = category_col) %>%
      mutate(prop = n / n_group_papers, enrichment = prop / subset_corpus_rate,
             group = factor(group_label_lookup[group], levels = group_order)) %>%
      group_by(.data[[category_col]]) %>%
      filter(sum(n) > 0) %>%
      ungroup()

    ggplot(enrichment, aes(x = .data[[category_col]], y = group, fill = enrichment)) +
      geom_tile(colour = "white") +
      geom_text(aes(label = n), size = 2.8) +
      scale_fill_gradient2(low = "#bf616a", mid = "white", high = "#5e81ac", midpoint = 1,
                            limits = c(0, 3), oob = scales::squish,
                            name = "Enrichment", labels = scales::label_number(suffix = "x")) +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      labs(x = "", y = "", title = fill_label)
  }

  topic_long <- nested %>%
    select(cite_key, group) %>%
    left_join(
      df %>% select(cite_key, contains("Topics")) %>%
        pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
        filter(!is.na(count)) %>%
        mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
      by = "cite_key"
    ) %>% filter(!is.na(topic))

  method_long <- nested %>%
    select(cite_key, group) %>%
    left_join(
      df %>% select(cite_key, contains("Methods")) %>%
        pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
        filter(!is.na(count)) %>%
        mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
      by = "cite_key"
    ) %>% filter(!is.na(method))

  source_long <- nested %>%
    select(cite_key, group) %>%
    left_join(
      df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
        pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
        filter(!is.na(count)) %>%
        mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
      by = "cite_key"
    ) %>% filter(!is.na(source))

  list(
    topic = make_one(topic_long, "topic", paste(group_col, "x topic")),
    method = make_one(method_long, "method", paste(group_col, "x method")),
    source = make_one(source_long, "source", paste(group_col, "x source"))
  )
}
```

::: {.panel-tabset}

#### By admin region

```{r}
#| label: region-nesting-admin
#| fig-width: 11
#| fig-height: 9
h <- make_nesting_heatmaps("admin_region")
if (!is.null(h)) h$topic
```

```{r}
#| label: region-nesting-admin-method
#| fig-width: 11
#| fig-height: 9
if (!is.null(h)) h$method
```

```{r}
#| label: region-nesting-admin-source
#| fig-width: 11
#| fig-height: 9
if (!is.null(h)) h$source
```

#### By country

```{r}
#| label: region-nesting-country
#| fig-width: 11
#| fig-height: 9
h <- make_nesting_heatmaps("country")
if (!is.null(h)) h$topic
```

```{r}
#| label: region-nesting-country-method
#| fig-width: 11
#| fig-height: 9
if (!is.null(h)) h$method
```

```{r}
#| label: region-nesting-country-source
#| fig-width: 11
#| fig-height: 9
if (!is.null(h)) h$source
```

#### By super-region

```{r}
#| label: region-nesting-superregion
#| fig-width: 11
#| fig-height: 9
h <- make_nesting_heatmaps("super_region")
if (!is.null(h)) h$topic
```

```{r}
#| label: region-nesting-superregion-method
#| fig-width: 11
#| fig-height: 9
if (!is.null(h)) h$method
```

```{r}
#| label: region-nesting-superregion-source
#| fig-width: 11
#| fig-height: 9
if (!is.null(h)) h$source
```

:::

### Change over time

Does the corpus's geographic focus shift across the years covered, more
Iceland/Greenland work recently versus more mainland Scandinavia
earlier, or the reverse? Same problem as the other region breakdowns:
most tokens have too few papers spread across too many years to support
a real per-year trend. Restricted to regions with at least 8 tagged
papers (a higher bar than the 5-paper threshold used for the
topic/method/source heatmaps above, since splitting an already-small
group by year further thins each cell), using the same Wilson-interval
and enrichment machinery already defined for Topic, Method, and Source.

```{r}
#| label: region-over-time-setup
region_min_shown_year <- 8

region_n_papers_year_eligible <- region_tokens %>%
  distinct(cite_key, region = Region_normalized) %>%
  count(region, name = "n_region_papers") %>%
  filter(n_region_papers >= region_min_shown_year)

region_long_with_key <- region_tokens %>%
  select(cite_key, region = Region_normalized) %>%
  filter(region %in% region_n_papers_year_eligible$region) %>%
  left_join(df %>% select(cite_key, publication_year), by = "cite_key")

region_corpus_rate <- region_long_with_key %>%
  distinct(cite_key, region) %>%
  count(region, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

region_year_wilson <- region_long_with_key %>%
  distinct(cite_key, publication_year, region) %>%
  count(publication_year, region, name = "n") %>%
  complete(publication_year, region, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(region_corpus_rate, by = "region") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(region, publication_year)
```

```{r}
#| label: region-over-time-count
#| fig-cap: >
#|   Papers per year, per region (regions with 8+ tagged papers only).
#|   Faint bars: few papers that year.
#| fig-width: 11
#| fig-height: 7
ggplot(region_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = "#bf616a") +
  facet_wrap(~ region, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
```

```{r}
#| label: region-over-time-enrichment
#| fig-cap: >
#|   Enrichment per year, per region (regions with 8+ tagged papers
#|   only) — that year's coverage of the region, divided by the region's
#|   own average coverage across the whole corpus (1.0x, dashed line =
#|   normal). Faint bars: few papers that year.
#| fig-width: 11
#| fig-height: 7
ggplot(region_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = "#bf616a") +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ region, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this region's own corpus-wide rate", title = "Enrichment") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
```
