---
title: "BibVik — Part 1: Annotation descriptives"
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

```{css, echo = FALSE}
p {
text-align: justify !important;
}
```

# Part 1 — Annotation descriptives

Strategy and reasoning for this analysis: see `00_plan.qmd`.

This section computes the annotation cross-tab entirely independently of
the graph work above — no reference to citation structure, communities, or
centrality. Per the no-baseline stance, this is not a check against Part 1;
it stands on its own, and Part 3 (below) is where the two are placed
side by side.

```{r}
#| label: annotation-libraries
library(here)
library(tidyverse)
library(patchwork)
library(nord)
library(tidytext)
library(khroma)
library(ggh4x)
```

```{r}
#| label: annotation-palettes
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868"
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]
trend_col <- "#3d3d3d"  # neutral, fixed across sections -- a rolling trend
                         # line/band is a different kind of thing from the
                         # section's own data, so it stays visually distinct
                         # from topic_col/method_col/source_col everywhere
gender_colors <- c("Woman" = nord_palettes$red_mountain[1],
                   "Man" = nord_palettes$algoma_forest[4])
```

```{r}
#| label: annotation-load
df <- read.csv(here("data/annotations.csv"), na.strings = "")

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))

# Exported so 03_joined_descriptions.qmd can load it when rendered
# standalone (outside the combined index.qmd include chain), without
# recomputing Part 1 — consistent with 03's own stated design ("does
# not recompute either" upstream part).
write_csv(df, here("analysis", "bibvik_annotation_df.csv"))
```

```{r}
#| label: annotation-shared-helpers
# Shared across every "by year" and Wilson-faded chart in this
# document (Gender, Topic, Method, Source of Evidence) -- defined once
# here so all of them draw on the same functions and the same
# per-year paper counts.
wilson_lower_bound <- function(x, n, conf = 0.95) {
  z <- qnorm(1 - (1 - conf) / 2)
  phat <- x / n
  denom <- 1 + z^2 / n
  centre <- phat + z^2 / (2 * n)
  adj <- z * sqrt(phat * (1 - phat) / n + z^2 / (4 * n^2))
  (centre - adj) / denom
}
wilson_upper_bound <- function(x, n, conf = 0.95) {
  z <- qnorm(1 - (1 - conf) / 2)
  phat <- x / n
  denom <- 1 + z^2 / n
  centre <- phat + z^2 / (2 * n)
  adj <- z * sqrt(phat * (1 - phat) / n + z^2 / (4 * n^2))
  (centre + adj) / denom
}

# Base-R centered rolling sum (window of 3, shrinking at the edges to
# whatever's available rather than requiring a full window) -- avoids
# adding a package dependency (slider) for one small operation.
roll_sum3 <- function(x) {
  n <- length(x)
  vapply(seq_len(n), function(i) sum(x[max(1, i - 1):min(n, i + 1)]), numeric(1))
}

n_per_year <- df %>% count(publication_year, name = "n_papers_year")
n_total_papers <- n_distinct(df$cite_key)
```

Corpus size, date range, and the publications-per-year chart moved to
`00_overview.qmd`.

## Gender

Author gender, examined two ways: first-author gender and all-author
gender by year, and gender composition across the full author list per
paper.

### By year

```{r}
#| label: gender-by-year
#| fig-cap: >
#|   Top: first author only, one bar per gender per year — faint bars
#|   mean few papers published that year. Bottom: every author
#|   position pooled (order not preserved), so a multi-author paper
#|   contributes more than one tally — faint bars there mean few
#|   author-slots that year, not few papers. Same Wilson-interval fade
#|   used throughout this document, computed against the relevant
#|   denominator for each panel.
#| fig-width: 10
#| fig-height: 8
gender_pubyears <- df %>%
  count(publication_year, Authors..Genders..Author.1., name = "n") %>%
  left_join(n_per_year, by = "publication_year") %>%
  mutate(
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  )

plt1 <-
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1., alpha = -ci_width)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = gender_colors) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = "First author gender by year") +
  theme(legend.position = "bottom")

all_authors_gender_year <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na(gender) %>%
  count(publication_year, gender)

# Denominator here is total author-slots that year, not paper count --
# a multi-author paper contributes more than one gender tally, so
# pooled counts routinely exceed n_papers_year and phat = n/n_papers_year
# would exceed 1, which breaks the Wilson formula (undefined variance).
n_author_slots_per_year <- all_authors_gender_year %>%
  group_by(publication_year) %>%
  summarise(n_slots_year = sum(n), .groups = "drop")

all_authors_gender_year <- all_authors_gender_year %>%
  left_join(n_author_slots_per_year, by = "publication_year") %>%
  mutate(
    ci_lower = wilson_lower_bound(n, n_slots_year),
    ci_upper = wilson_upper_bound(n, n_slots_year),
    ci_width = ci_upper - ci_lower
  )

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender, alpha = -ci_width)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender", title = "All authors' gender by year") +
  theme(legend.position = "bottom")

plt1 / plt2
```

### Composition across the full author list

First-author gender and the order-blind pool above don't capture gender
**composition per paper** (all-male vs. all-female vs. mixed) or **order**
within a paper.

```{r}
#| label: gender-sequence
#| fig-width: 6
#| fig-height: 3
author_gender_cols <- df %>%
  select(starts_with("Authors..Genders..Author.")) %>%
  names()

# Per-paper ordered gender sequence (e.g. "Man, Woman, Man"), preserving
# author order and author count, rather than pooling into an unordered set.
gender_sequences <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(
    cols = all_of(author_gender_cols),
    names_to = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  # Extract numeric author position from the column name for correct
  # ordering regardless of how many positions exist.
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(
    n_authors        = n(),
    gender_sequence  = paste(gender, collapse = ", "),
    first_author_gender = first(gender),
    last_author_gender  = last(gender),
    .groups = "drop"
  )

gender_sequences |>
  count(n_authors) |>
  ggplot(aes(x = factor(n_authors), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Authors per paper", y = "Papers") +
  theme_bw()
```

Several ways of collapsing a per-paper gender sequence into a single
cross-tabbable category are compared below — a strict all-same-vs-mixed
split, a majority-based split, and a version that separates solo
authorship out from multi-author teams (since "all-male" means something
different for n=1 than for a genuinely uniform team).

```{r}
#| label: gender-composition-candidates
#| fig-width: 10
#| fig-height: 3.5
composition_candidates <- gender_sequences %>%
  rowwise() %>%
  mutate(
    genders_list = list(str_split(gender_sequence, ", ")[[1]]),
    n_men       = sum(genders_list == "Man"),
    n_women     = sum(genders_list == "Woman"),
    n_uncertain = sum(genders_list == "Uncertain"),
  ) %>%
  ungroup() %>%
  mutate(
    # Strict: all-same vs. mixed
    composition_strict = case_when(
      n_uncertain > 0            ~ "includes uncertain",
      n_men > 0 & n_women == 0   ~ "all-male",
      n_women > 0 & n_men == 0   ~ "all-female",
      n_men > 0 & n_women > 0    ~ "mixed",
      TRUE                       ~ NA_character_
    ),
    # Majority-based (ties = balanced)
    composition_majority = case_when(
      n_uncertain > 0     ~ "includes uncertain",
      n_men > n_women     ~ "majority-male",
      n_women > n_men     ~ "majority-female",
      n_men == n_women    ~ "balanced",
      TRUE                ~ NA_character_
    ),
    # Single-author papers split out separately, since "all-male"/
    # "all-female" for n=1 is a different phenomenon than for n>1
    composition_singleauthor_split = case_when(
      n_authors == 1 & n_men == 1   ~ "solo-male",
      n_authors == 1 & n_women == 1 ~ "solo-female",
      n_authors == 1                ~ "solo-uncertain",
      n_uncertain > 0                ~ "includes uncertain",
      n_men > 0 & n_women == 0       ~ "all-male (multi-author)",
      n_women > 0 & n_men == 0       ~ "all-female (multi-author)",
      TRUE                           ~ "mixed"
    )
  ) %>%
  select(-genders_list)

bind_rows(
  composition_candidates |> count(value = composition_strict) |> mutate(scheme = "Strict"),
  composition_candidates |> count(value = composition_majority) |> mutate(scheme = "Majority"),
  composition_candidates |> count(value = composition_singleauthor_split) |> mutate(scheme = "Solo split out")
) |>
  mutate(scheme = factor(scheme, levels = c("Strict", "Majority", "Solo split out"))) |>
  ggplot(aes(x = fct_reorder(value, n), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), hjust = -0.3, size = 3) +
  facet_wrap(vars(scheme), scales = "free", ncol = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  coord_flip() +
  labs(x = "", y = "Papers") +
  theme_bw()
```

## Topic

A paper can carry a Primary-coded topic (the coder's judgment of its main
focus) and any number of Secondary-coded topics (also present, but not
the main focus) — see the codebook for the exact scope of each topic
category. Charts below count a paper toward a topic if it carries either
value, so a paper coded with one Primary and two Secondary topics
contributes to three bars — counts and proportions are not mutually
exclusive across topics and will not sum to the corpus total. This
applies throughout this document, not just to Topic.

```{r}
#| label: topics-setup
topic_count <- df %>%
  select(publication_year, contains("Topics")) %>%
  pivot_longer(!publication_year,
               names_to = "topic",
               values_to = "count")

totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n)
```

### Totals

```{r}
#| label: topics-totals
#| fig-cap: >
#|   A paper counts toward a topic once, regardless of how many values
#|   it carries (Primary or Secondary — see above).
#| fig-width: 8
#| fig-height: 7
merged_count <- topic_count %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic) %>%
  mutate(pct = n / sum(n))

ggplot(merged_count, aes(x = reorder(str_wrap(topic, width = 25), n), y = n)) +
  geom_col(fill = topic_col) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Topic", y = "Count") +
  theme_bw() +
  coord_flip()
ggsave("topics_aggregate.png", plot = last_plot(), width = 8, height = 7, dpi = 300)
```

### Over time

```{r}
#| label: topics-over-time-setup
# Built directly from df (not topic_count above, which was pivoted
# without cite_key and can't be used to count distinct papers per topic).
topic_long_with_key <- df %>%
  select(cite_key, publication_year, contains("Topics")) %>%
  pivot_longer(cols = contains("Topics"), names_to = "topic", values_to = "count") %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " "))

topic_corpus_rate <- topic_long_with_key %>%
  distinct(cite_key, topic) %>%
  count(topic, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

topic_year_wilson <- topic_long_with_key %>%
  distinct(cite_key, publication_year, topic) %>%
  count(publication_year, topic, name = "n") %>%
  complete(publication_year, topic, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(topic_corpus_rate, by = "topic") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(topic, publication_year)

# Rolling: pool counts over a 3-year window, then compute the Wilson CI
# on the POOLED proportion directly (not by averaging per-year CIs).
# ylim(0,4) on the enrichment charts below clips extreme ratios that
# come from topics with a very low corpus_rate (e.g. Heritage)
# inflating the enrichment-scaled interval -- clipping is done
# per-panel via coord_cartesian, not via a shared alpha scale, so one
# topic's wide interval cannot distort the visual reading of every
# other topic's panel.
topic_year_roll <- topic_year_wilson %>%
  arrange(topic, publication_year) %>%
  group_by(topic) %>%
  mutate(
    roll_n = roll_sum3(n),
    roll_n_papers_year = roll_sum3(n_papers_year),
    roll_enrichment = (roll_n / roll_n_papers_year) / corpus_rate,
    roll_ci_lower = wilson_lower_bound(roll_n, roll_n_papers_year) / corpus_rate,
    roll_ci_upper = wilson_upper_bound(roll_n, roll_n_papers_year) / corpus_rate
  ) %>%
  ungroup()
```

::: {.column-margin}
**Count.** Raw papers per year, nothing normalized.

**Confidence.** Bar and band opacity come from a 95% Wilson interval —
faint means too few papers behind that estimate to trust it,
solid means there's enough. Fades continuously, no hard cutoff.
:::

```{r}
#| label: topics-over-time-count
#| fig-cap: "Papers per year, per topic. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 10
ggplot(topic_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = "#6a9870") +
  facet_wrap(~ topic, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("topics_count_by_year.png", plot = last_plot(), width = 11, height = 10, dpi = 300)
```

::: {.panel-tabset}

#### By year

```{r}
#| label: topics-over-time-enrichment
#| fig-cap: "Enrichment per year, per topic. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 10
ggplot(topic_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = topic_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ topic, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this topic's own corpus-wide rate", title = "Enrichment — by year") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("topics_enrichment_by_year.png", plot = last_plot(), width = 11, height = 10, dpi = 300)
```

#### Smoothed trend

```{r}
#| label: topics-over-time-enrichment-smoothed
#| fig-cap: "Same enrichment bars, with a 3-year rolling average and its own confidence band."
#| fig-width: 11
#| fig-height: 10
ggplot(topic_year_roll, aes(x = publication_year)) +
  geom_col(aes(y = enrichment, alpha = -ci_width), fill = topic_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  geom_ribbon(aes(ymin = pmax(roll_ci_lower, 0), ymax = pmin(roll_ci_upper, 4), group = topic),
              fill = trend_col, alpha = 0.18) +
  geom_line(aes(y = roll_enrichment, group = topic), colour = trend_col, linewidth = 0.8) +
  facet_wrap(~ topic, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this topic's own corpus-wide rate", title = "Enrichment — smoothed trend") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("topics_enrichment_by_year_smoothed.png", plot = last_plot(), width = 11, height = 10, dpi = 300)
```

:::

::: {.column-margin}
**Enrichment.** That year's coverage of the topic, divided by the
topic's own average coverage across the whole corpus (1.0x, dashed
line = normal). Each topic is checked against its own average, not
against other topics.

**Smoothed trend.** Replaces individual years with a 3-year rolling
average and its own confidence band, to make direction easier to read
through the year-to-year noise.

**Y-axis.** Capped at 4x on both tabs so the bars stay readable — a
few bars run higher than that and get cut off at the top.
:::

### Co-occurrence

```{r}
#| label: topics-cooccurrence-setup
topics <- df %>%
  select(contains("Topics")) %>%
  rename_with(~str_remove(., "Topics.."))
```

::: {.column-margin}
**Reading a facet.** Each facet is one topic; its bars are every
other topic, ranked by how often it co-occurs with the facet topic.
Each facet's bars sum to 1 on their own, so proportions aren't
comparable across facets without also checking each facet's n.
:::

```{r}
#| label: topics-cooccurrence-directional
#| fig-cap: "Directional: given a paper's Primary-coded topic (the facet), how often is each other topic marked Secondary on the same paper."
#| fig-width: 30
#| fig-height: 20
coocurrences <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

```{r}
#| label: topics-cooccurrence-undirected
#| fig-cap: "Undirected: which topics appear together at all, regardless of rank — present if either value is carried, pairing symmetric."
#| fig-width: 30
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Topic.A", "Topic.B")) %>%
  filter(Topic.A != Topic.B) %>%
  mutate(Topic.A = str_wrap(str_replace_all(Topic.A, fixed("."), " "), width = 20)) %>%
  mutate(Topic.B = str_replace_all(Topic.B, fixed("."), " ")) %>%
  group_by(Topic.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Topic.A, Prop, Topic.B), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Topic.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

### By first-author gender

```{r}
#| label: topic-by-gender
#| fig-cap: >
#|   Proportion of papers by first-author gender, within each topic.
#|   Topics are labelled with their n; horizontal lines mark 25/50/75%
#|   for reference. Raw proportions, not corrected for the corpus-wide
#|   gender split.
#| fig-width: 10
#| fig-height: 8
topic_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Topics")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

merged_topic_gender <- topic_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, gender_first_author) %>%
  group_by(topic) %>%
  mutate(prop_merged = n / sum(n))

merged_topic_gender |>
  group_by(topic) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(topic = fct_reorder(topic, ratio),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("topic_by_gender_merged.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### By full-author-list gender composition

Every paper's authors, sorted into one of five groups: solo female
author, solo male author, an all-female team, an all-male team, or a
mixed-gender team. Solo authorship is kept separate from team
composition — a single female author isn't the same thing as a team
that happens to be all women, and lumping them together would hide
that difference.

```{r}
#| label: composition-by-topic
#| fig-cap: >
#|   Proportion of papers by author-list gender composition (solo
#|   female/male, all-female/all-male team, mixed team), within each
#|   topic. Horizontal lines mark 25/50/75% for reference.
#| fig-width: 10
#| fig-height: 8
composition_topic <- composition_candidates %>%
  select(cite_key, composition_singleauthor_split) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic), !is.na(composition_singleauthor_split))

comp_topic_summary <- composition_topic %>%
  count(topic, composition_singleauthor_split) %>%
  group_by(topic) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

comp_topic_female_share <- comp_topic_summary %>%
  filter(composition_singleauthor_split %in% c("solo-female", "all-female (multi-author)")) %>%
  group_by(topic) %>%
  summarise(female_share = sum(prop), .groups = "drop")

comp_topic_summary %>%
  left_join(comp_topic_female_share, by = "topic") %>%
  mutate(female_share = replace_na(female_share, 0),
         topic = fct_reorder(topic, female_share),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, female_share),
         composition_singleauthor_split = factor(composition_singleauthor_split,
           levels = c("solo-female", "all-female (multi-author)", "mixed",
                      "all-male (multi-author)", "solo-male"))) %>%
  ggplot(aes(x = topic_label, y = prop, fill = composition_singleauthor_split)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "black", linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "solo-female" = "#bf616a", "all-female (multi-author)" = "#d89898",
    "mixed" = "#a0a8b8",
    "all-male (multi-author)" = "#8fbc8f", "solo-male" = "#4f6a31")) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("composition_by_topic.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### Mixed-gender authorship only

::: {.column-margin}
**Green dot.** Share of that topic's mixed-gender papers with a woman
first author, sized by number of mixed papers.

**Tan dots.** Every distinct team size that occurs among that topic's
mixed papers, sized by how many papers share that size — no jitter,
no averaging, so a big tan dot at one team size means many papers,
not several dots scattered near it.
:::

```{r}
#| label: mixed-only-topic
#| fig-cap: >
#|   First-author gender within mixed-gender papers, by topic (green,
#|   bottom axis). Every distinct mixed-team size for that topic, and
#|   how many papers share it (tan, top axis). Dashed line at 50%.
#| fig-width: 11.5
#| fig-height: 9.5
mixed_topic_raw <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(all_of(author_gender_cols), names_to = "position", values_to = "gender") %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(n_authors = n(), n_women = sum(gender == "Woman"), n_men = sum(gender == "Man"),
            first = first(gender), .groups = "drop") %>%
  filter(n_women > 0, n_men > 0) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic)) %>%
  mutate(is_woman_first = as.numeric(first == "Woman"))

mixed_topic_summary <- mixed_topic_raw %>%
  group_by(topic) %>%
  summarise(prop = mean(is_woman_first), total = n(), .groups = "drop") %>%
  arrange(prop) %>%
  mutate(topic_label = paste0(str_wrap(topic, 22), " (n=", total, ")"),
         topic_label = factor(topic_label, levels = topic_label))

mixed_topic_teamsize_counts <- mixed_topic_raw %>%
  count(topic, n_authors, name = "n_at_size") %>%
  left_join(mixed_topic_summary %>% select(topic, topic_label), by = "topic")

mixed_topic_size_range <- range(mixed_topic_raw$n_authors)

ggplot() +
  geom_vline(xintercept = 0.5, colour = "grey70", linetype = "dashed") +
  geom_point(data = mixed_topic_teamsize_counts,
             aes(x = (n_authors - mixed_topic_size_range[1]) / diff(mixed_topic_size_range),
                 y = topic_label, size = n_at_size),
             colour = "#c8a87a", alpha = 0.8) +
  geom_point(data = mixed_topic_summary, aes(x = prop, y = topic_label, size = total), colour = topic_col) +
  scale_x_continuous(labels = scales::percent, limits = c(-0.02, 1.02),
    sec.axis = sec_axis(~ . * diff(mixed_topic_size_range) + mixed_topic_size_range[1],
                         name = "Team size (tan)",
                         breaks = c(mixed_topic_size_range[1], 5, 10, 15, mixed_topic_size_range[2]))) +
  scale_size_continuous(range = c(1.5, 8), name = "papers at\nthat size") +
  theme_bw() +
  labs(x = "Share of mixed papers with a woman first author (green)", y = "")
ggsave("mixed_firstauthor_by_topic.png", plot = last_plot(), width = 11.5, height = 9.5, dpi = 300)
```

## Method

```{r}
#| label: methods-setup
method_count <- df %>%
  select(publication_year, contains("Methods")) %>%
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")
```

### Totals

```{r}
#| label: methods-totals
#| fig-cap: "A paper counts toward a method once, regardless of Primary/Secondary status."
#| fig-width: 8
#| fig-height: 7
merged_count_method <- method_count %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method) %>%
  mutate(pct = n / sum(n))

ggplot(merged_count_method, aes(x = reorder(str_wrap(method, width = 25), n), y = n)) +
  geom_col(fill = method_col) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Method", y = "Count") +
  theme_bw() +
  coord_flip()
ggsave("methods_aggregate.png", plot = last_plot(), width = 8, height = 7, dpi = 300)
```

### Over time

```{r}
#| label: methods-over-time-setup
method_long_with_key <- df %>%
  select(cite_key, publication_year, contains("Methods")) %>%
  pivot_longer(cols = contains("Methods"), names_to = "method", values_to = "count") %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " "))

method_corpus_rate <- method_long_with_key %>%
  distinct(cite_key, method) %>%
  count(method, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

method_year_wilson <- method_long_with_key %>%
  distinct(cite_key, publication_year, method) %>%
  count(publication_year, method, name = "n") %>%
  complete(publication_year, method, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(method_corpus_rate, by = "method") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(method, publication_year)

method_year_roll <- method_year_wilson %>%
  arrange(method, publication_year) %>%
  group_by(method) %>%
  mutate(
    roll_n = roll_sum3(n),
    roll_n_papers_year = roll_sum3(n_papers_year),
    roll_enrichment = (roll_n / roll_n_papers_year) / corpus_rate,
    roll_ci_lower = wilson_lower_bound(roll_n, roll_n_papers_year) / corpus_rate,
    roll_ci_upper = wilson_upper_bound(roll_n, roll_n_papers_year) / corpus_rate
  ) %>%
  ungroup()
```

::: {.column-margin}
**Count.** Raw papers per year, nothing normalized.

**Confidence.** Bar and band opacity come from a 95% Wilson interval —
faint means too few papers behind that estimate to trust it,
solid means there's enough. Fades continuously, no hard cutoff.
:::

```{r}
#| label: methods-over-time-count
#| fig-cap: "Papers per year, per method. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 8
ggplot(method_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = method_col) +
  facet_wrap(~ method, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("methods_count_by_year.png", plot = last_plot(), width = 11, height = 8, dpi = 300)
```

::: {.panel-tabset}

#### By year

```{r}
#| label: methods-over-time-enrichment
#| fig-cap: "Enrichment per year, per method. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 8
ggplot(method_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = method_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ method, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this method's own corpus-wide rate", title = "Enrichment — by year") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("methods_enrichment_by_year.png", plot = last_plot(), width = 11, height = 8, dpi = 300)
```

#### Smoothed trend

```{r}
#| label: methods-over-time-enrichment-smoothed
#| fig-cap: "Same enrichment bars, with a 3-year rolling average and its own confidence band."
#| fig-width: 11
#| fig-height: 8
ggplot(method_year_roll, aes(x = publication_year)) +
  geom_col(aes(y = enrichment, alpha = -ci_width), fill = method_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  geom_ribbon(aes(ymin = pmax(roll_ci_lower, 0), ymax = pmin(roll_ci_upper, 4), group = method),
              fill = trend_col, alpha = 0.18) +
  geom_line(aes(y = roll_enrichment, group = method), colour = trend_col, linewidth = 0.8) +
  facet_wrap(~ method, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this method's own corpus-wide rate", title = "Enrichment — smoothed trend") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("methods_enrichment_by_year_smoothed.png", plot = last_plot(), width = 11, height = 8, dpi = 300)
```

:::

::: {.column-margin}
**Enrichment.** That year's coverage of the method, divided by the
method's own average coverage across the whole corpus (1.0x, dashed
line = normal). Each method is checked against its own average, not
against other methods.

**Smoothed trend.** Replaces individual years with a 3-year rolling
average and its own confidence band, to make direction easier to read
through the year-to-year noise.

**Y-axis.** Capped at 4x on both tabs so the bars stay readable — a
few bars run higher than that and get cut off at the top.
:::

### Co-occurrence

```{r}
#| label: methods-cooccurrence-setup
methods <- df %>%
  select(contains("Methods")) %>%
  rename_with(~str_remove(., "Methods.."))
```

::: {.column-margin}
**Reading a facet.** Each facet is one method; its bars are every
other method, ranked by how often it co-occurs with the facet method.
Each facet's bars sum to 1 on their own, so proportions aren't
comparable across facets without also checking each facet's n.
:::

```{r}
#| label: methods-cooccurrence-directional
#| fig-cap: "Directional: given a paper's Primary-coded method (the facet), how often is each other method marked Secondary on the same paper."
#| fig-width: 20
#| fig-height: 20
coocurrences <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

```{r}
#| label: methods-cooccurrence-undirected
#| fig-cap: "Undirected: which methods appear together at all, regardless of rank — present if either value is carried, pairing symmetric."
#| fig-width: 20
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Method.A", "Method.B")) %>%
  filter(Method.A != Method.B) %>%
  mutate(Method.A = str_wrap(str_replace_all(Method.A, fixed("."), " "), width = 20)) %>%
  mutate(Method.B = str_replace_all(Method.B, fixed("."), " ")) %>%
  group_by(Method.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Method.A, Prop, Method.B), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Method.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

### By first-author gender

```{r}
#| label: method-by-gender
#| fig-cap: >
#|   Proportion of papers by first-author gender, within each method.
#|   Methods are labelled with their n; horizontal lines mark 25/50/75%
#|   for reference. Raw proportions, not corrected for the corpus-wide
#|   gender split.
#| fig-width: 10
#| fig-height: 8
method_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Methods")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

merged_method_gender <- method_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, gender_first_author) %>%
  group_by(method) %>%
  mutate(prop_merged = n / sum(n))

merged_method_gender |>
  group_by(method) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(method = fct_reorder(method, ratio),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("method_by_gender_merged.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### By full-author-list gender composition

Every paper's authors, sorted into one of five groups: solo female
author, solo male author, an all-female team, an all-male team, or a
mixed-gender team. Solo authorship is kept separate from team
composition, so a single author's gender isn't treated as if it were
a team's composition.

```{r}
#| label: composition-by-method
#| fig-cap: >
#|   Proportion of papers by author-list gender composition (solo
#|   female/male, all-female/all-male team, mixed team), within each
#|   method. Horizontal lines mark 25/50/75% for reference.
#| fig-width: 10
#| fig-height: 8
composition_method <- composition_candidates %>%
  select(cite_key, composition_singleauthor_split) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method), !is.na(composition_singleauthor_split))

comp_method_summary <- composition_method %>%
  count(method, composition_singleauthor_split) %>%
  group_by(method) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

comp_method_female_share <- comp_method_summary %>%
  filter(composition_singleauthor_split %in% c("solo-female", "all-female (multi-author)")) %>%
  group_by(method) %>%
  summarise(female_share = sum(prop), .groups = "drop")

comp_method_summary %>%
  left_join(comp_method_female_share, by = "method") %>%
  mutate(female_share = replace_na(female_share, 0),
         method = fct_reorder(method, female_share),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, female_share),
         composition_singleauthor_split = factor(composition_singleauthor_split,
           levels = c("solo-female", "all-female (multi-author)", "mixed",
                      "all-male (multi-author)", "solo-male"))) %>%
  ggplot(aes(x = method_label, y = prop, fill = composition_singleauthor_split)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "black", linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "solo-female" = "#bf616a", "all-female (multi-author)" = "#d89898",
    "mixed" = "#a0a8b8",
    "all-male (multi-author)" = "#8fbc8f", "solo-male" = "#4f6a31")) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("composition_by_method.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### Mixed-gender authorship only

::: {.column-margin}
**Navy dot.** Share of that method's mixed-gender papers with a woman
first author, sized by number of mixed papers.

**Tan dots.** Every distinct team size that occurs among that method's
mixed papers, sized by how many papers share it — no jitter, no
averaging.
:::

```{r}
#| label: mixed-only-method
#| fig-cap: >
#|   First-author gender within mixed-gender papers, by method (navy,
#|   bottom axis). Every distinct mixed-team size for that method, and
#|   how many papers share it (tan, top axis). Dashed line at 50%.
#| fig-width: 11
#| fig-height: 7
mixed_method_raw <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(all_of(author_gender_cols), names_to = "position", values_to = "gender") %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(n_authors = n(), n_women = sum(gender == "Woman"), n_men = sum(gender == "Man"),
            first = first(gender), .groups = "drop") %>%
  filter(n_women > 0, n_men > 0) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method)) %>%
  mutate(is_woman_first = as.numeric(first == "Woman"))

mixed_method_summary <- mixed_method_raw %>%
  group_by(method) %>%
  summarise(prop = mean(is_woman_first), total = n(), .groups = "drop") %>%
  arrange(prop) %>%
  mutate(method_label = paste0(str_wrap(method, 22), " (n=", total, ")"),
         method_label = factor(method_label, levels = method_label))

mixed_method_teamsize_counts <- mixed_method_raw %>%
  count(method, n_authors, name = "n_at_size") %>%
  left_join(mixed_method_summary %>% select(method, method_label), by = "method")

mixed_method_size_range <- range(mixed_method_raw$n_authors)

ggplot() +
  geom_vline(xintercept = 0.5, colour = "grey70", linetype = "dashed") +
  geom_point(data = mixed_method_teamsize_counts,
             aes(x = (n_authors - mixed_method_size_range[1]) / diff(mixed_method_size_range),
                 y = method_label, size = n_at_size),
             colour = "#c8a87a", alpha = 0.8) +
  geom_point(data = mixed_method_summary, aes(x = prop, y = method_label, size = total), colour = method_col) +
  scale_x_continuous(labels = scales::percent, limits = c(-0.02, 1.02),
    sec.axis = sec_axis(~ . * diff(mixed_method_size_range) + mixed_method_size_range[1],
                         name = "Team size (tan)",
                         breaks = c(mixed_method_size_range[1], 5, 10, 15, mixed_method_size_range[2]))) +
  scale_size_continuous(range = c(1.5, 8), name = "papers at\nthat size") +
  theme_bw() +
  labs(x = "Share of mixed papers with a woman first author (navy)", y = "")
ggsave("mixed_firstauthor_by_method.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

## Source of evidence

```{r}
#| label: sources-setup
source_count <- df %>%
  select(publication_year, contains("Sources.of.Evidence")) %>%
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")
```

### Totals

```{r}
#| label: sources-totals
#| fig-cap: "A paper counts toward a source once, regardless of Primary/Secondary status."
#| fig-width: 8
#| fig-height: 7
merged_source_count <- source_count %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source) %>%
  mutate(pct = n / sum(n))

ggplot(merged_source_count, aes(x = reorder(str_wrap(source, width = 25), n), y = n)) +
  geom_col(fill = source_col) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Source of evidence", y = "Count") +
  theme_bw() +
  coord_flip()
ggsave("sources_aggregate.png", plot = last_plot(), width = 8, height = 7, dpi = 300)
```

### Over time

```{r}
#| label: sources-over-time-setup
source_long_with_key <- df %>%
  select(cite_key, publication_year, contains("Sources.of.Evidence")) %>%
  pivot_longer(cols = contains("Sources.of.Evidence"), names_to = "source", values_to = "count") %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " "))

source_corpus_rate <- source_long_with_key %>%
  distinct(cite_key, source) %>%
  count(source, name = "n_corpus") %>%
  mutate(corpus_rate = n_corpus / n_total_papers)

source_year_wilson <- source_long_with_key %>%
  distinct(cite_key, publication_year, source) %>%
  count(publication_year, source, name = "n") %>%
  complete(publication_year, source, fill = list(n = 0)) %>%
  left_join(n_per_year, by = "publication_year") %>%
  left_join(source_corpus_rate, by = "source") %>%
  mutate(
    prop = n / n_papers_year,
    enrichment = prop / corpus_rate,
    ci_lower = wilson_lower_bound(n, n_papers_year),
    ci_upper = wilson_upper_bound(n, n_papers_year),
    ci_width = ci_upper - ci_lower
  ) %>%
  arrange(source, publication_year)

source_year_roll <- source_year_wilson %>%
  arrange(source, publication_year) %>%
  group_by(source) %>%
  mutate(
    roll_n = roll_sum3(n),
    roll_n_papers_year = roll_sum3(n_papers_year),
    roll_enrichment = (roll_n / roll_n_papers_year) / corpus_rate,
    roll_ci_lower = wilson_lower_bound(roll_n, roll_n_papers_year) / corpus_rate,
    roll_ci_upper = wilson_upper_bound(roll_n, roll_n_papers_year) / corpus_rate
  ) %>%
  ungroup()
```

::: {.column-margin}
**Count.** Raw papers per year, nothing normalized.

**Confidence.** Bar and band opacity come from a 95% Wilson interval —
faint means too few papers behind that estimate to trust it,
solid means there's enough. Fades continuously, no hard cutoff.
:::

```{r}
#| label: sources-over-time-count
#| fig-cap: "Papers per year, per source. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 7
ggplot(source_year_wilson, aes(x = publication_year, y = n, alpha = -ci_width)) +
  geom_col(fill = source_col) +
  facet_wrap(~ source, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  labs(x = "", y = "Papers published that year", title = "Count") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("sources_count_by_year.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

::: {.panel-tabset}

#### By year

```{r}
#| label: sources-over-time-enrichment
#| fig-cap: "Enrichment per year, per source. Faint bars: few papers that year."
#| fig-width: 11
#| fig-height: 7
ggplot(source_year_wilson, aes(x = publication_year, y = enrichment, alpha = -ci_width)) +
  geom_col(fill = source_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  facet_wrap(~ source, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this source's own corpus-wide rate", title = "Enrichment — by year") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("sources_enrichment_by_year.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

#### Smoothed trend

```{r}
#| label: sources-over-time-enrichment-smoothed
#| fig-cap: "Same enrichment bars, with a 3-year rolling average and its own confidence band."
#| fig-width: 11
#| fig-height: 7
ggplot(source_year_roll, aes(x = publication_year)) +
  geom_col(aes(y = enrichment, alpha = -ci_width), fill = source_col) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40", linewidth = 0.3) +
  geom_ribbon(aes(ymin = pmax(roll_ci_lower, 0), ymax = pmin(roll_ci_upper, 4), group = source),
              fill = trend_col, alpha = 0.18) +
  geom_line(aes(y = roll_enrichment, group = source), colour = trend_col, linewidth = 0.8) +
  facet_wrap(~ source, ncol = 4) +
  scale_alpha_continuous(range = c(0.25, 1), guide = "none") +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "", y = "Enrichment vs. this source's own corpus-wide rate", title = "Enrichment — smoothed trend") +
  theme_bw() +
  theme(strip.text = element_text(size = 8), axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        plot.title = element_text(size = 11))
ggsave("sources_enrichment_by_year_smoothed.png", plot = last_plot(), width = 11, height = 7, dpi = 300)
```

:::

::: {.column-margin}
**Enrichment.** That year's coverage of the source, divided by the
source's own average coverage across the whole corpus (1.0x, dashed
line = normal). Each source is checked against its own average, not
against other sources.

**Smoothed trend.** Replaces individual years with a 3-year rolling
average and its own confidence band, to make direction easier to read
through the year-to-year noise.

**Y-axis.** Capped at 4x on both tabs so the bars stay readable — a
few bars run higher than that and get cut off at the top.
:::

### Co-occurrence

```{r}
#| label: sources-cooccurrence-setup
sources <- df %>%
  select(contains("Sources.of.Evidence")) %>%
  rename_with(~str_remove(., "Sources.of.Evidence.."))
```

::: {.column-margin}
**Reading a facet.** Each facet is one source; its bars are every
other source, ranked by how often it co-occurs with the facet source.
Each facet's bars sum to 1 on their own, so proportions aren't
comparable across facets without also checking each facet's n.
:::

```{r}
#| label: sources-cooccurrence-directional
#| fig-cap: "Directional: given a paper's Primary-coded source (the facet), how often is each other source marked Secondary on the same paper."
#| fig-width: 20
#| fig-height: 20
coocurrences <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

```{r}
#| label: sources-cooccurrence-undirected
#| fig-cap: "Undirected: which sources appear together at all, regardless of rank — present if either value is carried, pairing symmetric."
#| fig-width: 20
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Source.A", "Source.B")) %>%
  filter(Source.A != Source.B) %>%
  mutate(Source.A = str_wrap(str_replace_all(Source.A, fixed("."), " "), width = 20)) %>%
  mutate(Source.B = str_replace_all(Source.B, fixed("."), " ")) %>%
  group_by(Source.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Source.A, Prop, Source.B), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Source.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "")
```

### By first-author gender

```{r}
#| label: source-by-gender
#| fig-cap: >
#|   Proportion of papers by first-author gender, within each source.
#|   Sources are labelled with their n; horizontal lines mark 25/50/75%
#|   for reference. Raw proportions, not corrected for the corpus-wide
#|   gender split.
#| fig-width: 10
#| fig-height: 8
source_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Sources")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

merged_source_gender <- source_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, gender_first_author) %>%
  group_by(source) %>%
  mutate(prop_merged = n / sum(n))

merged_source_gender |>
  group_by(source) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(source = fct_reorder(source, ratio),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("source_by_gender_merged.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### By full-author-list gender composition

Every paper's authors, sorted into one of five groups: solo female
author, solo male author, an all-female team, an all-male team, or a
mixed-gender team. Solo authorship is kept separate from team
composition, so a single author's gender isn't treated as if it were
a team's composition.

```{r}
#| label: composition-by-source
#| fig-cap: >
#|   Proportion of papers by author-list gender composition (solo
#|   female/male, all-female/all-male team, mixed team), within each
#|   source of evidence. Horizontal lines mark 25/50/75% for reference.
#| fig-width: 10
#| fig-height: 8
composition_source <- composition_candidates %>%
  select(cite_key, composition_singleauthor_split) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source), !is.na(composition_singleauthor_split))

comp_source_summary <- composition_source %>%
  count(source, composition_singleauthor_split) %>%
  group_by(source) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

comp_source_female_share <- comp_source_summary %>%
  filter(composition_singleauthor_split %in% c("solo-female", "all-female (multi-author)")) %>%
  group_by(source) %>%
  summarise(female_share = sum(prop), .groups = "drop")

comp_source_summary %>%
  left_join(comp_source_female_share, by = "source") %>%
  mutate(female_share = replace_na(female_share, 0),
         source = fct_reorder(source, female_share),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, female_share),
         composition_singleauthor_split = factor(composition_singleauthor_split,
           levels = c("solo-female", "all-female (multi-author)", "mixed",
                      "all-male (multi-author)", "solo-male"))) %>%
  ggplot(aes(x = source_label, y = prop, fill = composition_singleauthor_split)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), colour = "black", linewidth = 0.4) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c(
    "solo-female" = "#bf616a", "all-female (multi-author)" = "#d89898",
    "mixed" = "#a0a8b8",
    "all-male (multi-author)" = "#8fbc8f", "solo-male" = "#4f6a31")) +
  coord_flip() +
  theme_bw() +
  labs(x = "", y = "Proportion", fill = NULL)
ggsave("composition_by_source.png", plot = last_plot(), width = 10, height = 8, dpi = 300)
```

### Mixed-gender authorship only

::: {.column-margin}
**Red dot.** Share of that source's mixed-gender papers with a woman
first author, sized by number of mixed papers.

**Tan dots.** Every distinct team size that occurs among that source's
mixed papers, sized by how many papers share it — no jitter, no
averaging.
:::

```{r}
#| label: mixed-only-source
#| fig-cap: >
#|   First-author gender within mixed-gender papers, by source of
#|   evidence (red, bottom axis). Every distinct mixed-team size for
#|   that source, and how many papers share it (tan, top axis).
#|   Dashed line at 50%.
#| fig-width: 11
#| fig-height: 6.5
mixed_source_raw <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(all_of(author_gender_cols), names_to = "position", values_to = "gender") %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(n_authors = n(), n_women = sum(gender == "Woman"), n_men = sum(gender == "Man"),
            first = first(gender), .groups = "drop") %>%
  filter(n_women > 0, n_men > 0) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source)) %>%
  mutate(is_woman_first = as.numeric(first == "Woman"))

mixed_source_summary <- mixed_source_raw %>%
  group_by(source) %>%
  summarise(prop = mean(is_woman_first), total = n(), .groups = "drop") %>%
  arrange(prop) %>%
  mutate(source_label = paste0(str_wrap(source, 22), " (n=", total, ")"),
         source_label = factor(source_label, levels = source_label))

mixed_source_teamsize_counts <- mixed_source_raw %>%
  count(source, n_authors, name = "n_at_size") %>%
  left_join(mixed_source_summary %>% select(source, source_label), by = "source")

mixed_source_size_range <- range(mixed_source_raw$n_authors)

ggplot() +
  geom_vline(xintercept = 0.5, colour = "grey70", linetype = "dashed") +
  geom_point(data = mixed_source_teamsize_counts,
             aes(x = (n_authors - mixed_source_size_range[1]) / diff(mixed_source_size_range),
                 y = source_label, size = n_at_size),
             colour = "#c8a87a", alpha = 0.8) +
  geom_point(data = mixed_source_summary, aes(x = prop, y = source_label, size = total), colour = source_col) +
  scale_x_continuous(labels = scales::percent, limits = c(-0.02, 1.02),
    sec.axis = sec_axis(~ . * diff(mixed_source_size_range) + mixed_source_size_range[1],
                         name = "Team size (tan)",
                         breaks = c(mixed_source_size_range[1], 5, 10, 15, mixed_source_size_range[2]))) +
  scale_size_continuous(range = c(1.5, 8), name = "papers at\nthat size") +
  theme_bw() +
  labs(x = "Share of mixed papers with a woman first author (red)", y = "")
ggsave("mixed_firstauthor_by_source.png", plot = last_plot(), width = 11, height = 6.5, dpi = 300)
```

## Region

```{r}
#| label: region
library(ggwordcloud)

region_counts <- df %>%
  pull(Region) %>%
  na.omit() %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Region", "n")) %>%
  mutate(n = as.numeric(n),
         n_scaled = n + 3) %>%
  arrange(desc(n))

ggplot(region_counts, aes(label = Region, size = n_scaled, color = n)) +
  geom_text_wordcloud(area_corr = TRUE, rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  scale_color_gradient(low = topic_col, high = "darkred") +
  theme_minimal()
```

## Not yet done

### Gender

- Composition scheme for the topic/method/source cross-tabs — three
  candidate schemes are computed (strict same/mixed, majority-based,
  and solo-vs-team split), and the third is used throughout this
  document's charts. Whether that's the right one to settle on hasn't
  been decided.
- Topic/method/source-by-gender proportions are raw, not normalized
  against the corpus-wide gender base rate.
- Gender's By-year and Composition subsections have captions and
  Wilson fade; no margin notes.
- Mixed-gender team size is shown per topic/method/source as every
  paper's own size (no averaging); the same has not been built for
  Gender's own corpus-level sections.

### Region

- Word cloud from the free-text Region field, no normalization —
  spelling variants and different specificity levels for the same
  place count as distinct.
- Semantic similarity matching against a reference gazetteer (e.g.
  Pleiades, https://pleiades.stoa.us) could consolidate variant place
  names and add coordinates for a map. Not started.
