---
title: "BibVik — Part 1: Annotation descriptives (F1 corpus)"
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

# Part 1 — Annotation descriptives (F1 corpus)

Strategy and reasoning for this analysis: see `00_plan.qmd`.

This section computes the annotation cross-tab entirely independently of
the graph work above — no reference to citation structure, communities, or
centrality. Per the no-baseline stance, this is not a check against Part 1;
it stands on its own, and Part 3 (below) is where the two are placed
side by side.

```{r}
#| label: annotation-libraries
library(here)
library(tidyverse)
library(patchwork)
library(nord)
library(tidytext)
library(khroma)
library(ggh4x)
```

```{r}
#| label: annotation-palettes
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868"
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]
gender_colors <- c("Woman" = nord_palettes$red_mountain[1],
                   "Man" = nord_palettes$algoma_forest[4])
```

```{r}
#| label: annotation-load
df <- read.csv(here("data/annotations.csv"), na.strings = "")

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))

# Exported so 03_joined_descriptions.qmd can load it when rendered
# standalone (outside the combined index.qmd include chain), without
# recomputing Part 1 — consistent with 03's own stated design ("does
# not recompute either" upstream part).
write_csv(df, here("analysis", "bibvik_annotation_df.csv"))
```

## Publications per year

```{r}
#| label: pubyears
pubyears <- df %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat="identity", fill = nord_palettes$aurora[1]) +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE ) +
  labs(x = "", y = "Count", title = "Publications by year")
```

## Gender

Author gender, examined three ways: first-author gender and all-author
gender by year; gender composition across the full author list per
paper; and (for a subset where a lead-author convention can be
established) lead-author gender.

### By year

```{r}
#| label: gender-by-year
#| fig-width: 10
#| fig-height: 8
gender_pubyears <- df %>%
  group_by(Authors..Genders..Author.1.) %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

plt1 <-
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1.)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = gender_colors) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = "First author gender by year") +
  theme(legend.position = "bottom")

all_authors_gender_year <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(publication_year, gender) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na()

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender", title = "All authors' gender by year") +
  theme(legend.position = "bottom")

plt1 / plt2
```

### Composition across the full author list

First-author gender and the order-blind pool above don't capture gender
**composition per paper** (all-male vs. all-female vs. mixed) or **order**
within a paper. Author order is genuinely ambiguous here: field
convention varies (first-author-as-lead for fieldwork, last-author-as-lead
for lab-based work), and which applies depends on a paper's own method —
see [Lead-author gender](#lead-author-gender) below, where that mapping
is still a placeholder.

```{r}
#| label: gender-sequence
#| fig-width: 6
#| fig-height: 3
author_gender_cols <- df %>%
  select(starts_with("Authors..Genders..Author.")) %>%
  names()

# Per-paper ordered gender sequence (e.g. "Man, Woman, Man"), preserving
# author order and author count, rather than pooling into an unordered set.
gender_sequences <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(
    cols = all_of(author_gender_cols),
    names_to = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  # Extract numeric author position from the column name for correct
  # ordering regardless of how many positions exist.
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(
    n_authors        = n(),
    gender_sequence  = paste(gender, collapse = ", "),
    first_author_gender = first(gender),
    last_author_gender  = last(gender),
    .groups = "drop"
  )

gender_sequences |>
  count(n_authors) |>
  ggplot(aes(x = factor(n_authors), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), vjust = -0.3, size = 3.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Authors per paper", y = "Papers") +
  theme_bw()
```

Several ways of collapsing a per-paper gender sequence into a single
cross-tabbable category are compared below — a strict all-same-vs-mixed
split, a majority-based split, and a version that separates solo
authorship out from multi-author teams (since "all-male" means something
different for n=1 than for a genuinely uniform team).

```{r}
#| label: gender-composition-candidates
#| fig-width: 10
#| fig-height: 3.5
composition_candidates <- gender_sequences %>%
  rowwise() %>%
  mutate(
    genders_list = list(str_split(gender_sequence, ", ")[[1]]),
    n_men       = sum(genders_list == "Man"),
    n_women     = sum(genders_list == "Woman"),
    n_uncertain = sum(genders_list == "Uncertain"),
  ) %>%
  ungroup() %>%
  mutate(
    # Strict: all-same vs. mixed
    composition_strict = case_when(
      n_uncertain > 0            ~ "includes uncertain",
      n_men > 0 & n_women == 0   ~ "all-male",
      n_women > 0 & n_men == 0   ~ "all-female",
      n_men > 0 & n_women > 0    ~ "mixed",
      TRUE                       ~ NA_character_
    ),
    # Majority-based (ties = balanced)
    composition_majority = case_when(
      n_uncertain > 0     ~ "includes uncertain",
      n_men > n_women     ~ "majority-male",
      n_women > n_men     ~ "majority-female",
      n_men == n_women    ~ "balanced",
      TRUE                ~ NA_character_
    ),
    # Single-author papers split out separately, since "all-male"/
    # "all-female" for n=1 is a different phenomenon than for n>1
    composition_singleauthor_split = case_when(
      n_authors == 1 & n_men == 1   ~ "solo-male",
      n_authors == 1 & n_women == 1 ~ "solo-female",
      n_authors == 1                ~ "solo-uncertain",
      n_uncertain > 0                ~ "includes uncertain",
      n_men > 0 & n_women == 0       ~ "all-male (multi-author)",
      n_women > 0 & n_men == 0       ~ "all-female (multi-author)",
      TRUE                           ~ "mixed"
    )
  ) %>%
  select(-genders_list)

bind_rows(
  composition_candidates |> count(value = composition_strict) |> mutate(scheme = "Strict"),
  composition_candidates |> count(value = composition_majority) |> mutate(scheme = "Majority"),
  composition_candidates |> count(value = composition_singleauthor_split) |> mutate(scheme = "Solo split out")
) |>
  mutate(scheme = factor(scheme, levels = c("Strict", "Majority", "Solo split out"))) |>
  ggplot(aes(x = fct_reorder(value, n), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), hjust = -0.3, size = 3) +
  facet_wrap(vars(scheme), scales = "free", ncol = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  coord_flip() +
  labs(x = "", y = "Papers") +
  theme_bw()
```

### Lead-author gender {#lead-author-gender}

**Placeholder — not yet meaningful.** `fieldwork_lead_methods` below maps
each coded method to whether first-author or last-author is the
conventional "lead" for that kind of work, and needs real method names
from the codebook before this result can be trusted. Every paper
currently falls under "ambiguous" until that mapping is filled in.

```{r}
#| label: lead-author-convention
#| fig-width: 6
#| fig-height: 2.5
# PLACEHOLDER — fill in with actual method category names from the
# codebook. Methods not listed default to "ambiguous".
fieldwork_lead_methods <- c(
  # "Excavation"        = "first",
  # "Survey"            = "first",
  # "Isotope analysis"  = "last",
  # "aDNA analysis"     = "last"
)

primary_method_lookup <- df %>%
  select(cite_key, contains("Methods")) %>%
  pivot_longer(!cite_key, names_to = "method", values_to = "value") %>%
  filter(value == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  select(cite_key, primary_method = method)

lead_author_gender <- gender_sequences %>%
  left_join(primary_method_lookup, by = "cite_key") %>%
  mutate(
    # Indexing an EMPTY named vector (fieldwork_lead_methods currently
    # has zero entries) returns an unnamed NA vector, which broke
    # replace_na() downstream; unname() + explicit fallback avoids that
    # while the mapping is unfilled.
    lead_convention = if (length(fieldwork_lead_methods) > 0) {
      unname(fieldwork_lead_methods[primary_method])
    } else {
      rep(NA_character_, nrow(.))
    },
    lead_convention = replace_na(lead_convention, "ambiguous"),
    lead_author_gender = case_when(
      lead_convention == "first" ~ first_author_gender,
      lead_convention == "last"  ~ last_author_gender,
      TRUE                        ~ NA_character_
    )
  )

lead_author_gender |>
  count(lead_convention) |>
  ggplot(aes(x = fct_reorder(lead_convention, n), y = n)) +
  geom_col(fill = topic_col) +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  coord_flip() +
  labs(x = "", y = "Papers") +
  theme_bw()
```

## Topic

Every plot in this section is shown two ways: **Unmerged**, where a
paper's Primary- and Secondary-coded topics are kept distinguished, and
**Merged**, where a paper counts toward a topic if it carries either
value. In the Merged view a paper with one primary and two secondary
topics contributes to three bars, so counts and proportions are no
longer mutually exclusive across topics and won't sum to the corpus
total — this applies throughout, not just here.

```{r}
#| label: topics-setup
topic_count <- df %>%
  select(publication_year, contains("Topics")) %>%
  pivot_longer(!publication_year,
               names_to = "topic",
               values_to = "count")

totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n)
```

### Totals

```{r}
#| label: topics-totals
#| fig-width: 10
#| fig-height: 6
unmerged_count <- topic_count %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, count) %>%
  group_by(topic) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

merged_count <- topic_count %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic) %>%
  mutate(pct = n / sum(n))

plt_unmerged <-
ggplot(unmerged_count |> mutate(topic = str_wrap(topic, width = 25)),
       aes(x = reorder(topic, n, sum), y = n, fill = count)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_fill_manual(values = c(Primary = topic_col, Secondary = "#d8d8a0")) +
  labs(x = "Topic", y = "Count", fill = "", title = "Unmerged") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(9, "cm"))

plt_merged <-
ggplot(merged_count |> mutate(topic = str_wrap(topic, width = 25)),
       aes(x = reorder(topic, n), y = n)) +
  geom_bar(stat="identity", fill = topic_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = "", y = "Count", title = "Merged") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(9, "cm"))

plt_unmerged + plt_merged
ggsave("topics_aggregate.png", plot = plt_unmerged, width = 10, height = 6, dpi = 300)
ggsave("topics_aggregate_merged.png", plot = plt_merged, width = 10, height = 6, dpi = 300)
```

### Over time

```{r}
#| label: topics-over-time
#| fig-width: 12
#| fig-height: 7
unmerged_count_year <- topic_count %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, publication_year, count) %>%
  group_by(publication_year, count) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

merged_count_year <- topic_count %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, publication_year) %>%
  group_by(publication_year) %>%
  mutate(prop_merged = n / sum(n))

ggplot(unmerged_count_year, aes(x = publication_year, y = n, fill = topic)) +
  geom_bar(stat="identity") +
  facet_wrap(vars(count), ncol = 1) +
  labs(x = "", y = "Count", fill = "Topic", title = "Unmerged — counts by year") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("topics_count_by_year.png", plot = last_plot(), width = 12, height = 8, dpi = 300)

ggplot(merged_count_year, aes(x = publication_year, y = n, fill = topic)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Count", fill = "Topic", title = "Merged — counts by year") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_count_by_year_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### By first-author gender

Every plot in this document that normalizes proportions against the
corpus-wide gender base rate is a stated follow-up, not yet implemented
— the raw proportions below do not control for the overall gender
imbalance in the corpus.

```{r}
#| label: topic-by-gender
#| fig-width: 12
#| fig-height: 8
topic_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Topics")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

unmerged_topic_gender <- topic_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, gender_first_author, count) %>%
  group_by(topic, count) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

unmerged_topic_gender |>
  group_by(topic, count) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n)) |>
  ungroup() |>
  mutate(topic = fct_relabel(factor(topic), ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  facet_wrap(vars(count)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(8, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL, title = "Unmerged")
ggsave("topic_by_gender.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

merged_topic_gender <- topic_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  count(topic, gender_first_author) %>%
  group_by(topic) %>%
  mutate(prop_merged = n / sum(n))

merged_topic_gender |>
  group_by(topic) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(topic = fct_reorder(topic, ratio),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL, title = "Merged")
ggsave("topic_by_gender_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### By full-author-list gender composition

Unlike first-author-only gender above, this uses the composition
categories from the Gender section — bringing multi-author, order-aware
gender data into the same comparison. `composition_strict` is used as a
starting point; swap in `composition_majority` or
`composition_singleauthor_split` to compare.

```{r}
#| label: composition-by-topic
#| fig-width: 12
composition_topic <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic), !is.na(composition_strict))

composition_topic %>%
  count(topic, composition_strict, count) %>%
  group_by(topic, count) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(topic = fct_relabel(factor(topic), ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")")) %>%
  ggplot(aes(x = topic_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  facet_wrap(vars(count)) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(8, "cm")) +
  labs(x = "", y = "Proportion", fill = "Gender composition", title = "Unmerged")
ggsave("composition_by_topic.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

composition_topic %>%
  count(topic, composition_strict) %>%
  group_by(topic) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(topic = fct_reorder(topic, prop * (composition_strict == "all-female"), .fun = sum),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")")) %>%
  ggplot(aes(x = topic_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Gender composition", title = "Merged")
ggsave("composition_by_topic_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### Co-occurrence

**Directional**: given a paper's Primary-coded topic, what proportion of
the time is each other topic marked Secondary on the same paper.
**Undirected**: which topics tend to appear together at all, regardless
of which one a coder ranked first — a topic counts as present if it
carries either value, and the row/column pairing is symmetric (relabelled
`Topic.A`/`Topic.B` to avoid implying an order that doesn't exist here).
These answer different questions and neither supersedes the other.

```{r}
#| label: topics-cooccurrence-directional
#| fig-width: 30
#| fig-height: 20
topics <- df %>%
  select(contains("Topics")) %>%
  rename_with(~str_remove(., "Topics.."))

coocurrences <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Directional: secondary topic given primary topic")
```

```{r}
#| label: topics-cooccurrence-undirected
#| fig-width: 30
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Topic.A", "Topic.B")) %>%
  filter(Topic.A != Topic.B) %>%
  mutate(Topic.A = str_wrap(str_replace_all(Topic.A, fixed("."), " "), width = 20)) %>%
  mutate(Topic.B = str_replace_all(Topic.B, fixed("."), " ")) %>%
  group_by(Topic.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Topic.A, Prop, Topic.B), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Topic.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Undirected: co-occurring topics")
```

## Method

Same Unmerged/Merged logic as the Topic section above.

```{r}
#| label: methods-setup
method_count <- df %>%
  select(publication_year, contains("Methods")) %>%
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")
```

### Totals

```{r}
#| label: methods-totals
#| fig-width: 10
#| fig-height: 6
unmerged_count_method <- method_count %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, count) %>%
  group_by(method) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

merged_count_method <- method_count %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method) %>%
  mutate(pct = n / sum(n))

plt_unmerged <-
ggplot(unmerged_count_method |> mutate(method = str_wrap(method, width = 25)),
       aes(x = reorder(method, n, sum), y = n, fill = count)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_fill_manual(values = c(Primary = method_col, Secondary = "#8eaa9c")) +
  labs(x = "Method", y = "Count", fill = "", title = "Unmerged") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(9, "cm"))

plt_merged <-
ggplot(merged_count_method |> mutate(method = str_wrap(method, width = 25)),
       aes(x = reorder(method, n), y = n)) +
  geom_bar(stat="identity", fill = method_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = "", y = "Count", title = "Merged") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(9, "cm"))

plt_unmerged + plt_merged
ggsave("methods_aggregate.png", plot = plt_unmerged, width = 10, height = 6, dpi = 300)
ggsave("methods_aggregate_merged.png", plot = plt_merged, width = 10, height = 6, dpi = 300)
```

### Over time

```{r}
#| label: methods-over-time
#| fig-width: 12
#| fig-height: 7
unmerged_method_year <- method_count %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, publication_year, count) %>%
  group_by(publication_year, count) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

merged_method_year <- method_count %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, publication_year) %>%
  group_by(publication_year) %>%
  mutate(prop_merged = n / sum(n))

ggplot(unmerged_method_year, aes(x = publication_year, y = n, fill = method)) +
  geom_bar(stat="identity") +
  facet_wrap(vars(count), ncol = 1) +
  labs(x = "", y = "Count", fill = "Method", title = "Unmerged — counts by year") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_count_by_year.png", plot = last_plot(), width = 12, height = 8, dpi = 300)

ggplot(merged_method_year, aes(x = publication_year, y = n, fill = method)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Count", fill = "Method", title = "Merged — counts by year") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_count_by_year_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### By first-author gender

```{r}
#| label: method-by-gender
#| fig-width: 12
#| fig-height: 8
method_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Methods")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

unmerged_method_gender <- method_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, gender_first_author, count) %>%
  group_by(method, count) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

unmerged_method_gender |>
  group_by(method, count) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n)) |>
  ungroup() |>
  mutate(method = fct_relabel(factor(method), ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  facet_wrap(vars(count)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(8, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL, title = "Unmerged")
ggsave("method_by_gender.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

merged_method_gender <- method_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  count(method, gender_first_author) %>%
  group_by(method) %>%
  mutate(prop_merged = n / sum(n))

merged_method_gender |>
  group_by(method) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(method = fct_reorder(method, ratio),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL, title = "Merged")
ggsave("method_by_gender_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### By full-author-list gender composition

```{r}
#| label: composition-by-method
#| fig-width: 12
composition_method <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method), !is.na(composition_strict))

composition_method %>%
  count(method, composition_strict, count) %>%
  group_by(method, count) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(method = fct_relabel(factor(method), ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")")) %>%
  ggplot(aes(x = method_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  facet_wrap(vars(count)) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(8, "cm")) +
  labs(x = "", y = "Proportion", fill = "Gender composition", title = "Unmerged")
ggsave("composition_by_method.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

composition_method %>%
  count(method, composition_strict) %>%
  group_by(method) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")")) %>%
  ggplot(aes(x = method_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Gender composition", title = "Merged")
ggsave("composition_by_method_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### Co-occurrence

```{r}
#| label: methods-cooccurrence-directional
#| fig-width: 20
#| fig-height: 20
methods <- df %>%
  select(contains("Methods")) %>%
  rename_with(~str_remove(., "Methods.."))

coocurrences <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Directional: secondary method given primary method")
```

```{r}
#| label: methods-cooccurrence-undirected
#| fig-width: 20
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Method.A", "Method.B")) %>%
  filter(Method.A != Method.B) %>%
  mutate(Method.A = str_wrap(str_replace_all(Method.A, fixed("."), " "), width = 20)) %>%
  mutate(Method.B = str_replace_all(Method.B, fixed("."), " ")) %>%
  group_by(Method.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Method.A, Prop, Method.B), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Method.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Undirected: co-occurring methods")
```

## Source of evidence

Same Unmerged/Merged logic as above.

```{r}
#| label: sources-setup
totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n) %>%
  mutate(publication_year = factor(publication_year, levels = sort(unique(as.numeric(as.character(publication_year))))))

source_count <- df %>%
  select(publication_year, contains("Sources.of.Evidence")) %>%
  mutate(publication_year = factor(publication_year, levels = levels(totals_by_year$publication_year))) %>%
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")
```

### Totals

```{r}
#| label: sources-totals
#| fig-width: 10
#| fig-height: 6
unmerged_count_source <- source_count %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, count) %>%
  group_by(source) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup()

merged_source_count <- source_count %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source) %>%
  mutate(pct = n / sum(n))

plt_unmerged <-
ggplot(unmerged_count_source |> mutate(source = str_wrap(source, width = 25)),
       aes(x = reorder(source, n, sum), y = n, fill = count)) +
  geom_bar(stat="identity") +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_fill_manual(values = c(Primary = source_col, Secondary = "#d89898")) +
  labs(x = "Source of evidence", y = "Count", fill = "", title = "Unmerged") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(9, "cm"))

plt_merged <-
ggplot(merged_source_count |> mutate(source = str_wrap(source, width = 25)),
       aes(x = reorder(source, n), y = n)) +
  geom_bar(stat="identity", fill = source_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = "", y = "Count", title = "Merged") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(9, "cm"))

plt_unmerged + plt_merged
ggsave("sources_aggregate.png", plot = plt_unmerged, width = 10, height = 6, dpi = 300)
ggsave("sources_aggregate_merged.png", plot = plt_merged, width = 10, height = 6, dpi = 300)
```

### Over time

```{r}
#| label: sources-over-time
#| fig-width: 12
#| fig-height: 6
unmerged_source_year <- source_count %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, publication_year, count) %>%
  group_by(publication_year, count) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(publication_year = droplevels(publication_year))

merged_source_year <- source_count %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, publication_year) %>%
  group_by(publication_year) %>%
  mutate(prop_merged = n / sum(n)) %>%
  ungroup() %>%
  mutate(publication_year = droplevels(publication_year))

ggplot(unmerged_source_year, aes(x = publication_year, y = n, fill = source)) +
  geom_bar(stat="identity") +
  facet_wrap(vars(count), ncol = 1) +
  labs(x = "", y = "Count", fill = "Source of evidence", title = "Unmerged — counts by year") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 3))
ggsave("sources_count_by_year.png", plot = last_plot(), width = 12, height = 8, dpi = 300)

ggplot(merged_source_year, aes(x = publication_year, y = n, fill = source)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Count", fill = "Source of evidence", title = "Merged — counts by year") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 2))
ggsave("sources_count_by_year_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### By first-author gender

```{r}
#| label: source-by-gender
#| fig-width: 12
#| fig-height: 8
source_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Sources")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

unmerged_source_gender <- source_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, gender_first_author, count) %>%
  group_by(source, count) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup()

unmerged_source_gender |>
  group_by(source, count) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n)) |>
  ungroup() |>
  mutate(source = fct_relabel(factor(source), ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  facet_wrap(vars(count)) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(8, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL, title = "Unmerged")
ggsave("source_by_gender.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

merged_source_gender <- source_count_gender %>%
  filter(!is.na(count)) %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  count(source, gender_first_author) %>%
  group_by(source) %>%
  mutate(prop_merged = n / sum(n))

merged_source_gender |>
  group_by(source) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(source = fct_reorder(source, ratio),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop_merged, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL, title = "Merged")
ggsave("source_by_gender_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### By full-author-list gender composition

```{r}
#| label: composition-by-source
#| fig-width: 12
composition_source <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(!is.na(count)) %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source), !is.na(composition_strict))

composition_source %>%
  count(source, composition_strict, count) %>%
  group_by(source, count) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(source = fct_relabel(factor(source), ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")")) %>%
  ggplot(aes(x = source_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  facet_wrap(vars(count)) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(8, "cm")) +
  labs(x = "", y = "Proportion", fill = "Gender composition", title = "Unmerged")
ggsave("composition_by_source.png", plot = last_plot(), width = 10, height = 8, dpi = 300)

composition_source %>%
  count(source, composition_strict) %>%
  group_by(source) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")")) %>%
  ggplot(aes(x = source_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Gender composition", title = "Merged")
ggsave("composition_by_source_merged.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### Co-occurrence

```{r}
#| label: sources-cooccurrence-directional
#| fig-width: 20
#| fig-height: 20
sources <- df %>%
  select(contains("Sources.of.Evidence")) %>%
  rename_with(~str_remove(., "Sources.of.Evidence.."))

coocurrences <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Directional: secondary source given primary source")
```

```{r}
#| label: sources-cooccurrence-undirected
#| fig-width: 20
#| fig-height: 20
coocurrences_merged <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_present <- names(current_row)
  current_row <- paste(rep(current_present, each = length(current_present)), current_present, sep = "_")
  coocurrences_merged <- append(coocurrences_merged, current_row)
}

coocdf_merged <- data.frame(table(coocurrences_merged)) %>%
  separate_wider_delim(coocurrences_merged, "_", names = c("Source.A", "Source.B")) %>%
  filter(Source.A != Source.B) %>%
  mutate(Source.A = str_wrap(str_replace_all(Source.A, fixed("."), " "), width = 20)) %>%
  mutate(Source.B = str_replace_all(Source.B, fixed("."), " ")) %>%
  group_by(Source.B) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf_merged, aes(x = reorder_within(Source.A, Prop, Source.B), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Source.B), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Undirected: co-occurring sources of evidence")
```

## Region

```{r}
#| label: region
library(ggwordcloud)

region_counts <- df %>%
  pull(Region) %>%
  na.omit() %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Region", "n")) %>%
  mutate(n = as.numeric(n),
         n_scaled = n + 3) %>%
  arrange(desc(n))

ggplot(region_counts, aes(label = Region, size = n_scaled, color = n)) +
  geom_text_wordcloud(area_corr = TRUE, rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  scale_color_gradient(low = topic_col, high = "darkred") +
  theme_minimal()
```
