---
format:
  html:
    toc: true
    toc-depth: 4
    code-fold: true
    self-contained: true
    include-in-header:
      text: |
        <style>
        /* kable() tables default to filling their container's full
           width regardless of column count — on this page's 1100px
           body width, a 2-3 column table stretches absurdly wide with
           padded-out columns. Scoped to size every table to its actual
           content instead, in one place, rather than passing a width
           override to each individual kable() call throughout the
           document. */
        table { width: auto; }
        </style>
execute:
  warning: false
  message: false
  eval: true
---

# Part 2 — Citation graph: structural descriptions

```{r}
#| label: install
#| eval: false
install.packages(c("igraph", "tidyverse", "ggraph", "tidygraph", "jsonlite",
                   "here", "ggrepel", "scales", "graphlayouts"))
```

```{r}
#| label: libraries
library(igraph)
library(tidyverse)
library(ggraph)
library(tidygraph)
library(jsonlite)
library(here)
library(ggrepel)
```

```{r}
#| label: graph-definitions
#| echo: false
# Two versions of the annotated graph, handled identically throughout.
#   annotated      the citations made by the annotated papers, plus the seed
#                  paper's own references. lund2021 is not annotated; it is the
#                  origin of the corpus and a source in this edge list.
#   seed_filtered  the same list with the seed's rows dropped.
# Every table and plot below is computed on both and shown in two tabs.
graph_keys   <- c("annotated", "seed_filtered")
graph_labels <- c(annotated = "Annotated", seed_filtered = "Annotated, seed filtered")
graph_edgelists <- c(
  annotated     = here("data", "citation_edgelist_annotated.csv"),
  seed_filtered = here("analysis", "gn_analysis", "edgelist_annotated_no_seed.csv")
)

# Girvan-Newman and cluster-labelling outputs are produced outside this
# render (see analysis/README.md). Each run is one folder with identical
# contents:
#   communities.csv, modularity_trace.csv, removal_log.csv, final_edgelist.csv
#   multi_cut/    communities_round_<N>.csv and multi_cut_summary.csv
#   dendrogram/   gn_dendrogram.png
#   subclusters/  subcluster_summary.csv
#   labels/       representatives.csv, cluster_labels.csv, node_labels.csv
#
# gn_root is the one place this document looks for them: the pipelines'
# own results folder. If finished runs are later copied into data/, the
# way the exported bibliography and edge lists are, change gn_root to
# here("data", "gn") and nothing else.
gn_runs <- c(annotated = "annotated", seed_filtered = "annotated_no_seed")
gn_root <- here("analysis", "gn_analysis", "results")
gn_path <- function(key, ...) file.path(gn_root, gn_runs[[key]], ...)

# How many of the largest communities the plots and tables single out.
# Set here because inline text refers to it and has to render even when
# no run has been installed yet.
gn_top_n <- 20
```

```{r}
#| label: load-graphs
#| echo: false
#| results: hide
edges_list <- map(graph_edgelists, read_csv, show_col_types = FALSE)
graphs <- map(edges_list, ~ graph_from_data_frame(.x, directed = TRUE))

bib <- fromJSON(here("data", "bibliography.json"), simplifyDataFrame = FALSE)

# Active entries only (tombstoned entries are excluded from the export,
# but guard here in case of an older bibliography file). Note: entries
# tombstoned via `delete` (not `merge`) that appear as a citer elsewhere
# are retained in the edge list so their outbound edges aren't lost — see
# exporter.py and docs/methods/corrections-system.md. Ghost nodes have no
# reliable bibliographic metadata; they are kept in the graph (their edges
# are real) but excluded from any display table or join that assumes real
# title/author/topic data.
bib <- bib[!sapply(bib, function(e) isTRUE(e[["_deleted"]]))]

node_attrs <- tibble(
  name            = names(bib),
  title           = map_chr(bib, ~ .x$title %||% ""),
  # year prefers the already-clean year field; date (a full string like
  # "2021-05-15") is only used as a fallback, and then only its leading
  # year digits, so a table column named "year" never shows a full date.
  year            = map_chr(bib, ~ coalesce(
    if (!is.null(.x$year) && nzchar(.x$year)) .x$year,
    str_extract(.x$date %||% "", "^\\d{4}"),
    ""
  )),
  generation      = map_chr(bib, ~ .x$generation %||% ""),
  entry_type      = map_chr(bib, ~ .x$entry_type %||% ""),
  resolution      = map_chr(bib, ~ .x[["_resolution_method"]] %||% "grobid"),
  completeness    = map_chr(bib, ~ .x$completeness$label %||% ""),
  # author is a list of {family, given} objects per entry (or NULL/empty
  # for entries GROBID/the LLM couldn't resolve an author for); flattened
  # here to a single "Family, Family" display string, since this is only
  # used for display (tooltips, tables) and never joined or filtered on.
  author          = map_chr(bib, function(e) {
    au <- e$author
    if (is.null(au) || length(au) == 0) return("")
    paste(map_chr(au, ~ .x$family %||% ""), collapse = ", ")
  }),
)

n_annotated_papers <- n_distinct(edges_list$seed_filtered$source)
```

```{r}
#| label: graph-metrics
#| echo: false
#| results: hide
# Every per-node measure for one graph, computed once. Betweenness is the
# slowest calculation in this document, so the raw value is computed once
# and the normalized value derived from it: for a directed graph igraph
# normalizes by (n - 1)(n - 2), so the two are the same up to that factor.
# Louvain, Leiden and K-means are computed here with no visible output;
# their agreement across the two graphs appears in the comparison section.
graph_metrics <- function(g) {
  set.seed(42)
  und <- as_undirected(g, mode = "collapse")
  n <- vcount(g)
  btw_raw <- betweenness(g, directed = TRUE, normalized = FALSE)
  weak <- components(g, mode = "weak")
  strong <- components(g, mode = "strong")

  m <- tibble(
    name              = V(g)$name,
    betweenness       = btw_raw / ((n - 1) * (n - 2)),
    betweenness_raw   = btw_raw,
    closeness         = closeness(g, mode = "in", normalized = TRUE),
    pagerank          = page_rank(g, directed = TRUE)$vector,
    hub               = hub_score(g)$vector,
    authority         = authority_score(g)$vector,
    in_degree         = degree(g, mode = "in"),
    out_degree        = degree(g, mode = "out"),
    degree            = degree(g, mode = "all")
  ) |>
    left_join(node_attrs, by = "name") |>
    # Ghost = a node present in the edge list (it cites or is cited by
    # something) but absent from the active bibliography — i.e. no real
    # metadata to join. Derived from edge-list membership rather than a
    # GraphML attribute, since we're not loading GraphML at all (libxml2
    # isn't installable here without sudo — see gn_analysis/README.md for
    # the same constraint on the GN side).
    mutate(is_ghost = if_else(name %in% node_attrs$name, "false", "true")) |>
    mutate(
      coreness          = coreness(und),
      component_weak    = weak$membership,
      component_strong  = strong$membership,
      community_louvain = as.integer(membership(cluster_louvain(und))),
      community_leiden  = as.integer(membership(cluster_leiden(und, objective_function = "modularity")))
    )

  features <- m |>
    select(betweenness, closeness, pagerank, hub, authority, in_degree) |>
    mutate(across(everything(), ~ replace_na(., 0))) |>
    mutate(across(everything(), ~ scale(.)[, 1]))
  set.seed(42)
  m$cluster_kmeans <- kmeans(features, centers = 6, nstart = 25)$cluster
  m
}

metrics <- map(graphs, graph_metrics)
```

```{r}
#| label: gn-load
#| echo: false
#| results: hide
# One list per graph holding that graph's Girvan-Newman run, or just
# available = FALSE when the run has not been installed. The working
# partition is the earliest fragmentation-onset cut, selected by sorting
# rather than a hardcoded round number, in case a future re-run shifts
# which rounds qualify.
gn_load <- function(key) {
  dir <- gn_path(key, "multi_cut")
  files <- if (dir.exists(dir)) {
    list.files(dir, pattern = "^communities_round_[0-9]+\\.csv$", full.names = TRUE)
  } else character(0)
  summary_path <- file.path(dir, "multi_cut_summary.csv")
  out <- list(key = key, available = length(files) > 0 && file.exists(summary_path))
  if (!out$available) return(out)

  rounds <- as.integer(gsub(".*communities_round_([0-9]+)\\.csv$", "\\1", files))
  out$cut_file <- files[which.min(rounds)]
  out$summary <- read_csv(summary_path, show_col_types = FALSE) |>
    mutate(pct_singleton = n_singletons / n_communities) |>
    arrange(round)
  out$earliest <- slice(out$summary, 1)
  out$partition <- read_csv(out$cut_file, show_col_types = FALSE) |>
    mutate(community_gn = as.character(community_id)) |>
    select(node_id, community_gn)

  trace_path <- gn_path(key, "modularity_trace.csv")
  out$trace <- if (file.exists(trace_path)) read_csv(trace_path, show_col_types = FALSE) else NULL

  # Share of nodes in a singleton community at the best-modularity cut
  # (communities.csv), quoted in the prose in place of a typed-in figure.
  best_path <- gn_path(key, "communities.csv")
  out$best_singleton_share <- if (file.exists(best_path)) {
    read_csv(best_path, show_col_types = FALSE) |>
      add_count(community_id) |>
      summarise(share = mean(n == 1)) |>
      pull(share)
  } else NA_real_
  out
}
gn <- map(set_names(graph_keys), gn_load)

gn_round_text <- function(key) {
  if (gn[[key]]$available) format(gn[[key]]$earliest$round, big.mark = ",") else "n/a"
}
gn_pct_phrase <- function(key) {
  s <- gn[[key]]$best_singleton_share
  if (is.null(s) || is.na(s)) "nearly all nodes" else paste0(scales::percent(s, accuracy = 0.1), " of nodes")
}
gn_label_lookup <- function(key) {
  path <- gn_path(key, "labels", "cluster_labels.csv")
  if (file.exists(path)) {
    read_csv(path, show_col_types = FALSE) |>
      mutate(community_id = as.character(community_id)) |>
      select(community_id, label)
  } else {
    tibble(community_id = character(), label = character())
  }
}
gn_label_or_id <- function(ids, lookup) {
  coalesce(lookup$label[match(ids, lookup$community_id)], paste("Community", ids))
}
```

```{r}
#| label: structural-functions
#| echo: false
# Everything below that is computed per graph goes through one of these,
# called once per tab, so the two tabs of a section cannot drift apart.
graph_summary <- function(g) {
  w <- components(g, mode = "weak")
  s <- components(g, mode = "strong")
  list(nodes = vcount(g), edges = ecount(g), density = edge_density(g),
       wcc = w$no, wcc_largest = max(w$csize),
       wcc_pct = round(100 * max(w$csize) / vcount(g), 1),
       scc = s$no, scc_largest = max(s$csize))
}
summ <- map(graphs, graph_summary)

top_table <- function(key, col, n = 20) {
  metrics[[key]] |>
    filter(is_ghost == "false") |>
    arrange(desc(.data[[col]])) |>
    head(n) |>
    select(name, title, year, generation, all_of(col)) |>
    knitr::kable()
}

plot_in_degree <- function(key) {
  ggplot(metrics[[key]], aes(x = in_degree)) +
    geom_histogram(aes(x = log10(in_degree)), bins = 20, fill = "steelblue") +
    scale_y_log10() +
    scale_x_continuous(
      labels = function(x) round(10^x),
      name = "In-degree (times cited)"
    ) +
    labs(
      title = paste0("In-degree distribution (log-log), ", graph_labels[[key]]),
      y = "Count"
    ) +
    theme_minimal()
}

tail_table <- function(key, threshold = 30) {
  metrics[[key]] |>
    filter(in_degree > threshold) |>
    select(name, title, year, in_degree) |>
    arrange(desc(in_degree)) |>
    knitr::kable()
}

# Neighbours' Girvan-Newman clusters for each of the top-betweenness works:
# what a work is bridging, not just where it sits. Ranked by how many of
# the work's neighbours fall in each cluster, and capped at the top 3, since
# a highly connected broker can touch a dozen clusters, most through a
# single incidental neighbour. Reads the same earliest-cut partition and
# labels as the Girvan-Newman section, so it does not depend on the
# labelling pipeline having been run.
bridges_table <- function(key, n = 20) {
  m <- metrics[[key]]
  g <- graphs[[key]]
  top <- m |> filter(is_ghost == "false") |> arrange(desc(betweenness)) |> head(n)
  bridges <- tibble(name = character(), bridges = character())

  if (gn[[key]]$available) {
    lookup <- gn_label_lookup(key)
    part <- gn[[key]]$partition |> mutate(label = gn_label_or_id(community_gn, lookup))
    bridges <- map_dfr(top$name, function(nm) {
      neighbor_ids <- unique(c(
        names(igraph::neighbors(g, nm, mode = "in")),
        names(igraph::neighbors(g, nm, mode = "out"))
      ))
      counts <- part |> filter(node_id %in% neighbor_ids) |> count(label, sort = TRUE)
      top_labels <- counts |> slice_head(n = 3) |> mutate(display = paste0(label, " (", n, ")"))
      n_more <- nrow(counts) - nrow(top_labels)
      text <- if (nrow(top_labels) == 0) {
        ""
      } else {
        paste0(paste(top_labels$display, collapse = "; "),
               if (n_more > 0) paste0("; +", n_more, " more") else "")
      }
      tibble(name = nm, bridges = text)
    })
  }

  top |>
    select(name, title, year, betweenness, betweenness_raw) |>
    left_join(bridges, by = "name") |>
    knitr::kable(digits = 4)
}

# year is free-text, sourced from bibliography metadata (dates, ranges, and
# circa notation all appear there). This pulls the first 4-digit run found,
# a pragmatic extraction, not a full date parser. Works with no recognizable
# 4-digit year are dropped rather than guessed at.
btw_year_data <- function(key) {
  metrics[[key]] |>
    filter(is_ghost == "false") |>
    mutate(year_numeric = suppressWarnings(as.integer(str_extract(year, "\\d{4}"))))
}

# Zero betweenness is the normal state for most works in a sparse citation
# graph (only a work on a shortest path between some other pair has nonzero
# betweenness at all), so the first plot keeps only the nonzero subset,
# where there is an actual distribution to look at. The second plot shows
# every work.
plot_btw_year_nonzero <- function(key) {
  d <- btw_year_data(key)
  n_total <- nrow(d)
  n_no_year <- sum(is.na(d$year_numeric))
  n_zero <- sum(!is.na(d$year_numeric) & d$betweenness_raw == 0)
  d <- d |> filter(!is.na(year_numeric), betweenness_raw > 0)

  ggplot(d, aes(x = year_numeric, y = betweenness_raw)) +
    geom_point(alpha = 0.4, colour = "steelblue") +
    geom_smooth(method = "loess", colour = "firebrick", se = TRUE) +
    scale_y_log10() +
    labs(
      title = paste0("Betweenness (raw) by publication year, nonzero works only, ", graph_labels[[key]]),
      subtitle = paste0(
        nrow(d), " of ", format(n_total, big.mark = ","),
        " works are shown here: works with zero betweenness (",
        format(n_zero, big.mark = ","), " of them, the normal state ",
        "for most works in a sparse citation graph, not excluded data) and ",
        "works with no parseable year (", n_no_year, ") are left out."
      ),
      x = "Publication year", y = "Betweenness (raw, log scale)"
    ) +
    theme_minimal()
}

# betweenness_raw + 1 is plotted because a true log scale can't display a
# zero: every zero maps to 1, a real position at the bottom of the axis, and
# every nonzero value keeps its relative position. Jittered in x only, since
# thousands of exact zeros would otherwise stack into one solid line.
plot_btw_year_full <- function(key) {
  ggplot(btw_year_data(key) |> filter(!is.na(year_numeric)),
         aes(x = year_numeric, y = betweenness_raw + 1)) +
    geom_jitter(width = 0.3, height = 0, alpha = 0.15, colour = "steelblue") +
    scale_y_log10() +
    labs(
      title = paste0("Betweenness (raw + 1) by publication year, every work, ", graph_labels[[key]]),
      x = "Publication year", y = "Betweenness (raw + 1, log scale)"
    ) +
    theme_minimal()
}

kcore_summary_table <- function(key) {
  metrics[[key]] |>
    group_by(coreness) |>
    summarise(n = n(), mean_in_degree = mean(in_degree)) |>
    arrange(desc(coreness)) |>
    head(15) |>
    knitr::kable(digits = 2)
}
kcore_top_table <- function(key) {
  metrics[[key]] |>
    filter(is_ghost == "false") |>
    arrange(desc(coreness), desc(in_degree)) |>
    head(20) |>
    select(name, title, year, coreness, in_degree) |>
    knitr::kable()
}

# Figures quoted in the prose, computed from the data so they follow a
# re-render instead of being typed in.
deg_stats <- map(metrics, function(m) {
  cited <- m$in_degree[m$in_degree >= 1]
  list(share_le8 = scales::percent(mean(cited <= 8), accuracy = 1),
       max_in = max(m$in_degree, na.rm = TRUE))
})
btw_stats <- map(graphs |> names() |> set_names(), function(key) {
  d <- btw_year_data(key) |> filter(!is.na(year_numeric), betweenness_raw > 0)
  list(n_nonzero = nrow(d), n_pre2000 = sum(d$year_numeric < 2000),
       pct_pre2000 = scales::percent(mean(d$year_numeric < 2000), accuracy = 1))
})
kcore_stats <- map(metrics, function(m) {
  real <- m |> filter(is_ghost == "false")
  kmax <- max(real$coreness)
  top_core <- real |> filter(coreness == kmax) |> pull(name)
  list(kmax = kmax, n_core = length(top_core),
       n_in = length(intersect(top_core, real |> arrange(desc(in_degree)) |> head(5) |> pull(name))),
       n_btw = length(intersect(top_core, real |> arrange(desc(betweenness)) |> head(5) |> pull(name))))
})
```

## The two graphs

This part describes the citation graph and its communities twice. Both versions come from the annotated filter, which keeps the citations made by the `r n_annotated_papers` annotated papers. **Annotated** also keeps the seed paper's own references. The seed, `lund2021`, is not annotated. It is the origin of the corpus and counts as one more citing paper in the edge list. **Annotated, seed filtered** drops the seed's rows. Each table and plot below is computed on both graphs and shown in two tabs, Annotated first. "What the seed filter changes" compares the two directly.

| Metric | Annotated | Annotated, seed filtered |
|---|---|---|
| Nodes | `r format(summ$annotated$nodes, big.mark = ",")` | `r format(summ$seed_filtered$nodes, big.mark = ",")` |
| Edges | `r format(summ$annotated$edges, big.mark = ",")` | `r format(summ$seed_filtered$edges, big.mark = ",")` |
| Directed | Yes | Yes |
| Density[^density] | `r format(summ$annotated$density, scientific = FALSE)` | `r format(summ$seed_filtered$density, scientific = FALSE)` |
| Weakly connected components[^wcc] | `r format(summ$annotated$wcc, big.mark = ",")` | `r format(summ$seed_filtered$wcc, big.mark = ",")` |
| Largest weak component | `r format(summ$annotated$wcc_largest, big.mark = ",")` nodes (`r summ$annotated$wcc_pct`% of graph) | `r format(summ$seed_filtered$wcc_largest, big.mark = ",")` nodes (`r summ$seed_filtered$wcc_pct`% of graph) |
| Strongly connected components[^scc] | `r format(summ$annotated$scc, big.mark = ",")` | `r format(summ$seed_filtered$scc, big.mark = ",")` |
| Largest strong component | `r format(summ$annotated$scc_largest, big.mark = ",")` nodes | `r format(summ$seed_filtered$scc_largest, big.mark = ",")` nodes |

[^density]: The proportion of possible edges that actually exist: edges
  present divided by every pair of nodes that *could* be connected. Low
  density is normal and expected for a citation network — most works
  don't cite most other works.

[^wcc]: A weakly connected component treats every edge as undirected: two
  nodes are in the same component if a path connects them in either
  citation direction. This is the more forgiving of the two measures
  here.

[^scc]: A strongly connected component requires an actual directed path
  in *both* directions between two nodes. Rare in citation networks,
  since citations are historically directed — a 2010 paper cannot cite
  a 2020 paper — so most strongly connected components are single
  isolated nodes.

## Degree and centrality

In-degree (times cited within the corpus) and out-degree (number of works
cited) capture different roles. High in-degree identifies widely-cited works
— the intellectual anchors of the field. High out-degree identifies works
that are heavily engaged with the literature, drawing on many sources.

Betweenness centrality identifies works that sit on many shortest paths
between other works — brokers between sub-literatures. Betweenness is
expensive to compute on large graphs; both normalized and raw values are
reported. Normalized divides by the theoretical maximum for a graph this
size (roughly (n-1)(n-2) for a directed graph), so on a graph with over
`r format(vcount(graphs$annotated), big.mark = ",")` nodes even a genuinely
high-betweenness work normalizes to a very small fraction — small in
absolute terms, not because the work is unimportant. Raw values, not
rescaled, are more intuitively legible for comparing works directly
against each other.

PageRank assigns authority based on incoming links weighted by the
authority of the linking node — conceptually similar to being cited by
works that are themselves widely cited. Hub and authority scores (HITS)
are complementary: hubs point to many authorities; authorities are pointed
to by many hubs.

Peripheral works — low degree, low centrality by every measure below — are
not treated as noise in this project. A work cited only once may be exactly
the interesting case (a niche, contested, or newly emerging reference) for
a study of citational politics. No node is excluded from any measure in
this section on the basis of low degree.

### Most cited (by in-degree)

::: {.panel-tabset}

#### Annotated

```{r}
#| label: most-cited-annotated
top_table("annotated", "in_degree")
```

#### Annotated, seed filtered

```{r}
#| label: most-cited-seed-filtered
top_table("seed_filtered", "in_degree")
```

:::

### Most citing (by out-degree)

::: {.panel-tabset}

#### Annotated

```{r}
#| label: most-citing-annotated
top_table("annotated", "out_degree")
```

#### Annotated, seed filtered

```{r}
#| label: most-citing-seed-filtered
top_table("seed_filtered", "out_degree")
```

:::

### Degree distribution

A power-law or scale-free degree distribution is characteristic of citation
networks: most works are cited rarely, a small number are cited frequently,
visible as a roughly straight-line decline on a log-log plot. The plot below
tests whether this corpus follows that pattern.

::: {.panel-tabset}

#### Annotated

```{r}
#| label: degree-dist-annotated
#| fig-width: 8
#| fig-height: 5
plot_in_degree("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: degree-dist-seed-filtered
#| fig-width: 8
#| fig-height: 5
plot_in_degree("seed_filtered")
```

:::

The bulk of the distribution follows a fairly straight decline on this log-log scale, consistent with the expected power-law pattern: most works are cited once or twice, a much smaller number several times. In Annotated, `r deg_stats$annotated$share_le8` of works with at least one citation are cited eight times or fewer, and `r deg_stats$seed_filtered$share_le8` in Annotated, seed filtered. The tail is real, not an artifact of a truncated plot axis: the maximum in-degree is `r deg_stats$annotated$max_in` in Annotated and `r deg_stats$seed_filtered$max_in` in Annotated, seed filtered. A handful of works drive that high-citation tail, listed below rather than relied on to show clearly as bars on the histogram above, since bars this sparse are easy to miss regardless of binning.

### Most-cited works (the high end of the tail)

The most-cited works in each graph, driving the sparse, high-in-degree tail visible in the plots above:

::: {.panel-tabset}

#### Annotated

```{r}
#| label: degree-dist-tail-annotated
tail_table("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: degree-dist-tail-seed-filtered
tail_table("seed_filtered")
```

:::

### Top by betweenness centrality

Works with high betweenness mediate between otherwise disconnected parts
of the citation network — likely works that bridge sub-fields or periods.
"Bridges" below lists the top 3 Girvan-Newman clusters (by count) found
among each work's direct citations and citers, with the number of
neighbors in each — a broker connecting mostly to one cluster and only
incidentally to others will show that concentration, rather than an
undifferentiated list of every cluster it happens to touch (see
"Girvan-Newman" and "Cluster labels" further down for how
those labels are produced; a self-contained lookup here, independent of
that later section, so this table isn't affected by whether cluster
labelling has been run).

::: {.panel-tabset}

#### Annotated

```{r}
#| label: betweenness-bridges-annotated
bridges_table("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: betweenness-bridges-seed-filtered
bridges_table("seed_filtered")
```

:::

### Betweenness and publication year

A negative relationship between publication year and betweenness is
expected as a baseline structural effect, independent of any specific
work's importance: an older work has had more time to accumulate
citations from both earlier and later literature, giving it more
opportunities to sit on shortest paths between clusters that didn't yet
exist when it was published. The interesting question is not whether
that baseline trend exists, but what departs from it: works that are old
without being especially well-connected, and, more notably, works that
are young but already bridging despite limited time to accumulate
citations — those are candidates for genuinely novel synthesis, not just
structurally advantaged age.

The trend line[^loess] in the first plot of each tab is fit to the nonzero subset, where the actual bridging signal is. In Annotated, `r btw_stats$annotated$n_pre2000` of the `r btw_stats$annotated$n_nonzero` works with nonzero betweenness (`r btw_stats$annotated$pct_pre2000`) were published before 2000. In Annotated, seed filtered, the figures are `r btw_stats$seed_filtered$n_pre2000` of `r btw_stats$seed_filtered$n_nonzero` (`r btw_stats$seed_filtered$pct_pre2000`). The second plot in each tab shows every work, including the zero-betweenness majority, so the shape of any boundary between old and recent works is visible directly.

::: {.panel-tabset}

#### Annotated

```{r}
#| label: betweenness-year-annotated
#| fig-width: 8
#| fig-height: 5
plot_btw_year_nonzero("annotated")
```

```{r}
#| label: betweenness-year-full-annotated
#| fig-width: 8
#| fig-height: 5
plot_btw_year_full("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: betweenness-year-seed-filtered
#| fig-width: 8
#| fig-height: 5
plot_btw_year_nonzero("seed_filtered")
```

```{r}
#| label: betweenness-year-full-seed-filtered
#| fig-width: 8
#| fig-height: 5
plot_btw_year_full("seed_filtered")
```

:::

Whether the boundary reflects something about the corpus itself (older
material may be primary sources, sagas, or editions with different
citation patterns than modern academic synthesis) or a genuine
historiographical shift (post-2000 archaeology being where cross-cutting
synthesis actually happens) can't be settled from either plot alone.

[^loess]: A LOESS fit, not a strict linear one, to allow for a
  non-constant relationship across the full year range rather than
  forcing a single slope.

## K-core decomposition

The k-core of a graph is the maximal subgraph in which every node has at
least k connections to other nodes within that subgraph. It is computed by
iteratively removing nodes with degree below k until no further removal
is possible. A node's **coreness** is the highest k for which it still
belongs to the k-core: a coreness of 1 means the node only survives the
weakest possible core (removed as soon as k reaches 2), while a high
coreness means the node stays embedded even as the graph is stripped down
to only its most densely-interconnected parts.

K-core decomposition is reported here purely as a **descriptive**
measure, showing which works sit in dense mutual-citation relationships
versus which sit at the periphery, following @alvarezhamelin2005. Every
work, regardless of its coreness, is still included in full in every
other analysis in this document: coreness only describes a work's
position, it doesn't filter anything out.

Coreness and in-degree measure genuinely different things, worth
comparing directly: in-degree simply counts how many times a work is
cited, with no regard for whether those citing works are themselves
well-connected. Coreness asks a structurally deeper question, whether a
work sits inside a neighborhood that is *itself* densely interconnected,
not just cited often in isolation. A work could have high in-degree
while sitting at low coreness, if most of its citations come from
otherwise peripheral, poorly-connected works; conversely, a work with
only moderate in-degree can have high coreness if it's embedded in a
tightly interlinked core. The tables below show whether these two
measures broadly agree on this corpus, or diverge in informative ways.

::: {.panel-tabset}

### Annotated

```{r}
#| label: kcore-table-annotated
#| echo: false
kcore_summary_table("annotated")
```

```{r}
#| label: kcore-top-annotated
kcore_top_table("annotated")
```

### Annotated, seed filtered

```{r}
#| label: kcore-table-seed-filtered
#| echo: false
kcore_summary_table("seed_filtered")
```

```{r}
#| label: kcore-top-seed-filtered
kcore_top_table("seed_filtered")
```

:::

At the highest coreness (k = `r kcore_stats$annotated$kmax`), Annotated holds `r kcore_stats$annotated$n_core` works. `r kcore_stats$annotated$n_in` of them are also among the five most cited and `r kcore_stats$annotated$n_btw` among the five highest-betweenness works. In Annotated, seed filtered, the highest coreness is k = `r kcore_stats$seed_filtered$kmax` with `r kcore_stats$seed_filtered$n_core` works, of which `r kcore_stats$seed_filtered$n_in` are among the five most cited and `r kcore_stats$seed_filtered$n_btw` among the five highest-betweenness works. Where in-degree, betweenness and k-core point to the same few works, that is a stronger, cross-validated signal than any one measure alone: those works are central by several different, genuinely distinct criteria at once.


```{r}
#| label: gn-functions
#| echo: false
# Everything per-run in the Girvan-Newman sections goes through these, called
# once per tab.

# Rebuilds the Girvan-Newman dendrogram image from a run's removal log. The log
# is replayed forward, recording every round where removing an edge splits a
# component. Read backwards, those splits are merges, which gives an hclust tree
# over whole pieces. max_round is the earliest cut round, so the tree ends where
# the labelled partition begins. Height is merge order, since the real rounds are
# spaced too unevenly for a linear axis.
build_gn_dendrogram <- function(edgelist_file, removal_log_file, output_png,
                                max_round, top_n_splits = 300L) {
  edges_df <- read.csv(edgelist_file, stringsAsFactors = FALSE)
  g <- igraph::graph_from_data_frame(edges_df[, c("source", "target")], directed = TRUE)
  if (igraph::is_directed(g)) g <- as_undirected(g, mode = "collapse")

  removal_log <- read.csv(removal_log_file, stringsAsFactors = FALSE)
  removal_log <- removal_log[order(removal_log$round), ]
  removal_log <- removal_log[removal_log$round <= max_round, ]

  # Replay the removals, recording each split with the node names of its pieces.
  # Components are tracked by an integer label per node.
  node_names <- igraph::V(g)$name
  comp_id <- rep(1L, length(node_names))
  names(comp_id) <- node_names
  next_comp_id <- 2L
  g_live <- g
  splits <- list()

  for (i in seq_len(nrow(removal_log))) {
    r <- removal_log$round[i]
    from <- removal_log$from[i]
    to <- removal_log$to[i]

    eid <- get_edge_ids(g_live, c(from, to))
    if (eid == 0) next
    g_live <- igraph::delete_edges(g_live, eid)

    # Only the component holding the removed edge can split.
    parent_comp <- comp_id[[from]]
    members <- names(comp_id)[comp_id == parent_comp]
    sub_g <- igraph::induced_subgraph(g_live, vids = members)
    sub_comp <- igraph::components(sub_g)

    if (sub_comp$no > 1) {
      pieces <- split(igraph::V(sub_g)$name, sub_comp$membership)
      for (k in seq_along(pieces)) {
        new_id <- if (k == 1) parent_comp else {
          id <- next_comp_id
          next_comp_id <- next_comp_id + 1L
          id
        }
        comp_id[pieces[[k]]] <- new_id
      }
      splits[[length(splits) + 1]] <- list(round = r, pieces = pieces)
    }
  }
  if (length(splits) == 0) {
    stop("No split events found. Check that the removal log matches the edge list.")
  }

  # Build the merge tree from the earliest top_n_splits splits (the coarsest
  # structure), walking them in decreasing round order. Read that way, each
  # split merges its pieces back into the piece that existed just before it,
  # which is what hclust's merge matrix needs. Each piece's hclust reference
  # (negative for a leaf, positive for the row that produced it) is recorded
  # when the piece is created.
  splits_to_use <- if (top_n_splits > 0 && top_n_splits < length(splits)) {
    splits[seq_len(top_n_splits)]
  } else {
    splits
  }
  splits_rev <- rev(splits_to_use)

  next_id <- 1L
  piece_size <- c()
  name_to_id <- new.env()
  is_leaf <- integer(0)
  hclust_ref <- new.env()
  leaf_counter <- 0L

  register_leaf <- function(members) {
    id <- next_id
    next_id <<- next_id + 1L
    piece_size[as.character(id)] <<- length(members)
    is_leaf <<- c(is_leaf, id)
    for (nm in members) assign(nm, id, envir = name_to_id)
    leaf_counter <<- leaf_counter + 1L
    assign(as.character(id), -leaf_counter, envir = hclust_ref)
    id
  }

  # Every member is checked for an existing id. Checking only the first member
  # fails because the large remaining component's first name changes each round.
  get_id_any_member <- function(piece_members) {
    for (nm in piece_members) {
      v <- mget(nm, envir = name_to_id, ifnotfound = NA, inherits = FALSE)[[1]]
      if (!is.na(v)) return(v)
    }
    NA_integer_
  }

  merge_a <- integer(0)
  merge_b <- integer(0)
  heights <- numeric(0)
  round_at_height <- integer(0)

  for (s in splits_rev) {
    ids_here <- integer(0)
    for (p in s$pieces) {
      id <- get_id_any_member(p)
      if (is.na(id)) {
        id <- register_leaf(p)
      } else {
        # Index members freshly separated off a lineage that is still shrinking.
        for (nm in p) {
          v <- mget(nm, envir = name_to_id, ifnotfound = NA, inherits = FALSE)[[1]]
          if (is.na(v)) assign(nm, id, envir = name_to_id)
        }
      }
      ids_here <- c(ids_here, id)
    }
    ids_here <- unique(ids_here)
    if (length(ids_here) < 2) next

    while (length(ids_here) >= 2) {
      a <- ids_here[1]
      b <- ids_here[2]
      parent_id <- next_id
      next_id <- next_id + 1L
      piece_size[as.character(parent_id)] <- piece_size[as.character(a)] + piece_size[as.character(b)]

      merge_a <- c(merge_a, get(as.character(a), envir = hclust_ref))
      merge_b <- c(merge_b, get(as.character(b), envir = hclust_ref))
      # Height is merge order. Raw rounds put almost every merge at nearly the
      # same height, because the splits in any window are a few hundred rounds
      # apart in a run of tens of thousands.
      heights <- c(heights, length(merge_a))
      round_at_height <- c(round_at_height, s$round)
      assign(as.character(parent_id), length(merge_a), envir = hclust_ref)

      for (nm in ls(name_to_id)) {
        v <- get(nm, envir = name_to_id)
        if (v == a || v == b) assign(nm, parent_id, envir = name_to_id)
      }
      ids_here <- c(parent_id, ids_here[-c(1, 2)])
    }
  }
  if (length(merge_a) == 0) {
    stop("Could not build any merges from the selected splits.")
  }

  leaf_size_by_pos <- vapply(is_leaf, function(id) piece_size[as.character(id)], numeric(1))
  leaf_pos_order <- vapply(is_leaf, function(id) -get(as.character(id), envir = hclust_ref), integer(1))
  leaf_labels <- character(length(is_leaf))
  leaf_labels[leaf_pos_order] <- sprintf("n=%d", leaf_size_by_pos)

  # hc$order must follow the tree's branch structure and must be unnamed,
  # because plot.hclust() rejects the names that order.dendrogram() inherits.
  hc <- list(
    merge = cbind(merge_a, merge_b),
    height = heights,
    order = seq_along(leaf_labels),
    labels = leaf_labels,
    method = "girvan-newman-divisive"
  )
  class(hc) <- "hclust"
  hc$order <- unname(order.dendrogram(as.dendrogram(hc)))

  grDevices::png(output_png, width = 4000, height = 2000, res = 200)
  on.exit(grDevices::dev.off(), add = TRUE)
  par(mar = c(5, 6, 4, 8))
  plot(hc, labels = FALSE, hang = -1,
       main = sprintf("Girvan-Newman dendrogram (top %d splits, %d leaves, rounds %d-%d)",
                      length(splits_to_use), length(is_leaf),
                      min(round_at_height), max(round_at_height)),
       xlab = "", sub = "", ylab = "Merge order (evenly spaced steps)")
  invisible(output_png)
}

# Replays the removal log inside each of the largest communities of one cut and
# records what happens to each afterward. The starting groups come straight from
# the cut's communities_round_N.csv, so the analysis cannot drift from the
# partition. Each community is replayed to natural completion, and one row per
# community is written to output_csv.
build_gn_subclusters <- function(edgelist_file, removal_log_file, cut_file, seed_round,
                                 output_csv, min_community_size = 3L, max_trees = 20L) {
  communities <- read.csv(cut_file, stringsAsFactors = FALSE)
  if (!all(c("node_id", "community_id") %in% names(communities))) {
    stop("Expected columns 'node_id' and 'community_id' in ", cut_file)
  }
  groups <- split(communities$node_id, communities$community_id)
  group_sizes <- sapply(groups, length)
  qualifying <- names(group_sizes)[group_sizes >= min_community_size]
  qualifying <- qualifying[order(-group_sizes[qualifying])]
  if (max_trees > 0 && length(qualifying) > max_trees) {
    qualifying <- qualifying[seq_len(max_trees)]
  }

  removal_log <- read.csv(removal_log_file, stringsAsFactors = FALSE)
  removal_log <- removal_log[order(removal_log$round), ]
  removal_after <- removal_log[removal_log$round > seed_round, ]

  edges_df <- read.csv(edgelist_file, stringsAsFactors = FALSE)
  g <- igraph::graph_from_data_frame(edges_df[, c("source", "target")], directed = TRUE)
  g <- as_undirected(g, mode = "collapse")

  # Split events inside one community: each round where removing an edge divides
  # the piece that held it, with the node names of the resulting pieces.
  find_splits <- function(member_names) {
    sub_edges <- removal_after[
      removal_after$from %in% member_names & removal_after$to %in% member_names,
    ]
    if (nrow(sub_edges) == 0) return(list())

    g_sub <- igraph::induced_subgraph(g, vids = member_names)
    comp_id <- rep(1L, length(member_names))
    names(comp_id) <- member_names
    next_comp_id <- 2L
    splits <- list()

    for (i in seq_len(nrow(sub_edges))) {
      r <- sub_edges$round[i]
      from <- sub_edges$from[i]
      to <- sub_edges$to[i]
      eid <- get_edge_ids(g_sub, c(from, to))
      if (eid == 0) next
      g_sub <- igraph::delete_edges(g_sub, eid)

      parent_comp <- comp_id[[from]]
      members_here <- names(comp_id)[comp_id == parent_comp]
      sub_g <- igraph::induced_subgraph(g_sub, vids = members_here)
      sub_comp <- igraph::components(sub_g)

      if (sub_comp$no > 1) {
        pieces <- split(igraph::V(sub_g)$name, sub_comp$membership)
        for (k in seq_along(pieces)) {
          new_id <- if (k == 1) parent_comp else {
            id <- next_comp_id
            next_comp_id <- next_comp_id + 1L
            id
          }
          comp_id[pieces[[k]]] <- new_id
        }
        splits[[length(splits) + 1]] <- list(round = r, pieces = pieces)
      }
    }
    splits
  }

  # Leaves of the community's split tree: pieces that never divide again, plus
  # members that no split touched. A piece's identity is resolved by checking
  # every member for an existing id, because the large remaining piece changes
  # its first member each round. Returns NULL when the community never divides.
  count_leaves <- function(splits, all_members) {
    if (length(splits) == 0) return(NULL)
    next_id <- 1L
    n_leaves <- 0L
    n_merges <- 0L
    name_to_id <- new.env()
    lookup <- function(nm) mget(nm, envir = name_to_id, ifnotfound = NA, inherits = FALSE)[[1]]
    get_id_any_member <- function(piece_members) {
      for (nm in piece_members) {
        v <- lookup(nm)
        if (!is.na(v)) return(v)
      }
      NA_integer_
    }

    for (s in rev(splits)) {
      ids_here <- integer(0)
      for (p in s$pieces) {
        id <- get_id_any_member(p)
        if (is.na(id)) {
          id <- next_id
          next_id <- next_id + 1L
          n_leaves <- n_leaves + 1L
          for (nm in p) assign(nm, id, envir = name_to_id)
        } else {
          for (nm in p) if (is.na(lookup(nm))) assign(nm, id, envir = name_to_id)
        }
        ids_here <- c(ids_here, id)
      }
      ids_here <- unique(ids_here)
      while (length(ids_here) >= 2) {
        a <- ids_here[1]
        b <- ids_here[2]
        parent_id <- next_id
        next_id <- next_id + 1L
        n_merges <- n_merges + 1L
        for (nm in ls(name_to_id)) {
          v <- get(nm, envir = name_to_id)
          if (v == a || v == b) assign(nm, parent_id, envir = name_to_id)
        }
        ids_here <- c(parent_id, ids_here[-c(1, 2)])
      }
    }
    n_leaves <- n_leaves + length(setdiff(all_members, ls(name_to_id)))
    if (n_merges == 0) return(NULL)
    n_leaves
  }

  rows <- lapply(qualifying, function(gid) {
    member_names <- as.character(groups[[gid]])
    splits <- find_splits(member_names)
    n_leaves <- count_leaves(splits, member_names)
    if (is.null(n_leaves)) {
      return(data.frame(community_id = gid, size = length(member_names), n_splits = 0,
                        n_leaves = 1, first_split_round = NA, first_split_sizes = NA))
    }
    data.frame(
      community_id = gid, size = length(member_names), n_splits = length(splits),
      n_leaves = n_leaves,
      first_split_round = vapply(splits, function(s) s$round, integer(1))[1],
      first_split_sizes = paste(sapply(splits[[1]]$pieces, length), collapse = "/")
    )
  })
  write.csv(do.call(rbind, rows), output_csv, row.names = FALSE)
  invisible(output_csv)
}

# The dendrogram and the subcluster summary are rebuilt from the run's removal
# log when they are missing or older than the file they come from, so a fresh
# GN run is reflected without a manual step, but re-rendering right after an
# unrelated edit does not re-run these multi-minute builds for no reason.
gn_dendrogram_png <- function(key) {
  dir <- gn_path(key, "dendrogram")
  png <- file.path(dir, "gn_dendrogram.png")
  log <- gn_path(key, "removal_log.csv")
  if (gn[[key]]$available && file.exists(log)) {
    stale <- !file.exists(png) ||
      file.info(png)$mtime < max(file.info(c(log, gn[[key]]$cut_file))$mtime)
    if (stale) {
      dir.create(dir, recursive = TRUE, showWarnings = FALSE)
      tryCatch(
        build_gn_dendrogram(graph_edgelists[[key]], log, png, gn[[key]]$earliest$round),
        error = function(e) {
          warning("The dendrogram for ", key, " could not be built: ", conditionMessage(e))
        }
      )
    }
  }
  png
}

gn_subclusters_csv <- function(key) {
  dir <- gn_path(key, "subclusters")
  csv <- file.path(dir, "subcluster_summary.csv")
  if (gn[[key]]$available) {
    stale <- !file.exists(csv) || file.info(csv)$mtime < file.info(gn[[key]]$cut_file)$mtime
    if (stale) {
      dir.create(dir, recursive = TRUE, showWarnings = FALSE)
      tryCatch(
        build_gn_subclusters(
          graph_edgelists[[key]], gn_path(key, "removal_log.csv"), gn[[key]]$cut_file,
          gn[[key]]$earliest$round, csv,
          min_community_size = 3L,
          max_trees = gn_top_n  # tied to gn_top_n so the prose's "largest communities"
                                # matches what is computed
        ),
        error = function(e) {
          warning("The subcluster summary for ", key, " could not be built: ", conditionMessage(e))
        }
      )
    }
  }
  csv
}

gn_ncomm_text <- function(key) {
  if (gn[[key]]$available) format(gn[[key]]$earliest$n_communities, big.mark = ",") else "an unknown number of"
}

# One row per Girvan-Newman round (round, edges_removed, n_components,
# modularity), written by run_gn_analysis.R. Two series on one plot with
# different scales (modularity is roughly 0-1, component count runs into the
# thousands), so component count is rescaled onto modularity's range and given
# its own right-hand axis with the real numbers.
gn_trace_plot <- function(key) {
  tr <- gn[[key]]$trace
  comp_scale <- max(tr$n_components) / max(tr$modularity, na.rm = TRUE)
  ggplot(tr, aes(x = round)) +
    geom_line(aes(y = modularity, colour = "Modularity")) +
    geom_line(aes(y = n_components / comp_scale, colour = "Connected components")) +
    geom_vline(xintercept = gn[[key]]$summary$round, linetype = "dashed",
               colour = "grey40", linewidth = 0.4) +
    scale_y_continuous(
      name = "Modularity",
      sec.axis = sec_axis(~ . * comp_scale, name = "Connected components",
                           labels = function(x) format(x, big.mark = ","))
    ) +
    scale_colour_manual(values = c("Modularity" = "firebrick", "Connected components" = "steelblue"),
                         name = NULL) +
    labs(
      title = paste0("Modularity and fragmentation across the full Girvan-Newman run, ", graph_labels[[key]]),
      subtitle = "Dashed lines mark the fragmentation-onset rounds in the table below",
      x = "Round"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")
}

gn_onset_table <- function(key) {
  gn[[key]]$summary |>
    transmute(
      Round = format(round, big.mark = ","),
      Communities = format(n_communities, big.mark = ","),
      Singletons = format(n_singletons, big.mark = ","),
      `% singleton` = scales::percent(pct_singleton, accuracy = 1),
      `Largest community` = largest_community
    ) |>
    knitr::kable()
}

# Highlights only the largest communities in distinct colours, with
# everything else grey. Colouring every community individually on a
# continuous scale is not legible: adjacent IDs land on near-identical
# shades. F1/F2 is a separate visual channel (a black ring), since it cuts
# across community membership rather than nesting inside it.
gn_prepare <- function(key) {
  part <- gn[[key]]$partition
  top <- part |> count(community_gn, sort = TRUE) |> slice_head(n = gn_top_n) |> pull(community_gn)
  lookup <- gn_label_lookup(key)
  tg <- as_tbl_graph(graphs[[key]]) |>
    activate(nodes) |>
    left_join(part, by = c("name" = "node_id")) |>
    left_join(metrics[[key]] |> select(name, betweenness, generation), by = "name") |>
    mutate(
      highlight = if_else(community_gn %in% top, community_gn, "other"),
      community_label = case_when(
        highlight == "other" ~ "Other",
        !is.na(lookup$label[match(highlight, lookup$community_id)]) ~ lookup$label[match(highlight, lookup$community_id)],
        TRUE ~ paste("Community", highlight)
      ),
      is_f1 = !is.na(generation) & generation == "F1"
    )
  nodes <- tg |> activate(nodes) |> as_tibble()
  n_top <- length(top)
  hues <- grDevices::hcl(h = seq(15, 375, length.out = n_top + 1)[-(n_top + 1)], c = 100, l = 65)
  colour_by_id <- c(other = "grey85", setNames(hues, top))
  # Palette keyed by the label text ggplot actually plots, built from the
  # graph's own data rather than re-keyed from community IDs.
  label_colours <- nodes |>
    distinct(highlight, community_label) |>
    mutate(colour = colour_by_id[as.character(highlight)]) |>
    select(community_label, colour) |>
    tibble::deframe()
  list(tg = tg, nodes = nodes, label_colours = label_colours, top = top)
}

# The plot shows every node. `crop = TRUE` trims the view to the 1st-99th
# percentile of each axis (coord_cartesian crops the view only, not the data),
# which the `lgl` trials in the appendix need because they throw a few outliers
# far out. The Fruchterman-Reingold layout places each connected component on
# its own and packs the components together, so the small disconnected pieces
# of the seed-filtered graph stay beside the main body instead of drifting off.
# `weights = NA` makes the layout ignore any edge weight attribute.
gn_plot_partition <- function(key, layout = "fr",
                              layout_args = list(niter = 500, grid = "grid", weights = NA),
                              crop = FALSE, f1_rings = TRUE) {
  d <- gn_prepare(key)
  tg <- d$tg |>
    activate(nodes) |>
    mutate(community_label = fct_reorder(community_label, highlight == "other"))
  set.seed(42)
  lay <- if (layout == "fr") {
    xy <- do.call(igraph::layout_components, c(list(tg, layout = igraph::layout_with_fr), layout_args))
    create_layout(tg, layout = "manual", x = xy[, 1], y = xy[, 2])
  } else if (layout == "lgl") {
    do.call(create_layout, c(list(tg, layout = "igraph", algorithm = "lgl"), layout_args))
  } else {
    do.call(create_layout, c(list(tg, layout = layout), layout_args))
  }
  n_f1 <- sum(d$nodes$is_f1 & d$nodes$highlight != "other", na.rm = TRUE)

  p <- ggraph(lay) +
    geom_edge_link(alpha = 0.04, colour = "grey75") +
    geom_node_point(aes(colour = community_label, size = betweenness), alpha = 0.8)
  if (f1_rings) {
    p <- p + geom_node_point(data = ~ dplyr::filter(.x, is_f1),
                             shape = 21, colour = "black", fill = NA, stroke = 0.5, size = 1.5)
  }
  p <- p +
    scale_colour_manual(values = d$label_colours, name = "Cluster",
                         guide = guide_legend(ncol = 1)) +
    scale_size_continuous(name = "Betweenness", range = c(0.3, 6)) +
    theme_graph() +
    theme(legend.text = element_text(size = 8)) +
    labs(
      title = paste0(graph_labels[[key]], ": ", format(vcount(graphs[[key]]), big.mark = ","),
                     " nodes, ", gn_top_n, " largest Girvan-Newman clusters, sized by betweenness"),
      subtitle = paste0(
        "(fragmentation-onset cut, round ", gn_round_text(key), ").",
        if (f1_rings) paste0(" Black-ringed dots mark F1 papers (", n_f1, " within these clusters).") else ""
      )
    )
  if (crop) {
    p <- p + coord_cartesian(xlim = quantile(lay$x, c(0.01, 0.99)),
                             ylim = quantile(lay$y, c(0.01, 0.99)))
  }
  p
}

gn_community_graph <- function(key) {
  d <- gn_prepare(key)
  part <- gn[[key]]$partition
  comm_nodes <- d$nodes |>
    filter(highlight != "other") |>
    count(highlight, community_label, name = "size")

  comm_edges <- edges_list[[key]] |>
    left_join(part, by = c("source" = "node_id")) |>
    rename(community_source = community_gn) |>
    left_join(part, by = c("target" = "node_id")) |>
    rename(community_target = community_gn) |>
    filter(
      community_source %in% comm_nodes$highlight,
      community_target %in% comm_nodes$highlight,
      community_source != community_target
    ) |>
    count(community_source, community_target, name = "weight") |>
    # Undirected for this summary: A-B and B-A collapsed into one real
    # crossing count, since the question is "how connected are these two
    # communities," not citation direction between them.
    mutate(pair_key = paste(pmin(community_source, community_target),
                             pmax(community_source, community_target))) |>
    group_by(pair_key) |>
    summarise(community_source = first(community_source),
              community_target = first(community_target),
              weight = sum(weight), .groups = "drop") |>
    select(-pair_key)

  g_comm <- graph_from_data_frame(comm_edges, directed = FALSE, vertices = comm_nodes)
  set.seed(42)
  ggraph(as_tbl_graph(g_comm), layout = "fr") +
    geom_edge_link(aes(width = weight), alpha = 0.4, colour = "steelblue") +
    scale_edge_width(name = "Cross-community\ncitation edges", range = c(0.2, 4)) +
    geom_node_point(aes(size = size, colour = community_label), alpha = 0.9) +
    geom_node_text(aes(label = community_label), repel = TRUE, size = 3, max.overlaps = 20) +
    scale_colour_manual(values = d$label_colours, guide = "none") +
    scale_size_continuous(name = "Cluster size\n(real membership)", range = c(3, 16)) +
    theme_graph() +
    labs(
      title = paste0("How the top ", gn_top_n, " Girvan-Newman clusters relate to each other, ",
                     graph_labels[[key]]),
      subtitle = paste0(
        "(fragmentation-onset cut, round ", gn_round_text(key), "). ",
        "Edge width is the real count of citation edges crossing between two clusters; ",
        "isolated clusters (no edges above) share no real citation ties with any other top-",
        gn_top_n, " cluster."
      )
    )
}

gn_first_splits_data <- function(key) {
  read_csv(gn_subclusters_csv(key), show_col_types = FALSE) |>
    filter(!is.na(first_split_sizes)) |>
    separate(first_split_sizes, into = c("piece_a", "piece_b"), sep = "/", convert = TRUE) |>
    mutate(larger_piece = pmax(piece_a, piece_b), smaller_piece = pmin(piece_a, piece_b))
}

gn_first_split_text <- function() {
  one <- function(key) {
    if (!gn[[key]]$available) return("the first-split sizes are not available until the run is in place.")
    d <- gn_first_splits_data(key)
    n <- nrow(d)
    k <- sum(d$smaller_piece == 1)
    if (n > 0 && k == n) {
      paste0("every one of the ", n, " largest communities sheds exactly one node at first split, with the remaining piece staying fully intact.")
    } else {
      paste0(k, " of the ", n, " largest communities shed exactly one node at first split, with the remaining piece staying intact.")
    }
  }
  paste0("In Annotated, ", one("annotated"), " In Annotated, seed filtered, ", one("seed_filtered"))
}
gn_first_splits_plot <- function(key) {
  first_splits <- gn_first_splits_data(key) |>
    arrange(desc(smaller_piece)) |>
    mutate(community_label = fct_reorder(paste("Community", community_id), smaller_piece))

  long <- first_splits |>
    select(community_label, larger_piece, smaller_piece) |>
    pivot_longer(c(larger_piece, smaller_piece), names_to = "piece", values_to = "size") |>
    mutate(piece = factor(piece, levels = c("larger_piece", "smaller_piece"),
                           labels = c("Larger piece", "Smaller piece")))

  ggplot(long, aes(x = size, y = community_label, fill = piece)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_text(aes(label = size), position = position_dodge(width = 0.7), hjust = -0.15, size = 3) +
    scale_x_log10(expand = expansion(mult = c(0, 0.15))) +
    scale_fill_manual(values = c("Larger piece" = "#4C72B0", "Smaller piece" = "#DD8452")) +
    labs(
      title = paste0("First-split piece sizes, round ", gn_round_text(key), "'s largest communities, ",
                     graph_labels[[key]]),
      subtitle = "Sorted by size of the smaller piece; most balanced at top",
      x = "Nodes in each piece at first split (log scale)",
      y = NULL, fill = NULL
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")
}

# One table per run, built from that run's own labels/ folder.
gn_cluster_labels_table <- function(key) {
  labels_full <- read_csv(gn_path(key, "labels", "cluster_labels.csv"), show_col_types = FALSE)

  # representatives.csv carries the within-cluster degree of each sampled
  # paper, in rank order (see select_representatives.py), reused here to derive
  # the same hub-dominated classification that script computes internally and
  # only prints to the console: rank-1's within-cluster degree at least
  # --hub-ratio-threshold (5.0, the script's own default) times rank-2's.
  # cluster_size, also from this file, is the figure the labelling pipeline
  # used, not a count recomputed here.
  reps_path <- gn_path(key, "labels", "representatives.csv")
  cluster_shape <- if (file.exists(reps_path)) {
    read_csv(reps_path, show_col_types = FALSE) |>
      group_by(community_id) |>
      summarise(
        cluster_size = first(cluster_size),
        shape = if (n() >= 2 && within_cluster_degree[2] > 0 &&
                    within_cluster_degree[1] / within_cluster_degree[2] >= 5.0) {
          "Hub-dominated"
        } else {
          "Distributed"
        },
        .groups = "drop"
      ) |>
      mutate(community_id = as.character(community_id))
  } else {
    tibble(community_id = character(), cluster_size = integer(), shape = character())
  }

  labels_full |>
    mutate(community_id = as.character(community_id)) |>
    left_join(cluster_shape, by = "community_id") |>
    arrange(desc(cluster_size)) |>
    transmute(
      `Cluster ID` = community_id,
      Size = cluster_size,
      Shape = coalesce(shape, "n/a"),
      Label = label,
      Description = description,
      `Secondary label` = coalesce(secondary_label, "")
    ) |>
    knitr::kable()
}
```

## Community detection

Community detection here produces several independent descriptions of how
the graph divides, not a single answer to be triangulated. Louvain
[@blondel2008] and Leiden [@traag2019] ask "which groups of nodes are more
densely interconnected than a random graph with the same degree sequence
would predict," a notion formalized as modularity [@newman2006].
Girvan-Newman [@girvan2002] asks "which edges, if removed, disconnect
otherwise-separate parts of the network." K-means asks a third,
differently-grounded question again: not about graph structure at all,
but about each work's centrality profile, grouping works that occupy
similar structural positions, for example highly-cited authorities vs.
high-betweenness brokers vs. peripheral works.

### Louvain, Leiden and K-means

::: {.callout-note}
Louvain, Leiden and K-means are computed for both graphs in `graph_metrics()` with no visible output. Their agreement across the two graphs is reported in "What the seed filter changes". The plotting code for each is kept in the appendix, disabled; toggle `eval: true` on any of those chunks to render it.
:::

### Girvan-Newman

#### How it works

Louvain and Leiden each partition the graph once, at one resolution,
and return a single result. Girvan-Newman [@girvan2002] instead removes
edges one at a time: at each step, it identifies the edge with the
highest betweenness centrality (the edge carrying the largest number of
shortest paths between all pairs of nodes in the graph), removes it,
and recomputes betweenness for the remaining edges before repeating.
Because a shortest path between two densely connected regions of a
graph usually has to cross whichever single edge joins them, that edge
tends to have disproportionately high betweenness and is removed
early, while edges internal to a well-connected region, which have
many alternative paths around them, are removed only much later. The
result is a record of the entire removal sequence, from the original,
fully connected graph down to every paper standing alone: a
dendrogram. The tree below is that record. Height represents the order
in which components split apart, not the raw round number, since the
actual rounds are spaced too unevenly to remain legible on a linear
axis; left-right position carries no information and is arranged only
to avoid crossing branches.

::: {.panel-tabset}

##### Annotated

```{r}
#| label: dendrogram-annotated
#| out-width: "100%"
#| eval: !expr gn[["annotated"]]$available && file.exists(gn_dendrogram_png("annotated"))
knitr::include_graphics(gn_dendrogram_png("annotated"))
```

##### Annotated, seed filtered

```{r}
#| label: dendrogram-seed-filtered
#| out-width: "100%"
#| eval: !expr gn[["seed_filtered"]]$available && file.exists(gn_dendrogram_png("seed_filtered"))
knitr::include_graphics(gn_dendrogram_png("seed_filtered"))
```

:::

#### Choosing a stopping point

Girvan-Newman, in its original form, does not select a stopping point
from this sequence [@fortunato2010]; the removal process continues
until no edges remain. A refinement published two years after the
original algorithm [@newman2004] selects the
round with the highest modularity score as the working partition, and
this is the conventional choice, but it fails on this corpus: modularity
continues to rise as the graph fragments, reaching its maximum only once
`r gn_pct_phrase("annotated")` are singleton communities (`r gn_pct_phrase("seed_filtered")` in the seed filtered run), which is not a usable partition. This is a documented limitation of modularity on sparse
networks, where it can lack a clear maximum and instead admit many
different high-scoring partitions [@good2010]. The criterion used
instead counts connected components directly: the point where that
count starts climbing steadily, one round after another, marks where
the graph is actually breaking into pieces, independent of whether
modularity happens to be rising or falling at the same time. The earliest such round is `r gn_round_text("annotated")` (Annotated) and `r gn_round_text("seed_filtered")` (Annotated, seed filtered), and it is used for every community grouping in the rest of this document.

::: {.panel-tabset}

##### Annotated

```{r}
#| label: stopping-1-annotated
#| eval: !expr gn[["annotated"]]$available
gn_trace_plot("annotated")
```

```{r}
#| label: stopping-2-annotated
#| eval: !expr gn[["annotated"]]$available
gn_onset_table("annotated")
```

##### Annotated, seed filtered

```{r}
#| label: stopping-1-seed-filtered
#| eval: !expr gn[["seed_filtered"]]$available
gn_trace_plot("seed_filtered")
```

```{r}
#| label: stopping-2-seed-filtered
#| eval: !expr gn[["seed_filtered"]]$available
gn_onset_table("seed_filtered")
```

:::

#### The working partition

Only the top `r gn_top_n` clusters are individually coloured and the rest are grey, since beyond roughly `r gn_top_n` colours a reader cannot reliably distinguish them on the plot. F1 papers carry a black ring except in the tabs marked no F1 rings. Node size is each node's betweenness centrality in that graph, so the largest dots in the dense core are the corpus's structural bridges. The dense core is real structure and no layout parameter changes what it is. Every plot uses the Fruchterman-Reingold layout, the clearest of the layouts tried (see the appendix).

::: {.panel-tabset}

##### Annotated

```{r}
#| label: partition-annotated
#| fig-width: 16
#| fig-height: 14
#| eval: !expr gn[["annotated"]]$available
gn_plot_partition("annotated")
```

##### Annotated, no F1 rings

```{r}
#| label: partition-annotated-plain
#| fig-width: 16
#| fig-height: 14
#| eval: !expr gn[["annotated"]]$available
gn_plot_partition("annotated", f1_rings = FALSE)
```

##### Annotated, seed filtered

```{r}
#| label: partition-seed-filtered
#| fig-width: 16
#| fig-height: 14
#| eval: !expr gn[["seed_filtered"]]$available
gn_plot_partition("seed_filtered")
```

##### Annotated, seed filtered, no F1 rings

```{r}
#| label: partition-seed-filtered-plain
#| fig-width: 16
#| fig-height: 14
#| eval: !expr gn[["seed_filtered"]]$available
gn_plot_partition("seed_filtered", f1_rings = FALSE)
```

:::

#### Community-level structure

The plots above show individual nodes coloured by
community; none of them show which communities are strongly connected
to which others, since edges are drawn identically whether they cross
community boundaries or stay within one. This collapses each of the
top `r gn_top_n` clusters to a single node, sized by its real
membership count, with an edge between two cluster-nodes weighted by
the real number of citation edges crossing between them in the
original graph — the genuine inter-community structure, not
individual-node detail.

::: {.panel-tabset}

##### Annotated

```{r}
#| label: community-graph-annotated
#| fig-width: 14
#| fig-height: 12
#| eval: !expr gn[["annotated"]]$available
gn_community_graph("annotated")
```

##### Annotated, seed filtered

```{r}
#| label: community-graph-seed-filtered
#| fig-width: 14
#| fig-height: 12
#| eval: !expr gn[["seed_filtered"]]$available
gn_community_graph("seed_filtered")
```

:::

A real edge here means real papers in one cluster actually cite, or
are cited by, real papers in the other; its width is that count, not
an estimate. A cluster with no edges at all in this plot is real too:
it means none of the top `r gn_top_n` clusters share a direct citation
tie with it, whatever thematic similarity a reader might expect from
the labels alone.

#### Internal structure of individual communities

Each graph's earliest-cut partition and its whole-graph dendrogram above both describe the corpus at a single level: `r gn_ncomm_text("annotated")` communities in Annotated and `r gn_ncomm_text("seed_filtered")` in Annotated, seed filtered, and one tree per graph covering the earliest few hundred splits across all of them together.
A separate question is whether any individual community has further
structure worth surfacing on its own, the same kind of check performed
earlier (see "Choosing a stopping point") for two clusters against
paper titles and annotation coding.

The subcluster analysis addresses this directly: seeded from the real community membership of each graph's earliest cut (not a re-derived or arbitrary starting point), it replays `removal_log.csv` separately for
each of the round's largest communities and records what happens to
each one afterward, run below using that same round so it can't drift
out of sync with the partition described above. An initial version
capped this replay at a fixed number of split events; that cap was
itself an arbitrary choice with no basis in the data, so it
now replays each community to natural completion (full fragmentation
or the end of the log) by default.

Doing so shows that, past each community's first split, every one of
the `r gn_top_n` largest communities collapses the same way the whole graph
does: a single dominant piece sheds one node at a time for hundreds of
further rounds, with no further balanced division anywhere in the
sequence. Plotted as a dendrogram, this produces the same long
single-branch shape for every community regardless of its size or its
first split, differing only in length, so the per-community dendrogram
plots turned out uninformative and are not included here.

The one real signal in this data is the size of the two pieces at each
community's first split, plotted below. `r gn_first_split_text()` Any community whose two
pieces are closer in size is a candidate for the same kind of
title/annotation check performed earlier (see "Choosing a stopping
point"), and any community where the *smaller* piece is the one that
stays intact — rather than the one shed — runs against the pattern
seen in every other community here and in the previous version of this
corpus, and would be worth investigating further. Reading the specific
communities and split sizes that meet either description off the plot,
rather than naming them in prose here, keeps this section accurate
across re-runs against different partitions.

::: {.panel-tabset}

##### Annotated

```{r}
#| label: first-splits-annotated
#| fig-width: 11
#| fig-height: 6.5
#| eval: !expr gn[["annotated"]]$available
gn_first_splits_plot("annotated")
```

##### Annotated, seed filtered

```{r}
#| label: first-splits-seed-filtered
#| fig-width: 11
#| fig-height: 6.5
#| eval: !expr gn[["seed_filtered"]]$available
gn_first_splits_plot("seed_filtered")
```

:::

### Cluster labels

This section identifies each cluster's most-cited papers by
within-cluster citation degree, then has a language model write a
label and description from their titles, independent of the F1
annotation codebook. Comparing these labels against the codebook
happens separately, in `03_joined_descriptions.qmd`.

For each cluster, `select_representatives.py` ranks the cluster's own
papers by how many other members of that cluster cite them, and takes
the top few as a representative sample. `label_clusters.py` sends
those titles to a language model, which sorts them into named groups
and writes a label and description from the largest group; a second
label is written only when a real second group exists with enough
titles behind it. Full detail on both steps, including the
hub-independence rule and the mechanical checks run on the model's
output, is in `cluster_labelling/README.md`.

Each row is one cluster.

| Column | Meaning |
|---|---|
| Cluster ID | Girvan-Newman community ID at the working partition |
| Size | Cluster size |
| Shape | `Hub-dominated` (top paper's within-cluster citation degree ≥ 5× the next-highest, `--hub-ratio-threshold` per the README), `Distributed` (otherwise), or `n/a` (no data for this cluster) |
| Label | From the language model's primary group for this cluster |
| Description | From the language model's primary group for this cluster |
| Secondary label | From a real second group, where one exists; blank otherwise |

::: {.panel-tabset}

#### Annotated

```{r}
#| label: cluster-labels-annotated
#| echo: false
#| eval: !expr file.exists(gn_path("annotated", "labels", "cluster_labels.csv"))
gn_cluster_labels_table("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: cluster-labels-seed-filtered
#| echo: false
#| eval: !expr file.exists(gn_path("seed_filtered", "labels", "cluster_labels.csv"))
gn_cluster_labels_table("seed_filtered")
```

:::


### Communities of communities

```{r}
#| label: grouping-functions
#| echo: false
# One run of a community-detection method is a weak basis for groups, because the
# result changes with the method, the resolution and what counts as a link between
# two clusters. The groups come from a consensus over several variants, and each
# group reports how often its members were grouped together.
gn_group_cut <- 0.5          # co-association needed on average to share a group
gn_group_seeds <- 1:5
gn_group_min_papers <- 10    # smallest cluster drawn in the figures

gn_leiden_groups <- function(g, resolution = 1) {
  args <- list(g, objective_function = "modularity", weights = igraph::E(g)$weight, n_iterations = 10)
  # igraph 2.x renamed resolution_parameter to resolution.
  args[[if ("resolution" %in% names(formals(igraph::cluster_leiden))) "resolution" else "resolution_parameter"]] <- resolution
  as.integer(igraph::membership(do.call(igraph::cluster_leiden, args)))
}

# One row per citation edge whose two papers both sit in a cluster, with the
# clusters of the citing paper (from) and the cited paper (to).
gn_cluster_edges <- function(key) {
  part <- gn[[key]]$partition
  edges_list[[key]] |>
    select(source, target) |>
    left_join(transmute(part, source = node_id, from = community_gn), by = "source") |>
    left_join(transmute(part, target = node_id, to = community_gn), by = "target") |>
    filter(!is.na(from), !is.na(to))
}

# Two clusters are linked by the citation edges that cross between them.
gn_cluster_link_graph <- function(ce, clusters, directed = FALSE) {
  x <- ce |> filter(from != to, from %in% clusters, to %in% clusters)
  if (!directed) x <- x |> mutate(a = pmin(from, to), b = pmax(from, to), from = a, to = b)
  x <- x |> count(from, to, name = "weight")
  igraph::graph_from_data_frame(x, directed = directed, vertices = data.frame(name = clusters))
}

# Two clusters are also linked when their citing papers cite the same works,
# measured as the cosine of the two sets of cited works.
gn_cluster_reference_graph <- function(ce, clusters) {
  refs <- ce |> filter(from %in% clusters) |> distinct(from, target)
  m <- Matrix::sparseMatrix(i = match(refs$from, clusters), j = as.integer(factor(refs$target)),
                            x = 1, dims = c(length(clusters), n_distinct(refs$target)))
  b <- as.matrix(Matrix::tcrossprod(m))
  d <- sqrt(Matrix::rowSums(m))
  b <- b / outer(d, d)
  b[!is.finite(b)] <- 0
  diag(b) <- 0
  g <- igraph::graph_from_adjacency_matrix(b, mode = "undirected", weighted = TRUE, diag = FALSE)
  igraph::V(g)$name <- clusters
  g
}

gn_group_variant_names <- c("Leiden, resolution 0.75", "Leiden, resolution 1", "Leiden, resolution 1.25",
                            "Leiden, resolution 1.5", "Louvain", "Infomap, undirected links",
                            "Infomap, directed links", "Leiden, shared references")
gn_group_n_variants <- length(gn_group_variant_names)

gn_cluster_groups <- function(key) {
  ce <- gn_cluster_edges(key)
  sizes <- count(gn[[key]]$partition, community_gn, name = "papers") |> arrange(desc(papers), community_gn)
  clusters <- sizes$community_gn
  g_links <- gn_cluster_link_graph(ce, clusters)
  if (length(clusters) < 3 || igraph::ecount(g_links) == 0) return(NULL)
  g_dir <- gn_cluster_link_graph(ce, clusters, directed = TRUE)
  g_refs <- gn_cluster_reference_graph(ce, clusters)
  w <- igraph::E(g_links)$weight
  wd <- igraph::E(g_dir)$weight
  membership_of <- function(x) as.integer(igraph::membership(x))
  variants <- list(
    function() gn_leiden_groups(g_links, 0.75),
    function() gn_leiden_groups(g_links, 1),
    function() gn_leiden_groups(g_links, 1.25),
    function() gn_leiden_groups(g_links, 1.5),
    function() membership_of(igraph::cluster_louvain(g_links, weights = w)),
    function() membership_of(igraph::cluster_infomap(g_links, e.weights = w)),
    function() membership_of(igraph::cluster_infomap(g_dir, e.weights = wd)),
    function() gn_leiden_groups(g_refs, 1))
  names(variants) <- gn_group_variant_names
  larger <- which(sizes$papers >= gn_group_min_papers)
  runs <- map(variants, function(f) {
    compact(map(gn_group_seeds, function(s) {
      set.seed(s)
      tryCatch(f(), error = function(e) NULL)
    }))
  })
  runs <- runs[lengths(runs) > 0]
  n <- length(clusters)
  coassoc <- matrix(0, n, n)
  for (r in runs) {
    together <- matrix(0, n, n)
    for (m in r) together <- together + outer(m, m, "==")
    coassoc <- coassoc + together / length(r)
  }
  coassoc <- coassoc / length(runs)
  diag(coassoc) <- 1
  raw <- cutree(hclust(as.dist(1 - coassoc), method = "average"), h = gn_group_cut)
  totals <- tapply(sizes$papers, raw, sum)
  group <- as.integer(rank(-totals, ties.method = "first")[as.character(raw)])
  agreement <- map_dbl(split(seq_len(n), group), function(ix) {
    if (length(ix) > 1) mean(coassoc[ix, ix][lower.tri(coassoc[ix, ix])]) else NA_real_
  })
  a <- as.matrix(igraph::as_adjacency_matrix(g_links, attr = "weight"))
  inside <- rowSums(a * outer(group, group, "==")) / rowSums(a)
  list(
    clusters = tibble(community_gn = clusters, papers = sizes$papers, group = group,
                      agreement = unname(agreement[as.character(group)]),
                      links = rowSums(a), share_links_inside = inside),
    variants = tibble(
      variant = names(runs),
      groups = unname(map_chr(runs, function(r) {
        k <- map_int(r, function(m) length(unique(m[larger])))
        if (min(k) == max(k)) as.character(min(k)) else paste0(min(k), " to ", max(k))
      })),
      agreement = unname(map_dbl(runs, function(r) {
        mean(map_dbl(r, function(m) igraph::compare(m[larger], group[larger], method = "adjusted.rand")))
      }))),
    links = g_links)
}

gn_has_group <- function(key) !is.null(gn_groups[[key]])

gn_group_note <- function(key) {
  d <- gn_groups[[key]]
  if (is.null(d)) return("No grouping is available for this graph.")
  sizes <- count(d$clusters, group)
  n_multi <- sum(sizes$n > 1)
  n_alone <- sum(sizes$n == 1)
  paste0(format(nrow(d$clusters), big.mark = ","), " clusters fall into ", n_multi,
         if (n_multi == 1) " group" else " groups", " of two or more clusters, and ", n_alone,
         if (n_alone == 1) " cluster stays" else " clusters stay", " alone. ",
         "The table lists the groups of two or more clusters, and larger clusters are those with ",
         gn_group_min_papers, " or more papers. The figures show only the larger clusters.")
}

gn_group_table <- function(key) {
  d <- gn_groups[[key]]
  lookup <- gn_label_lookup(key)
  multi <- d$clusters |> count(group) |> filter(n > 1) |> pull(group)
  d$clusters |>
    mutate(name = coalesce(lookup$label[match(community_gn, lookup$community_id)], paste0("cluster ", community_gn))) |>
    filter(group %in% multi) |>
    group_by(group) |>
    summarise(Clusters = n(), `Larger clusters` = sum(papers >= gn_group_min_papers), Papers = sum(papers), Agreement = first(agreement),
              `Largest clusters` = paste0(head(name[order(-papers)], 3), " (", head(sort(papers, decreasing = TRUE), 3), ")", collapse = "; "),
              .groups = "drop") |>
    arrange(desc(Papers)) |>
    mutate(Agreement = scales::percent(Agreement, accuracy = 1)) |>
    rename(Group = group) |>
    knitr::kable()
}

gn_group_graph <- function(key) {
  d <- gn_groups[[key]]
  shown <- filter(d$clusters, papers >= gn_group_min_papers)
  g <- igraph::induced_subgraph(d$links, shown$community_gn)
  link_counts <- igraph::E(g)$weight
  # The layout reads the weight attribute separately for each connected piece.
  igraph::E(g)$weight <- sqrt(link_counts)
  set.seed(42)
  xy <- igraph::layout_components(g, layout = igraph::layout_with_fr, niter = 1000)
  top <- shown |> group_by(group) |> summarise(p = sum(papers), .groups = "drop") |>
    arrange(desc(p)) |> slice_head(n = 12) |> pull(group)
  hues <- setNames(grDevices::hcl(h = seq(15, 375, length.out = length(top) + 1)[-(length(top) + 1)], c = 100, l = 62), top)
  nodes <- tibble(x = xy[, 1], y = xy[, 2], community_gn = igraph::V(g)$name) |>
    left_join(shown, by = "community_gn") |>
    mutate(colour = if_else(group %in% top, hues[as.character(group)], "grey75"))
  el <- igraph::as_edgelist(g, names = FALSE)
  segments <- tibble(x = xy[el[, 1], 1], y = xy[el[, 1], 2], xend = xy[el[, 2], 1], yend = xy[el[, 2], 2],
                     weight = link_counts)
  centres <- nodes |> filter(group %in% top) |> group_by(group) |>
    summarise(x = median(x), y = median(y), .groups = "drop")
  ggplot() +
    geom_segment(data = filter(segments, weight >= 2), aes(x, y, xend = xend, yend = yend, linewidth = weight), alpha = 0.2, colour = "grey50") +
    scale_linewidth(range = c(0.1, 2), guide = "none") +
    geom_point(data = nodes, aes(x, y, size = papers, colour = colour), alpha = 0.85) +
    scale_colour_identity() +
    scale_size_continuous(range = c(1.5, 10), guide = "none") +
    geom_text(data = centres, aes(x, y, label = group), size = 5, fontface = "bold") +
    theme_graph() +
    labs(title = paste0(graph_labels[[key]], ": clusters of ", gn_group_min_papers, " or more papers, coloured by group"),
         subtitle = "Dot size is the number of papers in the cluster. Edges are drawn where two clusters have at least two citation links, and their width is that count.")
}

gn_group_heatmap <- function(key) {
  d <- gn_groups[[key]]
  shown <- d$clusters |> filter(papers >= gn_group_min_papers)
  # Groups with fewer than three larger clusters share the last block.
  in_block <- shown |> count(group) |> filter(n >= 3) |> pull(group)
  shown <- shown |> mutate(block = if_else(group %in% in_block, group, Inf)) |> arrange(block, group, desc(papers))
  g <- igraph::induced_subgraph(d$links, shown$community_gn)
  a <- as.matrix(igraph::as_adjacency_matrix(g, attr = "weight"))[shown$community_gn, shown$community_gn]
  s <- rowSums(a)
  m <- a / sqrt(outer(s, s))
  m[!is.finite(m)] <- 0
  n <- nrow(m)
  long <- tibble(i = rep(seq_len(n), n), j = rep(seq_len(n), each = n), value = as.vector(m))
  bounds <- cumsum(table(factor(shown$block, levels = unique(shown$block)))) + 0.5
  ggplot(long, aes(j, i, fill = value)) +
    geom_raster() +
    scale_fill_gradient(low = "white", high = "black", trans = "sqrt", name = "Citation links,\nadjusted for size") +
    scale_y_reverse() +
    geom_vline(xintercept = head(bounds, -1), colour = "red", linewidth = 0.3) +
    geom_hline(yintercept = head(bounds, -1), colour = "red", linewidth = 0.3) +
    coord_equal() +
    theme_void() +
    labs(title = paste0(graph_labels[[key]], ": citation links between clusters, ordered by group"),
         subtitle = paste0("Clusters of ", gn_group_min_papers, " or more papers. Red lines separate groups, and the last block pools small groups."))
}

gn_group_variants_table <- function(key) {
  gn_groups[[key]]$variants |>
    rename(Variant = variant, `Groups among the larger clusters` = groups, `Agreement with the consensus` = agreement) |>
    knitr::kable(digits = 2)
}
```

```{r}
#| label: grouping-compute
#| echo: false
gn_groups <- map(c(annotated = "annotated", seed_filtered = "seed_filtered"), function(k) {
  if (!gn[[k]]$available) return(NULL)
  tryCatch(gn_cluster_groups(k), error = function(e) {
    warning("The grouping of clusters for ", k, " failed: ", conditionMessage(e))
    NULL
  })
})
```

The clusters can be grouped in turn. Two clusters are linked when papers in one cite papers in the other, and also when their citing papers cite many of the same works. A single run of one community-detection method is a weak basis for groups here, because the result changes with the method, the resolution and the definition of a link. The grouping below therefore runs `r gn_group_n_variants` variants, with `r length(gn_group_seeds)` random seeds each, on every cluster of every size. The variants are Leiden at four resolutions, Louvain, Infomap on undirected and on directed links, and Leiden on shared references. Two clusters share a group when they were grouped together in at least `r scales::percent(1 - gn_group_cut)` of the runs on average. Each group reports the share of runs in which its members stayed together. A high share is necessary but not enough, because weakly linked clusters can be grouped together consistently without sharing a theme, so the cluster labels in each row need reading as well.

::: {.panel-tabset}

#### Annotated

`r gn_group_note("annotated")`

```{r}
#| label: groups-table-annotated
#| echo: false
#| eval: !expr gn_has_group("annotated")
gn_group_table("annotated")
```

```{r}
#| label: groups-graph-annotated
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| eval: !expr gn_has_group("annotated")
gn_group_graph("annotated")
```

```{r}
#| label: groups-heatmap-annotated
#| echo: false
#| fig-width: 8
#| fig-height: 8
#| eval: !expr gn_has_group("annotated")
gn_group_heatmap("annotated")
```

```{r}
#| label: groups-variants-annotated
#| echo: false
#| eval: !expr gn_has_group("annotated")
gn_group_variants_table("annotated")
```

#### Annotated, seed filtered

`r gn_group_note("seed_filtered")`

```{r}
#| label: groups-table-seed-filtered
#| echo: false
#| eval: !expr gn_has_group("seed_filtered")
gn_group_table("seed_filtered")
```

```{r}
#| label: groups-graph-seed-filtered
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| eval: !expr gn_has_group("seed_filtered")
gn_group_graph("seed_filtered")
```

```{r}
#| label: groups-heatmap-seed-filtered
#| echo: false
#| fig-width: 8
#| fig-height: 8
#| eval: !expr gn_has_group("seed_filtered")
gn_group_heatmap("seed_filtered")
```

```{r}
#| label: groups-variants-seed-filtered
#| echo: false
#| eval: !expr gn_has_group("seed_filtered")
gn_group_variants_table("seed_filtered")
```

:::

The table before each figure lists the groups of two or more clusters, with the three largest clusters in each. In the cluster graph, each dot is a cluster sized by its papers, edges are citation links between clusters, and the numbers mark the twelve largest groups. In the heatmap, darker cells mean more citation links between two clusters after allowing for how many links each has, and the red lines separate groups. The last table shows how closely each variant agrees with the consensus on the larger clusters, as an adjusted Rand index where 1 is identical. A variant that agrees little finds different structure or none. The consensus is built from all of them regardless, so the share reported for each group is a cautious figure.


### Cluster shapes

```{r}
#| label: shape-functions
#| echo: false
gn_shape_min_papers <- 3
gn_shape_single_hub <- 0.95
gn_shape_shared <- 0.2

# One row per cluster. Every measure comes from the citation edges between
# papers, labelled with the cluster of each end (gn_cluster_edges).
gn_cluster_shapes <- function(key) {
  ce <- gn_cluster_edges(key)
  sizes <- count(gn[[key]]$partition, community_gn, name = "papers")
  internal <- filter(ce, from == to)
  cross <- filter(ce, from != to)
  out_total <- count(ce, source, name = "out_total")

  hubs <- internal |>
    count(from, source, name = "refs_inside") |>
    left_join(out_total, by = "source") |>
    arrange(from, desc(refs_inside)) |>
    group_by(from) |>
    summarise(citing = n(), edges_inside = sum(refs_inside),
              top_share = max(refs_inside) / sum(refs_inside),
              hub_paper = source[1], hub_refs_inside = refs_inside[1], hub_refs_total = out_total[1],
              .groups = "drop")
  out_all <- count(ce, from, name = "out_all")
  shared <- internal |> count(from, target, name = "n_citing") |>
    group_by(from) |> summarise(shared_share = mean(n_citing >= 2), .groups = "drop")
  one_tie <- bind_rows(transmute(internal, from, paper = source), transmute(internal, from, paper = target)) |>
    count(from, paper) |> group_by(from) |> summarise(one_tie_papers = sum(n == 1), .groups = "drop")
  outsiders <- ce |> group_by(to, target) |> summarise(outside = any(from != to), .groups = "drop") |>
    group_by(to) |> summarise(cited_from_outside = mean(outside), .groups = "drop")
  touching <- bind_rows(count(internal, cluster = from, name = "n"),
                        count(cross, cluster = from, name = "n"),
                        count(cross, cluster = to, name = "n")) |>
    group_by(cluster) |> summarise(touching = sum(n), .groups = "drop")
  neighbours <- bind_rows(transmute(cross, cluster = from, other = to),
                          transmute(cross, cluster = to, other = from)) |>
    group_by(cluster) |> summarise(neighbour_clusters = n_distinct(other), .groups = "drop")

  sizes |>
    filter(papers >= gn_shape_min_papers) |>
    left_join(hubs, by = c("community_gn" = "from")) |>
    left_join(out_all, by = c("community_gn" = "from")) |>
    left_join(shared, by = c("community_gn" = "from")) |>
    left_join(one_tie, by = c("community_gn" = "from")) |>
    left_join(outsiders, by = c("community_gn" = "to")) |>
    left_join(touching, by = c("community_gn" = "cluster")) |>
    left_join(neighbours, by = c("community_gn" = "cluster")) |>
    left_join(select(node_attrs, name, hub_title = title, hub_year = year), by = c("hub_paper" = "name")) |>
    mutate(
      shared_share = replace_na(shared_share, 0),
      neighbour_clusters = replace_na(neighbour_clusters, 0L),
      edges_per_paper = edges_inside / papers,
      one_tie_share = one_tie_papers / papers,
      kept = edges_inside / out_all,
      cohesion = edges_inside / touching,
      shape = case_when(top_share >= gn_shape_single_hub ~ "Single hub",
                        shared_share >= gn_shape_shared ~ "Shared references",
                        TRUE ~ "Several hubs")
    )
}

gn_has_shapes <- function(key) !is.null(gn_shapes[[key]])

gn_shape_note <- function(key) {
  if (!gn_has_shapes(key)) return("No cluster shapes are available for this graph.")
  s <- gn_shapes[[key]]
  n <- count(s, shape)
  part <- function(label) {
    k <- n$n[n$shape == label]
    if (length(k) == 0) 0L else k
  }
  paste0(nrow(s), " clusters have ", gn_shape_min_papers, " or more papers. ", part("Single hub"),
         " are single hubs, ", part("Shared references"), " share references and ", part("Several hubs"),
         " have several hubs without shared references.")
}

gn_shape_summary <- function(key) {
  s <- gn_shapes[[key]]
  q <- function(x, accuracy) {
    paste0(scales::number(median(x, na.rm = TRUE), accuracy = accuracy), " (",
           scales::number(quantile(x, 0.25, na.rm = TRUE), accuracy = accuracy), " to ",
           scales::number(quantile(x, 0.75, na.rm = TRUE), accuracy = accuracy), ")")
  }
  tibble(
    Measure = c("Papers", "Papers citing inside the cluster", "Internal edges per paper",
                "Share of internal edges from the top citing paper",
                "Share of cited works cited by two or more papers inside",
                "Share of papers with a single tie inside",
                "Share of the cluster's citing papers' references that stay inside",
                "Share of cited works also cited from other clusters",
                "Share of edges touching the cluster that stay inside",
                "Neighbouring clusters"),
    `Median (quartiles)` = c(q(s$papers, 1), q(s$citing, 1), q(s$edges_per_paper, 0.01), q(s$top_share, 0.01),
                             q(s$shared_share, 0.01), q(s$one_tie_share, 0.01), q(s$kept, 0.01),
                             q(s$cited_from_outside, 0.01), q(s$cohesion, 0.01), q(s$neighbour_clusters, 1))
  ) |> knitr::kable()
}

gn_shape_largest <- function(key, n = 15) {
  lookup <- gn_label_lookup(key)
  gn_shapes[[key]] |>
    arrange(desc(papers)) |>
    slice_head(n = n) |>
    transmute(Cluster = gn_label_or_id(community_gn, lookup), Papers = papers, `Citing papers` = citing,
              Shape = shape,
              `Top citing paper` = paste0(str_trunc(coalesce(hub_title, ""), 70), " (", hub_year, ")"),
              `Its references inside / total` = paste0(hub_refs_inside, " / ", hub_refs_total)) |>
    knitr::kable()
}

# Clusters whose two largest citing papers cite nearly the same works. Two
# different papers rarely do, and when both records point at the same source
# PDF the second record is probably attached to the wrong file.
gn_twin_table <- function(key, min_overlap = 0.7) {
  ce <- gn_cluster_edges(key)
  refs <- ce |> group_by(source) |> summarise(refs = list(target), .groups = "drop")
  refs <- setNames(refs$refs, refs$source)
  pairs <- ce |>
    filter(from == to) |>
    count(from, source, name = "n") |>
    group_by(from) |> slice_max(n, n = 2, with_ties = FALSE) |> filter(n() == 2) |>
    mutate(rank = row_number()) |> ungroup() |>
    select(from, source, rank) |>
    pivot_wider(names_from = rank, values_from = source, names_prefix = "paper_") |>
    rowwise() |>
    mutate(overlap = length(intersect(refs[[paper_1]], refs[[paper_2]])) /
             length(union(refs[[paper_1]], refs[[paper_2]]))) |>
    ungroup() |>
    filter(overlap >= min_overlap)
  if (nrow(pairs) == 0) return(knitr::kable(tibble(Result = "No cluster has two citing papers with near-identical reference lists.")))
  lookup <- gn_label_lookup(key)
  describe <- function(x) {
    i <- match(x, node_attrs$name)
    paste0(x, ": ", str_trunc(node_attrs$title[i], 60), " (", node_attrs$year[i], ")")
  }
  pdf_of <- function(x) map_chr(x, function(k) bib[[k]][["_source_pdf"]] %||% "")
  pairs |>
    arrange(desc(overlap)) |>
    transmute(Cluster = gn_label_or_id(from, lookup), `Paper A` = describe(paper_1), `Paper B` = describe(paper_2),
              `Reference lists in common` = scales::percent(overlap, accuracy = 1),
              `Same source PDF` = if_else(nzchar(pdf_of(paper_1)) & pdf_of(paper_1) == pdf_of(paper_2), "yes", "no")) |>
    knitr::kable()
}

# Twelve clusters drawn from the inside: single hubs and clusters with shared
# references, spread evenly through the ordering so the gallery shows the range.
gn_shape_gallery <- function(key) {
  s <- gn_shapes[[key]]
  part <- gn[[key]]$partition
  g <- as_undirected(graphs[[key]], mode = "collapse")
  lookup <- gn_label_lookup(key)
  spread <- function(x, k) x[unique(round(seq(1, length(x), length.out = min(k, length(x)))))]
  big <- filter(s, papers >= 50)
  singles <- big |> filter(shape == "Single hub") |> arrange(papers) |> pull(community_gn)
  others <- big |> filter(shape != "Single hub") |> arrange(desc(shared_share), top_share) |> pull(community_gn)
  pick <- c(if (length(singles) > 0) spread(singles, 4), if (length(others) > 0) spread(others, 8))
  set.seed(7)
  parts <- map(pick, function(cid) {
    members <- part$node_id[part$community_gn == cid]
    sg <- igraph::induced_subgraph(g, members)
    xy <- scale(igraph::layout_with_fr(sg, niter = 600))
    r <- filter(s, community_gn == cid)
    panel <- paste0(str_trunc(gn_label_or_id(cid, lookup), 30), " (", r$papers, " papers, ", tolower(r$shape), ")")
    gen <- node_attrs$generation[match(igraph::V(sg)$name, node_attrs$name)]
    el <- igraph::as_edgelist(sg, names = FALSE)
    list(nodes = tibble(x = xy[, 1], y = xy[, 2], ties = igraph::degree(sg), panel = panel,
                        role = if_else(gen %in% c("F1", "P"), "Citing paper", "Cited work")),
         edges = tibble(x = xy[el[, 1], 1], y = xy[el[, 1], 2], xend = xy[el[, 2], 1], yend = xy[el[, 2], 2], panel = panel))
  })
  levels <- map_chr(parts, ~ .x$nodes$panel[1])
  nodes <- bind_rows(map(parts, "nodes")) |> mutate(panel = factor(panel, levels = levels))
  edges <- bind_rows(map(parts, "edges")) |> mutate(panel = factor(panel, levels = levels))
  ggplot() +
    geom_segment(data = edges, aes(x, y, xend = xend, yend = yend), alpha = 0.3, colour = "grey60", linewidth = 0.25) +
    geom_point(data = arrange(nodes, role == "Citing paper"), aes(x, y, size = sqrt(ties), colour = role), alpha = 0.9) +
    scale_size_continuous(range = c(0.4, 4), guide = "none") +
    scale_colour_manual(values = c("Citing paper" = "#d73027", "Cited work" = "#4575b4"), name = NULL) +
    facet_wrap(~ panel, ncol = 4, scales = "free") +
    theme_void() +
    theme(strip.text = element_text(size = 6.5), legend.position = "bottom") +
    labs(title = paste0(graph_labels[[key]], ": twelve clusters seen from the inside"),
         subtitle = "Dot size grows with the number of citation ties inside the cluster.")
}
```

```{r}
#| label: shape-compute
#| echo: false
gn_shapes <- map(c(annotated = "annotated", seed_filtered = "seed_filtered"), function(k) {
  if (!gn[[k]]$available) return(NULL)
  tryCatch(gn_cluster_shapes(k), error = function(e) {
    warning("The cluster shapes for ", k, " failed: ", conditionMessage(e))
    NULL
  })
})
```

The grouping above treats each cluster as one unit. This section looks inside them. A paper counts as citing inside its cluster when it has citation edges to other members of the cluster. Each cluster of `r gn_shape_min_papers` or more papers is described by how its internal edges are spread across citing papers and by how much of its neighbourhood it shares with other clusters. Clusters of one or two papers have no internal structure and are left out.

| Shape | Definition |
|---|---|
| Single hub | One citing paper accounts for at least `r scales::percent(gn_shape_single_hub)` of the cluster's internal edges |
| Shared references | At least `r scales::percent(gn_shape_shared)` of the works cited inside the cluster are cited by two or more of its papers |
| Several hubs | Two or more citing papers, with fewer shared references than that |

::: {.panel-tabset}

#### Annotated

`r gn_shape_note("annotated")`

```{r}
#| label: shape-summary-annotated
#| echo: false
#| eval: !expr gn_has_shapes("annotated")
gn_shape_summary("annotated")
```

```{r}
#| label: shape-largest-annotated
#| echo: false
#| eval: !expr gn_has_shapes("annotated")
gn_shape_largest("annotated")
```

```{r}
#| label: shape-gallery-annotated
#| echo: false
#| fig-width: 12
#| fig-height: 9
#| eval: !expr gn_has_shapes("annotated")
gn_shape_gallery("annotated")
```

```{r}
#| label: shape-twins-annotated
#| echo: false
#| eval: !expr gn_has_shapes("annotated")
gn_twin_table("annotated")
```

#### Annotated, seed filtered

`r gn_shape_note("seed_filtered")`

```{r}
#| label: shape-summary-seed-filtered
#| echo: false
#| eval: !expr gn_has_shapes("seed_filtered")
gn_shape_summary("seed_filtered")
```

```{r}
#| label: shape-largest-seed-filtered
#| echo: false
#| eval: !expr gn_has_shapes("seed_filtered")
gn_shape_largest("seed_filtered")
```

```{r}
#| label: shape-gallery-seed-filtered
#| echo: false
#| fig-width: 12
#| fig-height: 9
#| eval: !expr gn_has_shapes("seed_filtered")
gn_shape_gallery("seed_filtered")
```

```{r}
#| label: shape-twins-seed-filtered
#| echo: false
#| eval: !expr gn_has_shapes("seed_filtered")
gn_twin_table("seed_filtered")
```

:::

The first table summarises every cluster with the median and the quartiles of each measure. The second table names the citing paper behind each of the largest clusters and how much of that paper's reference list stays inside the cluster. In the gallery, red dots are citing papers and blue dots are the works they cite. The last table lists clusters whose two largest citing papers cite nearly the same works. Two different papers with near-identical reference lists, especially when both records carry the same source PDF, point to a bibliography record attached to the wrong file, and the edges from that second record are suspect.

## What the seed filter changes

```{r}
#| label: seed-filter-nodes
#| echo: false
#| results: hide

shared <- inner_join(
  metrics$annotated |> filter(is_ghost == "false") |>
    select(name, title, year, in_degree, betweenness_raw, pagerank, coreness),
  metrics$seed_filtered |> filter(is_ghost == "false") |>
    select(name, in_degree, betweenness_raw, pagerank, coreness),
  by = "name", suffix = c("_a", "_f")
)
n_btw_lost   <- sum(shared$betweenness_raw_a > 0 & shared$betweenness_raw_f == 0)
n_btw_gained <- sum(shared$betweenness_raw_a == 0 & shared$betweenness_raw_f > 0)
n_btw_both   <- sum(shared$betweenness_raw_a > 0 & shared$betweenness_raw_f > 0)

wcc_f <- components(graphs$seed_filtered, mode = "weak")
frag_sizes <- sort(wcc_f$csize, decreasing = TRUE)
frag_text <- if (length(frag_sizes) > 1) paste(frag_sizes[-1], collapse = ", ") else "none"
separation_text <- if (wcc_f$no == 1) {
  paste0("Annotated has ", summ$annotated$wcc, " weakly connected component(s) and Annotated, seed filtered has 1, ",
         "so the seed's citations do not hold any fragment to the main body.")
} else {
  paste0("Annotated has ", summ$annotated$wcc, " weakly connected component(s) and Annotated, seed filtered has ",
         wcc_f$no, ": a main body of ", format(frag_sizes[1], big.mark = ","), " works and ",
         length(frag_sizes) - 1, " smaller fragments of ", frag_text, " works.",
         if (summ$annotated$wcc == 1) " In Annotated those fragments join the main body only through the seed's citations." else "")
}
fragments <- tibble(name = names(wcc_f$membership), component = wcc_f$membership) |>
  mutate(component_size = wcc_f$csize[component]) |>
  filter(component != which.max(wcc_f$csize)) |>
  left_join(node_attrs |> select(name, title, year), by = "name") |>
  arrange(desc(component_size), component, year)
```

### What separates

`r separation_text`

::: {.callout-note collapse="true"}
## The fragments

```{r}
#| label: seed-filter-fragments
#| echo: false
if (nrow(fragments) == 0) {
  cat("No fragments.")
} else {
  fragments |>
    transmute(`Fragment size` = component_size, Year = year, Title = title) |>
    knitr::kable()
}
```
:::

### Centrality

The comparison covers the works present in both graphs, with ghost entries excluded. A Spearman correlation near 1 and a large top-20 overlap mean the seed's citations barely change a measure's ranking. Betweenness is zero for most works. The seed filter takes nonzero betweenness away from `r n_btw_lost` works that had it in Annotated and gives it to `r n_btw_gained`. The second table lists the ten works whose betweenness rank moves the most among the `r n_btw_both` that have nonzero betweenness in both graphs.

```{r}
#| label: seed-filter-centrality
#| echo: false
measure_agreement <- function(col, label) {
  a <- shared[[paste0(col, "_a")]]
  f <- shared[[paste0(col, "_f")]]
  tibble(
    Measure = label,
    `Spearman correlation` = round(cor(a, f, method = "spearman"), 3),
    `Top 20 in both` = length(intersect(shared$name[order(-a)][1:20], shared$name[order(-f)][1:20]))
  )
}

bind_rows(
  measure_agreement("in_degree", "In-degree"),
  measure_agreement("betweenness_raw", "Betweenness"),
  measure_agreement("pagerank", "PageRank"),
  measure_agreement("coreness", "Coreness")
) |>
  knitr::kable()
```

```{r}
#| label: seed-filter-movers
#| echo: false
shared |>
  filter(betweenness_raw_a > 0, betweenness_raw_f > 0) |>
  mutate(
    rank_a = rank(-betweenness_raw_a, ties.method = "min"),
    rank_f = rank(-betweenness_raw_f, ties.method = "min"),
    change = rank_a - rank_f
  ) |>
  arrange(desc(abs(change))) |>
  head(10) |>
  transmute(
    Work = name, Year = year, Title = title,
    `Rank, Annotated` = rank_a,
    `Rank, seed filtered` = rank_f,
    `Places gained` = change
  ) |>
  knitr::kable()
```

### Communities

Partition agreement is measured on the nodes both graphs share. The adjusted Rand index and normalized mutual information both measure how much two partitions agree, from 0 for no more agreement than chance to 1 for identical groupings. Louvain and Leiden are rerun on each graph, and the Girvan-Newman partitions are the earliest fragmentation-onset cuts of the two runs. The last two rows rerun each algorithm on the Annotated graph with a different random seed. They show how much of the disagreement comes from the algorithm's own randomness and not from the seed filter.

```{r}
#| label: seed-filter-partitions
#| echo: false
gn_both <- gn$annotated$available && gn$seed_filtered$available

agree_row <- function(method, x, y) {
  d <- inner_join(x, y, by = "name")
  a <- as.integer(factor(d$x))
  b <- as.integer(factor(d$y))
  tibble(
    Method = method,
    `Adjusted Rand index` = round(igraph::compare(a, b, method = "adjusted.rand"), 3),
    `Normalized mutual information` = round(igraph::compare(a, b, method = "nmi"), 3),
    `Nodes compared` = format(nrow(d), big.mark = ",")
  )
}

rerun_agreement <- function(method, cluster_fn) {
  und <- as_undirected(graphs$annotated, mode = "collapse")
  set.seed(1); a <- as.integer(membership(cluster_fn(und)))
  set.seed(2); b <- as.integer(membership(cluster_fn(und)))
  agree_row(paste0(method, ", Annotated against its own rerun"),
            tibble(name = V(und)$name, x = a), tibble(name = V(und)$name, y = b))
}

bind_rows(
  if (gn_both) agree_row("Girvan-Newman, earliest cut",
                         gn$annotated$partition |> transmute(name = node_id, x = community_gn),
                         gn$seed_filtered$partition |> transmute(name = node_id, y = community_gn)),
  agree_row("Louvain", metrics$annotated |> transmute(name, x = community_louvain),
                       metrics$seed_filtered |> transmute(name, y = community_louvain)),
  agree_row("Leiden", metrics$annotated |> transmute(name, x = community_leiden),
                      metrics$seed_filtered |> transmute(name, y = community_leiden)),
  rerun_agreement("Louvain", cluster_louvain),
  rerun_agreement("Leiden", function(g) cluster_leiden(g, objective_function = "modularity"))
) |>
  knitr::kable()
```

```{r}
#| label: gn-trace-overlay
#| echo: false
#| fig-width: 8
#| fig-height: 5
#| eval: !expr gn$annotated$available && gn$seed_filtered$available && !is.null(gn$annotated$trace) && !is.null(gn$seed_filtered$trace)
trace_both <- bind_rows(
  gn$annotated$trace |> mutate(run = graph_labels[["annotated"]]),
  gn$seed_filtered$trace |> mutate(run = graph_labels[["seed_filtered"]])
)
onsets <- tibble(
  run = unname(graph_labels[c("annotated", "seed_filtered")]),
  round = c(gn$annotated$earliest$round, gn$seed_filtered$earliest$round)
)
ggplot(trace_both, aes(x = round, y = n_components, colour = run)) +
  geom_line() +
  geom_vline(data = onsets, aes(xintercept = round, colour = run),
             linetype = "dashed", linewidth = 0.4, show.legend = FALSE) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",")) +
  scale_colour_manual(values = setNames(c("steelblue", "firebrick"), unname(graph_labels[c("annotated", "seed_filtered")])),
                      name = NULL) +
  labs(
    title = "Connected components across both Girvan-Newman runs",
    subtitle = "Dashed lines mark each run's earliest fragmentation onset",
    x = "Round", y = "Connected components"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

Each of the largest Annotated clusters, with the seed-filtered cluster that shares the most nodes with it. The share is the fraction of the Annotated cluster that lands in that match, and the Jaccard index is the overlap divided by the union of the two clusters.

```{r}
#| label: seed-filter-matches
#| echo: false
#| eval: !expr gn$annotated$available && gn$seed_filtered$available
gn_shared <- inner_join(
  gn$annotated$partition |> rename(community_a = community_gn),
  gn$seed_filtered$partition |> rename(community_f = community_gn),
  by = "node_id"
)
gn_overlap <- gn_shared |>
  count(community_a, community_f, name = "n_both") |>
  group_by(community_a) |> mutate(size_a = sum(n_both)) |> ungroup() |>
  group_by(community_f) |> mutate(size_f = sum(n_both)) |> ungroup() |>
  mutate(jaccard = n_both / (size_a + size_f - n_both))
gn_best_match <- gn_overlap |>
  group_by(community_a) |>
  slice_max(n_both, n = 1, with_ties = FALSE) |>
  ungroup()
gn_switched <- gn_shared |>
  left_join(gn_best_match |> select(community_a, best_f = community_f), by = "community_a") |>
  summarise(n = sum(community_f != best_f), share = mean(community_f != best_f))

lookup_a <- gn_label_lookup("annotated")
lookup_f <- gn_label_lookup("seed_filtered")
gn$annotated$partition |>
  count(community_gn, name = "size", sort = TRUE) |>
  slice_head(n = gn_top_n) |>
  left_join(gn_best_match, by = c("community_gn" = "community_a")) |>
  transmute(
    `Cluster (Annotated)` = gn_label_or_id(community_gn, lookup_a),
    Size = size,
    `Best match (seed filtered)` = gn_label_or_id(community_f, lookup_f),
    `Shared nodes` = n_both,
    `Share of cluster` = scales::percent(n_both / size, accuracy = 1),
    Jaccard = round(jaccard, 2)
  ) |>
  knitr::kable()
```

`r if (gn$annotated$available && gn$seed_filtered$available) paste0(format(gn_switched$n, big.mark = ","), " of the ", format(nrow(gn_shared), big.mark = ","), " nodes in both partitions (", scales::percent(gn_switched$share, accuracy = 0.1), ") sit outside the best-matching cluster of their Annotated cluster.") else ""`

## Appendix: layout trials

The partition plots above use the Fruchterman-Reingold layout (`fr`, 500 iterations, with igraph's grid speed-up). The layouts below were tried and are kept for reference with `eval: false`. Each calls the same plotting function as the main plots, so any of them can be switched on one at a time and run on either graph.

- `fr` spreads the papers that cite a hub around it, so each community reads as a region of the plot. It runs in seconds on either graph.
- `lgl` with `coolexp` lowered to 0.5, the layout used before, stretches the graph into one diagonal band and leaves the highlighted communities overlapping through its middle. It is also far slower than `fr` on the seed-filtered graph.
- `drl` stacks the papers that cite a hub on top of it, so every community collapses into a single dot and the plot shows nothing of its shape.
- `stress` (`graphlayouts::layout_with_stress()`) is faster than `drl` and built to preserve graph distance. It produced a concentric-ring pattern and was rejected.
- Untuned `lgl` places a handful of true outliers far from the rest. That is why the main plots crop the view to the 1st to 99th percentile.
- `lgl` with `repulserad` doubled spread the periphery out and left the inner core dense. `repulserad` sets the distance at which repulsion stops acting and does not change its strength, so it is the wrong lever for core density.

```{r}
#| label: layout-trial-drl
#| eval: false
#| fig-width: 16
#| fig-height: 14
gn_plot_partition("annotated", layout = "drl", layout_args = list(), crop = FALSE)
```

```{r}
#| label: layout-trial-stress
#| eval: false
#| fig-width: 16
#| fig-height: 14
library(graphlayouts)
gn_plot_partition("annotated", layout = "stress", layout_args = list(), crop = FALSE)
```

```{r}
#| label: layout-trial-lgl-default
#| eval: false
#| fig-width: 16
#| fig-height: 14
gn_plot_partition("annotated", layout = "lgl", layout_args = list(), crop = TRUE)
```

```{r}
#| label: layout-trial-lgl-repulserad
#| eval: false
#| fig-width: 16
#| fig-height: 14
lgl_area <- vcount(graphs$annotated)^2
gn_plot_partition("annotated", layout = "lgl",
                  layout_args = list(area = lgl_area, repulserad = 2 * lgl_area * vcount(graphs$annotated)),
                  crop = TRUE)
```

```{r}
#| label: layout-trial-lgl-coolexp-annotated
#| eval: false
#| fig-width: 16
#| fig-height: 14
gn_plot_partition("annotated", layout = "lgl", layout_args = list(coolexp = 0.5), crop = TRUE)
```

```{r}
#| label: layout-trial-lgl-coolexp-seed-filtered
#| eval: false
#| fig-width: 16
#| fig-height: 14
gn_plot_partition("seed_filtered", layout = "lgl", layout_args = list(coolexp = 0.5), crop = TRUE)
```

The Louvain, Leiden and K-means plots, also disabled. Each saves a PNG and a PDF next to this document.

```{r}
#| label: membership-viz
#| eval: false
#| fig-width: 12
#| fig-height: 12
#| cache: true
plot_membership <- function(key, column, title, file) {
  tg <- as_tbl_graph(graphs[[key]]) |>
    activate(nodes) |>
    left_join(metrics[[key]] |> select(name, value = all_of(column)), by = "name") |>
    mutate(group = as.factor(value))
  p <- ggraph(tg, layout = "drl") +
    geom_edge_link(alpha = 0.05, colour = "grey60") +
    geom_node_point(aes(colour = group), size = 0.3, alpha = 0.7) +
    scale_colour_viridis_d(guide = "none") +
    theme_graph() +
    labs(title = title)
  ggsave(here("analysis", paste0(file, ".png")), p, width = 16, height = 16, dpi = 300)
  ggsave(here("analysis", paste0(file, ".pdf")), p, width = 16, height = 16, device = cairo_pdf)
  p
}
plot_membership("annotated", "community_louvain", "Citation graph, Louvain communities", "bibvik_louvain")
plot_membership("annotated", "community_leiden", "Citation graph, Leiden communities", "bibvik_leiden")
plot_membership("annotated", "cluster_kmeans", "Citation graph, K-means clusters (by centrality profile)", "bibvik_kmeans")
```

## Node-table export

`analysis/bibvik_node_table.csv`, which `03_joined_descriptions.qmd` reads, holds the Annotated graph's measures with one row per node, plus `community_gn` from the Annotated run's earliest cut and `community_gn_seed_filtered` from the seed-filtered run. `group_gn` and `group_gn_seed_filtered` hold the consensus group of each paper's cluster. The Girvan-Newman columns are left out when a run has not been installed.

```{r}
#| label: node-table-export
#| echo: false
#| results: hide
node_table <- metrics$annotated |>
  select(name, betweenness, betweenness_raw, closeness, pagerank, hub, authority,
         in_degree, out_degree, degree,
         title, year, generation, entry_type, resolution, completeness, author, is_ghost,
         coreness, cluster_kmeans, component_weak, component_strong,
         community_louvain, community_leiden)

gn_column <- function(key, column) {
  gn[[key]]$partition |>
    transmute(name = node_id, "{column}" := as.integer(community_gn))
}
if (gn$annotated$available) {
  node_table <- node_table |> left_join(gn_column("annotated", "community_gn"), by = "name")
  cat("Joined Girvan-Newman communities from", gn$annotated$cut_file,
      "(fragmentation-onset round", gn$annotated$earliest$round, ")\n")
} else {
  cat("No Girvan-Newman run found at", gn_path("annotated"),
      "- run the pipeline first (see analysis/README.md). Skipping join.\n")
}
if (gn$seed_filtered$available) {
  node_table <- node_table |>
    left_join(gn_column("seed_filtered", "community_gn_seed_filtered"), by = "name")
}

gn_group_column <- function(key, column) {
  if (is.null(gn_groups[[key]])) return(NULL)
  gn[[key]]$partition |>
    left_join(select(gn_groups[[key]]$clusters, community_gn, group), by = "community_gn") |>
    transmute(name = node_id, "{column}" := group)
}
for (spec in list(c("annotated", "group_gn"), c("seed_filtered", "group_gn_seed_filtered"))) {
  group_column <- gn_group_column(spec[1], spec[2])
  if (!is.null(group_column)) node_table <- node_table |> left_join(group_column, by = "name")
}

write_csv(node_table, here("analysis", "bibvik_node_table.csv"))
cat("Written bibvik_node_table.csv\n")
cat("Rows:", nrow(node_table), "\n")
```

## References

::: {#refs}
:::
