---
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

# Part 3 — Joining structural and annotation descriptions

Everything in Part 1 describes the annotation data on its own terms; everything in Part 2 describes the graph on its own terms. Neither is a baseline the other is checked against. This section places them side by side and reads the relationship, which is where the actual analytical content of this project lives.

```{r}
#| label: setup
# Loaded here (not assumed already in scope) so this file can render
# standalone, reading 01_annotation.qmd and 02_network_structure.qmd's
# saved outputs rather than requiring the combined index.qmd include
# chain. If rendered as part of that chain instead, df and
# centrality_df/node_table are already in scope from the included
# upstream parts, and these loads are redundant but harmless (they
# will just re-read the same values from disk).
library(here)
library(tidyverse)
if (!requireNamespace("ggrepel", quietly = TRUE)) {
  install.packages("ggrepel", repos = "https://cloud.r-project.org")
}
library(ggrepel)

# Standard Wilson score interval on a proportion, base R only (matches
# the method underlying stats::prop.test, computed directly here for
# clarity). Used throughout this document as the single, consistent
# criterion for "is a cluster's coverage of a category meaningfully
# above the corpus baseline" — the lower bound of the interval must
# exceed the corpus rate. Replaces earlier, separately-chosen fixed
# floors (an arbitrary minimum count and enrichment ratio) that were
# used in different places in this document without being reconciled
# with each other; one method, defined once, used everywhere below.
wilson_lower_bound <- function(x, n, conf = 0.95) {
  z <- qnorm(1 - (1 - conf) / 2)
  phat <- x / n
  denom <- 1 + z^2 / n
  centre <- phat + z^2 / (2 * n)
  adj <- z * sqrt(phat * (1 - phat) / n + z^2 / (4 * n^2))
  (centre - adj) / denom
}
```

```{r}
#| label: join
# df (Part 1's annotation dataframe, from 01_annotation.qmd) is keyed by
# cite_key; centrality_df / node_table (Part 2's structural data, from
# 02_network_structure.qmd) is keyed by name (citekey). F1-only:
# annotations exist for F1 papers only. F2 works cited by multiple F1
# papers can still be described using the annotation profile of their
# *citer set*, even though the F2 works themselves are not annotated.
# See "per-F2-node citer profiles" below.

if (!exists("df", inherits = FALSE)) {
  df <- read_csv(here("analysis", "bibvik_annotation_df.csv"), show_col_types = FALSE)
  cat("Loaded df from disk:", nrow(df), "rows\n")
} else {
  cat("df already in scope:", nrow(df), "rows\n")
}
if (!exists("node_table", inherits = FALSE)) {
  node_table <- read_csv(here("analysis", "bibvik_node_table.csv"), show_col_types = FALSE)
  cat("Loaded node_table from disk:", nrow(node_table), "rows\n")
} else {
  cat("node_table already in scope:", nrow(node_table), "rows\n")
}
# node_table, not centrality_df, is the object that actually carries
# community_gn: in 02_network_structure.qmd, node_table is built FROM
# centrality_df with the community assignment joined on afterward
# (node_table <- centrality_df |> left_join(...)), so centrality_df
# itself never has community_gn even when it's in scope. Preferring
# centrality_df here (an earlier version of this join checked for it
# first) silently grabbed the wrong, earlier object in the combined
# index.qmd render, where centrality_df exists but lacks the column
# this file needs — always use node_table.
structural_df <- node_table
cat("structural_df:", nrow(structural_df), "rows. is_ghost values:",
    paste(unique(structural_df$is_ghost), collapse = ", "),
    "(class:", class(structural_df$is_ghost), ")\n")

# is_ghost is written as an R logical (TRUE/FALSE) by 02_network_structure.qmd
# internally, but a round-trip through write_csv()/read_csv() serializes
# logicals as the bare words TRUE/FALSE, not the quoted string "false" the
# original in-memory comparison assumed. Compare against both to be safe
# regardless of whether structural_df came from disk (logical after
# read_csv's type inference) or was passed in memory as character "false".
cluster_table_annotated <- structural_df |>
  filter(is_ghost == FALSE | is_ghost == "false") |>
  left_join(df, by = c("name" = "cite_key"))

cat("Joined cluster_table_annotated:", nrow(cluster_table_annotated), "rows\n")
cat("Rows with a matching annotation:", sum(!is.na(cluster_table_annotated$publication_year)), "\n")
```

```{r}
#| label: graph-definitions
#| echo: false
# The same two versions of the graph as in Part 2. node_table carries one
# Girvan-Newman column for each: community_gn (Annotated) and
# community_gn_seed_filtered (Annotated, seed filtered).
graph_labels <- c(annotated = "Annotated", seed_filtered = "Annotated, seed filtered")
gn_columns <- c(annotated = "community_gn", seed_filtered = "community_gn_seed_filtered")
has_gn <- function(key) {
  col <- gn_columns[[key]]
  col %in% names(node_table) && any(!is.na(node_table[[col]]))
}
```


## Cluster structure vs. topics, methods, and sources of evidence

```{r}
#| label: cluster-functions
#| echo: false
# Everything below that is computed per graph goes through these functions,
# called once per tab, so the two tabs of a section cannot drift apart.

# The annotated works with a cluster in this graph, and the true size of every cluster.
cluster_context <- function(key) {
  col <- gn_columns[[key]]
  assigned <- cluster_table_annotated |>
    mutate(community_gn = as.character(.data[[col]])) |>
    filter(!is.na(community_gn))
  list(
    annotated_only = filter(assigned, !is.na(publication_year)),
    true_sizes = node_table |>
      filter(!is.na(.data[[col]])) |>
      count(community_gn = as.character(.data[[col]]), name = "true_size")
  )
}

# Coverage and enrichment of one category family (topic, method or source of
# evidence) in the 20 clusters with the most annotated works.
category_data <- function(ctx, pattern, kind) {
  long <- ctx$annotated_only |>
    select(name, community_gn, contains(pattern)) |>
    pivot_longer(contains(pattern), names_to = "category", values_to = "value") |>
    filter(!is.na(value)) |>
    mutate(category = str_replace_all(str_remove(category, pattern), fixed("."), " "))
  annotated_n <- long |> distinct(community_gn, name) |> count(community_gn, name = "n_annotated")
  keep <- annotated_n |> slice_max(n_annotated, n = 20) |> pull(community_gn)
  labels <- annotated_n |>
    filter(community_gn %in% keep) |>
    left_join(ctx$true_sizes, by = "community_gn") |>
    arrange(desc(true_size)) |>
    mutate(label = sprintf("Cluster %s (%d of %d annotated)", community_gn, n_annotated, true_size)) |>
    mutate(label = fct_rev(fct_inorder(label))) |>
    select(community_gn, label) |>
    deframe()
  # Corpus-wide base rate per category: what fraction of ALL annotated works
  # (not just the 20 shown clusters) carry it. Dividing a cluster's coverage by
  # this rate gives enrichment, the signal of thematic distinctiveness.
  corpus_rate <- long |>
    distinct(name, category) |>
    count(category, name = "n_corpus") |>
    mutate(corpus_rate = n_corpus / n_distinct(long$name))
  enrichment <- long |>
    filter(community_gn %in% keep) |>
    distinct(community_gn, name, category) |>
    count(community_gn, category) |>
    left_join(annotated_n, by = "community_gn") |>
    left_join(corpus_rate, by = "category") |>
    mutate(prop = n / n_annotated, enrichment = prop / corpus_rate,
           ci_lower = wilson_lower_bound(n, n_annotated), category_type = kind)
  list(kind = kind, long = long, annotated_n = annotated_n, keep = keep, labels = labels,
       corpus_rate = corpus_rate, enrichment = enrichment)
}

cluster_data <- map(set_names(names(gn_columns)), function(key) {
  if (!has_gn(key)) return(NULL)
  ctx <- cluster_context(key)
  ctx$topic <- category_data(ctx, "Topics..", "topic")
  ctx$method <- category_data(ctx, "Methods..", "method")
  ctx$source <- category_data(ctx, "Sources.of.Evidence..", "source")
  ctx$n_matched <- nrow(ctx$annotated_only)
  sizes <- arrange(ctx$true_sizes, desc(true_size))
  ctx$n_clusters <- nrow(sizes)
  ctx$size_drop <- sizes$true_size[1] - sizes$true_size[2]
  ctx$n_pass_3plus <- ctx$topic$long |> distinct(community_gn, name) |> count(community_gn) |> filter(n >= 3) |> nrow()
  ctx
})

category_note <- function(key, kind) {
  d <- cluster_data[[key]]
  cat("Works with a cluster assignment and any ", kind, " coded: ", n_distinct(d[[kind]]$long$name), "\n\n", sep = "")
}

category_heatmap <- function(key, kind) {
  d <- cluster_data[[key]][[kind]]
  family <- c(topic = "Topic", method = "Method", source = "Source-of-evidence")[[kind]]
  tail_text <- c(
    topic = paste0("a cluster at 1x carries the topic at exactly the corpus average; ubiquitous topics (e.g. Environment) will not dominate every row here the way they do on raw coverage"),
    method = "ubiquitous methods (e.g. Spatial analysis) will not dominate every row here the way they do on raw coverage",
    source = "ubiquitous sources (e.g. Written sources, Graves and monuments) will not dominate every row here the way they do on raw coverage")[[kind]]
  d$enrichment |>
    mutate(cluster_label = d$labels[as.character(community_gn)]) |>
    ggplot(aes(x = category, y = cluster_label, fill = enrichment)) +
    geom_tile(colour = "white") +
    geom_text(aes(label = n), size = 3, colour = "grey20") +
    # Diverging scale centred on 1.0 (cluster rate == corpus rate): blue means
    # this cluster carries the category more than the corpus baseline
    # (distinctive), red means less. Capped at 3x so a handful of extreme
    # small-n ratios don't wash out the rest of the scale.
    scale_fill_gradient2(low = "firebrick", mid = "white", high = "steelblue",
                         midpoint = 1, limits = c(0, 3), oob = scales::squish,
                         labels = scales::label_number(suffix = "x")) +
    labs(x = "", y = "", fill = "Enrichment vs.\ncorpus base rate",
         title = paste0(family, " enrichment within each Girvan-Newman cluster, ", graph_labels[[key]]),
         subtitle = paste0("Cell numbers are raw counts. Fill shows coverage relative to the corpus-wide rate for that ",
                           kind, ", not raw coverage — ", tail_text)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# A peak requires the Wilson lower bound (defined in setup) on a cluster's
# coverage rate to exceed the corpus baseline rate for that category.
cluster_peaks <- function(key) {
  d <- cluster_data[[key]]
  bind_rows(d$topic$enrichment, d$method$enrichment, d$source$enrichment) |>
    filter(ci_lower > corpus_rate)
}

cluster_signatures <- function(key) {
  cluster_peaks(key) |>
    group_by(community_gn) |>
    summarise(
      n_peaks = n(),
      mean_peak_enrichment = mean(enrichment),
      peak_categories = paste0(category_type, ": ", category, " (", round(enrichment, 1), "x)", collapse = "; "),
      .groups = "drop"
    ) |>
    arrange(desc(n_peaks), desc(mean_peak_enrichment))
}

# No composite "score" here. Peak count x mean enrichment is not a real unit and
# a single number invites over-interpretation, so the two are plotted directly.
signature_plot <- function(key) {
  cat("Peaks found (Wilson lower bound exceeds corpus baseline): ", nrow(cluster_peaks(key)), "\n\n", sep = "")
  cluster_signatures(key) |>
    ggplot(aes(x = n_peaks, y = mean_peak_enrichment, label = community_gn)) +
    geom_point(size = 3, colour = "steelblue") +
    geom_text_repel(size = 3.5, seed = 1) +
    scale_x_continuous(breaks = scales::breaks_pretty()) +
    labs(x = "Number of peaks (categories where coverage exceeds\ncorpus baseline, Wilson 95% lower bound)",
         y = "Mean enrichment among those peaks",
         title = paste0("Clusters by peak count and peak strength, ", graph_labels[[key]]),
         subtitle = "Upper right: several strong, corroborating signals. Lower right: one very strong but isolated signal.") +
    theme_minimal()
}

signature_detail <- function(key) {
  top_communities <- cluster_signatures(key) |> slice_max(n_peaks, n = 8, with_ties = FALSE) |> pull(community_gn)
  cluster_peaks(key) |>
    filter(community_gn %in% top_communities) |>
    mutate(
      cluster_label = fct_relevel(paste("Cluster", community_gn), paste("Cluster", top_communities)),
      category_label = fct_reorder(paste0(category, " (", category_type, ")"), enrichment)
    ) |>
    ggplot(aes(x = enrichment, y = category_label)) +
    geom_col(fill = "steelblue") +
    geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
    facet_wrap(vars(cluster_label), scales = "free_y", ncol = 2) +
    labs(x = "Enrichment vs. corpus base rate", y = "",
         title = paste0("Peak categories for the clusters with the most peaks, ", graph_labels[[key]]),
         subtitle = "Dashed line = corpus baseline (1x). Bars are that cluster's individual peaks.") +
    theme_minimal() +
    theme(strip.text = element_text(face = "bold"))
}
```


`community_gn` (Girvan-Newman, fragmentation-onset cut) is checked here
against the topics, methods, and sources of evidence coded in Part 1,
for both versions of the graph. Annotated uses the cut of the Annotated
run and Annotated, seed filtered uses the cut of the seed-filtered run.
The comparison is restricted to works with a matching annotation
(F1-only; see the join above), which is
`r if (has_gn("annotated")) cluster_data$annotated$n_matched else "n/a"`
works in Annotated and
`r if (has_gn("seed_filtered")) cluster_data$seed_filtered$n_matched else "n/a"`
in Annotated, seed filtered. Louvain and Leiden cluster assignments also
exist from Part 2 but are not used in this section.

Coverage is thin relative to the graph as a whole: in Annotated,
`r if (has_gn("annotated")) cluster_data$annotated$n_matched else "n/a"` annotated
works out of `r nrow(node_table)` nodes, and within any single cluster
that annotated count is often a small fraction of the cluster's real
size (the "other" grouping below hides clusters with fewer than 3
annotated works, but even the clusters shown may have only a handful of
annotated works out of a much larger true membership). A prior
exploratory pass on an earlier GN cut of this same corpus found 7-13
annotated nodes per cluster typical, out of hundreds of members, and
treated agreement between structure and annotation as directionally
suggestive rather than comprehensively verified for that reason. The
same caution applies here: what follows can show that structure and
coding line up somewhere, but cannot rule out that the uncoded majority
of any cluster tells a different story.

The coded columns are wide, one column per topic/method/source
category, valued `"Primary"`, `"Secondary"`, or blank. Primary and
secondary are merged throughout what follows: a work counts as coded
with a category at either level, matching the method used in a prior
exploratory pass on an earlier GN cut of this corpus. A work can carry
more than one topic, method, or source, so per-cluster percentages
below do not sum to 100%; each cell shows the raw count behind its
percentage for transparency.

**How to read the enrichment heatmaps below.** Raw coverage (what
fraction of a cluster's annotated works carry a given category) is
misleading on its own: some categories are common across the whole
corpus regardless of cluster — "Spatial analysis" as a method, or
"Environment" as a topic, show up in most clusters simply because
most works in this corpus use them, not because any one cluster is
distinctively characterized by them. To separate genuine
distinctiveness from corpus-wide popularity, each heatmap cell instead
shows **enrichment**: a cluster's coverage of a category divided by
that category's rate across the whole annotated corpus. A value of 1x
means the cluster carries the category at exactly the corpus
average; 2x means twice the corpus rate; 0.5x means half. White is 1x
(baseline); blue is above baseline (the category is over-represented in
that cluster, a real distinguishing signal); red is below baseline
(under-represented). For example, if "Palaeoenvironment reconstruction"
is coded in 23% of the whole corpus but 100% of a given cluster's
annotated works, that cell shows roughly 4.4x — a real signal that
environmental method is disproportionately concentrated there, not
just present because it's common everywhere.

### By topic

**Girvan-Newman clusters by topic.** A threshold of 3+ annotated
works lets through only a small share of the clusters in this corpus (see
the counts below), so it does not give a workable set. Shown here
instead, in the heatmap below: the 20 clusters with the most annotated works. This is
a readability choice, not a derived breakpoint.

At this cut, Annotated has
`r if (has_gn("annotated")) format(cluster_data$annotated$n_clusters, big.mark = ",") else "n/a"` clusters,
`r if (has_gn("annotated")) cluster_data$annotated$n_pass_3plus else "n/a"` of which pass a
3+ annotated-works threshold, and its largest cluster is
`r if (has_gn("annotated")) cluster_data$annotated$size_drop else "n/a"` members larger than
the second-largest. Annotated, seed filtered has
`r if (has_gn("seed_filtered")) format(cluster_data$seed_filtered$n_clusters, big.mark = ",") else "n/a"` clusters,
`r if (has_gn("seed_filtered")) cluster_data$seed_filtered$n_pass_3plus else "n/a"` of which pass
the threshold, and its largest cluster is
`r if (has_gn("seed_filtered")) cluster_data$seed_filtered$size_drop else "n/a"` members larger than
the second-largest.

Coverage varies a lot by cluster (annotation is thin relative to true
cluster size, per the caveat above), so each row label shows how many
annotated works its percentages are based on, out of the cluster's
true total size.

::: {.panel-tabset}

#### Annotated

```{r}
#| label: cluster-vs-topic-annotated
#| echo: false
#| fig-width: 10
#| fig-height: 7
#| results: asis
#| eval: !expr has_gn("annotated")
category_note("annotated", "topic")
category_heatmap("annotated", "topic")
```

#### Annotated, seed filtered

```{r}
#| label: cluster-vs-topic-seed-filtered
#| echo: false
#| fig-width: 10
#| fig-height: 7
#| results: asis
#| eval: !expr has_gn("seed_filtered")
category_note("seed_filtered", "topic")
category_heatmap("seed_filtered", "topic")
```

:::

### By method

Same approach as the topic comparison above, applied to method.

::: {.panel-tabset}

#### Annotated

```{r}
#| label: cluster-vs-method-annotated
#| echo: false
#| fig-width: 10
#| fig-height: 7
#| results: asis
#| eval: !expr has_gn("annotated")
category_note("annotated", "method")
category_heatmap("annotated", "method")
```

#### Annotated, seed filtered

```{r}
#| label: cluster-vs-method-seed-filtered
#| echo: false
#| fig-width: 10
#| fig-height: 7
#| results: asis
#| eval: !expr has_gn("seed_filtered")
category_note("seed_filtered", "method")
category_heatmap("seed_filtered", "method")
```

:::

### By source of evidence

Same approach again, applied to source of evidence.

::: {.panel-tabset}

#### Annotated

```{r}
#| label: cluster-vs-source-annotated
#| echo: false
#| fig-width: 10
#| fig-height: 7
#| results: asis
#| eval: !expr has_gn("annotated")
category_note("annotated", "source")
category_heatmap("annotated", "source")
```

#### Annotated, seed filtered

```{r}
#| label: cluster-vs-source-seed-filtered
#| echo: false
#| fig-width: 10
#| fig-height: 7
#| results: asis
#| eval: !expr has_gn("seed_filtered")
category_note("seed_filtered", "source")
category_heatmap("seed_filtered", "source")
```

:::

## Cluster signatures: peak count and peak strength

Throughout this section, "cluster" means a Girvan-Newman graph
cluster (`community_gn`), not a scholarly or social cluster.

The topic, method, and source enrichment findings above surface
individual standout cells one at a time. This section combines them
into two directly interpretable numbers per cluster, so the
clusters most worth a closer look surface first instead of
requiring a manual scan across three separate tables.

A **peak** is any category (topic, method, or source) whose 95% Wilson
confidence interval on a cluster's coverage rate lies entirely above
the corpus-wide rate for that category — meaning the cluster's
coverage is higher than baseline with reasonable statistical
confidence, not just a higher raw count. This adapts to sample size
automatically: a cluster with only 3 annotated works needs a much
larger raw enrichment to register as a peak than a cluster with 13,
because a small sample's interval is wider. This is the same criterion
used in the topic, method, and source findings above (`ci_lower >
corpus_rate`); "peak" here just names a category that already met that
bar, so this section's peak count and mean enrichment are computed
directly from the same standouts listed there, not a separate pass.

For each cluster, two numbers: how many peaks it has, and how strong
those peaks are on average. Neither is combined into a single score —
peak count times mean enrichment is not a real, interpretable unit, and
collapsing two different kinds of signal into one number invites
over-reading it. Plotted directly instead: clusters with many peaks
AND high average enrichment are worth reading first; a cluster with
one very large but isolated peak is a different kind of finding (a
strong single signal) than one with several moderate peaks (a broader,
possibly more coherent pattern) — the plot keeps that distinction
visible rather than flattening it. Whether a cluster's peak
categories actually cohere thematically is a judgment that requires
reading them against domain knowledge of this corpus, which this plot
does not attempt.

### Peak count against peak strength

::: {.panel-tabset}

#### Annotated

```{r}
#| label: cluster-signature-annotated
#| echo: false
#| results: asis
#| eval: !expr has_gn("annotated")
signature_plot("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: cluster-signature-seed-filtered
#| echo: false
#| results: asis
#| eval: !expr has_gn("seed_filtered")
signature_plot("seed_filtered")
```

:::

### Peaks behind the clusters with the most peaks

The peak categories behind the clusters with the most peaks, shown
directly rather than as a summary number:

::: {.panel-tabset}

#### Annotated

```{r}
#| label: cluster-signature-detail-annotated
#| echo: false
#| fig-width: 9
#| fig-height: 7
#| eval: !expr has_gn("annotated")
signature_detail("annotated")
```

#### Annotated, seed filtered

```{r}
#| label: cluster-signature-detail-seed-filtered
#| echo: false
#| fig-width: 9
#| fig-height: 7
#| eval: !expr has_gn("seed_filtered")
signature_detail("seed_filtered")
```

:::
