---
title: "BibVik — Part 3: Joining structural and annotation descriptions"
format:
  html:
    toc: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: true
---

# Part 3 — Joining structural and annotation descriptions

Strategy and reasoning for this analysis: see `00_plan.qmd`. Reads outputs
of `01_annotation.qmd` and `02_network_structure.qmd`. Does not recompute
either.

Everything in Part 1 describes the annotation data on its own terms;
everything in Part 2 describes the graph on its own terms. Neither is a
baseline the other is checked against. This section places them side by
side and reads the relationship, which is where the actual analytical
content of this project lives.

```{r}
#| label: setup
# Loaded here (not assumed already in scope) so this file can render
# standalone, reading 01_annotation.qmd and 02_network_structure.qmd's
# saved outputs rather than requiring the combined index.qmd include
# chain. If rendered as part of that chain instead, df and
# centrality_df/node_table are already in scope from the included
# upstream parts, and these loads are redundant but harmless (they
# will just re-read the same values from disk).
library(here)
library(tidyverse)
if (!requireNamespace("ggrepel", quietly = TRUE)) {
  install.packages("ggrepel", repos = "https://cloud.r-project.org")
}
library(ggrepel)

# Standard Wilson score interval on a proportion, base R only (matches
# the method underlying stats::prop.test, computed directly here for
# clarity). Used throughout this document as the single, consistent
# criterion for "is a cluster's coverage of a category meaningfully
# above the corpus baseline" — the lower bound of the interval must
# exceed the corpus rate. Replaces earlier, separately-chosen fixed
# floors (an arbitrary minimum count and enrichment ratio) that were
# used in different places in this document without being reconciled
# with each other; one method, defined once, used everywhere below.
wilson_lower_bound <- function(x, n, conf = 0.95) {
  z <- qnorm(1 - (1 - conf) / 2)
  phat <- x / n
  denom <- 1 + z^2 / n
  centre <- phat + z^2 / (2 * n)
  adj <- z * sqrt(phat * (1 - phat) / n + z^2 / (4 * n^2))
  (centre - adj) / denom
}
```

```{r}
#| label: join
# df (Part 1's annotation dataframe, from 01_annotation.qmd) is keyed by
# cite_key; centrality_df / node_table (Part 2's structural data, from
# 02_network_structure.qmd) is keyed by name (citekey). F1-only:
# annotations exist for F1 papers only. F2 works cited by multiple F1
# papers can still be described using the annotation profile of their
# *citer set*, even though the F2 works themselves are not annotated.
# See "per-F2-node citer profiles" below.

if (!exists("df", inherits = FALSE)) {
  df <- read_csv(here("analysis", "bibvik_annotation_df.csv"), show_col_types = FALSE)
  cat("Loaded df from disk:", nrow(df), "rows\n")
} else {
  cat("df already in scope:", nrow(df), "rows\n")
}
if (!exists("node_table", inherits = FALSE)) {
  node_table <- read_csv(here("analysis", "bibvik_node_table.csv"), show_col_types = FALSE)
  cat("Loaded node_table from disk:", nrow(node_table), "rows\n")
} else {
  cat("node_table already in scope:", nrow(node_table), "rows\n")
}
# node_table, not centrality_df, is the object that actually carries
# community_gn: in 02_network_structure.qmd, node_table is built FROM
# centrality_df with the community assignment joined on afterward
# (node_table <- centrality_df |> left_join(...)), so centrality_df
# itself never has community_gn even when it's in scope. Preferring
# centrality_df here (an earlier version of this join checked for it
# first) silently grabbed the wrong, earlier object in the combined
# index.qmd render, where centrality_df exists but lacks the column
# this file needs — always use node_table.
structural_df <- node_table
cat("structural_df:", nrow(structural_df), "rows. is_ghost values:",
    paste(unique(structural_df$is_ghost), collapse = ", "),
    "(class:", class(structural_df$is_ghost), ")\n")

# is_ghost is written as an R logical (TRUE/FALSE) by 02_network_structure.qmd
# internally, but a round-trip through write_csv()/read_csv() serializes
# logicals as the bare words TRUE/FALSE, not the quoted string "false" the
# original in-memory comparison assumed. Compare against both to be safe
# regardless of whether structural_df came from disk (logical after
# read_csv's type inference) or was passed in memory as character "false".
cluster_table_annotated <- structural_df |>
  filter(is_ghost == FALSE | is_ghost == "false") |>
  left_join(df, by = c("name" = "cite_key"))

cat("Joined cluster_table_annotated:", nrow(cluster_table_annotated), "rows\n")
cat("Rows with a matching annotation:", sum(!is.na(cluster_table_annotated$publication_year)), "\n")
```

## Cluster structure vs. topics, methods, and sources of evidence

`community_gn` (Girvan-Newman, fragmentation-onset cut) is checked here
against the topics, methods, and sources of evidence coded in Part 1,
restricted to the 294 works with a matching annotation (F1-only; see
the join above). Louvain and Leiden cluster assignments also exist
from Part 2 but are not used in this section.

Coverage is thin relative to the graph as a whole: 294 annotated works
out of 14,331 nodes, and within any single cluster that annotated
count is often a small fraction of the cluster's real size (the
"other" grouping below hides clusters with fewer than 3 annotated
works, but even the clusters shown may have only a handful of
annotated works out of a much larger true membership). A prior
exploratory pass on an earlier GN cut of this same corpus found 7-13
annotated nodes per cluster typical, out of hundreds of members, and
treated agreement between structure and annotation as directionally
suggestive rather than comprehensively verified for that reason. The
same caution applies here: what follows can show that structure and
coding line up somewhere, but cannot rule out that the uncoded majority
of any cluster tells a different story.

The coded columns are wide, one column per topic/method/source
category, valued `"Primary"`, `"Secondary"`, or blank. Primary and
secondary are merged throughout what follows: a work counts as coded
with a category at either level, matching the method used in a prior
exploratory pass on an earlier GN cut of this corpus. A work can carry
more than one topic, method, or source, so per-cluster percentages
below do not sum to 100%; each cell shows the raw count behind its
percentage for transparency.

**How to read the enrichment heatmaps below.** Raw coverage (what
fraction of a cluster's annotated works carry a given category) is
misleading on its own: some categories are common across the whole
corpus regardless of cluster — "Spatial analysis" as a method, or
"Environment" as a topic, show up in most clusters simply because
most works in this corpus use them, not because any one cluster is
distinctively characterized by them. To separate genuine
distinctiveness from corpus-wide popularity, each heatmap cell instead
shows **enrichment**: a cluster's coverage of a category divided by
that category's rate across the whole annotated corpus. A value of 1x
means the cluster carries the category at exactly the corpus
average; 2x means twice the corpus rate; 0.5x means half. White is 1x
(baseline); blue is above baseline (the category is over-represented in
that cluster, a real distinguishing signal); red is below baseline
(under-represented). For example, if "Palaeoenvironment reconstruction"
is coded in 23% of the whole corpus but 100% of cluster 52's
annotated works, that cell shows roughly 4.4x — a real signal that
environmental method is disproportionately concentrated there, not
just present because it's common everywhere.

### By topic

```{r}
#| label: cluster-vs-topic
annotated_only <- cluster_table_annotated |>
  filter(!is.na(publication_year))

cat("Works with both a cluster assignment and an annotation:", nrow(annotated_only), "\n")

gn_true_sizes <- node_table |>
  filter(!is.na(community_gn)) |>
  count(community_gn, name = "true_size")

topic_long <- annotated_only |>
  select(name, community_gn, contains("Topics..")) |>
  pivot_longer(contains("Topics.."), names_to = "topic", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " "))

cat("Works with any topic coded:", n_distinct(topic_long$name), "\n")
```

**Girvan-Newman clusters by topic.** A threshold of 3+ annotated
works let through 118 of 199 clusters, unworkable at any scale.
Shown here instead, in the heatmap below: the 20 clusters with the
most annotated works. This is a readability choice, not a derived
breakpoint; the underlying cluster sizes have a clear natural break
only between the first and second-largest clusters (a drop of over
600 members), after which the size distribution tapers smoothly with
no further break to justify any particular cutoff.

```{r}
#| label: gn-vs-topic-table
gn_topic_annotated_n <- topic_long |>
  distinct(community_gn, name) |>
  count(community_gn, name = "n_annotated")
gn_topic_keep <- gn_topic_annotated_n |> slice_max(n_annotated, n = 20) |> pull(community_gn)
```

Coverage varies a lot by cluster (annotation is thin relative to true
cluster size, per the caveat above), so each row label shows how many
annotated works its percentages are based on, out of the cluster's
true total size.

```{r}
#| label: gn-vs-topic-heatmap
#| fig-width: 10
#| fig-height: 7
gn_topic_labels <- gn_topic_annotated_n |>
  filter(community_gn %in% gn_topic_keep) |>
  left_join(gn_true_sizes, by = "community_gn") |>
  arrange(desc(true_size)) |>
  mutate(label = sprintf("Cluster %s (%d of %d annotated)", community_gn, n_annotated, true_size)) |>
  mutate(label = fct_rev(fct_inorder(label))) |>
  select(community_gn, label) |>
  deframe()

# Corpus-wide base rate per topic: what fraction of ALL annotated works
# (not just the 20 shown clusters) carry this topic. Some categories
# (e.g. broad topics like "Environment") are common across the whole
# corpus regardless of cluster, so raw per-cluster coverage lights
# up almost everywhere for them without saying anything distinctive
# about any one cluster. Dividing by the corpus rate instead shows
# enrichment: whether a cluster carries a topic MORE than the corpus
# baseline, which is the actual signal of thematic distinctiveness.
topic_corpus_rate <- topic_long |>
  distinct(name, topic) |>
  count(topic, name = "n_corpus") |>
  mutate(corpus_rate = n_corpus / n_distinct(topic_long$name))

topic_long |>
  filter(community_gn %in% gn_topic_keep) |>
  distinct(community_gn, name, topic) |>
  count(community_gn, topic) |>
  left_join(gn_topic_annotated_n, by = "community_gn") |>
  left_join(topic_corpus_rate, by = "topic") |>
  mutate(
    prop = n / n_annotated,
    enrichment = prop / corpus_rate
  ) |>
  # Drop topics with no coverage in any of the shown clusters, so the
  # chart doesn't carry empty columns for categories nobody here was
  # coded with.
  group_by(topic) |>
  filter(sum(n) > 0) |>
  ungroup() |>
  mutate(cluster_label = gn_topic_labels[as.character(community_gn)]) |>
  ggplot(aes(x = topic, y = cluster_label, fill = enrichment)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = n), size = 3, colour = "grey20") +
  # Diverging scale centred on 1.0 (cluster rate == corpus rate): blue
  # means this cluster carries the topic more than the corpus baseline
  # (distinctive), red means less. Capped at 3x so a handful of extreme
  # small-n ratios don't wash out the rest of the scale.
  scale_fill_gradient2(low = "firebrick", mid = "white", high = "steelblue",
                        midpoint = 1, limits = c(0, 3), oob = scales::squish,
                        labels = scales::label_number(suffix = "x")) +
  labs(x = "", y = "", fill = "Enrichment vs.\ncorpus base rate",
       title = "Topic enrichment within each Girvan-Newman cluster",
       subtitle = "Cell numbers are raw counts. Fill shows coverage relative to the corpus-wide rate for that topic, not raw coverage — a cluster at 1x carries the topic at exactly the corpus average; ubiquitous topics (e.g. Environment) will not dominate every row here the way they do on raw coverage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### By method

Same approach as the topic comparison above, applied to method.

```{r}
#| label: cluster-vs-method
method_long <- annotated_only |>
  select(name, community_gn, contains("Methods..")) |>
  pivot_longer(contains("Methods.."), names_to = "method", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " "))

cat("Works with any method coded:", n_distinct(method_long$name), "\n")
```

```{r}
#| label: gn-vs-method-table
gn_method_annotated_n <- method_long |>
  distinct(community_gn, name) |>
  count(community_gn, name = "n_annotated")
gn_method_keep <- gn_method_annotated_n |> slice_max(n_annotated, n = 20) |> pull(community_gn)
```

```{r}
#| label: gn-vs-method-heatmap
#| fig-width: 10
#| fig-height: 7
gn_method_labels <- gn_method_annotated_n |>
  filter(community_gn %in% gn_method_keep) |>
  left_join(gn_true_sizes, by = "community_gn") |>
  arrange(desc(true_size)) |>
  mutate(label = sprintf("Cluster %s (%d of %d annotated)", community_gn, n_annotated, true_size)) |>
  mutate(label = fct_rev(fct_inorder(label))) |>
  select(community_gn, label) |>
  deframe()

method_corpus_rate <- method_long |>
  distinct(name, method) |>
  count(method, name = "n_corpus") |>
  mutate(corpus_rate = n_corpus / n_distinct(method_long$name))

method_long |>
  filter(community_gn %in% gn_method_keep) |>
  distinct(community_gn, name, method) |>
  count(community_gn, method) |>
  left_join(gn_method_annotated_n, by = "community_gn") |>
  left_join(method_corpus_rate, by = "method") |>
  mutate(
    prop = n / n_annotated,
    enrichment = prop / corpus_rate
  ) |>
  group_by(method) |>
  filter(sum(n) > 0) |>
  ungroup() |>
  mutate(cluster_label = gn_method_labels[as.character(community_gn)]) |>
  ggplot(aes(x = method, y = cluster_label, fill = enrichment)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = n), size = 3, colour = "grey20") +
  scale_fill_gradient2(low = "firebrick", mid = "white", high = "steelblue",
                        midpoint = 1, limits = c(0, 3), oob = scales::squish,
                        labels = scales::label_number(suffix = "x")) +
  labs(x = "", y = "", fill = "Enrichment vs.\ncorpus base rate",
       title = "Method enrichment within each Girvan-Newman cluster",
       subtitle = "Cell numbers are raw counts. Fill shows coverage relative to the corpus-wide rate for that method, not raw coverage — ubiquitous methods (e.g. Spatial analysis) will not dominate every row here the way they do on raw coverage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### By source of evidence

Same approach again, applied to source of evidence.

```{r}
#| label: cluster-vs-source
source_long <- annotated_only |>
  select(name, community_gn, contains("Sources.of.Evidence..")) |>
  pivot_longer(contains("Sources.of.Evidence.."), names_to = "source", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " "))

cat("Works with any source of evidence coded:", n_distinct(source_long$name), "\n")
```

```{r}
#| label: gn-vs-source-table
gn_source_annotated_n <- source_long |>
  distinct(community_gn, name) |>
  count(community_gn, name = "n_annotated")
gn_source_keep <- gn_source_annotated_n |> slice_max(n_annotated, n = 20) |> pull(community_gn)
```

```{r}
#| label: gn-vs-source-heatmap
#| fig-width: 10
#| fig-height: 7
gn_source_labels <- gn_source_annotated_n |>
  filter(community_gn %in% gn_source_keep) |>
  left_join(gn_true_sizes, by = "community_gn") |>
  arrange(desc(true_size)) |>
  mutate(label = sprintf("Cluster %s (%d of %d annotated)", community_gn, n_annotated, true_size)) |>
  mutate(label = fct_rev(fct_inorder(label))) |>
  select(community_gn, label) |>
  deframe()

source_corpus_rate <- source_long |>
  distinct(name, source) |>
  count(source, name = "n_corpus") |>
  mutate(corpus_rate = n_corpus / n_distinct(source_long$name))

source_long |>
  filter(community_gn %in% gn_source_keep) |>
  distinct(community_gn, name, source) |>
  count(community_gn, source) |>
  left_join(gn_source_annotated_n, by = "community_gn") |>
  left_join(source_corpus_rate, by = "source") |>
  mutate(
    prop = n / n_annotated,
    enrichment = prop / corpus_rate
  ) |>
  group_by(source) |>
  filter(sum(n) > 0) |>
  ungroup() |>
  mutate(cluster_label = gn_source_labels[as.character(community_gn)]) |>
  ggplot(aes(x = source, y = cluster_label, fill = enrichment)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = n), size = 3, colour = "grey20") +
  scale_fill_gradient2(low = "firebrick", mid = "white", high = "steelblue",
                        midpoint = 1, limits = c(0, 3), oob = scales::squish,
                        labels = scales::label_number(suffix = "x")) +
  labs(x = "", y = "", fill = "Enrichment vs.\ncorpus base rate",
       title = "Source-of-evidence enrichment within each Girvan-Newman cluster",
       subtitle = "Cell numbers are raw counts. Fill shows coverage relative to the corpus-wide rate for that source, not raw coverage — ubiquitous sources (e.g. Written sources, Graves and monuments) will not dominate every row here the way they do on raw coverage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Cluster signatures: peak count and peak strength

Throughout this section, "cluster" means a Girvan-Newman graph
cluster (`community_gn`), not a scholarly or social cluster.

The topic, method, and source enrichment findings above surface
individual standout cells one at a time. This section combines them
into two directly interpretable numbers per cluster, so the
clusters most worth a closer look surface first instead of
requiring a manual scan across three separate tables.

A **peak** is any category (topic, method, or source) whose 95% Wilson
confidence interval on a cluster's coverage rate lies entirely above
the corpus-wide rate for that category — meaning the cluster's
coverage is higher than baseline with reasonable statistical
confidence, not just a higher raw count. This adapts to sample size
automatically: a cluster with only 3 annotated works needs a much
larger raw enrichment to register as a peak than a cluster with 13,
because a small sample's interval is wider. This is the same criterion
used in the topic, method, and source findings above (`ci_lower >
corpus_rate`); "peak" here just names a category that already met that
bar, so this section's peak count and mean enrichment are computed
directly from the same standouts listed there, not a separate pass.

For each cluster, two numbers: how many peaks it has, and how strong
those peaks are on average. Neither is combined into a single score —
peak count times mean enrichment is not a real, interpretable unit, and
collapsing two different kinds of signal into one number invites
over-reading it. Plotted directly instead: clusters with many peaks
AND high average enrichment are worth reading first; a cluster with
one very large but isolated peak is a different kind of finding (a
strong single signal) than one with several moderate peaks (a broader,
possibly more coherent pattern) — the plot keeps that distinction
visible rather than flattening it. Whether a cluster's peak
categories actually cohere thematically is a judgment that requires
reading them against domain knowledge of this corpus, which this plot
does not attempt.

```{r}
#| label: gn-cluster-signature
# A peak requires the Wilson lower bound (defined in setup) on a
# cluster's coverage rate to exceed the corpus baseline rate for
# that category.
topic_peaks <- topic_long |>
  filter(community_gn %in% gn_topic_keep) |>
  distinct(community_gn, name, topic) |>
  count(community_gn, topic) |>
  left_join(gn_topic_annotated_n, by = "community_gn") |>
  left_join(topic_corpus_rate, by = "topic") |>
  mutate(prop = n / n_annotated, enrichment = prop / corpus_rate,
         ci_lower = wilson_lower_bound(n, n_annotated),
         category_type = "topic") |>
  rename(category = topic)

method_peaks <- method_long |>
  filter(community_gn %in% gn_method_keep) |>
  distinct(community_gn, name, method) |>
  count(community_gn, method) |>
  left_join(gn_method_annotated_n, by = "community_gn") |>
  left_join(method_corpus_rate, by = "method") |>
  mutate(prop = n / n_annotated, enrichment = prop / corpus_rate,
         ci_lower = wilson_lower_bound(n, n_annotated),
         category_type = "method") |>
  rename(category = method)

source_peaks <- source_long |>
  filter(community_gn %in% gn_source_keep) |>
  distinct(community_gn, name, source) |>
  count(community_gn, source) |>
  left_join(gn_source_annotated_n, by = "community_gn") |>
  left_join(source_corpus_rate, by = "source") |>
  mutate(prop = n / n_annotated, enrichment = prop / corpus_rate,
         ci_lower = wilson_lower_bound(n, n_annotated),
         category_type = "source") |>
  rename(category = source)

all_peaks <- bind_rows(topic_peaks, method_peaks, source_peaks) |>
  filter(ci_lower > corpus_rate)

cat("Peaks found (Wilson lower bound exceeds corpus baseline):", nrow(all_peaks), "\n")

cluster_signatures <- all_peaks |>
  group_by(community_gn) |>
  summarise(
    n_peaks = n(),
    mean_peak_enrichment = mean(enrichment),
    peak_categories = paste0(category_type, ": ", category, " (", round(enrichment, 1), "x)", collapse = "; "),
    .groups = "drop"
  ) |>
  arrange(desc(n_peaks), desc(mean_peak_enrichment))

# No composite "score" here — peak count x mean enrichment is not a real
# unit and a single number invites over-interpretation. Plotted directly
# instead: clusters with many peaks AND high average enrichment sit
# in the upper right, which is where to look first; clusters with one
# very large peak but few others sit lower-right (a strong single
# signal, not necessarily a broad pattern); clusters with several
# modest peaks sit upper-left.
cluster_signatures |>
  ggplot(aes(x = n_peaks, y = mean_peak_enrichment, label = community_gn)) +
  geom_point(size = 3, colour = "steelblue") +
  geom_text_repel(size = 3.5, seed = 1) +
  scale_x_continuous(breaks = scales::breaks_pretty()) +
  labs(x = "Number of peaks (categories where coverage exceeds\ncorpus baseline, Wilson 95% lower bound)",
       y = "Mean enrichment among those peaks",
       title = "Clusters by peak count and peak strength",
       subtitle = "Upper right: several strong, corroborating signals. Lower right: one very strong but isolated signal.") +
  theme_minimal()
```

The peak categories behind the clusters with the most peaks, shown
directly rather than as a summary number:

```{r}
#| label: gn-cluster-signature-detail
#| fig-width: 9
#| fig-height: 7
top_communities <- cluster_signatures |> slice_max(n_peaks, n = 8, with_ties = FALSE) |> pull(community_gn)

all_peaks |>
  filter(community_gn %in% top_communities) |>
  mutate(
    cluster_label = fct_relevel(paste("Cluster", community_gn),
                                    paste("Cluster", top_communities)),
    category_label = fct_reorder(paste0(category, " (", category_type, ")"), enrichment)
  ) |>
  ggplot(aes(x = enrichment, y = category_label)) +
  geom_col(fill = "steelblue") +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  facet_wrap(vars(cluster_label), scales = "free_y", ncol = 2) +
  labs(x = "Enrichment vs. corpus base rate", y = "",
       title = "Peak categories for the clusters with the most peaks",
       subtitle = "Dashed line = corpus baseline (1x). Bars are that cluster's individual peaks.") +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```

## Not yet done

### Remaining planned comparisons

- Whether centrality (`in_degree`, `betweenness`, `pagerank`) within a
  cluster correlates with author gender — i.e. whether the most central
  or bridging works within a cluster are disproportionately authored by
  one gender even when the cluster's dominant topic is not itself
  gender-skewed in the annotation data.
- Direct citation homophily: whether papers (or, at a future author-level
  aggregation, authors) disproportionately cite others sharing their coded
  gender / topic / method characteristics. Computable directly from the
  edge list and the join above; does not require any cluster-detection
  output.
- Per-F2-node citer profiles: for F2 works with `in_degree` above some
  threshold, characterize the annotation profile (gender, topic, method)
  of their F1 citer set, to see whether particular sources function as
  shared touchstones across the field's internal divisions or as
  in-group markers cited mainly within one part of it.

### Open decisions

- Whether and how to implement bounded-depth, parallelized Girvan-Newman;
  no benchmarking done yet on how many splits are needed for a stable,
  interpretable first division, or on runtime for the full-graph
  betweenness computation as currently implemented.
- Whether directed variants of GN/Louvain/Leiden are worth the added
  complexity for this project's specific questions.
- Which alternative unit of analysis (author, venue, kind-of-work), if any,
  to pursue first, and on what timeline relative to the paper-level work
  in this document.
- How explicitly to frame the seed-paper-testing angle in eventual
  writeups, given the PI/co-author relationship.
- Normalizing topic/method/source proportions against the corpus-wide
  gender base rate, rather than reporting raw proportions.
- Which gender-composition categorization (`composition_strict`,
  `composition_majority`, `composition_singleauthor_split`, or another)
  is actually the most useful for the topic/method/source cross-tabs —
  computed side by side, not yet chosen.
- `fieldwork_lead_methods` (the mapping from coded method category to
  first-vs-last lead-author convention) is a placeholder and must be
  filled in with real codebook category names before the lead-author
  analysis is meaningful; currently everything defaults to "ambiguous."
- Secondary topic/method/source values are not yet included in any
  gender cross-tab (composition-based or first-author-only) — only
  primary values are currently compared against gender.
- Region is not yet cross-tabbed against gender/topic/method/source
  (word cloud only, by design for now).
