---
title: "BibVik Citation Analysis: Camps, Communities, and Citational Politics"
subtitle: "Strategy, method, and analysis — working document, 2026-07-03"
format:
  html:
    toc: false
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
  eval: false
---

## Status

This is a single working document combining the analytical plan with the
code and outputs that carry it out. Several conceptual questions below are
resolved; several concrete decisions (final scope of Girvan-Newman, whether
to add directionality, which alternative unit of analysis to pursue first)
are deliberately still open and marked as such. Sections without code are
strategy and reasoning; sections with code are live analysis.

## Starting point

The original motivating question, from the seed paper (Lund & Sindbæk 2022):
does the field divide into something like two camps around the origin of the
Viking Age, along lines that co-vary across **gender, topic, and method**?
The seed paper proposes this division without drawing firm boundaries and
without assigning most authors to a side — it names at least one individual
as an anchor point, but the claim itself is about the tripartite
co-variation, not about any specific person.

This is a substantive, testable proposition from the literature — worth
distinguishing from the research team's own fieldwork-based instincts about
where the data might be heading (e.g. topic/gender associations noticed
informally in the field). Both are legitimate to bring to this analysis, but
they carry different epistemic weight and should be kept distinct in any
eventual writeup: one is a claim to confirm or complicate; the other is an
interpretive lens the team brings to reading whatever the data shows.

Because Julie Lund is both PI on this project and a co-author of the seed
paper, testing the seed paper's own supposition carries some obvious
sensitivity. The recommendation is to be transparent about this in whatever
methods writeup accompanies the analysis, rather than obscuring whose claim
is under examination.

## Core methodological stance: no baseline, just independent descriptions

This is not a search for deviation from an expected or random pattern. The
dataset is historically and socially situated — citation clustering along
social, institutional, or intellectual lines is the expected,
produced-by-history outcome, not an anomaly standing out against a null
model. This is a real objection to modularity-based community detection
(Louvain, Leiden), whose optimization target is explicitly "more clustered
than a random graph with the same degree sequence would be." That framing
is not a neutral default for this kind of data.

The working alternative, used throughout this document: treat every
computed structure — citation-graph community structure, the annotation
cross-tab, per-work citer profiles, direct homophily rates — as an
**independent, self-contained description** of the corpus, computed on its
own terms with no reference point implied. None is privileged as ground
truth against which another is checked. The actual analytical content is in
the *relationship* between these descriptions once more than one exists:
agreement, disagreement, partial overlap, or cross-cutting are all just
facts about that relationship, not confirmations or violations of anything.
This is the logic behind treating the annotation data as something to
"bolt on" to a citation-structure analysis, rather than using one to
validate the other, and it governs how every result below should be read.

## Why Girvan-Newman over modularity methods, for the "camps" question

Louvain and Leiden are fast (near-linear time, tractable at any scale
without special hardware) and were considered as a way around
Girvan-Newman's poor scalability. Both are computed below as additional
independent descriptions, but neither is treated as the primary method for
the "camps" question specifically, for reasons distinct from performance:

- Their optimization target (deviation from a random-graph null model) is
  the wrong lens for socially/historically produced structure, per above.
- The "camps" question is specifically about **whether the graph splits
  cleanly, and in what order** — a first split into two, each of which may
  or may not split further into more nuanced sub-divisions. Girvan-Newman's
  hierarchical, edge-removal structure encodes exactly this nesting as a
  byproduct of how it runs; a flat modularity partition does not answer
  "is two-way division the first-order truth" as directly.

Louvain/Leiden are computed as **additional, independent descriptions** in
their own right — not as a faster substitute for Girvan-Newman, and not as
a "cross-check" implying agreement would validate either method. The two
families of method ask different structural questions (dense-cluster
membership vs. load-bearing bridges) and can legitimately produce
different, non-comparable pictures of the same graph. Their disagreement,
if any, is reported as a finding, not a discrepancy to resolve.

## Girvan-Newman tractability

**The problem.** Full Girvan-Newman decomposition is expensive because edge
betweenness centrality must be recomputed after every single edge removal —
this is inherently sequential across iterations (removal *i+1* depends on
the graph state after removal *i*), so parallelizing across iterations is
not possible. An earlier assessment that GN "can't really be parallelized"
on an available 84-core server is only half right.

**What does parallelize.** The betweenness calculation *within* a single
iteration (Brandes' algorithm) sums independent per-source-node
contributions, and is embarrassingly parallel across those source nodes.
`graph-tool` supports this; a `networkx`-based multiprocessing wrapper could
achieve something similar. This would meaningfully speed up the expensive
per-iteration cost, without approximating anything — every node, including
low-degree ones, remains fully part of the calculation. **Not yet
implemented in the code below** — the current implementation uses
`igraph::cluster_edge_betweenness()` as-is, which does not parallelize the
per-iteration cost. This should be revisited if runtime on the full graph
proves impractical.

**Rejected optimization: peeling degree-1 (leaf) nodes, or pre-filtering to
a k-core, before running GN.** Considered and explicitly rejected. Leaves
(or low-coreness nodes generally) are cheap to strip and their community
assignment could be trivially inferred afterward from their highest-degree
neighbor — but this is a real approximation, not a free win:

- Leaves attached to a neighbor whose own community assignment is itself
  marginal/unstable would inherit an arbitrary decision.
- Leaf-peeling cascades: removing one leaf can drop a neighbor to degree 1,
  triggering further removal, which can strip small but genuine citation
  chains, not just pendant noise.
- Removing leaves changes the betweenness values of the edges that remain,
  since leaves still contribute path endpoints to the full-graph
  calculation — the pruned-graph result is not guaranteed to approximate
  the full-graph result closely.
- Most importantly for this project: **a paper cited only once may be
  exactly the interesting case** (niche, contested, or newly-emerging
  reference) for a study of citational politics, not noise to discard.
  Peripheral/rarely-cited nodes are data, not overhead, given the research
  question explicitly concerns how peripheral material relates to
  dominant positions.

**Current direction: bounded-depth GN on the full, unfiltered graph.**
Rather than decomposing to individual papers, or filtering the input graph
first, the edge-removal process is read off at the first several splits
(e.g. first split, then the splits within each resulting side), computed on
every node in the graph. This is not merely a performance compromise — the
first few splits are the more interpretable structural fact for the "camps"
question; full decomposition to singleton papers stops being a meaningful
description of factional structure at all. This is implemented below;
whether it is *fast enough* on the full graph without the parallelization
described above has not yet been benchmarked.

## Time and directionality

The F1 corpus spans a single decade — too short to usefully time-slice
without each slice becoming too sparse to support reliable community
detection or homophily estimates. Time-sliced structural analysis and
homophily-as-a-trend-over-time were both considered and set aside for this
reason, not because time is uninteresting in principle.

**Directionality remains open** and is a separate matter from time-slicing.
The citation graph is directed (citer → cited); the community-detection
methods below are run on the undirected projection (standard practice for
this class of algorithm, and consistent with the centrality measures, which
are computed in directed form separately). Whether "who is drawn upon vs.
who does the drawing" should be built into the community-detection step
itself, using directed variants of GN/Louvain/Leiden, is not yet decided.

**F2 publication date at time of citation** is a separate small thread worth
keeping in mind: F2 works are not bound by the F1 decade and may span a much
longer history. Whether the age of cited material varies systematically
depending on which part of the F1 corpus is doing the citing is a fact
already present in the data (publication years) without requiring any
time-slicing of the corpus itself. Not yet implemented below.

## The independent descriptions computed in this document

Consistent with the no-baseline stance, the analysis is organized as a set
of independently computable, independently meaningful descriptions of the
corpus, to be placed alongside each other rather than run in a required
sequence. This document computes:

1. **Structural descriptions of the citation graph** — degree, centrality
   (betweenness, closeness, PageRank, HITS), k-core, connected components,
   and community detection (Girvan-Newman, Louvain, Leiden, k-means on
   centrality profiles). See below.

2. **The annotation cross-tab** — author gender × primary/secondary topic ×
   primary/secondary method, across the F1 corpus. Tests whether the seed
   paper's supposed tripartite co-variation is visible in the coded data at
   all. Computed independently of any graph work.

Not yet implemented in this document, recorded here so the plan is not
lost:

3. **Per-F2-node citer profiles.** For F2 works cited by multiple F1 papers,
   characterize the citer set using annotation data that exists only on the
   F1 side: gender composition, topic/method diversity, and whether the
   citer set is itself structurally clustered or scattered. Does not
   require resolving the community-detection method question and is
   tractable immediately once prioritized. Speaks to a genuinely different
   citational-politics phenomenon than the corpus-wide cross-tab: whether
   particular sources function as shared/foundational touchstones cited
   across the field's internal divisions, or as in-group markers cited
   mainly within one part of it.

4. **Direct citation homophily.** Do papers (or, at the author level — see
   "alternative units" below — authors) disproportionately cite others
   sharing their coded gender/topic/method characteristics? A direct join
   of two already-independent descriptions (citation edges, annotation
   data) with no algorithmic mediation in between. Likely the most
   immediately tractable of the not-yet-built analyses.

5. **Joining structural and annotation descriptions.** Whether
   Girvan-Newman's top-level split(s), Louvain communities, and Leiden
   communities correspond to, cut across, or ignore the coded categories;
   whether centrality within a community correlates with author gender
   even when a community's dominant topic is not itself gender-skewed.

## Extension for future consideration: alternative units of analysis

Everything in this document treats the paper as the unit of the graph
(papers cite papers). This is one particular projection of a richer
structure, not the only one, and is worth extending in future work — not
resolved or prioritized now, but recorded so it isn't lost:

- **Author-level graphs.** Collapsing papers into authors changes what
  centrality means (a prolific author becomes a hub independent of any
  single paper's standing) and makes author-to-author citation homophily a
  more natural unit than paper-to-paper for gender-citation questions.
  Co-authorship itself becomes an additional structural question: do
  co-authorships cross or reinforce whatever divisions appear elsewhere.

- **Venue-level graphs** (journal, edited volume, conference proceedings,
  grey literature/excavation report). Likely to reveal institutional or
  disciplinary structure rather than intellectual structure directly, and
  could complicate a purely topic/gender reading — a "camp" might partly
  reflect which kind of venue a sub-field mostly publishes in, particularly
  given the corpus's substantial non-English, non-journal material
  (excavation reports, conference proceedings).

- **Kind-of-work as a unit** (monograph vs. article vs. excavation report
  vs. edited-volume chapter). Related to venue but not identical; may track
  disciplinary/generational norms that could explain or cross-cut part of
  any gender/topic pattern rather than sitting alongside it as a fully
  independent axis.

Each alternative unit would be, again, an independent description rather
than a more "real" or authoritative one. A useful way to stress-test any
finding at the paper level is to check whether it persists, sharpens, or
dissolves when re-aggregated to authors or venues — if a division only
appears at one level of aggregation, that is itself informative about
whether the division is fundamentally about individual works, about
people, or about where people publish.

---

# Part 1 — Annotation descriptives (F1 corpus)

This section computes the annotation cross-tab entirely independently of
the graph work above — no reference to citation structure, communities, or
centrality. Per the no-baseline stance, this is not a check against Part 1;
it stands on its own, and Part 3 (below) is where the two are placed
side by side.

```{r}
#| label: annotation-libraries
library(patchwork)
library(nord)
library(tidytext)
library(khroma)
library(ggh4x)
```

```{r}
#| label: annotation-palettes
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868"
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]
gender_colors <- c("Woman" = nord_palettes$red_mountain[1],
                   "Man" = nord_palettes$algoma_forest[4])
```

```{r}
#| label: annotation-load
df <- read.csv(here("data/BibVik_Data_Entry_cite_keys.csv"), na.strings = "")

df <- df %>%
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))
```

## Publications per year

```{r}
#| label: pubyears
pubyears <- df %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat="identity", fill = nord_palettes$aurora[1]) +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE ) +
  labs(x = "", y = "Count", title = "Publications by year")
```

## Gender by first author and all authors

Publications by gender for each year. Single-author work vs. collaborations
have not yet been examined separately.

```{r}
#| label: gender-by-year
#| fig-width: 10
#| fig-height: 8
gender_pubyears <- df %>%
  group_by(Authors..Genders..Author.1.) %>%
  count(publication_year) %>%
  mutate(proportion_year = n / sum(n))

men_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Man",]) / nrow(gender_pubyears) * 100)
women_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Woman",]) / nrow(gender_pubyears) * 100)

plt1 <-
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1.)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = gender_colors) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = paste0("First author gender: Women = ", women_prcnt, "% Men = ", men_prcnt, "%")) +
  theme(legend.position = "bottom")

all_authors_gender_year <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(publication_year, gender) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na()

all_authors_gender <- df %>%
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(gender) %>%
  mutate(gender = na_if(gender, "")) %>%
  drop_na()

men_aa <- round(all_authors_gender[all_authors_gender$gender == "Man", "n"] / sum(all_authors_gender$n) * 100)
women_aa <- round(all_authors_gender[all_authors_gender$gender == "Woman", "n"] / sum(all_authors_gender$n) * 100)
uncertain_aa <- round(all_authors_gender[all_authors_gender$gender == "Uncertain", "n"] / sum(all_authors_gender$n) * 100)

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender",
       title = paste0("All authors gender: Uncertain = ", uncertain_aa, "% Women = ", women_aa, "% Men = ", men_aa, "%")) +
  theme(legend.position = "bottom")

plt1 / plt2
```

## Multi-author gender composition and author order

The analyses above use first-author gender only (for the topic/method/source
cross-tabs later in this document) or a flattened, order-blind pool of all
author genders (the "all authors" plot). Neither captures gender
**composition per paper** (all-male vs. all-female vs. mixed) or **order**
within a paper, both of which are needed for the co-authorship questions
raised in planning (do co-authorships cross or reinforce whatever divisions
appear elsewhere; does the most senior/lead author's gender matter
independently of the full author-list composition).

**Author order is genuinely ambiguous in this corpus and must be handled
carefully, not with a single fixed rule.** Field convention varies:
first-author-as-lead is standard for excavation/fieldwork-based
publications, but last-author-as-lead (PI convention) is standard for
lab-based work. Which convention applies depends on the paper's own
method — one of the coded variables, not something inferable from the
author list alone. The lookup below (`fieldwork_lead_methods`) is a
placeholder mapping specific coded method categories to the convention
that applies; **it must be filled in with the actual method category
names from the codebook before this analysis is meaningful.** Left
unmapped, methods default to "ambiguous" and are excluded from any
lead-author-specific analysis (though still included in the
order-blind composition analysis below).

This section works generically across however many
`Authors..Genders..Author.N.` columns exist in the data — no fixed
column count is assumed.

```{r}
#| label: author-gender-cols
author_gender_cols <- df %>%
  select(starts_with("Authors..Genders..Author.")) %>%
  names()

cat("Detected", length(author_gender_cols), "author-position gender columns:\n")
print(author_gender_cols)
```

```{r}
#| label: gender-sequence
# Per-paper ordered gender sequence (e.g. "Man, Woman, Man"), preserving
# author order and author count, rather than pooling into an unordered set.
gender_sequences <- df %>%
  select(cite_key, all_of(author_gender_cols)) %>%
  pivot_longer(
    cols = all_of(author_gender_cols),
    names_to = "position",
    values_to = "gender"
  ) %>%
  mutate(gender = na_if(gender, "")) %>%
  filter(!is.na(gender)) %>%
  # Extract numeric author position from the column name for correct
  # ordering regardless of how many positions exist.
  mutate(author_n = as.integer(str_extract(position, "\\d+$"))) %>%
  arrange(cite_key, author_n) %>%
  group_by(cite_key) %>%
  summarise(
    n_authors        = n(),
    gender_sequence  = paste(gender, collapse = ", "),
    first_author_gender = first(gender),
    last_author_gender  = last(gender),
    .groups = "drop"
  )

gender_sequences |> count(n_authors) |> knitr::kable(caption = "Author-count distribution")
```

### Candidate gender-composition categorizations

Several ways of collapsing a per-paper gender sequence into a single
cross-tabbable category are computed side by side below, since it is not
yet decided which is most useful. This is a case of comparing several
independent categorization schemes against the same underlying data, not
picking the "correct" one — inspect the resulting distributions before
choosing.

```{r}
#| label: gender-composition-candidates
composition_candidates <- gender_sequences %>%
  rowwise() %>%
  mutate(
    genders_list = list(str_split(gender_sequence, ", ")[[1]]),
    n_men       = sum(genders_list == "Man"),
    n_women     = sum(genders_list == "Woman"),
    n_uncertain = sum(genders_list == "Uncertain"),
  ) %>%
  ungroup() %>%
  mutate(
    # Candidate 1: strict all-same vs. mixed
    composition_strict = case_when(
      n_uncertain > 0            ~ "includes uncertain",
      n_men > 0 & n_women == 0   ~ "all-male",
      n_women > 0 & n_men == 0   ~ "all-female",
      n_men > 0 & n_women > 0    ~ "mixed",
      TRUE                       ~ NA_character_
    ),
    # Candidate 2: majority-based (ties = balanced)
    composition_majority = case_when(
      n_uncertain > 0     ~ "includes uncertain",
      n_men > n_women     ~ "majority-male",
      n_women > n_men     ~ "majority-female",
      n_men == n_women    ~ "balanced",
      TRUE                ~ NA_character_
    ),
    # Candidate 3: single-author papers split out separately, since
    # "all-male"/"all-female" for n=1 is a different phenomenon
    # (individual authorship) than for n>1 (a genuinely uniform team)
    composition_singleauthor_split = case_when(
      n_authors == 1 & n_men == 1   ~ "solo-male",
      n_authors == 1 & n_women == 1 ~ "solo-female",
      n_authors == 1                ~ "solo-uncertain",
      n_uncertain > 0                ~ "includes uncertain",
      n_men > 0 & n_women == 0       ~ "all-male (multi-author)",
      n_women > 0 & n_men == 0       ~ "all-female (multi-author)",
      TRUE                           ~ "mixed"
    )
  ) %>%
  select(-genders_list)

composition_candidates |> count(composition_strict) |> knitr::kable(caption = "Candidate 1: strict (all-same vs. mixed)")
composition_candidates |> count(composition_majority) |> knitr::kable(caption = "Candidate 2: majority-based")
composition_candidates |> count(composition_singleauthor_split) |> knitr::kable(caption = "Candidate 3: solo-authorship split out")
```

### Lead-author gender, conditioned on field convention

```{r}
#| label: lead-author-convention
# PLACEHOLDER — fill in with actual method category names from the
# codebook. Maps each coded primary method to whether first-author or
# last-author is the conventional "lead" position for that kind of work.
# Methods not listed default to "ambiguous" and are excluded from the
# lead-author-specific comparison below (they remain in every other
# analysis in this document).
fieldwork_lead_methods <- c(
  # "Excavation"        = "first",
  # "Survey"            = "first",
  # "Isotope analysis"  = "last",
  # "aDNA analysis"     = "last"
)

primary_method_lookup <- df %>%
  select(cite_key, contains("Methods")) %>%
  pivot_longer(!cite_key, names_to = "method", values_to = "value") %>%
  filter(value == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  select(cite_key, primary_method = method)

lead_author_gender <- gender_sequences %>%
  left_join(primary_method_lookup, by = "cite_key") %>%
  mutate(
    lead_convention = fieldwork_lead_methods[primary_method],
    lead_convention = replace_na(lead_convention, "ambiguous"),
    lead_author_gender = case_when(
      lead_convention == "first" ~ first_author_gender,
      lead_convention == "last"  ~ last_author_gender,
      TRUE                        ~ NA_character_
    )
  )

cat("Papers by lead convention (based on placeholder mapping above — fill in before trusting this):\n")
lead_author_gender |> count(lead_convention) |> knitr::kable()
```

### Gender composition by primary topic, method, and source of evidence

Unlike the first-author-only cross-tabs later in this document, this uses
the composition categories above — bringing multi-author, order-aware
gender data into the same kind of comparison already done for first-author
gender. `composition_strict` is used here as a starting point; swap in
`composition_majority` or `composition_singleauthor_split` to compare.

```{r}
#| label: composition-by-topic
#| fig-width: 8
composition_topic <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Topics")) %>%
      pivot_longer(!cite_key, names_to = "topic", values_to = "count") %>%
      filter(count == "Primary") %>%
      mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(topic), !is.na(composition_strict))

composition_topic %>%
  count(topic, composition_strict) %>%
  group_by(topic) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(topic = fct_reorder(topic, prop * (composition_strict == "all-female"), .fun = sum),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")")) %>%
  ggplot(aes(x = topic_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Author gender composition",
       title = "Primary topic by full-author-list gender composition")
```

```{r}
#| label: composition-by-method
#| fig-width: 8
composition_method <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Methods")) %>%
      pivot_longer(!cite_key, names_to = "method", values_to = "count") %>%
      filter(count == "Primary") %>%
      mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(method), !is.na(composition_strict))

composition_method %>%
  count(method, composition_strict) %>%
  group_by(method) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")")) %>%
  ggplot(aes(x = method_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Author gender composition",
       title = "Primary method by full-author-list gender composition")
```

```{r}
#| label: composition-by-source
#| fig-width: 8
composition_source <- composition_candidates %>%
  select(cite_key, composition_strict) %>%
  left_join(
    df %>% select(cite_key, contains("Sources.of.Evidence")) %>%
      pivot_longer(!cite_key, names_to = "source", values_to = "count") %>%
      filter(count == "Primary") %>%
      mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")),
    by = "cite_key"
  ) %>%
  filter(!is.na(source), !is.na(composition_strict))

composition_source %>%
  count(source, composition_strict) %>%
  group_by(source) %>%
  mutate(prop = n / sum(n), total = sum(n)) %>%
  ungroup() %>%
  mutate(source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")")) %>%
  ggplot(aes(x = source_label, y = prop, fill = composition_strict)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = "Author gender composition",
       title = "Primary source of evidence by full-author-list gender composition")
```

**Not yet done, worth noting:** the topic/method/source cross-tabs above
use *primary* values only, same limitation as the first-author-only
cross-tabs later in this document — secondary topic/method/source values
are not yet brought into any gender comparison. Extending
`composition_topic`/`composition_method`/`composition_source` to include
secondary values would mean joining on both Primary and Secondary counts
rather than filtering to Primary only, and deciding how to weight or
display a paper that contributes to multiple topics at once.

## Primary topic, in aggregate and over time

```{r}
#| label: topics
#| fig-width: 10
#| fig-height: 6
topic_count <- df %>%
  select(publication_year, contains("Topics")) %>%
  pivot_longer(!publication_year,
               names_to = "topic",
               values_to = "count")

totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n)

primary_count_year <- topic_count %>%
  group_by(topic, publication_year) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(publication_year) %>%
  mutate(prop_primary = n / sum(n))

primary_count <- topic_count %>%
  group_by(topic) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_count_year, aes(x = publication_year, y = n, fill = topic)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count_year, aes(x = publication_year, y = prop_primary, fill = topic)) +
  geom_bar(stat="identity")+
  geom_text(data = totals_by_year,
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count |> mutate(topic = str_wrap(topic, width = 25)),
       aes(x = reorder(topic, n), y = n)) +
  geom_bar(stat="identity", fill = topic_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary topic", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("topics_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

## Primary method, in aggregate and over time

```{r}
#| label: methods
#| fig-width: 10
#| fig-height: 6
method_count <- df %>%
  select(publication_year, contains("Methods")) %>%
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")

primary_method_year <- method_count %>%
  group_by(method, publication_year) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(publication_year) %>%
  mutate(prop_primary = n / sum(n))

primary_count_method <- method_count %>%
  group_by(method) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_method_year, aes(x = publication_year, y = n, fill = method)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Count", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_method_year, aes(x = publication_year, y = prop_primary, fill = method)) +
  geom_bar(stat="identity") +
  geom_text(data = totals_by_year,
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count_method |> mutate(method = str_wrap(method, width = 25)),
       aes(x = reorder(method, n), y = n)) +
  geom_bar(stat="identity", fill = method_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary method", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("methods_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

## Primary sources of evidence, in aggregate and over time

```{r}
#| label: sources
#| fig-width: 10
#| fig-height: 6
totals_by_year <- df %>%
  count(publication_year) %>%
  rename(total = n) %>%
  mutate(publication_year = factor(publication_year, levels = sort(unique(as.numeric(as.character(publication_year))))))

source_count <- df %>%
  select(publication_year, contains("Sources.of.Evidence")) %>%
  mutate(publication_year = factor(publication_year, levels = levels(totals_by_year$publication_year))) %>%
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")

primary_source_year <- source_count %>%
  group_by(source, publication_year) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(publication_year) %>%
  mutate(prop_primary = n / sum(n)) %>%
  ungroup() %>%
  mutate(publication_year = droplevels(publication_year))

primary_source_count <- source_count %>%
  group_by(source) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_source_year, aes(x = publication_year, y = n, fill = source)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 2))
ggsave("sources_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_source_year, aes(x = publication_year, y = prop_primary, fill = source)) +
  geom_bar(stat="identity")+
  geom_text(data = totals_by_year %>% filter(publication_year %in% unique(primary_source_year$publication_year)),
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("sources_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_source_count |> mutate(source = str_wrap(source, width = 25)),
       aes(x = reorder(source, n), y = n)) +
  geom_bar(stat="identity", fill = source_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary source of evidence", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("sources_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

## Secondary topics by primary topic

```{r}
#| label: topics-secondary
#| fig-width: 30
#| fig-height: 20
topics <- df %>%
  select(contains("Topics")) %>%
  rename_with(~str_remove(., "Topics.."))

coocurrences <- c()
for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary topic by primary topic")
```

## Secondary methods by primary method

```{r}
#| label: methods-secondary
#| fig-width: 20
#| fig-height: 20
methods <- df %>%
  select(contains("Methods")) %>%
  rename_with(~str_remove(., "Methods.."))

coocurrences <- c()
for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary method by primary method")
```

## Secondary sources of evidence by primary source of evidence

```{r}
#| label: sources-secondary
#| fig-width: 20
#| fig-height: 20
sources <- df %>%
  select(contains("Sources.of.Evidence")) %>%
  rename_with(~str_remove(., "Sources.of.Evidence.."))

coocurrences <- c()
for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>%
  filter(substr(coocurrences, 1, 1) != "_") %>%
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>%
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>%
  group_by(Primary) %>%
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary source of evidence by primary source of evidence")
```

## Primary topic by first author gender

This is the first of the three plots most directly relevant to the seed
paper's tripartite (gender/topic/method) supposition. Normalizing each
topic's proportions against the corpus-wide gender base rate has been
discussed as a follow-up refinement but not yet implemented — the raw
proportions below do not yet control for the overall gender imbalance in
the corpus.

```{r}
#| label: topic-by-gender
#| fig-width: 8
topic_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Topics")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_topic_gender <- topic_count_gender %>%
  group_by(topic, gender_first_author) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(topic) %>%
  mutate(prop_primary = n / sum(n))

primary_topic_gender |>
  group_by(topic) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(topic = fct_reorder(topic, ratio),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary topic by first author gender")
```

## Primary method by first author gender

```{r}
#| label: method-by-gender
#| fig-width: 8
method_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Methods")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_method_gender <- method_count_gender %>%
  group_by(method, gender_first_author) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(method) %>%
  mutate(prop_primary = n / sum(n))

primary_method_gender |>
  group_by(method) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(method = fct_reorder(method, ratio),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary method by first author gender")
```

## Primary source of evidence by first author gender

```{r}
#| label: source-by-gender
#| fig-width: 8
source_count_gender <- df %>%
  select(Authors..Genders..Author.1., contains("Sources")) %>%
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>%
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_source_gender <- source_count_gender %>%
  group_by(source, gender_first_author) %>%
  count(count) %>%
  filter(count == "Primary") %>%
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(source) %>%
  mutate(prop_primary = n / sum(n))

primary_source_gender |>
  group_by(source) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(source = fct_reorder(source, ratio),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary source of evidence by first author gender")
```

## Region

```{r}
#| label: region
library(ggwordcloud)

region_counts <- df %>%
  pull(Region) %>%
  na.omit() %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Region", "n")) %>%
  mutate(n = as.numeric(n),
         n_scaled = n + 3) %>%
  arrange(desc(n))

ggplot(region_counts, aes(label = Region, size = n_scaled, color = n)) +
  geom_text_wordcloud(area_corr = TRUE, rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  scale_color_gradient(low = topic_col, high = "darkred") +
  theme_minimal()
```

---

# Part 2 — Citation graph: structural descriptions

## Setup

```{r}
#| label: install
#| eval: false
install.packages(c("igraph", "tidyverse", "ggraph", "tidygraph", "jsonlite"))
```

```{r}
#| label: libraries
library(igraph)
library(tidyverse)
library(ggraph)
library(tidygraph)
library(jsonlite)
library(here)
```

```{r}
#| label: load-graph
g <- read_graph("/home/zack/models/BibVik_output/citation_graph.graphml", format = "graphml")
g <- as.directed(g, mode = "arbitrary")

bib <- fromJSON("/home/zack/models/BibVik_output/bibliography.json", simplifyDataFrame = FALSE)

# Active entries only (tombstoned entries are excluded from the export,
# but guard here in case of an older bibliography file). Note: entries
# tombstoned via `delete` (not `merge`) that appear as a citer elsewhere
# are retained in the export as flagged "ghost" nodes so their outbound
# edges aren't lost — see exporter.py and docs/methods/corrections-system.md.
# Ghost nodes carry is_ghost = "true" in the GraphML attributes and have
# no reliable bibliographic metadata; they are kept in the graph (their
# edges are real) but excluded from any display table or join that assumes
# real title/author/topic data.
bib <- bib[!sapply(bib, function(e) isTRUE(e[["_deleted"]]))]

node_attrs <- tibble(
  name            = names(bib),
  title           = map_chr(bib, ~ .x$title %||% ""),
  year            = map_chr(bib, ~ .x$date %||% .x$year %||% ""),
  generation      = map_chr(bib, ~ .x$generation %||% ""),
  entry_type      = map_chr(bib, ~ .x$entry_type %||% ""),
  resolution      = map_chr(bib, ~ .x[["_resolution_method"]] %||% "grobid"),
  completeness    = map_chr(bib, ~ .x$completeness$label %||% ""),
)

is_ghost <- tibble(
  name     = V(g)$name,
  is_ghost = V(g)$is_ghost %||% rep("false", vcount(g))
)

cat("Nodes:", vcount(g), "\n")
cat("Edges:", ecount(g), "\n")
cat("Directed:", is_directed(g), "\n")
cat("Density:", edge_density(g), "\n")
cat("Ghost nodes (tombstoned entries retained for edge preservation):",
    sum(is_ghost$is_ghost == "true"), "\n")
```

## Degree and centrality

In-degree (times cited within the corpus) and out-degree (number of works
cited) capture different roles. High in-degree identifies widely-cited works
— the intellectual anchors of the field. High out-degree identifies works
that are heavily engaged with the literature, drawing on many sources.

Betweenness centrality identifies works that sit on many shortest paths
between other works — brokers between sub-literatures. Betweenness is
expensive to compute on large graphs; normalized values are reported.

PageRank assigns authority based on incoming links weighted by the
authority of the linking node — conceptually similar to being cited by
works that are themselves widely cited. Hub and authority scores (HITS)
are complementary: hubs point to many authorities; authorities are pointed
to by many hubs.

Peripheral works — low degree, low centrality by every measure below — are
not treated as noise in this project. A work cited only once may be exactly
the interesting case (a niche, contested, or newly emerging reference) for
a study of citational politics. No node is excluded from any measure in
this section on the basis of low degree.

```{r}
#| label: degree
V(g)$degree_in  <- degree(g, mode = "in")
V(g)$degree_out <- degree(g, mode = "out")
V(g)$degree_all <- degree(g, mode = "all")
```

```{r}
#| label: centrality
V(g)$betweenness  <- betweenness(g, directed = TRUE, normalized = TRUE)
V(g)$closeness_in <- closeness(g, mode = "in", normalized = TRUE)
V(g)$pagerank     <- page_rank(g, directed = TRUE)$vector
V(g)$hub          <- hub_score(g)$vector
V(g)$authority    <- authority_score(g)$vector

centrality_df <- tibble(
  name        = V(g)$name,
  betweenness = V(g)$betweenness,
  closeness   = V(g)$closeness_in,
  pagerank    = V(g)$pagerank,
  hub         = V(g)$hub,
  authority   = V(g)$authority,
  in_degree   = V(g)$degree_in,
  out_degree  = V(g)$degree_out,
  degree      = V(g)$degree_all,
) |> left_join(node_attrs, by = "name") |> left_join(is_ghost, by = "name")
```

### Most cited (by in-degree)

```{r}
centrality_df |>
  filter(is_ghost == "false") |>
  arrange(desc(in_degree)) |>
  head(20) |>
  select(name, title, year, generation, in_degree) |>
  knitr::kable()
```

### Most citing (by out-degree)

```{r}
centrality_df |>
  filter(is_ghost == "false") |>
  arrange(desc(out_degree)) |>
  head(20) |>
  select(name, title, year, generation, out_degree) |>
  knitr::kable()
```

### Degree distribution

A power-law or scale-free degree distribution is characteristic of citation
networks: most works are cited rarely, a small number are cited frequently.
The log-log plot tests this.

```{r}
#| label: degree-dist
#| fig-width: 8
#| fig-height: 5
ggplot(centrality_df, aes(x = in_degree)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "In-degree distribution (log-log)",
    x = "In-degree (times cited)",
    y = "Count"
  ) +
  theme_minimal()
```

### Top by betweenness centrality

Works with high betweenness mediate between otherwise disconnected parts
of the citation network — likely works that bridge sub-fields or periods.

```{r}
centrality_df |>
  filter(is_ghost == "false") |>
  arrange(desc(betweenness)) |>
  head(20) |>
  select(name, title, year, betweenness) |>
  knitr::kable(digits = 4)
```

## Connected components

A weakly connected component (WCC) treats edges as undirected. A strongly
connected component (SCC) requires paths in both directions — rare in
citation networks, since citations are historically directed (a 2010 paper
cannot cite a 2020 paper). The WCC structure shows how fragmented the
network is; a dominant giant component indicates a well-connected literature.

```{r}
#| label: components
wcc <- components(g, mode = "weak")
scc <- components(g, mode = "strong")

cat("Weakly connected components:", wcc$no, "\n")
cat("Largest WCC:", max(wcc$csize), "nodes (",
    round(100 * max(wcc$csize) / vcount(g), 1), "% of graph)\n")
cat("Strongly connected components:", scc$no, "\n")
cat("Largest SCC:", max(scc$csize), "nodes\n")

V(g)$component_weak   <- wcc$membership
V(g)$component_strong <- scc$membership
```

### WCC size distribution

```{r}
tibble(size = wcc$csize) |>
  count(size) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

## K-core decomposition

The k-core of a graph is the maximal subgraph in which every node has at
least k connections to other nodes within that subgraph. It is computed by
iteratively removing nodes with degree below k until no further removal
is possible. Nodes are assigned a coreness value equal to the highest k
for which they remain in the k-core.

K-core decomposition is reported here purely as a **descriptive** measure —
identifying which works are structurally embedded in dense mutual-citation
relationships versus which sit at the periphery. It is not used below as a
preprocessing filter for any other method; every peripheral node (coreness
= 1, or excluded entirely from any k ≥ 2 core) remains a full participant in
every other analysis in this document, including Girvan-Newman.

Reference: Alvarez-Hamelin et al. (2005). Large scale networks
fingerprinting and visualization using the k-core decomposition.
*Advances in Neural Information Processing Systems*, 18.

```{r}
#| label: kcore
g_undirected <- as.undirected(g, mode = "collapse")
V(g)$coreness <- coreness(g_undirected)

cat("Coreness distribution:\n")
table(V(g)$coreness) |> print()

centrality_df <- centrality_df |>
  left_join(tibble(name = V(g)$name, coreness = V(g)$coreness), by = "name")

centrality_df |>
  group_by(coreness) |>
  summarise(n = n(), mean_in_degree = mean(in_degree)) |>
  arrange(desc(coreness)) |>
  head(15) |>
  knitr::kable(digits = 2)
```

```{r}
#| label: kcore-top
centrality_df |>
  filter(is_ghost == "false") |>
  arrange(desc(coreness), desc(in_degree)) |>
  head(20) |>
  select(name, title, year, coreness, in_degree) |>
  knitr::kable()
```

## Community detection

Community detection here produces several independent descriptions of how
the graph divides, not a single answer to be triangulated (see "Core
methodological stance" above). Louvain and Leiden ask "which groups of
nodes are more densely interconnected than a random graph with the same
degree sequence would predict." Girvan-Newman asks "which edges, if
removed, disconnect otherwise-separate parts of the network." These are
genuinely different questions; neither validates the other.

References:

- Blondel et al. (2008). Fast unfolding of communities in large networks.
  *Journal of Statistical Mechanics*, P10008.
- Traag et al. (2019). From Louvain to Leiden. *Scientific Reports*, 9, 5233.
- Girvan & Newman (2002). Community structure in social and biological
  networks. *PNAS*, 99(12), 7821–7826.
- Newman (2006). Modularity and community structure in networks.
  *PNAS*, 103(23), 8577–8582.

```{r}
#| label: louvain
louvain <- cluster_louvain(g_undirected)
cat("Louvain communities:", length(louvain), "\n")
cat("Modularity:", round(modularity(louvain), 4), "\n")
V(g)$community_louvain <- membership(louvain)

tibble(community = seq_along(sizes(louvain)), size = sizes(louvain)) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

```{r}
#| label: louvain-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(community = as.factor(V(g)$community_louvain))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = community), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = "Citation graph — Louvain communities")

ggsave("bibvik_louvain.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_louvain.pdf", width = 16, height = 16)
```

```{r}
#| label: leiden
leiden <- cluster_leiden(g_undirected, objective_function = "modularity")
cat("Leiden communities:", length(leiden), "\n")
cat("Modularity:", round(modularity(leiden), 4), "\n")
V(g)$community_leiden <- membership(leiden)

tibble(community = seq_along(sizes(leiden)), size = sizes(leiden)) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

```{r}
#| label: leiden-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(community = as.factor(V(g)$community_leiden))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = community), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = "Citation graph — Leiden communities")

ggsave("bibvik_leiden.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_leiden.pdf", width = 16, height = 16)
```

### Girvan-Newman, bounded depth, full graph, no pre-filtering

Run on the complete undirected graph — no k-core or degree-based
pre-filtering (see "Girvan-Newman tractability" above for why). `n_splits`
controls how many top-level divisions to read off, rather than decomposing
to full singleton dendrogram depth. `igraph::cluster_edge_betweenness()`
computes the full dendrogram internally; `cutat()` reads off the partition
at a given number of communities without altering the underlying
(expensive) betweenness computation — this means the implementation below
does **not** yet save the compute cost of full decomposition; it defers
that optimization (parallelized per-iteration betweenness) to future work
and should be benchmarked on the full graph before assuming it is
tractable as-is.

```{r}
#| label: girvan-newman
#| cache: true
n_splits <- 4  # number of top-level communities to report; placeholder — inspect the dendrogram below before trusting this

gn <- cluster_edge_betweenness(g_undirected)
cat("Girvan-Newman (full dendrogram computed):\n")
cat("  Total merge steps:", length(gn$merges), "\n")
cat("  Modularity-optimal partition:", length(gn), "communities, modularity =",
    round(modularity(gn), 4), "\n")

# Bounded-depth partition: read off the first n_splits communities from the
# dendrogram, rather than using the modularity-optimal cut. This is the
# structure relevant to "does the graph divide into a small number of
# camps" — not necessarily the same cut igraph would pick by modularity.
gn_bounded <- cutat(gn, no = n_splits)
V(g)$community_gn <- gn_bounded

tibble(community = gn_bounded) |>
  count(community) |>
  arrange(desc(n)) |>
  knitr::kable()
```

```{r}
#| label: gn-dendrogram
#| fig-width: 12
#| fig-height: 8
# The first few merges/splits from the top are the analytically relevant
# structure for a "camps" question; deeper structure corresponds to
# increasingly fine-grained sub-divisions.
plot_dendrogram(gn, mode = "hclust")
```

```{r}
#| label: gn-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(community = as.factor(V(g)$community_gn))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = community), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = paste0("Citation graph — Girvan-Newman, top ", n_splits, " communities"))

ggsave("bibvik_gn.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_gn.pdf", width = 16, height = 16)
```

## K-means clustering

K-means clusters nodes based on their centrality feature profiles rather
than graph structure. It groups works that occupy similar structural
positions — for example, highly-cited authorities vs. high-betweenness
brokers vs. peripheral works — rather than groups that densely cite each
other. This is a further independent description, not a validation of the
community-detection methods above.

```{r}
#| label: km
features <- centrality_df |>
  select(betweenness, closeness, pagerank, hub, authority, in_degree) |>
  mutate(across(everything(), ~ replace_na(., 0))) |>
  mutate(across(everything(), ~ scale(.)[,1]))

set.seed(42)
k <- 6
km <- kmeans(features, centers = k, nstart = 25)
centrality_df$cluster_kmeans <- km$cluster
V(g)$cluster_kmeans <- km$cluster

cat("Cluster sizes:\n")
table(km$cluster) |> print()
```

```{r}
#| label: km-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(cluster = as.factor(V(g)$cluster_kmeans))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = cluster), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = "Citation graph — K-means clusters (by centrality profile)")

ggsave("bibvik_kmeans.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_kmeans.pdf", width = 16, height = 16)
```

### Cluster profiles

```{r}
#| label: km-profiles
centrality_df |>
  group_by(cluster_kmeans) |>
  summarise(
    n                = n(),
    mean_pagerank    = mean(pagerank, na.rm = TRUE),
    mean_betweenness = mean(betweenness, na.rm = TRUE),
    mean_authority   = mean(authority, na.rm = TRUE),
    mean_in_degree   = mean(in_degree, na.rm = TRUE),
  ) |>
  arrange(desc(mean_in_degree)) |>
  knitr::kable(digits = 5)
```

## Structural node table export

```{r}
#| label: export-structural
node_table <- centrality_df |>
  left_join(
    tibble(
      name              = V(g)$name,
      coreness          = V(g)$coreness,
      component_weak    = V(g)$component_weak,
      component_strong  = V(g)$component_strong,
      community_louvain = V(g)$community_louvain,
      community_leiden  = V(g)$community_leiden,
      community_gn      = V(g)$community_gn,
      cluster_kmeans    = V(g)$cluster_kmeans,
    ),
    by = "name"
  )

write_csv(node_table, "bibvik_node_table.csv")
cat("Written bibvik_node_table.csv\n")
cat("Rows:", nrow(node_table), "\n")
```


# Part 3 — Joining structural and annotation descriptions

Everything in Part 1 describes the graph on its own terms; everything in
Part 2 describes the annotation data on its own terms. Neither is a
baseline the other is checked against — this section places them side by
side and reads the relationship, which is where the actual analytical
content of this project lives.

```{r}
#| label: join
#| eval: false
# Adjust path once the annotation export location is finalized. df (from
# Part 2) is keyed by cite_key; centrality_df / node_table (from Part 1)
# is keyed by name (citekey). F1-only: annotations exist for F1 papers
# only. F2 works cited by multiple F1 papers can still be described using
# the annotation profile of their *citer set*, even though the F2 works
# themselves are not annotated — see "per-F2-node citer profiles" above.

node_table_annotated <- centrality_df |>
  filter(is_ghost == "false") |>
  left_join(df, by = c("name" = "cite_key"))
```

Planned comparisons, not yet implemented:

- Whether Girvan-Newman's top-level split(s) (`community_gn`), Louvain
  communities (`community_louvain`), and Leiden communities
  (`community_leiden`) correspond to, cut across, or ignore the coded
  topic / method / source-of-evidence / author-gender categories from
  Part 2.
- Whether centrality (`in_degree`, `betweenness`, `pagerank`) within a
  community correlates with author gender — i.e. whether the most central
  or bridging works within a cluster are disproportionately authored by
  one gender even when the cluster's dominant topic is not itself
  gender-skewed in the annotation data.
- Direct citation homophily: whether papers (or, at a future author-level
  aggregation, authors) disproportionately cite others sharing their coded
  gender / topic / method characteristics. Computable directly from the
  edge list and the join above; does not require any community-detection
  output.
- Per-F2-node citer profiles: for F2 works with `in_degree` above some
  threshold, characterize the annotation profile (gender, topic, method)
  of their F1 citer set, to see whether particular sources function as
  shared touchstones across the field's internal divisions or as
  in-group markers cited mainly within one part of it.

## Open decisions (not yet resolved)

- Whether and how to implement bounded-depth, parallelized Girvan-Newman;
  no benchmarking done yet on how many splits are needed for a stable,
  interpretable first division, or on runtime for the full-graph
  betweenness computation as currently implemented.
- Whether directed variants of GN/Louvain/Leiden are worth the added
  complexity for this project's specific questions.
- Which alternative unit of analysis (author, venue, kind-of-work), if any,
  to pursue first, and on what timeline relative to the paper-level work
  in this document.
- How explicitly to frame the seed-paper-testing angle in eventual
  writeups, given the PI/co-author relationship.
- Normalizing topic/method/source proportions against the corpus-wide
  gender base rate, rather than reporting raw proportions.
- Which gender-composition categorization (`composition_strict`,
  `composition_majority`, `composition_singleauthor_split`, or another)
  is actually the most useful for the topic/method/source cross-tabs —
  computed side by side, not yet chosen.
- `fieldwork_lead_methods` (the mapping from coded method category to
  first-vs-last lead-author convention) is a placeholder and must be
  filled in with real codebook category names before the lead-author
  analysis is meaningful; currently everything defaults to "ambiguous."
- Secondary topic/method/source values are not yet included in any
  gender cross-tab (composition-based or first-author-only) — only
  primary values are currently compared against gender.
- Region is not yet cross-tabbed against gender/topic/method/source
  (word cloud only, by design for now).
