---
title: "BibVik Citation Graph Analysis"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
---

## Setup

```{r}
#| label: install
#| eval: false
install.packages(c("igraph", "tidyverse", "ggraph", "tidygraph", "jsonlite"))
```

```{r}
#| label: libraries
library(igraph)
library(tidyverse)
library(ggraph)
library(tidygraph)
library(jsonlite)
```

```{r}
#| label: load-graph
g <- read_graph("/home/zack/models/BibVik_output/citation_graph.graphml", format = "graphml")
g <- as.directed(g, mode = "arbitrary")

bib <- fromJSON("/home/zack/models/BibVik_output/bibliography.json", simplifyDataFrame = FALSE)

# Active entries only (tombstoned entries are excluded from the export,
# but guard here in case of an older bibliography file)
bib <- bib[!sapply(bib, function(e) isTRUE(e[["_deleted"]]))]

node_attrs <- tibble(
  name            = names(bib),
  title           = map_chr(bib, ~ .x$title %||% ""),
  year            = map_chr(bib, ~ .x$date %||% .x$year %||% ""),
  generation      = map_chr(bib, ~ .x$generation %||% ""),
  entry_type      = map_chr(bib, ~ .x$entry_type %||% ""),
  resolution      = map_chr(bib, ~ .x[["_resolution_method"]] %||% "grobid"),
  completeness    = map_chr(bib, ~ .x$completeness$label %||% ""),
)

cat("Nodes:", vcount(g), "\n")
cat("Edges:", ecount(g), "\n")
cat("Directed:", is_directed(g), "\n")
cat("Density:", edge_density(g), "\n")
```

## Degree and centrality

In-degree (times cited within the corpus) and out-degree (number of works
cited) capture different roles. High in-degree identifies widely-cited works
— the intellectual anchors of the field. High out-degree identifies works
that are heavily engaged with the literature, drawing on many sources.

Betweenness centrality identifies works that sit on many shortest paths
between other works — brokers between sub-literatures. Betweenness is
expensive to compute on large graphs; normalized values are reported.

PageRank assigns authority based on incoming links weighted by the
authority of the linking node — conceptually similar to being cited by
works that are themselves widely cited. Hub and authority scores (HITS)
are complementary: hubs point to many authorities; authorities are pointed
to by many hubs.

```{r}
#| label: degree
V(g)$degree_in  <- degree(g, mode = "in")
V(g)$degree_out <- degree(g, mode = "out")
V(g)$degree_all <- degree(g, mode = "all")
```

```{r}
#| label: centrality
V(g)$betweenness  <- betweenness(g, directed = TRUE, normalized = TRUE)
V(g)$closeness_in <- closeness(g, mode = "in", normalized = TRUE)
V(g)$pagerank     <- page_rank(g, directed = TRUE)$vector
V(g)$hub          <- hub_score(g)$vector
V(g)$authority    <- authority_score(g)$vector

centrality_df <- tibble(
  name        = V(g)$name,
  betweenness = V(g)$betweenness,
  closeness   = V(g)$closeness_in,
  pagerank    = V(g)$pagerank,
  hub         = V(g)$hub,
  authority   = V(g)$authority,
  in_degree   = V(g)$degree_in,
  out_degree  = V(g)$degree_out,
  degree      = V(g)$degree_all,
) |> left_join(node_attrs, by = "name")
```

### Most cited (by in-degree)

```{r}
centrality_df |>
  arrange(desc(in_degree)) |>
  head(20) |>
  select(name, title, year, generation, in_degree) |>
  knitr::kable()
```

### Most citing (by out-degree)

```{r}
centrality_df |>
  arrange(desc(out_degree)) |>
  head(20) |>
  select(name, title, year, generation, out_degree) |>
  knitr::kable()
```

### Degree distribution

A power-law or scale-free degree distribution is characteristic of citation
networks: most works are cited rarely, a small number are cited frequently.
The log-log plot tests this.

```{r}
#| label: degree-dist
#| fig-width: 8
#| fig-height: 5
ggplot(centrality_df, aes(x = in_degree)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "In-degree distribution (log-log)",
    x = "In-degree (times cited)",
    y = "Count"
  ) +
  theme_minimal()
```

### Top by betweenness centrality

Works with high betweenness mediate between otherwise disconnected parts
of the citation network — likely works that bridge sub-fields or periods.

```{r}
centrality_df |>
  arrange(desc(betweenness)) |>
  head(20) |>
  select(name, title, year, betweenness) |>
  knitr::kable(digits = 4)
```

## Connected components

A weakly connected component (WCC) treats edges as undirected. A strongly
connected component (SCC) requires paths in both directions — rare in
citation networks, since citations are historically directed (a 2010 paper
cannot cite a 2020 paper). The WCC structure shows how fragmented the
network is; a dominant giant component indicates a well-connected literature.

```{r}
#| label: components
wcc <- components(g, mode = "weak")
scc <- components(g, mode = "strong")

cat("Weakly connected components:", wcc$no, "\n")
cat("Largest WCC:", max(wcc$csize), "nodes (",
    round(100 * max(wcc$csize) / vcount(g), 1), "% of graph)\n")
cat("Strongly connected components:", scc$no, "\n")
cat("Largest SCC:", max(scc$csize), "nodes\n")

V(g)$component_weak   <- wcc$membership
V(g)$component_strong <- scc$membership
```

### WCC size distribution

```{r}
tibble(size = wcc$csize) |>
  count(size) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

## K-core decomposition

The k-core of a graph is the maximal subgraph in which every node has at
least k connections to other nodes within that subgraph. It is computed by
iteratively removing nodes with degree below k until no further removal
is possible. Nodes are assigned a coreness value equal to the highest k
for which they remain in the k-core.

K-core decomposition identifies the structurally central works in the
citation network — those that are mutually embedded in dense citation
relationships. Works in higher cores are not simply widely cited; they
are cited by works that are themselves central. Peripheral works (coreness
= 1) are connected to the network by a single edge and fall outside all
k > 1 cores.

Reference: Alvarez-Hamelin et al. (2005). Large scale networks
fingerprinting and visualization using the k-core decomposition.
*Advances in Neural Information Processing Systems*, 18.

```{r}
#| label: kcore
g_undirected <- as.undirected(g, mode = "collapse")
V(g)$coreness <- coreness(g_undirected)

cat("Coreness distribution:\n")
table(V(g)$coreness) |> print()

centrality_df <- centrality_df |>
  left_join(tibble(name = V(g)$name, coreness = V(g)$coreness), by = "name")

# Core summary
centrality_df |>
  group_by(coreness) |>
  summarise(n = n(), mean_in_degree = mean(in_degree)) |>
  arrange(desc(coreness)) |>
  head(15) |>
  knitr::kable(digits = 2)
```

```{r}
#| label: kcore-top
# Most central works by coreness
centrality_df |>
  arrange(desc(coreness), desc(in_degree)) |>
  head(20) |>
  select(name, title, year, coreness, in_degree) |>
  knitr::kable()
```

## Community detection

Community detection partitions the citation network into groups of works
that cite each other more densely than they cite works outside the group.
Communities in a citation network correspond roughly to sub-literatures or
research traditions.

All community detection algorithms operate on the undirected projection of
the graph, treating citation relationships symmetrically. This is standard
practice in citation network analysis (Newman 2006; Waltman & van Eck 2012).

Two algorithms are compared:

**Louvain** (Blondel et al. 2008) maximizes modularity — the fraction of
edges within communities minus the expected fraction under a null model.
It is fast (O(n log n)) and scales to large graphs. The standard choice
for citation networks at this scale.

**Leiden** (Traag et al. 2019) improves on Louvain by guaranteeing that
communities are internally well-connected (no disconnected communities).
It is the current best practice for large citation networks.

**Girvan-Newman** (Girvan & Newman 2002) identifies communities by
iteratively removing edges with the highest betweenness centrality. It
produces a dendrogram showing hierarchical community structure and has
a clear narrative interpretation. However, it requires recomputing
betweenness after each edge removal, making it O(m²n) — prohibitively
slow on graphs of this size without preprocessing.

The k-core subgraph (nodes with coreness ≥ threshold) is used as input
to Girvan-Newman to reduce the problem to a tractable size while
preserving the structurally significant portion of the network. The
threshold is a methodological choice; results should be reported with
the threshold value.

References:
- Blondel et al. (2008). Fast unfolding of communities in large networks.
  *Journal of Statistical Mechanics*, P10008.
- Traag et al. (2019). From Louvain to Leiden. *Scientific Reports*, 9, 5233.
- Girvan & Newman (2002). Community structure in social and biological
  networks. *PNAS*, 99(12), 7821–7826.
- Newman (2006). Modularity and community structure in networks.
  *PNAS*, 103(23), 8577–8582.

```{r}
#| label: louvain
louvain <- cluster_louvain(g_undirected)
cat("Louvain communities:", length(louvain), "\n")
cat("Modularity:", round(modularity(louvain), 4), "\n")
V(g)$community_louvain <- membership(louvain)

tibble(community = seq_along(sizes(louvain)), size = sizes(louvain)) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

```{r}
#| label: louvain-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
# DRL (Distributed Recursive Layout) is a force-directed layout designed
# for large graphs (> 1000 nodes). It is appropriate here given the size
# of the BibVik citation graph.

tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(community = as.factor(V(g)$community_louvain))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = community), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = "Citation graph — Louvain communities")

ggsave("bibvik_louvain.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_louvain.pdf", width = 16, height = 16)
```

```{r}
#| label: leiden
leiden <- cluster_leiden(g_undirected, objective_function = "modularity")
cat("Leiden communities:", length(leiden), "\n")
cat("Modularity:", round(modularity(leiden), 4), "\n")
V(g)$community_leiden <- membership(leiden)

tibble(community = seq_along(sizes(leiden)), size = sizes(leiden)) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

```{r}
#| label: leiden-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(community = as.factor(V(g)$community_leiden))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = community), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = "Citation graph — Leiden communities")

ggsave("bibvik_leiden.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_leiden.pdf", width = 16, height = 16)
```

### Girvan-Newman on k-core subgraph

Girvan-Newman is applied to the k-core subgraph to make the computation
tractable. The threshold k is set to retain the structurally dense core
of the network. Peripheral nodes (coreness < k) are excluded from community
detection and assigned post-hoc to the community of their highest-coreness
neighbour.

The choice of k involves a tradeoff: higher k produces a smaller, denser
subgraph with faster computation and cleaner communities, but excludes more
of the network. Lower k is more inclusive but slower. The k = 2 core retains
all nodes with at least two mutual connections — a minimal filter that
removes only true singletons.

```{r}
#| label: girvan-newman
#| cache: true
# Adjust k_threshold to balance coverage vs. computation time.
# k = 2: removes only degree-1 nodes (singletons in the undirected graph)
# k = 3: removes nodes connected to the core by only one or two edges
k_threshold <- 2

core_nodes <- which(V(g)$coreness >= k_threshold)
g_core <- induced_subgraph(g_undirected, core_nodes)

cat("K-core subgraph (k >=", k_threshold, "):\n")
cat("  Nodes:", vcount(g_core), "(", round(100 * vcount(g_core) / vcount(g), 1), "% of full graph)\n")
cat("  Edges:", ecount(g_core), "\n")

gn <- cluster_edge_betweenness(g_core)
cat("\nGirvan-Newman results:\n")
cat("  Communities:", length(gn), "\n")
cat("  Modularity:", round(modularity(gn), 4), "\n")

# Assign community membership back to full graph
# Nodes not in the core are assigned NA
V(g)$community_gn <- NA_integer_
V(g)$community_gn[core_nodes] <- membership(gn)

tibble(community = seq_along(sizes(gn)), size = sizes(gn)) |>
  arrange(desc(size)) |>
  head(20) |>
  knitr::kable()
```

```{r}
#| label: gn-dendrogram
#| fig-width: 12
#| fig-height: 8
# Dendrogram showing hierarchical community structure
plot_dendrogram(gn, mode = "hclust")
```

## K-means clustering

K-means clusters nodes based on their centrality feature profiles rather
than graph structure. It groups works that occupy similar structural
positions — for example, highly-cited authorities vs. high-betweenness
brokers vs. peripheral works — rather than groups that densely cite each
other. The results are complementary to community detection rather than
equivalent.

Features used: betweenness, closeness, PageRank, hub score, authority
score, and in-degree. All features are standardized (z-scored) before
clustering. The number of clusters k is set manually; results should be
compared across values of k.

```{r}
#| label: km
features <- centrality_df |>
  select(betweenness, closeness, pagerank, hub, authority, in_degree) |>
  mutate(across(everything(), ~ replace_na(., 0))) |>
  mutate(across(everything(), ~ scale(.)[,1]))

set.seed(42)
k <- 6
km <- kmeans(features, centers = k, nstart = 25)
centrality_df$cluster_kmeans <- km$cluster
V(g)$cluster_kmeans <- km$cluster

cat("Cluster sizes:\n")
table(km$cluster) |> print()
```

```{r}
#| label: km-viz
#| fig-width: 12
#| fig-height: 12
#| cache: true
tg <- as_tbl_graph(g) |>
  activate(nodes) |>
  mutate(cluster = as.factor(V(g)$cluster_kmeans))

ggraph(tg, layout = "drl") +
  geom_edge_link(alpha = 0.05, colour = "grey60") +
  geom_node_point(aes(colour = cluster), size = 0.3, alpha = 0.7) +
  scale_colour_viridis_d(guide = "none") +
  theme_graph() +
  labs(title = "Citation graph — K-means clusters (by centrality profile)")

ggsave("bibvik_kmeans.png", width = 16, height = 16, dpi = 300)
ggsave("bibvik_kmeans.pdf", width = 16, height = 16)
```

### Cluster profiles

```{r}
#| label: km-profiles
centrality_df |>
  group_by(cluster_kmeans) |>
  summarise(
    n                = n(),
    mean_pagerank    = mean(pagerank, na.rm = TRUE),
    mean_betweenness = mean(betweenness, na.rm = TRUE),
    mean_authority   = mean(authority, na.rm = TRUE),
    mean_in_degree   = mean(in_degree, na.rm = TRUE),
  ) |>
  arrange(desc(mean_in_degree)) |>
  knitr::kable(digits = 5)
```

## Export

All node-level attributes — centrality measures, coreness, community
memberships — are written to a CSV for use in Gephi, R, or other tools.

```{r}
#| label: export
node_table <- centrality_df |>
  left_join(
    tibble(
      name              = V(g)$name,
      coreness          = V(g)$coreness,
      component_weak    = V(g)$component_weak,
      component_strong  = V(g)$component_strong,
      community_louvain = V(g)$community_louvain,
      community_leiden  = V(g)$community_leiden,
      community_gn      = V(g)$community_gn,
      cluster_kmeans    = V(g)$cluster_kmeans,
    ),
    by = "name"
  )

write_csv(node_table, "bibvik_node_table.csv")
cat("Written bibvik_node_table.csv\n")
cat("Rows:", nrow(node_table), "\n")
```
