# Basic descriptive statistics

```{r}
library(tidyverse)
library(patchwork)
library(nord)
library(here)
library(tidytext)
library(khroma)
library(ggh4x)
```

Set up some initial color palettes based on the nord package.

```{r}
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868" 
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

# For maintaining consistent colors for plots with single categories
topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]
gender_colors <- c("Woman" = nord_palettes$red_mountain[1], 
                   "Man" = nord_palettes$algoma_forest[4])
```

Read in the data.

```{r}
df <- read.csv(here("data/BibVik_Data_Entry_cite_keys.csv"), na.strings = "") 

# Set up years as factor variable for plotting, with empty level for 2003 where
# we don't have any publications
df <- df %>% 
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))
```

## Intial plots

### Number of publications per year

```{r}
# Plot by number of publications per year
pubyears <- df %>% 
  count(publication_year) %>% 
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat="identity", fill = nord_palettes$aurora[1]) +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE ) +
  labs(x = "", y = "Count", title = "Publications by year")
```

## Gender by first author and all authors 

Publications by gender for each year. We also discussed looking at single author work and collaborations separately, but have not gotten around to doing that yet.

```{r}
# First author gender by year
gender_pubyears <- df %>% 
  group_by(Authors..Genders..Author.1.) %>% 
  count(publication_year) %>% 
  mutate(proportion_year = n / sum(n))

# Total proportions, first author gender
men_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Man",]) / nrow(gender_pubyears) * 100)
women_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Woman",]) / nrow(gender_pubyears) * 100)

plt1 <- 
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1.)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = gender_colors) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = paste0("First author gender: Women = ", women_prcnt, "% Men = ", men_prcnt, "%")) +
  theme(legend.position = "bottom")


# Repeat for all authors (author poistion 1-5)
all_authors_gender_year <- df %>% 
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(publication_year, gender) %>% 
  mutate(gender = na_if(gender, "")) %>% 
  drop_na()

# Proportion of author genders across all authors
all_authors_gender <- df %>% 
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(gender) %>% 
  mutate(gender = na_if(gender, "")) %>% 
  drop_na()

# Men (55%)
men_aa <- round(all_authors_gender[all_authors_gender$gender == "Man", "n"] / sum(all_authors_gender$n) * 100)
# Women (43%)
women_aa <- round(all_authors_gender[all_authors_gender$gender == "Woman", "n"] / sum(all_authors_gender$n) * 100) 
# Uncertain (2%)
uncertain_aa <- round(all_authors_gender[all_authors_gender$gender == "Uncertain", "n"] / sum(all_authors_gender$n) * 100) 

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender",
       title = paste0("All authors gender: Uncertain = ", uncertain_aa, "% Women = ", women_aa, "% Men = ", men_aa, "%")) +
  theme(legend.position = "bottom")

plt1 / plt2
```

### Primary topic, in aggregate and over time

```{r}
topic_count <- df %>% 
  select(publication_year, contains("Topics")) %>% 
  pivot_longer(!publication_year,
               names_to = "topic",
               values_to = "count")

totals_by_year <- df %>% 
  count(publication_year) %>% 
  rename(total = n)

primary_count_year <- topic_count %>%
  group_by(topic, publication_year) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(publication_year) %>% 
  mutate(prop_primary = n / sum(n))

primary_count <- topic_count %>%
  group_by(topic) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_count_year, aes(x = publication_year, y = n, fill = topic)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count_year, aes(x = publication_year, y = prop_primary, fill = topic)) +
  geom_bar(stat="identity")+
  geom_text(data = totals_by_year,
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 6))
ggsave("topics_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count |> mutate(topic = str_wrap(topic, width = 25)),
       aes(x = reorder(topic, n), y = n)) +
  geom_bar(stat="identity", fill = topic_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)), 
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary topic", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("topics_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### Primary method, in aggregate and over time

```{r}
method_count <- df %>% 
  select(publication_year, contains("Methods")) %>% 
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")

primary_method_year <- method_count %>%
  group_by(method, publication_year) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(publication_year) %>% 
  mutate(prop_primary = n / sum(n))

primary_count_method <- method_count %>%
  group_by(method) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_method_year, aes(x = publication_year, y = n, fill = method)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Count", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_method_year, aes(x = publication_year, y = prop_primary, fill = method)) +
  geom_bar(stat="identity") +
  geom_text(data = totals_by_year,
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("methods_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_count_method |> mutate(method = str_wrap(method, width = 25)),
       aes(x = reorder(method, n), y = n)) +
  geom_bar(stat="identity", fill = method_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)),
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary method", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("methods_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### Primary sources of evidence, in aggregate and over time

```{r}
totals_by_year <- df %>% 
  count(publication_year) %>% 
  rename(total = n) %>%
  mutate(publication_year = factor(publication_year, levels = sort(unique(as.numeric(as.character(publication_year))))))

source_count <- df %>% 
  select(publication_year, contains("Sources.of.Evidence")) %>%
  mutate(publication_year = factor(publication_year, levels = levels(totals_by_year$publication_year))) %>%
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")

primary_source_year <- source_count %>%
  group_by(source, publication_year) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(publication_year) %>% 
  mutate(prop_primary = n / sum(n)) %>%
  ungroup() %>%
  mutate(publication_year = droplevels(publication_year))

primary_source_count <- source_count %>%
  group_by(source) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  ungroup() %>%
  mutate(pct = n / sum(n))

ggplot(primary_source_year, aes(x = publication_year, y = n, fill = source)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 2))
ggsave("sources_count_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_source_year, aes(x = publication_year, y = prop_primary, fill = source)) +
  geom_bar(stat="identity")+
  geom_text(data = totals_by_year %>% filter(publication_year %in% unique(primary_source_year$publication_year)),
            aes(x = publication_year, y = 1.05, label = total),
            inherit.aes = FALSE, size = 3) +
  labs(x = "", y = "Proportion by year", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
ggsave("sources_proportion_by_year.png", plot = last_plot(), width = 10, height = 6, dpi = 300)

ggplot(primary_source_count |> mutate(source = str_wrap(source, width = 25)),
       aes(x = reorder(source, n), y = n)) +
  geom_bar(stat="identity", fill = source_col) +
  geom_text(aes(label = scales::percent(pct, accuracy = 1)), 
            hjust = -0.2, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Primary source of evidence", y = "Count") +
  theme_bw() +
  coord_flip() +
  force_panelsizes(cols = unit(10, "cm"))
ggsave("sources_aggregate.png", plot = last_plot(), width = 10, height = 6, dpi = 300)
```

### Seconday topics by primary topic

```{r fig.width=30, fig.height=20}
topics <- df %>% 
  select(contains("Topics")) %>% 
  rename_with(~str_remove(., "Topics..")) 

coocurrences <- c()

for(i in 1:nrow(topics)) {
  current_row <- topics[i, !is.na(topics[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>% 
  filter(substr(coocurrences, 1, 1) != "_") %>% 
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>% 
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>% 
  group_by(Primary) %>% 
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary topic by primary topic")
```

### Secondary methods by primary method

```{r fig.width=20, fig.height=20}
methods <- df %>% 
  select(contains("Methods")) %>% 
  rename_with(~str_remove(., "Methods..")) 

coocurrences <- c()

for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>% 
  filter(substr(coocurrences, 1, 1) != "_") %>% 
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>% 
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>% 
  group_by(Primary) %>% 
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary method by primary method")
```

### Seconday sources of evidence by primary source of evidence

```{r fig.width=20, fig.height=20}
sources <- df %>% 
  select(contains("Sources.of.Evidence")) %>% 
  rename_with(~str_remove(., "Sources.of.Evidence..")) 

coocurrences <- c()

for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>% 
  filter(substr(coocurrences, 1, 1) != "_") %>% 
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>% 
  mutate(Secondary = str_wrap(str_replace_all(Secondary, fixed("."), " "), width = 20)) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>% 
  group_by(Primary) %>% 
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = reorder_within(Secondary, Prop, Primary), y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary), scales = "free_x", nrow = 6) +
  scale_x_reordered() +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary source of evidence by primary source of evidence")
```

### Primary topic by first author gender

For the subsequent three plots we could also try normalizing the proportions for each topic against the total proportion men/women in the entire corpus.

```{r fig.width=8}
# Count topics by occurrence by gender
topic_count_gender <- df %>% 
  select(Authors..Genders..Author.1., contains("Topics")) %>% 
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>% 
  rename(gender_first_author = Authors..Genders..Author.1.)

# Counts of primary topics by gender
primary_topic_gender <- topic_count_gender %>%
  group_by(topic, gender_first_author) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(topic) %>% 
  mutate(prop_primary = n / sum(n))

primary_topic_gender |>
  group_by(topic) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(topic = fct_reorder(topic, ratio),
         topic = fct_relabel(topic, ~str_wrap(., width = 25)),
         topic_label = paste0(topic, " (n=", total, ")"),
         topic_label = fct_reorder(topic_label, ratio)) |>
  ggplot(aes(x = topic_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary topic by first author gender")
```

## Primary method by first author gender

```{r fig.width=8}
# Count methods by occurrence by gender
method_count_gender <- df %>% 
  select(Authors..Genders..Author.1., contains("Methods")) %>% 
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>% 
  rename(gender_first_author = Authors..Genders..Author.1.)

# Counts of primary methods by gender
primary_method_gender <- method_count_gender %>%
  group_by(method, gender_first_author) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(method) %>% 
  mutate(prop_primary = n / sum(n))

primary_method_gender |>
  group_by(method) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(method = fct_reorder(method, ratio),
         method = fct_relabel(method, ~str_wrap(., width = 25)),
         method_label = paste0(method, " (n=", total, ")"),
         method_label = fct_reorder(method_label, ratio)) |>
  ggplot(aes(x = method_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary method by first author gender")
```

## Primary source of evidence by first author gender
```{r fig.width=8}
# Count sources by occurrence by gender
source_count_gender <- df %>% 
  select(Authors..Genders..Author.1., contains("Sources")) %>% 
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>% 
  rename(gender_first_author = Authors..Genders..Author.1.)

# Counts of primary sources by gender
primary_source_gender <- source_count_gender %>%
  group_by(source, gender_first_author) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(source) %>% 
  mutate(prop_primary = n / sum(n))

primary_source_gender |>
  group_by(source) |>
  mutate(ratio = n[gender_first_author == "Woman"] / sum(n),
         total = sum(n)) |>
  ungroup() |>
  mutate(source = fct_reorder(source, ratio),
         source = fct_relabel(source, ~str_wrap(., width = 25)),
         source_label = paste0(source, " (n=", total, ")"),
         source_label = fct_reorder(source_label, ratio)) |>
  ggplot(aes(x = source_label, y = prop_primary, fill = gender_first_author)) +
  geom_col(position = "fill", linewidth = 0, width = 1) +
  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = gender_colors) +
  coord_flip() +
  theme_bw() +
  force_panelsizes(cols = unit(10, "cm")) +
  labs(x = "", y = "Proportion", fill = NULL,
       title = "Primary source of evidence by first author gender")
```

## Region
```{r}
library(ggwordcloud)

region_counts <- df %>%
  pull(Region) %>%
  na.omit() %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  table() %>%
  as.data.frame() %>%
  setNames(c("Region", "n")) %>%
  mutate(n = as.numeric(n),
         n_scaled = n + 3) %>%
  arrange(desc(n))

ggplot(region_counts, aes(label = Region, size = n_scaled, color = n)) +
  geom_text_wordcloud(area_corr = TRUE, rm_outside = TRUE) +
  scale_size_area(max_size = 20) +
  scale_color_gradient(low = topic_col, high = "darkred") +
  theme_minimal()
```