# Basic descriptive statistics

```{r}
library(tidyverse)
library(patchwork)
library(nord)
library(here)
```

Set up some initial color palettes based on the nord package.

```{r}
custom_palette <- c(
  "#a84848", "#bf616a", "#d89898", "#d08770", "#c8a87a", "#ebcb8b", "#d8d8a0",
  "#a3be8c", "#6a9870", "#7eaaaa", "#90c0cc", "#6888b0","#8080b8", "#b48ead",
  "#d8b8d8","#b8b8a0", "#a0a8b8","#787868" 
)
custom_palette2 <- c(nord_palettes$aurora, "#8eaa9c", nord_palettes$baie_mouton)

# For maintaining consistent colors for plots with single categories
topic_col <- nord_palettes$algoma_forest[5]
method_col <- nord_palettes$victory_bonds[5]
source_col <- nord_palettes$red_mountain[1]

```

Read in the data.

```{r}
df <- read.csv(here("data/BibVik_Data_Entry_cite_keys.csv"), na.strings = "") 

# Set up years as factor variable for plotting, with empty level for 2003 where
# we don't have any publications
df <- df %>% 
  mutate(publication_year = as.numeric(str_extract_all(cite_key, "\\d+")))

df$publication_year <- factor(df$publication_year,
                              levels = as.character(seq(min(df$publication_year,
                                       na.rm = TRUE),
                                       max(df$publication_year, na.rm = TRUE), 1)))
```

## Intial plots

### Number of publications per year

```{r}
# Plot by number of publications per year
pubyears <- df %>% 
  count(publication_year) %>% 
  mutate(proportion_year = n / sum(n))

ggplot(data = pubyears, aes(x = publication_year, y = n)) +
  geom_bar(stat="identity", fill = nord_palettes$aurora[1]) +
  theme_bw() +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE ) +
  labs(x = "", y = "Count", title = "Publications by year")
```

## Gender by first author and all authors 

Publications by gender for each year. We also discussed looking at single author work and collaborations separately, but have not gotten around to doing that yet.

```{r}
# First author gender by year
gender_pubyears <- df %>% 
  group_by(Authors..Genders..Author.1.) %>% 
  count(publication_year) %>% 
  mutate(proportion_year = n / sum(n))

# Total proportions, first author gender
men_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Man",]) / nrow(gender_pubyears) * 100)
women_prcnt <- round(nrow(gender_pubyears[gender_pubyears$Authors..Genders..Author.1. == "Woman",]) / nrow(gender_pubyears) * 100)

plt1 <- 
ggplot(data = gender_pubyears, aes(x = publication_year, y = n, fill = Authors..Genders..Author.1.)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3,4)]) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "First author gender", title = paste0("First author gender: Women = ", women_prcnt, "% Men = ", men_prcnt, "%")) +
  theme(legend.position = "bottom")


# Repeat for all authors (author poistion 1-5)
all_authors_gender_year <- df %>% 
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(publication_year, gender) %>% 
  mutate(gender = na_if(gender, "")) %>% 
  drop_na()

# Proportion of author genders across all authors
all_authors_gender <- df %>% 
  pivot_longer(
    cols = starts_with("Authors.."),
    names_to  = "position",
    values_to = "gender"
  ) %>%
  count(gender) %>% 
  mutate(gender = na_if(gender, "")) %>% 
  drop_na()

# Men (55%)
men_aa <- round(all_authors_gender[all_authors_gender$gender == "Man", "n"] / sum(all_authors_gender$n) * 100)
# Women (43%)
women_aa <- round(all_authors_gender[all_authors_gender$gender == "Woman", "n"] / sum(all_authors_gender$n) * 100) 
# Uncertain (2%)
uncertain_aa <- round(all_authors_gender[all_authors_gender$gender == "Uncertain", "n"] / sum(all_authors_gender$n) * 100) 

plt2 <-
ggplot(data = all_authors_gender_year, aes(x = publication_year, y = n, fill = gender)) +
  geom_bar(stat="identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$victory_bonds[c(3, 1, 4)]) +
  scale_x_discrete(breaks = seq(2000, 2022, 1), drop = FALSE) +
  labs(x = "", y = "Count", fill = "Author gender",
       title = paste0("All authors gender: Uncertain = ", uncertain_aa, "% Women = ", women_aa, "% Men = ", men_aa, "%")) +
  theme(legend.position = "bottom")

plt1 / plt2
```

### Primary topic, in aggregate and over time

```{r}
# Count topics by occurrence by year
topic_count <- df %>% 
  select(publication_year, contains("Topics")) %>% 
  pivot_longer(!publication_year,
    names_to = "topic",
    values_to = "count")

# Counts of primary topics by year
primary_count_year <- topic_count %>%
  group_by(topic, publication_year) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(publication_year) %>% 
  mutate(prop_primary = n / sum(n))

# Plot by count and proportion by year
ggplot(primary_count_year, aes(x = publication_year, y = n, fill = topic)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  theme_bw() 
ggplot(primary_count_year, aes(x = publication_year, y = prop_primary, fill = topic)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Proportion by year", fill = "Primary topic") +
  scale_fill_manual(values = custom_palette) +
  theme_bw()

# Plot counts of primary topics
primary_count <- topic_count %>%
  group_by(topic) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) 

ggplot(primary_count, aes(x = topic, y = n)) +
  geom_bar(stat="identity", fill = topic_col)+
  labs(x = "Primary topic", y = "Count") +
  theme_bw() +
  coord_flip()
```

### Primary method, in aggregate and over time

```{r}
# Count methods by occurrence by year
method_count <- df %>% 
  select(publication_year, contains("Methods")) %>% 
  pivot_longer(!publication_year,
               names_to = "method",
               values_to = "count")

primary_method_year <- method_count %>%
  group_by(method, publication_year) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(publication_year) %>% 
  mutate(prop_primary = n / sum(n))

# Plot by count and proportion by year
ggplot(primary_method_year, aes(x = publication_year, y = n, fill = method)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  theme_bw() 
ggplot(primary_method_year, aes(x = publication_year, y = prop_primary, fill = method)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Proportion by year", fill = "Primary method") +
  scale_fill_manual(values = custom_palette2) +
  theme_bw()

# Plot counts of primary methods
primary_count_method <- method_count %>%
  group_by(method) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) 

ggplot(primary_count_method, aes(x = method, y = n)) +
  geom_bar(stat="identity", fill = method_col)+
  labs(x = "Primary method", y = "Count") +
  theme_bw() +
  coord_flip()
```

### Primary sources of evidence, in aggregate and over time

```{r}
source_count <- df %>% 
  select(publication_year, contains("Sources.of.Evidence")) %>% 
  pivot_longer(!publication_year,
               names_to = "source",
               values_to = "count")

primary_source_year <- source_count %>%
  group_by(source, publication_year) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(publication_year) %>% 
  mutate(prop_primary = n / sum(n))

# Plot by count and proportion by year
ggplot(primary_source_year, aes(x = publication_year, y = n, fill = source)) +
  geom_bar(stat="identity")+
  labs(x = "", y = "Count", fill = "Primary source of evidence") +
  # scale_fill_nord("aurora") +
  scale_fill_manual(values = custom_palette) +
  theme_bw() +
  theme(legend.position = "bottom") 
ggplot(primary_source_year, aes(x = publication_year, y = prop_primary, fill = source)) +
  geom_bar(stat="identity") +
  labs(x = "", y = "Proportion by year", fill = "Primary source of evidence") +
  scale_fill_manual(values = custom_palette) +
  theme_bw()+
  theme(legend.position = "none")

# Plot counts of primary sources of evidence
primary_source_count <- source_count %>%
  group_by(source) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " "))

ggplot(primary_source_count, aes(x = source, y = n)) +
  geom_bar(stat="identity", fill = source_col)+
  labs(x = "Primary source of evidence", y = "Count") +
  theme_bw() +
  coord_flip()
```

### Seconday topics by primary topic

```{r}
# Select only the topics columns, and strip the "Topics.." from each column name
topics <- df %>% 
  select(contains("Topics")) %>% 
  rename_with(~str_remove(., "Topics..")) 

# Gave up on doing this with tidyverse
coocurrences <- c()

# For each row
for(i in 1:nrow(topics)) {
  # Select row and keep only values with either "Primary" or "Secondary in topics columns
  current_row <- topics[i, !is.na(topics[i,])]
  # Retrieve names of primary topics for this row
  current_primaries <- names(current_row[which(current_row == "Primary")])
  # Retrieve names of secondary topics for this row
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  # Combine each secondary topic with each primary topic in the row
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  # Append to results vector
  coocurrences <- append(coocurrences, current_row)
}

# Set up data frame counting unique combinations
# (number of times each secondary topic occurs on the same row as each primary topic)
coocdf <- data.frame(table(coocurrences)) %>% 
  # Remove all instances of there only being a primary topic 
  # (starts with "_" as introduced with the paste() above)
  filter(substr(coocurrences, 1, 1) != "_") %>% 
  # Separate secondary from primary topics into one column each
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>% 
  # Clean up names for plotting
  mutate(Secondary = str_replace_all(Secondary, fixed("."), " ")) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>% 
  # Find proportion of each secondary topics for each primary topic
  group_by(Primary) %>% 
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = Secondary, y = Prop)) +
  geom_bar(stat = "identity", fill = topic_col) +
  facet_wrap(vars(Primary)) +
  coord_flip() +
  theme_bw() +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary topic by primary topic")
```

### Secondary methods by primary method

```{r}
# Repeats code from secondary topics above
methods <- df %>% 
  select(contains("Methods")) %>% 
  rename_with(~str_remove(., "Methods..")) 

coocurrences <- c()

for(i in 1:nrow(methods)) {
  current_row <- methods[i, !is.na(methods[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>% 
  filter(substr(coocurrences, 1, 1) != "_") %>% 
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>% 
  mutate(Secondary = str_replace_all(Secondary, fixed("."), " ")) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>% 
  group_by(Primary) %>% 
  mutate(Prop = Freq / sum(Freq))
 
ggplot(data = coocdf, aes(x = Secondary, y = Prop)) +
  geom_bar(stat = "identity", fill = method_col) +
  facet_wrap(vars(Primary)) +
  coord_flip() +
  theme_bw() +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary method by primary method")

```

### Seconday sources of evidence by primary source of evidence

```{r}
# Repeats code from secondary topics and secondary methods above
sources <- df %>% 
  select(contains("Sources.of.Evidence")) %>% 
  rename_with(~str_remove(., "Sources.of.Evidence..")) 

coocurrences <- c()

for(i in 1:nrow(sources)) {
  current_row <- sources[i, !is.na(sources[i,])]
  current_primaries <- names(current_row[which(current_row == "Primary")])
  current_secondaries <- names(current_row[which(current_row == "Secondary")])
  current_row <- paste(rep(current_secondaries, each = length(current_primaries)), current_primaries, sep = "_")
  coocurrences <- append(coocurrences, current_row)
}

coocdf <- data.frame(table(coocurrences)) %>% 
  filter(substr(coocurrences, 1, 1) != "_") %>% 
  separate_wider_delim(coocurrences, "_", names = c("Secondary", "Primary")) %>% 
  mutate(Secondary = str_replace_all(Secondary, fixed("."), " ")) %>%
  mutate(Primary = str_replace_all(Primary, fixed("."), " ")) %>% 
  group_by(Primary) %>% 
  mutate(Prop = Freq / sum(Freq))

ggplot(data = coocdf, aes(x = Secondary, y = Prop)) +
  geom_bar(stat = "identity", fill = source_col) +
  facet_wrap(vars(Primary)) +
  coord_flip() +
  theme_bw() +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "", title = "Proportion of secondary source of evidence by primary source of evidence")
```

### Primary topic by first author gender

For the subsequent three plots we could also try normalizing the proportions for each topic against the total proportion men/women in the entire corpus.

```{r}
# Count topics by occurrence by gender
topic_count_gender <- df %>% 
  select(Authors..Genders..Author.1., contains("Topics")) %>% 
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "topic",
               values_to = "count") %>% 
  rename(gender_first_author = Authors..Genders..Author.1.)

# Counts of primary topics by gender
primary_topic_gender <- topic_count_gender %>%
  group_by(topic, gender_first_author) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(topic = str_replace_all(str_remove(topic, "Topics.."), fixed("."), " ")) %>%
  group_by(topic) %>% 
  mutate(prop_primary = n / sum(n))

## Commented out the raw counts versions for now
# ggplot(data = primary_topic_gender, aes(x = topic, y = n, fill = gender_first_author)) +
#   geom_bar(stat = "identity", position = 'dodge') +
#   theme_bw() +
#   scale_fill_manual(values = c(nord_palettes$red_mountain[1], nord_palettes$algoma_forest[4])) +
#   labs(x = "Primary topic", y = "Count", fill = "First author gender") +
#   theme(legend.position = "bottom") +
#   coord_flip() +
ggplot(data = primary_topic_gender, aes(x = topic, y = prop_primary, fill = gender_first_author)) +
  geom_bar(stat = "identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = c(nord_palettes$red_mountain[1], nord_palettes$algoma_forest[4])) +
  labs(x = "Primary topic", y = "Proportion", fill = "First author gender") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Primary method by first author gender

```{r}
method_count_gender <- df %>% 
  select(Authors..Genders..Author.1., contains("Methods")) %>% 
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "method",
               values_to = "count") %>% 
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_method_gender <- method_count_gender %>%
  group_by(method, gender_first_author) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(method = str_replace_all(str_remove(method, "Methods.."), fixed("."), " ")) %>%
  group_by(method) %>% 
  mutate(prop_primary = n / sum(n))

# ggplot(data = primary_method_gender, aes(x = method, y = n, fill = gender_first_author)) +
#   geom_bar(stat = "identity", position = 'dodge') +
#   theme_bw() +
#   scale_fill_manual(values = nord_palettes$baie_mouton[c(6,1)]) +
#   labs(x = "Primary method", y = "Count", fill = "First author gender") +
#   theme(legend.position = "bottom") +
#   coord_flip() +
ggplot(data = primary_method_gender, aes(x = method, y = prop_primary, fill = gender_first_author)) +
  geom_bar(stat = "identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$baie_mouton[c(6,1)]) +
  labs(x = "Primary method", y = "Proportion", fill = "First author gender") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Primary source of evidence by first author gender
```{r}
source_count_gender <- df %>% 
  select(Authors..Genders..Author.1., contains("Sources")) %>% 
  pivot_longer(!Authors..Genders..Author.1.,
               names_to = "source",
               values_to = "count") %>% 
  rename(gender_first_author = Authors..Genders..Author.1.)

primary_source_gender <- source_count_gender %>%
  group_by(source, gender_first_author) %>% 
  count(count) %>% 
  filter(count == "Primary") %>% 
  mutate(source = str_replace_all(str_remove(source, "Sources.of.Evidence.."), fixed("."), " ")) %>%
  group_by(source) %>% 
  mutate(prop_primary = n / sum(n))

# ggplot(data = primary_source_gender, aes(x = source, y = n, fill = gender_first_author)) +
#   geom_bar(stat = "identity", position = 'dodge') +
#   theme_bw() +
#   scale_fill_manual(values = nord_palettes$halifax_harbor[c(2,4)]) +
#   labs(x = "Primary source", y = "Count", fill = "First author gender") +
#   theme(legend.position = "bottom") +
#   coord_flip() +
ggplot(data = primary_source_gender, aes(x = source, y = prop_primary, fill = gender_first_author)) +
  geom_bar(stat = "identity", position = 'dodge') +
  theme_bw() +
  scale_fill_manual(values = nord_palettes$halifax_harbor[c(2,4)]) +
  labs(x = "Primary source", y = "Proportion", fill = "First author gender") +
  theme(legend.position = "bottom") +
  coord_flip()
```

